### 2019-02-27 manuscript/main.tex
GM-HMM

\documentclass{article}

\usepackage{hyperref} %BUG if put after
\usepackage{tikz}
\usetikzlibrary{calc,shapes,positioning}
\usetikzlibrary{arrows}
\newcommand{\midarrow}{\tikz \draw[-triangle 90] (0,0) -- +(.1,0);}
% Be sure to use PDxF Latex
\pdfoutput=1

\usepackage[latin1]{inputenc}

\usepackage{url}
\usepackage{fullpage}
\usepackage{cite}
\usepackage{caption}
\usepackage{bm}
\newcommand{\ubar}[1]{\mkern2mu\underline{\mkern-2mu #1\mkern-2mu}\mkern2mu}
% \allowdisplaybreaks
\usepackage{mystyle}
\newcommand{\ubm}[1]{\ubar{\bm{#1}}}
\newcommand{\ubmr}[2]{\ubar{\bm{#1}}^{(#2)}}

% \newcommand{\bmtr}[3]{\bm{#1}^{(#3)}_{#2}}

\newcommand{\bmtr}[3]{\bm{#1}^{(#3)}_{#2}}
\newcommand{\smtr}[3]{{#1}^{(#3)}_{#2}}

\usepackage{amsmath,graphicx}
% format A4
% \usepackage{vmargin}
% \setpapersize{A4}

\graphicspath{{./images/}}

\hypersetup{  
  bookmarks=true,
  backref=true,
  pagebackref=false,
  colorlinks=true,
  linkcolor=blue,
  citecolor=red,
  urlcolor=blue,
  pdftitle={Generative Model HMM},
  pdfauthor={Dong Liu},
  pdfsubject={}
}

\title{Powering Hidden Markov Model by Generative Models}
\author{
  Firsthand Scientists
}

% \name{
% Dong Liu,
% Minh Thành Vu,
% Saikat Chatterjee,
% and Lars~K. Rasmussen
% }

%   \address{
%   KTH Royal Institute of Technology, Stockholm, Sweden \\
%   E-mail: \{doli, mtvu, sach, lkra\}@kth.se}

\begin{document}

\maketitle
\section{Notation}
Time is indexed by subscript and sequence is denoted by underline. $\bm{x}_t$ is signal at time $t$. The sequential time is denoted by $\ubar{\bm{x}} = \left[ \bm{x}_1,
  \cdots, \bm{x}_T\right]^{\intercal}$, where $[\cdot]^{\intercal}$ means transpose and $T$ is the length of the sequence.  Sequential signal or clip uses underline notation and is indexed by superscript, for instance $\ubar{\bm{x}}^{(r)}$ means the $r$-th sequential signal, where $r = 1, 2, \cdots, R.$, and $\ubmr{x}{r} = \left[ \bmtr{x}{1}{r}, \bmtr{x}{2}{r}, \cdots, \bmtr{x}{T^{(r)}}{r} \right]$ with length ${T^{(r)}}$. Note different sequential signal $\ubmr{x}{r}$ could have different lengths.

The hypothesis of Hidden Markov Model (HMM): $\Hh := \{\bm{H} | \{\Ss, \bm{q}, A, p(\ubar{\bm{x}}|\ubar{\bm{s}}; \bm{\Phi})\}$,
\begin{itemize}
\item $\Ss$ is the set of states of HMM $\bm{H}$;
\item $\bm{q} = \left[ q_1, q_2, \cdots, q_{|\Ss|}\right]^\intercal$ initial distribution of HMM $\bm{H}$ with $|\Ss|$ is cardinality of $\Ss$, $q_k = p(s=k)$ for random state variable $s$.
\item $A$ is the transition matrix for the HMM $\bm{H}$ of size $|\Ss| \times |\Ss|$.
\item Observable signal density $p(\ubar{\bm{x}}|\ubar{\bm{s}};\bm{\Phi})$ given hidden state sequence, where $\bm{\Phi}$ is the parameter set that defines this conditional probabilistic model.
\end{itemize}

\section{Problem Statement}
Given a empirical distribution $\hat{p}(\ubm{x}) = \frac{1}{R}\sum_{r=1}^{R} \delta_{\ubmr{x}{r}}(\ubm{x})$. We want to find a probabilistic model such that:
\begin{equation}
  \min KL(\hat{p}(\ubm{x})\| p(\ubm{x}))
\end{equation}
where $KL(\cdot\|\cdot)$ denotes the Kullback-Leibler divergence.

When we use HMM to model the empirical distribution and approach the unknown true distribution, the problem boils down to:
\begin{equation}
  \uargmax{\bm{H} \in \Hh} p(\ubm{X}; \bm{H})
\end{equation}
where $\ubm{X} = \left[ \ubmr{x}{1}, \ubmr{x}{2}, \cdots, \ubmr{x}{R} \right]$

The problem can be reformulated as
\begin{equation}\label{eq:ml-of-hmm}
  \uargmax{\bm{H} \in \Hh} \sum_{r=1}^{R}\log\,p(\ubmr{x}{r}; \bm{H})
\end{equation}
for independent identical distributed assumption of $\ubm{x}$.

\section{Proposal}

Since model $\bm{H}$ contains hidden sequential variable $\ubm{s}$, we can not directly solve the maximum likelihood problem in $\autoref{eq:ml-of-hmm}$. We use expectation maximization (EM) to address the hidden variable problem by
\begin{itemize}
\item E-step: % the posterior probability of $\ubm{s}$:
  % \begin{equation}
  %   p(\ubm{s}|\ubm{x})
  % \end{equation}
  The ``expected likelihood'' function:
  \begin{equation}\label{eq:em-q-funciton}
    \Qq(\bm{H}; \bm{H}^{\mathrm{old}}) = \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}\left[ \sum_{r=1}^{R}\log\,p(\ubmr{x}{r}, \ubmr{s}{r}; \bm{H}) \right]
  \end{equation}
\item M-step: the optimization step:
  \begin{equation}\label{eq:em-m-opt}
    \umax{\bm{H}} \Qq(\bm{H}; \bm{H}^{\mathrm{old}})
  \end{equation}
\end{itemize}

The \autoref{eq:em-m-opt} can be reformulated as:
\begin{equation}\label{eq:m-step-subs}
  \umax{\bm{H}} \Qq(\bm{H}; \bm{H}^{\mathrm{old}}) = \umax{\bm{q}}\Qq(\bm{q}; \bm{H}^{\mathrm{old}}) + \umax{{A}}\Qq({A}; \bm{H}^{\mathrm{old}}) + \umax{\bm{\Phi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})
\end{equation}
where
\begin{align}
  \Qq(\bm{q}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,p(\smtr{s}{1}{r}; \bm{q})\right] \label{eq:init-distribution-update}\\
  \Qq({A}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})\right] \label{eq:transition-update}\\
  \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,p(\ubmr{x}{r} | \ubmr{s}{r}; \bm{\Phi})\right]\label{eq:generative-model-update}
\end{align}

We can see that the solution of $H$ depends on the posterior probability $p(\ubm{s}| \ubm{x}; \bm{H})$. Though the evaluation of posterior according to Bayesian theorem is simple, the computation complexity of $p(\ubm{s}| \ubm{x}; \bm{H})$ grows exponentially with the length of $\ubm{s}$. Therefore, we would employ Forward/Backward algorithm \cite{} to do the posterior computation efficiently. The marginal $p(s_t| \ubm{x}; \bm{H})$ is also efficiently computed as the joint posterior.

\subsection{Initial Probability Update}
\autoref{eq:init-distribution-update} can be written as:
\begin{align}
  \Qq(\bm{q}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R}\sum_{\ubmr{s}{r}} {p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r}; \bm{q}) \nonumber \\
                                     & = \sum_{r=1}^{R}\sum_{\smtr{s}{1}{r}=1}^{|\Ss|}\sum_{\smtr{s}{2}{r}=1}^{|\Ss|}\cdots \sum_{\smtr{s}{T^{r}}{r}}^{{|\Ss|}} {p(\smtr{s}{1}{r}, \smtr{s}{2}{r}, \cdots, \smtr{s}{T^{r}}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r}; \bm{q}) \\
                                     & = \sum_{r=1}^{R}\sum_{\smtr{s}{1}{r}=1}^{|\Ss|}{p(\smtr{s}{1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r}; \bm{q}) 
\end{align}

Since $p(\smtr{s}{1}{r}; \bm{H})$ is the probability of initial state of HMM $\bm{H}$ for $r$-th sequential, actually $q_i = p(\smtr{s}{1}{r} =i;\bm{H} ) $ for $i= 1, 2, \cdots, |\Ss|$. Solution to problem:
\begin{align}
  \bm{q}^{\mathrm{new}} &= \uargmax{\bm{q}} \Qq(\bm{q}; \bm{H}^{\mathrm{old}}), \nonumber \\
                        &\mathrm{s.t.} \sum_{i=1}^{ |\Ss| }q_i = 1 \nonumber\\
  &q_i \geq 0, \forall s.
\end{align}
is
\begin{equation}
  q_i = \frac{1}{R} \sum_{r=1}^{R} p(\smtr{s}{1}{r}=i | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}), \forall\; i = 1, 2, \cdots, |\Ss|.
\end{equation}

\subsection{Transition Probability Update}
\autoref{eq:transition-update} can be written as
\begin{align}
  \Qq({A}; \bm{H}^{\mathrm{old}})
  &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})\right] \nonumber\\
  &= \sum_{r=1}^{R} \sum_{\ubmr{s}{r}}{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A}) \nonumber \\
  &= \sum_{r=1}^{R} \sum_{t=1}^{T^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}\sum_{\smtr{s}{t+1}{r}=1}^{|\Ss|}{p(\smtr{s}{t}{r}, \smtr{s}{t+1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})
\end{align}

Since $A_{i, j}  = p(\smtr{s}{t+1}{r}=j|\smtr{s}{t}{r}=i; {A})$ where $A_{i, j}$ is the element of transition matrix $A$, the solution to problem:
\begin{align}
  {A}^{\mathrm{new}} = &\uargmax{{A}} \Qq({A}; \bm{H}^{\mathrm{old}}), \nonumber \\
  \mathrm{s.t.} &\hspace{0.2cm} A \cdot \bm{1} = \bm{1} \nonumber \\
                       &  A^{\intercal} \cdot \bm{1} = \bm{1} \nonumber \\
  & A_{i,j} \geq 0.
\end{align}
is
\begin{equation}
  A_{i,j}^{\mathrm{new}} = \frac{\bar{\xi}_{i,j}}{\sum_{k = 1}^{|\Ss|} \bar{\xi}_{i,k}},
\end{equation}
where
\begin{equation}
  \bar{\xi}_{i,j} = \sum_{r= 1}^{R} \sum_{t= 1}^{T^{(r)}-1}{p(\smtr{s}{t}{r}=i, \smtr{s}{t+1}{r}=j| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}
\end{equation}

\subsection{Generative Model Update}
\autoref{eq:generative-model-update} can be rewritten as
\begin{align}\label{eq:gm-update}
  \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})
  &= \sum_{r=1}^{R}\sum_{\ubmr{s}{r}}{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\ubmr{x}{r} | \ubmr{s}{r}; \bm{\Phi}) \nonumber \\
  &= \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \log\, p(\bmtr{x}{t}{r} | \smtr{s}{t}{r}; \bm{\Phi}).
\end{align}

Then the third subproblem of \autoref{eq:m-step-subs} becomes:
\begin{align}\label{eq:sub-gm}
  &\uargmax{\bm{\Phi}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}), \nonumber \\
  \mathrm{s.t.} &\hspace{0.2cm} p(\bm{x} | {s}; \bm{\Phi}) \mathrm{~is~our~ general~model}
\end{align}

It could be seen from \autoref{eq:gm-update} that the key to update generate  model is to evaluate $p(\bm{x}|{s}; \bm{\Phi})$ for all $s \in \Ss$. In Forward/Backward algorithm, evaluation of $p(\bm{x}|{s}; \bm{\Phi})$ is also all what is needed to compute $p(s|\bm{x};\bm{\Phi})$. In the following two subsections, we will provide two neural network based generative models that fulfill this requirement and also have high capability for complex signal modeling.

\subsubsection{Generator Mixed HMM (GenM-HMM)}
\begin{figure}[!h]
  \centering
  \begin{tikzpicture}
    \tikzstyle{enode} = [thick, draw=blue, ellipse, inner sep = 1pt,  align=center]
    % \tikzstyle{nnode} = [thick, rectangle, rounded corners = 1pt,minimum size = 1.2cm,draw,inner sep = 2pt]
    \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]
    \node[enode] (z1) at (-1.2,1.8) {$\bm{z}\sim p_1(\bm{z})$};
    \node[nnode] (g1) at (1,1.8) {$\bm{g}_1$};

    \node[enode] (z2) at (-1.2,0.5){$\bm{z}\sim p_{2}(\bm{z})$};
    \node[nnode] (g2) at (1,0.5) {$\bm{g}_2$};

    \node[enode] (zK) at (-1.2,-1.8) {$\bm{z}\sim p_{|\Ss|}(\bm{z})$};
    \node[nnode] (gs) at (1, -1.8) {$\bm{g}_{|\Ss|}$};

    \node[enode] (x) at (4.5,1.5){$\ubm{x}\sim p(\ubm{x};\bm{H})$};
    % \node at (3,-1){$p(\bm{z}) = \sum_k\pi_kp_k(\bm{z})$};

    \draw[dotted,line width=2pt] (0,-0.3) -- (0,-1.2);
    \filldraw[->] (1.9, 0.5)circle (2pt) --  (x) ;
    \draw[->] (z1) -- (g1);
    \draw[->] (g1) -- (1.8, 1.8);

    \draw[->] (z2) -- (g2);
    \draw[->] (g2) -- (1.8, 0.5);

    % \draw[->] (1,-0.5) -- (1.8, -0.5);
    \draw[->] (zK) -- (gs);
    \draw[->] (gs) -- (1.8, -1.8);

    % \draw[->] (z1) [in=180,out=0]  to node[right]{$\pi_1$} (g);
    % \draw[->] (z2) [in=180,out=0]  to node[above]{$\pi_2$} (g);
    % \draw[->,dashed] (zK) [in=180,out=0]  to node[right]{$\pi_K$} (g);
    %\draw[->] (g) to (x);
    \begin{scope}[xshift=0.5cm, thick, every node/.style={sloped,allow upside down}]
      \node[nnode] (m) at (3.5,-2) {Memory};
      \node[nnode] (a) at (3.5,-0.5) {$A$};

      \draw (2.1,0.9)-- (2.2, 0.);
      \draw (2.2,0.)-- node {\midarrow} (2.2,-2);
      \draw (2.2,-2)-- (m);
      \draw (m)-- (5, -2);
      \draw (5, -2)-- node {\midarrow} (5 ,-0.5);
      \draw (5, -0.5) -- (a);
      \draw (a)-- node {\midarrow} (2.2, -0.5);
      \node at (4.8, -1) {$s_{t}$};
      \node at (2.56, -0.25) {$s_{t+1}$};
    \end{scope}

  \end{tikzpicture}
  \caption{GenM-HMM Model defined by $\bm{H} = \{\Ss, \bm{q}, A, p({\bm{x}}|{\bm{s}}; \bm{\Phi}) \}$}
  \vspace{0.1cm}
\end{figure}

For this proposal, we seek to use a generator mixed HMM scheme, termed as GenM-HMM. We define a set of generators for GenM-HMM:
\begin{equation}
  \{\bm{g}_{s}| s \in \Ss, \bm{g}_s: \bm{z}\rightarrow \bm{x}, \bm{z}\sim p_{s}(\bm{z})\}.
\end{equation}
Thus there are total $|\Ss|$ generators. $p(\bm{x}|{s}; \bm{\Phi})$ is induced as $\bm{g}_s(\bm{z}) \sim p(\bm{x}|{s}; \bm{\Phi})$ where $\bm{z} \sim p_s(\bm{z})$ for $s \in \Ss$. Let us denote the inverse of $\bm{g}_s$ as $\bm{f}_s = \bm{g}_s^{-1}$. We have the $s$-th component of the GenM-HMM model as
\begin{align}
  p(\bm{x} | s; \bm{\Phi})
  &= p_s(\bm{z}) \bigg| \det \left( \pd{\bm{z}}{\bm{x}} \right) \bigg| \nonumber \\
  &= p_s(\bm{f}_s(\bm{x})) \bigg| \det \left( \pd{\bm{f}_s(\bm{x})}{\bm{x}} \right) \bigg|
\end{align}
where $p_s(\bm{z})$ is the latent source distribution for $s=1, 2, \cdots, |\Ss|$.

Let us denote the parameter set that defines latent distribution $p_s(z)$ by $\bm{\omega}_s$ and the parameter set that defines generator $\bm{g}_s$ by $\bm{\theta}_s$. Then $\bm{\Phi} = \left\{ \bm{\theta}_s,
  \bm{\omega}_s, \forall s\in \Ss\right\}$. The problem in \autoref{eq:sub-gm} can be reformulated as:
\begin{align}
  &\umax{\bm{\Phi}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) \nonumber \\
  =&\umax{ \bm{\theta}_s, \bm{\omega}_s, \forall s\in \Ss} \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \left[\log\,p_{\smtr{s}{t}{r}}(\bm{f}_{\smtr{s}{t}{r}}(\bmtr{x}{t}{r})) + \log\,\bigg| \det\, \left( \pd{\bm{f}_{\smtr{s}{t}{r}}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg| \right].
\end{align}

The diagram of GenM-HMM is shown as follows.

\subsubsection{Latent-source Mixed HMM (LatM-HMM)}
\begin{figure}[!h]
  \centering
  \begin{tikzpicture}
    \tikzstyle{enode} = [thick, draw=blue, ellipse, inner sep = 1pt,  align=center]
    % \tikzstyle{nnode} = [thick, rectangle, rounded corners = 1pt,minimum size = 1.2cm,draw,inner sep = 2pt]
    \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]
    \node[enode] (z1) at (0,1.8) {$\bm{z}\sim p_1(\bm{z})$};
    \node[enode] (z2) at (0,0.5){$\bm{z}\sim p_{2}(\bm{z})$};
    \node[enode] (zK) at (0,-1.8) {$\bm{z}\sim p_{|\Ss|}(\bm{z})$}; 
    \node[enode] (x) at (5.8,1){$\ubm{x}\sim p(\ubm{x};\bm{H})$};
    % \node at (3,-1){$p(\bm{z}) = \sum_k\pi_kp_k(\bm{z})$};
    \node[nnode] (g) at (3.5,1) {$\bm{g}$};

    \node[nnode] (m) at (3.5,-2) {Memory};
    \node[nnode] (a) at (3.5,-0.5) {$A$};

    \draw[dotted,line width=2pt] (0,-0.3) -- (0,-1.2);
    \filldraw[->] (1.8, 0.5)circle (2pt) --  (g) ;
    \draw[->] (z1) -- (1.6, 1.8);
    \draw[->] (z2) -- (1.6, 0.5);
    % \draw[->] (1,-0.5) -- (1.8, -0.5);
    \draw[->] (zK) -- (1.6, -1.8);
    % \draw[->] (z1) [in=180,out=0]  to node[right]{$\pi_1$} (g);
    % \draw[->] (z2) [in=180,out=0]  to node[above]{$\pi_2$} (g);
    % \draw[->,dashed] (zK) [in=180,out=0]  to node[right]{$\pi_K$} (g);
    \draw[->] (g) to (x);
    \begin{scope}[ thick, every node/.style={sloped,allow upside down}]
      \draw (2.1,0.9)-- (2.2, 0.);
      \draw (2.2,0.)-- node {\midarrow} (2.2,-2);
      \draw (2.2,-2)-- (m);
      \draw (m)-- (5, -2);
      \draw (5, -2)-- node {\midarrow} (5 ,-0.5);
      \draw (5, -0.5) -- (a);
      \draw (a)-- node {\midarrow} (2.2, -0.5);
    \end{scope}
    \node at (4.8, -1) {$s_{t}$};
    \node at (2.56, -0.25) {$s_{t+1}$};

  \end{tikzpicture}
  \caption{LatM-HMM Model defined by $\bm{H} = \{\Ss, \bm{q}, A, p({\bm{x}}|{\bm{s}}; \bm{\Phi}) \}$}
  \vspace{0.1cm}
\end{figure}

Alternatively, we can use a latent-source mixed HMM (LatM-HMM) where different latent source share the same generator functioning as feature mapping. Then the generator of the LatM-HMM is defined as
\begin{equation}
  \left\{\bm{g}| \bm{g}: \bm{z}\rightarrow \bm{x}, s \in \Ss, \bm{z}\sim p_{s}(\bm{z})\right\}.
\end{equation}
We use $\bm{f} = \bm{g}^{-1}$ to denote inverse of $\bm{g}$ and use $\bm{\theta}$ to denote the parameter set of $\bm{g}$. Then the conditional probability for LatM-HMM is modeled as
\begin{align}
  p(\bm{x} | s; \bm{\Phi})
  &= p_s(\bm{z}) \bigg| \det \left( \pd{\bm{z}}{\bm{x}} \right) \bigg| \nonumber \\
  &= p_s(\bm{f}(\bm{x})) \bigg| \det \left( \pd{\bm{f}(\bm{x})}{\bm{x}} \right) \bigg|
\end{align}

The parameter set for this model to be decide is $\bm{\Phi} = \left\{ \bm{\theta},  \bm{\omega}_s, \forall s\in \Ss\right\}$.
Then the problem in \autoref{eq:sub-gm} can be reformulated as:
\begin{align}
  &\umax{\bm{\Phi}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) \nonumber \\
  =&\umax{ \bm{\theta}, \bm{\omega}_s, \forall s\in \Ss} \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \left[\log\,p_{\smtr{s}{t}{r}}(\bm{f}(\bmtr{x}{t}{r})) + \log\,\bigg| \det\, \left( \pd{\bm{f}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg| \right].
\end{align}

\textbf{To be continued}...
\section{On Implementation}
Found a HMM python lib that basics provide needed API for us, see \href{https://hmmlearn.readthedocs.io/en/stable/index.html}{hmmlearn}. Saikat also has suggestion.

For problem \autoref{eq:sub-gm} we are going to use our generative models to solve. I have the following consideration to revised our LatMM and GenMM for this application:
\begin{itemize}
\item Use factorized model instead of additive mixture model, to make likelihood computation logarithm domain compatible; 
\item Use full EM fashion instead of mini-batch fashion for training: store generative model as old for EM, there are always two neural networks working, one old for probability evaluation and one new for optimization.
\end{itemize}
\bibliographystyle{IEEEbib}

% \bibliography{bibliography}
\bibliography{myref}

% \appendix

% \input{section/sec-appendixA}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2019-02-27 code/glow/builder.py
GenM and LatM

@@ -1,0 +1,124 @@
+import re, os
+import copy
+import torch
+from collections import defaultdict
+from . import learning_rate_schedule
+from .config import JsonConfig
+from .models import GenMM, LatMM
+from .utils import load, save, get_proper_device
+
+
+def build_adam(params, args):
+    return torch.optim.Adam(params, **args)
+
+
+__build_optim_dict = {
+    "adam": build_adam
+}
+
+# custom weights initialization called on netG and netD
+def weights_init(m):
+    classname = m.__class__.__name__
+    if classname.find('Conv') != -1:
+        pass
+        #m.weight.data.normal_(0.0, 0.01)
+    elif classname.find('BatchNorm') != -1:
+        m.weight.data.normal_(1.0, 0.02)
+        m.bias.data.fill_(0)
+
+
+def build(hparams, is_training):
+    if isinstance(hparams, str):
+        hparams = JsonConfig(hparams)
+    # get graph and criterions from build function
+    graph, optim, lrschedule, criterion_dict = None, None, None, None  # init with None
+    graph_prior = None
+    cpu, devices = "cpu", None
+    get_loss = None
+    # 1. build graph and criterion_dict, (on cpu)
+    # build and append `device attr` to graph
+    if hparams.Mixture.naive:
+        graph = GenMM(hparams)
+        
+    else:
+        graph = LatMM(hparams)
+        
+        
+    graph.device = hparams.Device.glow
+    if graph is not None:
+        # get device
+        devices = get_proper_device(graph.device)
+        graph.device = devices
+        graph.to(cpu)
+    # 2. get optim (on cpu)
+    try:
+        if graph is not None and is_training:
+            optim_name = hparams.Optim.name
+            optim = __build_optim_dict[optim_name](graph.parameters(), hparams.Optim.args.to_dict())
+            print("[Builder]: Using optimizer `{}`, with args:{}".format(optim_name, hparams.Optim.args))
+            # get lrschedule
+            schedule_name = "default"
+            schedule_args = {}
+            if "Schedule" in hparams.Optim:
+                schedule_name = hparams.Optim.Schedule.name
+                schedule_args = hparams.Optim.Schedule.args.to_dict()
+            if not ("init_lr" in schedule_args):
+                schedule_args["init_lr"] = hparams.Optim.args.lr
+            assert schedule_args["init_lr"] == hparams.Optim.args.lr,\
[…]

### 2019-02-27 code/glow/models.py

@@ -1,0 +1,521 @@
+import torch
+import torch.nn as nn
+import torch.nn.functional as F
+import numpy as np
+from tqdm import tqdm
+from . import thops
+from . import modules
+from . import utils
+from scipy import stats
+
+def prior_sampler(pk, size=64):
+        pk = tuple(pk.numpy())
+        xk = np.arange(len(pk))
+        custm = stats.rv_discrete(name='custm', values=(xk, pk))
+        prior_samples  = custm.rvs(size=size)
+        prior_samples = torch.from_numpy(prior_samples)
+        return prior_samples
+    
+
+
+def f(in_channels, out_channels, hidden_channels):
+    return nn.Sequential(
+        modules.Conv2d(in_channels, hidden_channels), nn.ReLU(inplace=False),
+        modules.Conv2d(hidden_channels, hidden_channels, kernel_size=[1, 1]), nn.ReLU(inplace=False),
+        modules.Conv2dZeros(hidden_channels, out_channels))
+
+
+class FlowStep(nn.Module):
+    FlowCoupling = ["additive", "affine"]
+    FlowPermutation = {
+        "reverse": lambda obj, z, logdet, rev: (obj.reverse(z, rev), logdet),
+        "shuffle": lambda obj, z, logdet, rev: (obj.shuffle(z, rev), logdet),
+        "invconv": lambda obj, z, logdet, rev: obj.invconv(z, logdet, rev)
+    }
+
+    def __init__(self, in_channels, hidden_channels,
+                 actnorm_scale=1.0,
+                 flow_permutation="invconv",
+                 flow_coupling="additive",
+                 LU_decomposed=False):
+        # check configures
+        assert flow_coupling in FlowStep.FlowCoupling,\
+            "flow_coupling should be in `{}`".format(FlowStep.FlowCoupling)
+        assert flow_permutation in FlowStep.FlowPermutation,\
+            "float_permutation should be in `{}`".format(
+                FlowStep.FlowPermutation.keys())
+        super().__init__()
+        self.flow_permutation = flow_permutation
+        self.flow_coupling = flow_coupling
+        # 1. actnorm
+        self.actnorm = modules.ActNorm2d(in_channels, actnorm_scale)
+        # 2. permute
+        if flow_permutation == "invconv":
+            self.invconv = modules.InvertibleConv1x1(
+                in_channels, LU_decomposed=LU_decomposed)
+        elif flow_permutation == "shuffle":
+            self.shuffle = modules.Permute2d(in_channels, shuffle=True)
+        else:
[…]

### 2019-02-27 code/glow/modules.py

@@ -1,0 +1,372 @@
+import torch
+import torch.nn as nn
+import torch.nn.functional as F
+import numpy as np
+import scipy.linalg
+from . import thops
+
+
+class _ActNorm(nn.Module):
+    """
+    Activation Normalization
+    Initialize the bias and scale with a given minibatch,
+    so that the output per-channel have zero mean and unit variance for that.
+
+    After initialization, `bias` and `logs` will be trained as parameters.
+    """
+
+    def __init__(self, num_features, scale=1.):
+        super().__init__()
+        # register mean and scale
+        size = [1, num_features, 1, 1]
+        self.register_parameter("bias", nn.Parameter(torch.zeros(*size)))
+        self.register_parameter("logs", nn.Parameter(torch.zeros(*size)))
+        self.num_features = num_features
+        self.scale = float(scale)
+        self.inited = False
+
+    def _check_input_dim(self, input):
+        return NotImplemented
+
+    def initialize_parameters(self, input):
+        self._check_input_dim(input)
+        if not self.training:
+            return
+        assert input.device == self.bias.device
+        with torch.no_grad():
+            bias = thops.mean(input.clone(), dim=[0, 2, 3], keepdim=True) * -1.0
+            vars = thops.mean((input.clone() + bias) ** 2, dim=[0, 2, 3], keepdim=True)
+            logs = torch.log(self.scale/(torch.sqrt(vars)+1e-6))
+            self.bias.data.copy_(bias.data)
+            self.logs.data.copy_(logs.data)
+            self.inited = True
+
+    def _center(self, input, reverse=False):
+        if not reverse:
+            return input + self.bias
+        else:
+            return input - self.bias
+
+    def _scale(self, input, logdet=None, reverse=False):
+        logs = self.logs
+        if not reverse:
+            input = input * torch.exp(logs)
+        else:
+            input = input * torch.exp(-logs)
+        if logdet is not None:
+            """
+            logs is log_std of `mean of channels`
+            so we need to multiply pixels
+            """
+            dlogdet = thops.sum(logs) * thops.pixels(input)
+            if reverse:
+                dlogdet *= -1
+            logdet = logdet + dlogdet
+        return input, logdet
+
+    def forward(self, input, logdet=None, reverse=False):
+        if not self.inited:
[…]

### 2019-02-27 code/glow/trainer.py

@@ -1,0 +1,222 @@
+import re
+import os
+import torch
+import torch.nn.functional as F
+
+import datetime
+import numpy as np
+from tqdm import tqdm
+from tensorboardX import SummaryWriter
+from torch.utils.data import DataLoader
+from .utils import save, load, plot_prob
+from .config import JsonConfig
+from .models import Glow
+from . import thops
+
+
+class Trainer(object):
+    def __init__(self, graph, graph_prior, optim, lrschedule, loaded_step,
+                 devices, data_device,
+                 dataset, hparams):
+        if isinstance(hparams, str):
+            hparams = JsonConfig(hparams)
+        # set members
+        # append date info
+        date = str(datetime.datetime.now())
+        date = date[:date.rfind(":")].replace("-", "")\
+                                     .replace(":", "")\
+                                     .replace(" ", "_")
+        self.log_dir = os.path.join(hparams.Dir.log_root, "log_" + date)
+        self.checkpoints_dir = os.path.join(self.log_dir, "checkpoints")
+        if not os.path.exists(self.log_dir):
+            os.makedirs(self.log_dir)
+        # write hparams
+        hparams.dump(self.log_dir)
+        if not os.path.exists(self.checkpoints_dir):
+            os.makedirs(self.checkpoints_dir)
+        self.checkpoints_gap = hparams.Train.checkpoints_gap
+        self.prior_gap = hparams.Train.prior_gap
+        self.max_checkpoints = hparams.Train.max_checkpoints
+        # set the mixture prior
+        self.num_component = hparams.Mixture.num_component
+        self.mix_prior = torch.ones(self.num_component)
+        # model relative
+        self.graph = graph
+
+        self.optim = optim
+        self.weight_y = hparams.Train.weight_y
+        # grad operation
+        self.max_grad_clip = hparams.Train.max_grad_clip
+        self.max_grad_norm = hparams.Train.max_grad_norm
+        # copy devices from built graph
+        self.devices = devices
+        self.data_device = data_device
+        # number of training batches
+        self.batch_size = hparams.Train.batch_size
+        self.data_loader = DataLoader(dataset,
+                                      batch_size=self.batch_size,
+                                    #   num_workers=8,
+                                      shuffle=True,
+                                      drop_last=True)
+        self.n_epoches = (hparams.Train.num_batches+len(self.data_loader)-1)
+        self.n_epoches = self.n_epoches // len(self.data_loader)
+        self.global_step = 0
+        # lr schedule
+        self.lrschedule = lrschedule
[…]

### 2019-02-27 code/glow/utils.py

@@ -1,0 +1,243 @@
+import os
+import re
+import copy
+import torch
+import numpy as np
+import matplotlib
+matplotlib.use("Agg")
+import matplotlib.pyplot as plt
+from shutil import copyfile
+
+def save_image(true_dist, fake_dist, img_dir, step):
+    """
+    Generates and saves a plot of the true distribution, the generator, and the
+    critic.
+    """
+    true_dist = true_dist.squeeze()
+    fake_dist = fake_dist.squeeze()
+    fake_dist = fake_dist.cpu().data.numpy()
+    true_dist = true_dist.cpu().data.numpy()
+    plt.clf()
+    plt.scatter(true_dist[:, 0], true_dist[:, 1], c='orange', marker='+')
+    plt.scatter(fake_dist[:, 0], fake_dist[:, 1], c='green', marker='+')
+    plt.savefig(img_dir + '/frame' + '_' + str(step) + '.jpg')
+
+
+def get_proper_cuda_device(device, verbose=True):
+    if not isinstance(device, list):
+        device = [device]
+    count = torch.cuda.device_count()
+    if verbose:
+        print("[Builder]: Found {} gpu".format(count))
+    for i in range(len(device)):
+        d = device[i]
+        did = None
+        if isinstance(d, str):
+            if re.search("cuda:[\d]+", d):
+                did = int(d[5:])
+        elif isinstance(d, int):
+            did = d
+        if did is None:
+            raise ValueError("[Builder]: Wrong cuda id {}".format(d))
+        if did < 0 or did >= count:
+            if verbose:
+                print("[Builder]: {} is not found, ignore.".format(d))
+            device[i] = None
+        else:
+            device[i] = did
+    device = [d for d in device if d is not None]
+    return device
+
+
+def get_proper_device(devices, verbose=True):
+    origin = copy.copy(devices)
+    devices = copy.copy(devices)
+    if not isinstance(devices, list):
+        devices = [devices]
+    use_cpu = any([d.find("cpu")>=0 for d in devices])
+    use_gpu = any([(d.find("cuda")>=0 or isinstance(d, int)) for d in devices])
+    assert not (use_cpu and use_gpu), "{} contains cpu and cuda device.".format(devices)
+    if use_gpu:
+        devices = get_proper_cuda_device(devices, verbose)
+        if len(devices) == 0:
+            if verbose:
[…]

### 2019-02-27 code/infer_mix.py

@@ -1,0 +1,234 @@
+"""Train script.
+Usage:
+    infer.py <hparams> <dataset_root> <mode>
+"""
+import os
+import cv2
+import random
+import torch
+import torchvision.datasets as dset
+import torch
+import numpy as np
+from docopt import docopt
+from torchvision import transforms
+from glow.builder import build
+from glow.config import JsonConfig
+from glow import modules
+import torchvision.utils as vutils
+from scipy import stats
+
+
+def prior_sampler(pk, size=64):
+    """ Define the sample generator for mixture distribution
+    xk = np.arrange(number of components)
+    pk = (0.1, 0.2, 0.3, 0.1, 0.1, 0.0, 0.2), tuple
+    R = custm.rvs(size=100)
+    """
+    xk = np.arange(len(pk))
+    custm = stats.rv_discrete(name='custm', values=(xk, pk))
+    prior_samples  = custm.rvs(size=size)
+    return prior_samples
+
+def GenMMmap_encoding(graph, x):
+    z, nll = graph(x=x, y_onehot=None, reverse=False)
+       
+    nlog_joint_prob =nll - torch.log(pk.unsqueeze(1).expand_as(nll)+1e-6).cuda()
+    tmp_sum = torch.log( torch.sum( torch.exp(-nlog_joint_prob), dim=[0]) ).cuda
+    nlog_gamma = nlog_joint_prob + tmp_sum().expand_as(nlog_joint_prob)
+    _, imin = nlog_gamma.min(dim=0)
+    r_c_index = torch.stack([imin.data.cpu().type(torch.int64), torch.arange(0,64).type(torch.int64)])
+    
+    return z[r_c_index[0], r_c_index[1]], r_c_index
+
+def LatMMmap_encoding(graph, x):
+    graph = graph.get_component()
+    z, _, _ = graph(x=x, y_onehot=None, reverse=False)
+       
+    return z
+
+def interpolation(start, end, points=9):
+    weight = np.linspace(0.1, 0.9, points)
+    inter = []
+    for i in weight:
+        inter.append(torch.lerp(start, end, i))
+    return inter
+
+if __name__ == "__main__":
+    args = docopt(__doc__)
+    hparams = args["<hparams>"]
+    dataset_root = args["<dataset_root>"]
+    mode = args["<mode>"]
+    assert mode in ["Generating", "Interpolation"]
+
+    assert os.path.exists(dataset_root), (
[…]

### 2019-02-27 code/train.py

@@ -1,0 +1,47 @@
+"""Train script.
+
+Usage:
+    train.py <hparams> <dataset> <dataset_root>
+"""
+import os
+import torchvision.datasets as dset
+from docopt import docopt
+from torchvision import transforms
+from glow.builder import build
+from glow.trainer import Trainer
+from glow.config import JsonConfig
+
+
+if __name__ == "__main__":
+    args = docopt(__doc__)
+    hparams = args["<hparams>"]
+    dataset = args["<dataset>"]
+
+    assert os.path.exists(hparams), (
+        "Failed to find hparams josn `{}`".format(hparams))
+
+    dataset_root = args["<dataset_root>"]
+    assert os.path.exists(dataset_root), (
+        "Failed to find root dir `{}` of dataset.".format(dataset_root))
+    hparams = JsonConfig(hparams)
+
+    if dataset == "mnist":
+        transform = transforms.Compose([
+            transforms.Resize(hparams.Data.resize),
+            transforms.ToTensor(),
+            transforms.Normalize((0.5,0.5,0.5),(0.5,0.5,0.5))
+        ])
+        dataset = dset.MNIST(root=dataset_root,
+                         download=True,
+                         transform=transform)
+    elif dataset == "fashion-mnist":
+        transform = transforms.Compose([
+            transforms.Resize(hparams.Data.resize),
+            transforms.ToTensor(),transforms.Normalize((0.5,0.5,0.5),(0.5,0.5,0.5))])
+        dataset = dset.FashionMNIST(root=dataset_root,
+                         download=True,
+                         transform=transform)
+    # build graph and dataset
+    built = build(hparams, True)
+    trainer = Trainer(**built, dataset=dataset, hparams=hparams)
+    trainer.train()

### 2019-03-05 manuscript/main.tex
adding pseudo algorithms in manuscript

@@
 % \setpapersize{A4}
+
+\RequirePackage{algorithm}
+\RequirePackage{algorithmic}
+
+% Attempt to make hyperref and algorithmic work together better:
+% \newcommand{\theHalgorithm}{\arabic{algorithm}}
 
@@
 
-The hypothesis of Hidden Markov Model (HMM): $\Hh := \{\bm{H} | \{\Ss, \bm{q}, A, p(\ubar{\bm{x}}|\ubar{\bm{s}}; \bm{\Phi})\}$,
+The hypothesis of Hidden Markov Model (HMM): $\Hh := \{\bm{H} | \{\Ss, \bm{q}, A, p(\bm{x}|{s}; \bm{\Phi})\}$,
 \begin{itemize}
@@
 \item $A$ is the transition matrix for the HMM $\bm{H}$ of size $|\Ss| \times |\Ss|$.
-\item Observable signal density $p(\ubar{\bm{x}}|\ubar{\bm{s}};\bm{\Phi})$ given hidden state sequence, where $\bm{\Phi}$ is the parameter set that defines this conditional probabilistic model.
+\item Observable signal density $p({\bm{x}}|{s};\bm{\Phi})$ given hidden state sequence, where $\bm{\Phi}$ is the parameter set that defines this conditional probabilistic model.
 \end{itemize}
@@
 
-Since model $\bm{H}$ contains hidden sequential variable $\ubm{s}$, we can not directly solve the maximum likelihood problem in $\autoref{eq:ml-of-hmm}$.
+\begin{figure}[!h]
+  \centering
+  \begin{tikzpicture}
+    \tikzstyle{enode} = [thick, draw=blue, ellipse, inner sep = 1pt,  align=center]
+    % \tikzstyle{nnode} = [thick, rectangle, rounded corners = 1pt,minimum size = 1.2cm,draw,inner sep = 2pt]
+    \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]
+    % \node[enode] (z1) at (-1.2,1.8) {Source 1}; 
+    \node[nnode] (g1) at (-0.5,1.8) {Source 1};
+
+    % \node[enode] (z2) at (-1.2,0.5){$\bm{z}\sim p_{2}(\bm{z})$};
+    \node[nnode] (g2) at (-0.5,0.5) {Source 2};
+
+    % \node[enode] (zK) at (-1.2,-1.8) {$\bm{z}\sim p_{|\Ss|}(\bm{z})$};
+    \node[nnode] (gs) at (-0.5, -1.8) {Source $|\Ss|$};
+
+    \node[enode] (x) at (4.5,1.5){$\ubm{x}\sim p(\ubm{x};\bm{H})$};
+    % \node at (3,-1){$p(\bm{z}) = \sum_k\pi_kp_k(\bm{z})$};
+
+    \draw[dotted,line width=2pt] (0,-0.3) -- (0,-1.2);
+    \filldraw[->] (1.9, 0.5)circle (2pt) --  (x) ;
+
+    \draw[->] (g1) -- (1.8, 1.8);
+
+    \draw[->] (g2) -- (1.8, 0.5);
+
+    % \draw[->] (1,-0.5) -- (1.8, -0.5);
+
+    \draw[->] (gs) -- (1.8, -1.8);
+
+    % \draw[->] (z1) [in=180,out=0]  to node[right]{$\pi_1$} (g);
+    % \draw[->] (z2) [in=180,out=0]  to node[above]{$\pi_2$} (g);
+    % \draw[->,dashed] (zK) [in=180,out=0]  to node[right]{$\pi_K$} (g);
+    % \draw[->] (g) to (x);
+    \begin{scope}[xshift=0.5cm, thick, every node/.style={sloped,allow upside down}]
+      \node[nnode] (m) at (3.5,-2) {Memory};
+      \node[nnode] (a) at (3.5,-0.5) {$\bm{A}$};
+
+      \draw (2.1,0.9)-- (2.2, 0.);
+      \draw (2.2,0.)-- node {\midarrow} (2.2,-2);
+      \draw (2.2,-2)-- (m);
+      \draw (m)-- (5, -2);
+      \draw (5, -2)-- node {\midarrow} (5 ,-0.5);
+      \draw (5, -0.5) -- (a);
+      \draw (a)-- node {\midarrow} (2.2, -0.5);
+      \node at (4.8, -1) {$s_{t}$};
+      \node at (2.56, -0.25) {$s_{t+1}$};
+    \end{scope}
+
+  \end{tikzpicture}
+  \caption{HMM Model defined by $\bm{H} = \{\Ss, \bm{q}, \bm{A}, p({\bm{x}}|{\bm{s}}; \bm{\Phi}) \}$}
+  \vspace{0.1cm}
+\end{figure}
+
+Since model $\bm{H}$ contains hidden sequential variable $\ubm{s}$, we can not directly solve the maximum likelihood problem in \autoref{eq:ml-of-hmm}.
 We use expectation maximization (EM) to address the hidden variable problem by
@@
   \Qq(\bm{q}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,p(\smtr{s}{1}{r}; \bm{q})\right] \label{eq:init-distribution-update}\\
-  \Qq({A}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})\right] \label{eq:transition-update}\\
+  \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})\right] \label{eq:transition-update}\\
   \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,p(\ubmr{x}{r} | \ubmr{s}{r}; \bm{\Phi})\right]\label{eq:generative-model-update}
@@
 The marginal $p(s_t| \ubm{x}; \bm{H})$ is also efficiently computed as the joint posterior.
+
+We summarize the optimization algorithm as:
+
+\begin{algorithm}[H]
+  \caption{Meta algorithm for HMM powered by Generative Models}
+  \begin{algorithmic}[1]
+    \STATE {\bfseries Input:}{
+      Building $\bm{H}^{\mathrm{old}}, \bm{H} \in \Hh$ gives: \\
+      $\bm{H}^{\mathrm{old}} = \{\Ss, \bm{q}^{\mathrm{old}}, A^{\mathrm{old}}, p(\bm{x}|s; \bm{\Phi}^{\mathrm{old}})\}$, $\bm{H} = \{\Ss, \bm{q}, A, p(\bm{x}|s; \bm{\Phi})\}$
+    }, \\
+    \STATE Initialize $\bm{H}$
+    \STATE $\bm{H}^{\mathrm{old}} \gets \bm{H}$
+    \FOR { $\bm{H}$ not converge}
+    \STATE Sample a batch of data $\left\{ \ubmr{x}{r} \right\}_{r=1}^{R_b}$ from the dataset $\hat{p}(\ubm{x})$
+    \STATE Compute $p(\smtr{s}{t}{r} | \ubmr{x}{r}; \bm{H}^{\mathrm{old}})$, $p(\smtr{s}{t}{r}, \smtr{s}{t+1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})$ by forward/backward algorithm;
+    \STATE $\bm{q} \gets \uargmin{\bm{q}}\, \Qq(\bm{q}; \bm{H}^{\mathrm{old}})$ by \autoref{eq:update-initial-state-prob};
+    \STATE $\bm{A} \gets \uargmin{\bm{A}}\Qq(\bm{A}; \bm{H}^{\mathrm{old}})$ by \autoref{eq:update-transition-prob};
+    \STATE $\bm{\Phi} \gets \uargmin{\bm{\Phi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})$ by calling Algorithm \autoref{algo:opt-Phi-GenHMM} or \autoref{algo:opt-Phi-LatHMM};
+    \STATE $\bm{H}^{\mathrm{old}} \gets \bm{H}$
+
+    \ENDFOR
+  \end{algorithmic}
+\end{algorithm}
 
@@
 is
-\begin{equation}
+\begin{equation}\label{eq:update-initial-state-prob}
   q_i = \frac{1}{R} \sum_{r=1}^{R} p(\smtr{s}{1}{r}=i | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}), \forall\; i = 1, 2, \cdots, |\Ss|.
@@
 \begin{align}
-  \Qq({A}; \bm{H}^{\mathrm{old}})
+  \Qq(\bm{A}; \bm{H}^{\mathrm{old}})
   &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})\right] \nonumber\\
@@
 
-Since $A_{i, j}  = p(\smtr{s}{t+1}{r}=j|\smtr{s}{t}{r}=i; {A})$ where $A_{i, j}$ is the element of transition matrix $A$, the solution to problem:
-\begin{align}
-  {A}^{\mathrm{new}} = &\uargmax{{A}} \Qq({A}; \bm{H}^{\mathrm{old}}), \nonumber \\
-  \mathrm{s.t.} &\hspace{0.2cm} A \cdot \bm{1} = \bm{1} \nonumber \\
-                       &  A^{\intercal} \cdot \bm{1} = \bm{1} \nonumber \\
-  & A_{i,j} \geq 0.
+Since $\bm{A}_{i, j}  = p(\smtr{s}{t+1}{r}=j|\smtr{s}{t}{r}=i; {A})$ where $A_{i, j}$ is the element of transition matrix $A$, the solution to problem:
+\begin{align}\label{eq:update-transition-prob}
+  \bm{A}^{\mathrm{new}} = &\uargmax{\bm{A}} \Qq(\bm{A}; \bm{H}^{\mathrm{old}}), \nonumber \\
+  \mathrm{s.t.} &\hspace{0.2cm} \bm{A} \cdot \bm{1} = \bm{1} \nonumber \\
+                          &  \bm{A}^{\intercal} \cdot \bm{1} = \bm{1} \nonumber \\
+                          & \bm{A}_{i,j} \geq 0.
 \end{align}
@@
 \begin{equation}
-  A_{i,j}^{\mathrm{new}} = \frac{\bar{\xi}_{i,j}}{\sum_{k = 1}^{|\Ss|} \bar{\xi}_{i,k}},
+  \bm{A}_{i,j}^{\mathrm{new}} = \frac{\bar{\xi}_{i,j}}{\sum_{k = 1}^{|\Ss|} \bar{\xi}_{i,k}},
 \end{equation}
@@
 
-\subsubsection{Generator Mixed HMM (GenM-HMM)}
+\subsubsection{Generator Mixed HMM (GenHMM)}\label{subsubsec:GenHMM}
 \begin{figure}[!h]
@@
   \begin{tikzpicture}
-    \tikzstyle{enode} = [thick, draw=blue, ellipse, inner sep = 1pt,  align=center]
-    % \tikzstyle{nnode} = [thick, rectangle, rounded corners = 1pt,minimum size = 1.2cm,draw,inner sep = 2pt]
+    % \tikzstyle{enode} = [thick, draw=blue, circle, inner sep = 3pt,
+    % align=center]
+    \tikzstyle{enode} = [thick, draw=blue, ellipse, inner sep = 2pt,  align=center]
     \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]
-    \node[enode] (z1) at (-1.2,1.8) {$\bm{z}\sim p_1(\bm{z})$};
-    \node[nnode] (g1) at (1,1.8) {$\bm{g}_1$};
 
-    \node[enode] (z2) at (-1.2,0.5){$\bm{z}\sim p_{2}(\bm{z})$};
-    \node[nnode] (g2) at (1,0.5) {$\bm{g}_2$};
+    \node[enode] (z1) at (-1.2,1.8) {$\bm{z}\sim p_{s,1}(\bm{z})$};
+    \node[nnode] (g1) at (1,1.8) {$\bm{g}_{s,1}$};
 
-    \node[enode] (zK) at (-1.2,-1.8) {$\bm{z}\sim p_{|\Ss|}(\bm{z})$};
-    \node[nnode] (gs) at (1, -1.8) {$\bm{g}_{|\Ss|}$};
+    \node[enode] (z2) at (-1.2,0.5){$\bm{z}\sim p_{s,2}(\bm{z})$};
+    \node[nnode] (g2) at (1,0.5) {$\bm{g}_{s,2}$};
 
-    \node[enode] (x) at (4.5,1.5){$\ubm{x}\sim p(\ubm{x};\bm{H})$};
+    \node[enode] (zK) at (-1.2,-1.8) {$\bm{z}\sim p_{s,K}(\bm{z})$};
+    \node[nnode] (gs) at (1, -1.8) {$\bm{g}_{s,K}$};
+
+    \node[enode] (x) at (4.5,0){$\bm{x}\sim p(\bm{x}| s; \bm{\Phi}_{s})$};
     % \node at (3,-1){$p(\bm{z}) = \sum_k\pi_kp_k(\bm{z})$};
@@
     \draw[dotted,line width=2pt] (0,-0.3) -- (0,-1.2);
-    \filldraw[->] (1.9, 0.5)circle (2pt) --  (x) ;
+    \filldraw[->] (1.9, 0.5)circle (2pt) --  node[above=0.2]{${\kappa}\sim \bm{\pi}_{s}$} (x)  ;
     \draw[->] (z1) -- (g1);
@@
 
+    % \node[enode] (z) at (0,0) {$\bm{z}\sim p(\bm{z})$};
+    % \node[enode] (x) at (5.5,0){$\bm{x}\sim p(\bm{x}; \bm{\Phi})$};
+    %               % \node at (5.2,-1) {$p(\bm{x};\bm{\Phi}) = \textstyle\sum_{k=1}^K \pi_k  p_k(\bm{x})$};
+    % \node[nnode] (g1) at (2.6,1.8) {$\bm{g}_1$};
+    % \node[nnode] (g2) at (2.6,0.5) {$\bm{g}_2$};
+    % \node[nnode] (gk) at (2.6,-1.8) {$\bm{g}_K$};
+    % \draw[dotted,line width=2pt] (2.6,-0.3) -- (2.6,-1.2);
+    % \draw[->] (z) [in= 180, out =0] to (g1);
+    % \draw[->] (z) [in= 180, out =0] to (g2);
+    % \draw[->] (z) [in= 180, out =0] to (gk);
+    % \filldraw[->] (3.7, 0.5)circle (2pt) -- node[above=0.2]{$\bm{s}\sim \bm{\pi}$} (x) ;
+    %               % \draw[->] (3,-0.8) -- (3.5, -0.8);
+    % \draw[->] (g1) -- (3.5,1.8);
+    % \draw[->] (g2) -- (3.5, 0.5);
+    % \draw[->] (gk) -- (3.5, -1.8);
+  \end{tikzpicture}
+  \caption{Source $s$ powered by generator mixed generative model (GenHMM).}
+  \vspace{0.1cm}
+\end{figure}
+
+Base on the generative model assumption:
+
+\begin{equation}
+  p(\bm{x}| s; \bm{\Phi}_{s}) = \sum_{\kappa=1}^{K}\pi_{s, \kappa} p(\bm{x}| s, \kappa; \bm{\Phi}_{s, \kappa})
+\end{equation}
+where $\bm{\Phi}_s = \left\{ \bm{\Phi}_{s, \kappa}| \kappa = 1, 2, \cdots, K. \right\}$, and 
+\begin{equation}
+  \sum_{\kappa = 1}^{K} \pi_{s, \kappa}= 1
+\end{equation}
+
+% Let us denote the inverse of $\bm{g}_s$ as $\bm{f}_s = \bm{g}_s^{-1}$.
+% For this case we assume the latent sources are Gaussian. We write the emission probability model as a $K$ generator mixed model:
+% \begin{align}
+    %     p(\bm{x} | s; \bm{\Phi}) &= \sum_{\kappa = 1}^{K} \pi_{s, \kappa} \Nn\left( \bm{f}_{s,\kappa}(\bm{x}); \bm{\mu}_{s,\kappa}, \bm{C}_{s,\kappa}  \right)\bigg|\det\,\left( \pd{\bm{f_{s,\kappa}(\bm{x})}}{\bm{x}} \right)\bigg| \nonumber \\
+    %     \sum_{\kappa = 1}^{K} \pi_{s, \kappa}&= 1
+                                                 %   \end{align}
+
+The help function should be revised to deal with the new latent variable $\kappa$ into:
+\begin{align}\label{eq:obj-q-gen-mix}
+  \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) & = \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|} \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})  \log\,p(\smtr{\kappa}{t}{r}, \bmtr{x}{t}{r}| \smtr{s}{t}{r}; \bm{\Phi})\nonumber \\
+  =& \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})p(\smtr{\kappa}{t}{r}|\smtr{s}{t}{r}, \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \bigg[\log\,\pi_{\smtr{s}{t}{r},\smtr{\kappa}{t}{r}} \nonumber \\
+                                        & + \log\, p(\bmtr{x}{t}{r} | \smtr{s}{t}{r}, \smtr{\kappa}{t}{r}; \bm{\Phi}) \bigg]
+\end{align}
+                                          %                                           \Nn\left( \bm{f}_{\smtr{s}{t}{r},\smtr{\kappa}{t}{r}}(\bm{x}); \bm{\mu}_{\smtr{s}{t}{r},\smtr{\kappa}{t}{r}}, \bm{C}_{\smtr{s}{t}{r},\smtr{\kappa}{t}{r}} \right) \log\,\bigg| \det\, \left( \pd{\bm{f}_{\smtr{s}{t}{r},\smtr{\kappa}{t}{r}}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg| \bigg].
+                                          %   \end{align}
+In \autoref{eq:obj-q-gen-mix}, $p(s_t| \ubm{x}, \bm{H}^{\mathrm{old}})$ is computed by forward/backward algorithm. The posterior of $\kappa$ is:
+\begin{align}\label{eq:kappa-posterior}
+  p(\kappa| s, \ubm{x}; \bm{H}^{\mathrm{old}})
+  &=  \frac{p(\kappa, \ubm{x}| s; \bm{H}^{\mathrm{old}})}{p(\ubm{x}| s,\bm{H}^{\mathrm{old}})} \nonumber \\
+  & = \frac{\pi_{s, \kappa} p(\ubm{x}| s, \kappa, \bm{H}^{\mathrm{old}})}{\sum_{\kappa=1}^{K} \pi_{s, \kappa} p(\ubm{x}| s, \kappa,\bm{H}^{\mathrm{old}})} \nonumber \\
+  & = \frac{\pi_{s, \kappa} p(\bm{x}| s, \kappa, \bm{H}^{\mathrm{old}})}{\sum_{\kappa=1}^{K}  \pi_{s, \kappa} p(\bm{x}| s, \kappa,\bm{H}^{\mathrm{old}})} 
+  % & = \frac{\pi_{s_t, \kappa_t} \Nn\left( \bm{f}_{{s}_{t},{\kappa}_{t}}(\bm{x}); \bm{\mu}_{{s}_{t},{\kappa}_{t}}, \bm{C}_{{s}_{t},{\kappa}_{t}} \right) \bigg| \det\, \left( \pd{\bm{f}_{{s}_{t},{\kappa}_{t}}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg| }{\sum_{\kappa_t=1}^{K}\pi_{s_t, \kappa_t} \Nn\left( \bm{f}_{{s}_{t},{\kappa}_{t}}(\bm{x}); \bm{\mu}_{{s}_{t},{\kappa}_{t}}, \bm{C}_{{s}_{t},{\kappa}_{t}} \right) \bigg| \det\, \left( \pd{\bm{f}_{{s}_{t},{\kappa}_{t}}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg|} \nonumber \\
+\end{align}
+where the last equation is due to the fact that only $\bm{x}_t$ among sequence $\ubm{x}$ depends on $s_t, \kappa_t$.
+
+    %     \begin{align}
+
+    %   \end{align}
+
+The latent prior for mixture of each source $s$ is obtained by solving the following problem:
+\begin{align}
+  \pi_{s, \kappa} & = \uargmax{pi_{s, \kappa}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) \\ \nonumber
+                  & s.t. \, \sum_{\kappa=1}^{K} \pi_{s, \kappa}= 1
+\end{align}
+which gives the solution:
+\begin{equation}\label{eq:mix-latent-parameter-solution}
+  \pi_{s, \kappa} = \frac{\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }{\sum_{k =1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=k | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }
+\end{equation}
+where $p(s, \kappa | \ubm{x}; \bm{H}^{\mathrm{old}}) = p(s| \ubm{x}; \bm{H}^{\mathrm{old}}) p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$ that can be computed by results of forward/backward and \autoref{eq:kappa-posterior}.
+
+In implementation, GenHMM uses the generator mixed emission model with $\kappa$-th component as:
+
+\begin{align}
+  &p(\bm{x}| s, \kappa; \bm{\Phi}_{s, \kappa}) \nonumber \\
+  =&p_{s, \kappa}(\bm{z}) \bigg| \det \left( \pd{\bm{z}}{\bm{x}} \right) \bigg| \nonumber \\
+  = &\Nn\left( \bm{f}_{{s},{\kappa}}(\bm{x}); \bm{\mu}_{{s},{\kappa}}, \bm{C}_{{s},{\kappa}} \right) \bigg| \det\, \left( \pd{\bm{f}_{{s},{\kappa}}(\bm{x})}{\bm{x}} \right) \bigg|
+\end{align}
+where $\bm{f}_{{s},{\kappa}} = \bm{g^{-1}_{{s},{\kappa}}}$ is defined by parameter set $\bm{\theta}_{s, \kappa}$, and $\bm{\Phi}_{s, \kappa} = \left\{\bm{\mu}_{{s},{\kappa}}, \bm{C}_{{s},{\kappa}} , \bm{\theta}_{s, \kappa} \right\}, \kappa = 1, 2, \cdots, K$.
+We can start by setting $\bm{\mu}_{{s},{\kappa}} = \bm{0}$ and $\bm{C}_{{s},{\kappa}} = diag(\bm{1})$.
+\begin{algorithm}[H]
+  \caption{M-step w.r.t. $\bm{\Phi}$ powered by GenMM}\label{algo:opt-Phi-GenHMM}
+  \begin{algorithmic}[1]
+    \STATE {\bfseries Input:}
+    Latent mixture distribution: $\Nn\left(\bm{z}; \bm{0}, \mathrm{diag}(\bm{1})\right)$, $\forall \, s \in \Ss, \kappa = 1, 2, \cdots, K$;\\
+    Empirical distribution $\hat{p}(\bm{x})$ of dataset; \\
+    \STATE Set a total number of epochs $T$ of training as stop criterion. A learning rate $\eta$.
+    \FOR {epoch $t < T$}
+    \STATE Sample a batch of data $\left\{ \ubmr{x}{r} \right\}_{r=1}^{R_b}$ from dataset $P_d(\ubm{x})$
+
+    \STATE Compute $p(\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})$  by Forward/backward and \autoref{eq:kappa-posterior}.
+    %     = \frac{\pi_k \Nn\left(f(X); \bm{\mu}_k^{t},\bm{\sigma}_k^{t}\right)}{\sum_{k=1}^K\; \pi_k\Nn\left(f(X); \bm{\mu}_k^{t},\bm{\sigma}_k^{t}\right) }, \forall \left\{ x^{(i)}\right\}_{i=1}^{n_b}$
+    % \Comment{Calculate posterior \autoref{eq-posterior-gamma}}
+    \STATE Compute loss ${\Qq}\left({\bm{\Phi}}, {\bm{H}}^{\mathrm{old}}\right)$ in \autoref{eq:obj-q-gen-mix}
+
+    \STATE $\partial{\bm{\th}_s} \gets  \nabla_{\bm{\theta}} -
+    \frac{1}{R_{b}}{\Qq}\left({\bm{\Phi}},{\bm{H}}^{\mathrm{old}}\right)$
+    % \Comment{E-step}
+    % \STATE $\partial{g_k} \gets \nabla_{\th_k}\widetilde{\Qq}\left(\left\{ \th_k
+    %   \right\}_{k=1}^{K};\left\{ \th_k^t\right\}_{k=1}^{K}\right)$,
+    % $\forall \th_k \in\left\{ \th_k\right\}_{k=1}^{K}$
+    \STATE $\bm{\th}_s \gets \bm{\th}_s - \eta \cdot \partial{\bm{\theta}_s}$, $\forall s \in \Ss$
+    \ENDFOR
+
+    \STATE Update $\pi_{s, \kappa}, \forall, s\in \Ss, \kappa= 1, 2, \cdots, K$, according to \autoref{eq:mix-latent-parameter-solution}
+
+    \STATE Assemble $\bm{\Phi} = \left\{  \bm{\Phi}_s| s \in \Ss\right\}$ for $\bm{\Phi}_s = \left\{ \bm{\theta}_{s, \kappa}, \bm{\mu}_{{s},{\kappa}}, \bm{C}_{{s},{\kappa}}| \kappa = 1, 2, \cdots, K \right\}$.
+
+  \end{algorithmic}
+\end{algorithm}
+
+\subsubsection{Latent-source Mixed HMM (LatMM)}
+
+\begin{figure}[h]
+  \centering
+  \begin{tikzpicture}
+    \tikzstyle{enode} = [thick, draw=blue, ellipse, inner sep = 1pt,  align=center]
+    % \tikzstyle{nnode} = [thick, rectangle, rounded corners = 1pt,minimum size = 1.2cm,draw,inner sep = 2pt]
+    \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]
+    \node[enode] (z1) at (0,1.8) {$\bm{z}\sim p_{s,1}(\bm{z})$};
+    \node[enode] (z2) at (0,0.5){$\bm{z}\sim p_{s,2}(\bm{z})$};
+    \node[enode] (zK) at (0,-1.8) {$\bm{z}\sim p_{s, K}(\bm{z})$}; 
+    \node[enode] (x) at (5.5,0){$\bm{x}\sim p(\bm{x}|s;{\bm{\Phi}}_{s})$};
+    % \node at (3,-1){$p(\bm{z}) = \sum_k\pi_kp_k(\bm{z})$};
+    \node[nnode] (g) at (3.2,0) {$\bm{g}_s$};
+    \draw[dotted,line width=2pt] (0,-0.3) -- (0,-1.2);
+    \filldraw[->] (1.8, 0.5)circle (2pt) -- node[above=0.2]{${\kappa}\sim \bm{\pi}_s$} (g) ;
+    \draw[->] (z1) -- (1.6, 1.8);
+    \draw[->] (z2) -- (1.6, 0.5);
+    % \draw[->] (1,-0.5) -- (1.8, -0.5);
+    \draw[->] (zK) -- (1.6, -1.8);
     % \draw[->] (z1) [in=180,out=0]  to node[right]{$\pi_1$} (g);
@@
     % \draw[->,dashed] (zK) [in=180,out=0]  to node[right]{$\pi_K$} (g);
-    %\draw[->] (g) to (x);
-    \begin{scope}[xshift=0.5cm, thick, every node/.style={sloped,allow upside down}]
-      \node[nnode] (m) at (3.5,-2) {Memory};
-      \node[nnode] (a) at (3.5,-0.5) {$A$};
-
-      \draw (2.1,0.9)-- (2.2, 0.);
-      \draw (2.2,0.)-- node {\midarrow} (2.2,-2);
-      \draw (2.2,-2)-- (m);
-      \draw (m)-- (5, -2);
-      \draw (5, -2)-- node {\midarrow} (5 ,-0.5);
-      \draw (5, -0.5) -- (a);
-      \draw (a)-- node {\midarrow} (2.2, -0.5);
-      \node at (4.8, -1) {$s_{t}$};
-      \node at (2.56, -0.25) {$s_{t+1}$};
-    \end{scope}
-
+    \draw[->] (g) to (x);
   \end{tikzpicture}
-  \caption{GenM-HMM Model defined by $\bm{H} = \{\Ss, \bm{q}, A, p({\bm{x}}|{\bm{s}}; \bm{\Phi}) \}$}
+  \caption{Source $s$ powered by Latent mixed generative model (LatHMM).}
   \vspace{0.1cm}
@@
 
-For this proposal, we seek to use a generator mixed HMM scheme, termed as GenM-HMM. We define a set of generators for GenM-HMM:
-\begin{equation}
-  \{\bm{g}_{s}| s \in \Ss, \bm{g}_s: \bm{z}\rightarrow \bm{x}, \bm{z}\sim p_{s}(\bm{z})\}.
+LatHMM, (\textbf{benching marking}), the latent source mixed emission model is built by allowing different latent source sharing the same generator $\bm{g}_s$:
+\begin{align}
+  & p(\bm{x}| s, \kappa; \bm{\Phi}_{s}) \nonumber \\
+  =&p_{s, \kappa}(\bm{z}) \bigg| \det \left( \pd{\bm{z}}{\bm{x}} \right) \bigg| \nonumber \\
+  =& \Nn\left( \bm{f}_{{s}}(\bm{x}); \bm{\mu}_{{s},{\kappa}}, \bm{C}_{{s},{\kappa}} \right) \bigg| \det\, \left( \pd{\bm{f}_{{s}}(\bm{x})}{\bm{x}} \right) \bigg|
+\end{align}
+where $\bm{f}_{{s}} = \bm{g^{-1}}_{s}$ is defined by $\bm{\theta}$, and $\bm{\Phi}_s = \left\{ \bm{\theta}_s, \bm{\mu}_{{s},{\kappa}}, \bm{C}_{{s},{\kappa}}| \kappa = 1, 2, \cdots, K \right\}$.
+
+This emission generator sharing makes the posterior computation w.r.t. of $\kappa$ easier:
+
+\begin{equation}\label{eq:kappa-posterior-LatHMM}
+  p(\kappa| s, \ubm{x}; \bm{H}^{\mathrm{old}})
+  = \frac{\pi_{s, \kappa} p(\bm{x}| s, \kappa, \bm{H}^{\mathrm{old}})}{\sum_{\kappa=1}^{K}  \pi_{s, \kappa} p(\bm{x}| s, \kappa,\bm{H}^{\mathrm{old}})}
+  = \frac{\pi_{s, \kappa} \Nn\left( \bm{z}; \bm{\mu}_{s, \kappa}, \bm{C}_{s, \kappa} \right)}{\sum_{\kappa=1}^{K}  \pi_{s, \kappa} \Nn\left( \bm{z}; \bm{\mu}_{s, \kappa}, \bm{C}_{s, \kappa}\right)} \bigg|_{\bm{z} = \bm{f}_s(\bm{x})}.
+  % & = \frac{\pi_{s_t, \kappa_t} \Nn\left( \bm{f}_{{s}_{t},{\kappa}_{t}}(\bm{x}); \bm{\mu}_{{s}_{t},{\kappa}_{t}}, \bm{C}_{{s}_{t},{\kappa}_{t}} \right) \bigg| \det\, \left( \pd{\bm{f}_{{s}_{t},{\kappa}_{t}}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg| }{\sum_{\kappa_t=1}^{K}\pi_{s_t, \kappa_t} \Nn\left( \bm{f}_{{s}_{t},{\kappa}_{t}}(\bm{x}); \bm{\mu}_{{s}_{t},{\kappa}_{t}}, \bm{C}_{{s}_{t},{\kappa}_{t}} \right) \bigg| \det\, \left( \pd{\bm{f}_{{s}_{t},{\kappa}_{t}}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg|} \nonumber \\
 \end{equation}
-Thus there are total $|\Ss|$ generators.
+Computation of $p(s_t| \ubm{x}, \bm{H}^{\mathrm{old}})$ remains the same, relying on forward/backward algorithm.
+
+In implementation, $\bm{C}_{s, \kappa} = diag\left( \bm{\sigma}_{s, \kappa} \right)$ for simplicity.
+To avoid the singularity problem of Gaussian mixture, we put a Gamma distribution as prior for $\bm{\sigma}_{s, \kappa}$, i.e.
+$\Gamma(\bm{\sigma}_{s, \kappa}^{-1}, a, b)$ where $a$ and $b$ are hyperparameter for Gamma distribution.
+Then the problem can be reformulated as:
+\begin{equation}\label{eq:loss-on-Phi-LatHMM}
+  \uargmax{\bm{\Phi}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) + \frac{1}{K}\log\prod_{k=1}^K\Gamma(\bm{\sigma}_{s, k}^{-1}; a, b)
+\end{equation}
+
+Apart from the emission prboability model difference, rest computation is the same as \autoref{subsubsec:GenHMM}. We summarize the algorithm as:
+
+\begin{algorithm}[H]
+  \caption{M-step w.r.t. $\bm{\Phi}$ powered by LatMM}\label{algo:opt-Phi-LatHMM}
+  \begin{algorithmic}[1]
+    \STATE {\bfseries Input:}
+    Latent mixture distribution: $\sum_{k=1}^{K}\pi_{s, \kappa} \Nn\left(\bm{z}; \bm{\mu}_{s, \kappa}, \mathrm{diag}(\bm{\sigma}_{s, \kappa}^2)\right)$\\
+    Empirical distribution $\hat{p}(\bm{x})$ of dataset; \\
+    \STATE Set a total number of epochs $T$ of training as stop criterion. A learning rate $\eta$. Set hyperparameter $a$ , $b$ for prior of $\bm{\sigma}_{s, \kappa}^{-1}, \forall k$.
+    \FOR {epoch $t < T$}
+    \STATE Sample a batch of data $\left\{ \ubmr{x}{r} \right\}_{r=1}^{R_b}$ from dataset $P_d(\ubm{x})$
+    \STATE Compute $p(\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})$  by Forward/backward and \autoref{eq:kappa-posterior-LatHMM}.
+%= \frac{\pi_k \Nn\left(f(X); \bm{\mu}_k^{t},\bm{\sigma}_k^{t}\right)}{\sum_{k=1}^K\; \pi_k\Nn\left(f(X); \bm{\mu}_k^{t},\bm{\sigma}_k^{t}\right) }, \forall \left\{ x^{(i)}\right\}_{i=1}^{n_b}$
+    % \Comment{Calculate posterior \autoref{eq-posterior-gamma}}
+    \STATE Compute loss in \autoref{eq:loss-on-Phi-LatHMM}%\Comment{Expectation}
+
+    \STATE $\partial{\bm{\th}_s}, \partial{\bm{\mu}_{s, \kappa}}, \partial{\bm{\sigma}_{s, \kappa}}\gets 
+    \nabla_{\bm{\theta}, \bm{\mu}_k, \bm{\sigma}_k} -
+    \frac{1}{R_{b}}{\Qq}\left({\bm{\Phi}},\ubar{\bm{H}}^{\mathrm{old}}\right)
+    -\frac{1}{K}\sum_{k=1}^K \log \, \Gamma(\bm{\sigma}_{s, \kappa}^{-1}; a, b)$ %\Comment{E-step}
+    % \STATE $\partial{g_k} \gets \nabla_{\th_k}\widetilde{\Qq}\left(\left\{ \th_k
+    %   \right\}_{k=1}^{K};\left\{ \th_k^t\right\}_{k=1}^{K}\right)$,
+    % $\forall \th_k \in\left\{ \th_k\right\}_{k=1}^{K}$
+    \STATE $\bm{\th}_s \gets \bm{\th}_s - \eta \cdot \partial{\bm{\theta}_s}$, $\forall s \in \Ss$
+    \STATE $\bm{\mu}_{s, \kappa} \gets \bm{\mu}_{s, \kappa} - \eta \cdot \partial{\bm{\mu}_{s, \kappa}}, \forall \kappa, s$
+    \STATE $\bm{\sigma}_{s, \kappa} \gets \bm{\sigma}_{s, \kappa} - \eta \cdot \partial{\bm{\sigma}_{s, \kappa}}, \forall \kappa, s$
+    \ENDFOR
+
+    \STATE Update $\pi_k$ according to \autoref{eq:mix-latent-parameter-solution}
+
+    \STATE Assemble $\bm{\Phi} = \left\{  \bm{\Phi}_s| s \in \Ss\right\}$ for $\bm{\Phi}_s = \left\{ \bm{\theta}_s, \bm{\mu}_{{s},{\kappa}}, \bm{C}_{{s},{\kappa}}| \kappa = 1, 2, \cdots, K \right\}$.
+
+  \end{algorithmic}
+\end{algorithm}
+
+    %     This mixture emission probability model introduces anther latent variable $\kappa$, indicating the actual component of the mixture probability model that corresponding to the signal $\bm{x}$.
+
+    % %     For this proposal, we seek to use a generator mixed HMM scheme, termed as GenM-HMM. We define a set of generators for GenM-HMM:
+    % %     \begin{equation}
+    % %       \{\bm{g}_{s}| s \in \Ss, \bm{g}_s: \bm{z}\rightarrow \bm{x}, \bm{z}\sim p_{s}(\bm{z})\}.
+    % %     \end{equation}
+    %     Thus there are total $|\Ss|$ generators.
 $p(\bm{x}|{s}; \bm{\Phi})$ is induced as $\bm{g}_s(\bm{z}) \sim p(\bm{x}|{s}; \bm{\Phi})$ where $\bm{z} \sim p_s(\bm{z})$ for $s \in \Ss$.
-Let us denote the inverse of $\bm{g}_s$ as $\bm{f}_s = \bm{g}_s^{-1}$.
 We have the $s$-th component of the GenM-HMM model as
-\begin{align}
-  p(\bm{x} | s; \bm{\Phi})
-  &= p_s(\bm{z}) \bigg| \det \left( \pd{\bm{z}}{\bm{x}} \right) \bigg| \nonumber \\
-  &= p_s(\bm{f}_s(\bm{x})) \bigg| \det \left( \pd{\bm{f}_s(\bm{x})}{\bm{x}} \right) \bigg|
-\end{align}
-where $p_s(\bm{z})$ is the latent source distribution for $s=1, 2, \cdots, |\Ss|$.
+    %     \begin{align}
+    %     p(\bm{x} | s; \bm{\Phi})
+    %                                    &= p_s(\bm{z}) \bigg| \det \left( \pd{\bm{z}}{\bm{x}} \right) \bigg| \nonumber \\
+    %                                    &= p_s(\bm{f}_s(\bm{x})) \bigg| \det \left( \pd{\bm{f}_s(\bm{x})}{\bm{x}} \right) \bigg|
+                                           %   \end{align}
+                                           %                                            where $p_s(\bm{z})$ is the latent source distribution for $s=1, 2, \cdots, |\Ss|$.
 
-Let us denote the parameter set that defines latent distribution $p_s(z)$ by $\bm{\omega}_s$ and the parameter set that defines generator $\bm{g}_s$ by $\bm{\theta}_s$.
+                                           %                                            Let us denote the parameter set that defines latent distribution $p_s(z)$ by $\bm{\omega}_s$ and the parameter set that defines generator $\bm{g}_s$ by $\bm{\theta}_s$.
 Then $\bm{\Phi} = \left\{ \bm{\theta}_s,
-  \bm{\omega}_s, \forall s\in \Ss\right\}$. The problem in \autoref{eq:sub-gm} can be reformulated as:
-\begin{align}
-  &\umax{\bm{\Phi}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) \nonumber \\
-  =&\umax{ \bm{\theta}_s, \bm{\omega}_s, \forall s\in \Ss} \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \left[\log\,p_{\smtr{s}{t}{r}}(\bm{f}_{\smtr{s}{t}{r}}(\bmtr{x}{t}{r})) + \log\,\bigg| \det\, \left( \pd{\bm{f}_{\smtr{s}{t}{r}}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg| \right].
-\end{align}
+                                           %                                            \bm{\omega}_s, \forall s\in \Ss\right\}$. The problem in \autoref{eq:sub-gm} can be reformulated as:
 
-The diagram of GenM-HMM is shown as follows.
-
-\subsubsection{Latent-source Mixed HMM (LatM-HMM)}
+\subsubsection{Generator-shared HMM (GSHMM)...to be continued}
 \begin{figure}[!h]
@@
     \node[nnode] (m) at (3.5,-2) {Memory};
-    \node[nnode] (a) at (3.5,-0.5) {$A$};
+    \node[nnode] (a) at (3.5,-0.5) {${A}$};
 
@@
 
+\textcolor{blue}{to be continued...}
+
 Alternatively, we can use a latent-source mixed HMM (LatM-HMM) where different latent source share the same generator functioning as feature mapping. Then the generator of the LatM-HMM is defined as
@@
 
+\subsubsection{Generator Shared HMM (GSHMM)}
+
 \textbf{To be continued}...
-\section{On Implementation}
+
+\section{On Implementation of acoustic signal}
 Found a HMM python lib that basics provide needed API for us, see \href{https://hmmlearn.readthedocs.io/en/stable/index.html}{hmmlearn}. Saikat also has suggestion.
@@
 \end{itemize}
-\bibliographystyle{IEEEbib}
+
+\section{On Implementation of Planning}
+Refer to \cite{kurutach2018learning_plan} and its \href{https://github.com/thanard/causal-infogan}{experiments}.
+
+\bibliographystyle{plain}
 

### 2019-03-13 code/timit-preprocessor/make_dataset.py
Now formats a database for phoneme reco

@@ -9,6 +9,9 @@ if __name__ == "__main__":
     file = sys.argv[1]
     out_file = sys.argv[2]
 
+    timit_folder = "/home/antoine/Documents/projects/asr/gm-hmm/data/timit"
+    fname_cb = "TIMIT-61.codebook"
+
     out = kaldi_io.read_mat_scp(file)
 
     n = 0
@@ -21,12 +24,75 @@ if __name__ == "__main__":
     lengths = []
     keys = []
     seq_start = 0
-
+    out = kaldi_io.read_mat_scp(file)
     for key, mat in out:
         l = mat.shape[0]
         lengths.append(l)
         keys.append(key)
         X[seq_start:seq_start+l] = mat
+        seq_start += l
 
-    pkl.dump([X, lengths, keys], open(out_file, "wb"))
-    sys.exit(0)
+    fe = 16e3
+    wsize_ms = 10
+    nsamples = len(lengths)
+    PHN = [0 for _ in range(nsamples)]
+    DATA = [0 for _ in range(nsamples)]
+
+    seq_start = 0
+    DATA_TYPE = "TEST" if "test" in out_file else "TRAIN"
+
+    for isample in range(nsamples):
+
+        # Get phone labels
+        fname = timit_folder + "/"+DATA_TYPE+"/" + keys[isample].replace("-", "/") + ".PHN"
+        #print(isample, "/", nsamples, ":", fname)
+        with open(fname, "r") as f:
+            lines = f.read().split("\n")[:-1]
+
+        PHN_file_data = np.array(list(map(lambda x: x.split(" "), lines)))
+        loc = PHN_file_data[:, :2]
+        y = PHN_file_data[:, -1]
+        del PHN_file_data
+        loc = loc.astype(np.int64)
+        x = X[seq_start:seq_start + lengths[isample]]
+
+        feats = np.arange(x.shape[0] * wsize_ms).reshape(-1, wsize_ms)
+
+
+        loc_ms = loc / fe * 1000
+        phones = np.copy(feats).reshape(-1)
+        # Clean loc_ms
+        if loc_ms[-1][1] < phones[-1]:
+            loc_ms[-1][1] = phones[-1]
+
+        if loc_ms[0][0] > 0:
+            loc_ms[0][0] = 0
+
+        for iy, irange in enumerate(loc_ms.tolist()):
+            phones[int(np.floor(irange[0])):int(np.floor(irange[1]))] = iy
+
[…]

### 2019-03-13 code/timit-preprocessor/extract_mfcc.sh
improve generalizability

@@ -1,32 +1,36 @@
 #!/usr/bin/env bash
 
-if [ ${#} != 1 ];then
-    echo "Usage: ./mfcc_extract.sh [Kaldi's relative path]"
+if [ ${#} != 3 ];then
+    echo "Usage: ./mfcc_extract.sh [Kaldi's relative path] [TIMIT relative path] [output .scp filename]"
     exit -1
 fi
 
 kaldi=${1}
-path=data/processed
+timit=${2}
+target_file=${3}
+
+path=.data/processed
 options="--use-energy=false"
 
 mkdir -p $path
 
 
-echo "Extracting train set"
-target=train
+fname=$(basename "$target_file")
+fname_trunk="${fname%.*}"
+target=(${fname_trunk//./ })
+
+echo $fname
+
 log=$path/$target.extract.log
+if [[ $fname == *".13."* ]]; then
+	${kaldi}/src/featbin/compute-mfcc-feats --verbose=2 $options scp:.data/material/$target.wav.scp ark,t,scp:$path/$target.13.ark,$target_file 2> $log
+fi
 
-${kaldi}/src/featbin/compute-mfcc-feats --verbose=2 $options scp:data/material/$target.wav.scp ark,t,scp:$path/$target.13.ark,$path/$target.13.scp 2> $log
-${kaldi}/src/featbin/add-deltas ark:$path/$target.13.ark ark,t:$path/$target.13_deltas.ark
-${kaldi}/src/featbin/compute-cmvn-stats ark:$path/$target.13_deltas.ark ark,t:$path/$target.cmvn_results.ark
-${kaldi}/src/featbin/apply-cmvn ark:$path/$target.cmvn_results.ark ark:$path/$target.13_deltas.ark ark,t,scp:$path/$target.39.cmvn.ark,$path/$target.39.cmvn.scp
+if [[ $fname == *".39."* ]]; then
+	${kaldi}/src/featbin/compute-mfcc-feats --verbose=2 $options scp:.data/material/$target.wav.scp ark,t,scp:$path/$target.13.ark,$target_file 2> $log
+	${kaldi}/src/featbin/add-deltas ark:$path/$target.13.ark ark,t:$path/$target.13_deltas.ark
+	${kaldi}/src/featbin/compute-cmvn-stats ark:$path/$target.13_deltas.ark ark,t:$path/$target.cmvn_results.ark
[…]

### 2019-03-13 code/timit-preprocessor/parsing.py

@@ -4,7 +4,6 @@ import os
 import sys
 import argparse
 
-
 def main():
     target_folder = os.path.join(args.datapath, args.folder)
     target = args.folder
@@ -38,7 +37,7 @@ if __name__ == "__main__":
     parser.add_argument('folder', metavar="<train|test>", type=str)
     args = parser.parse_args()
 
-    MATERIAL_PATH = 'data/material'
+    MATERIAL_PATH = '.data/material'
     if not os.path.exists(MATERIAL_PATH):
         os.makedirs(MATERIAL_PATH)
 

### 2019-03-14 code/glow/builder.py
move files: previous location deletion

@@ -1,137 +1,0 @@
-import re, os
-import copy
-import torch
-from collections import defaultdict
-from . import learning_rate_schedule
-from .config import JsonConfig
-from .models import LatMM, GenMM
-from .utils import load, save, get_proper_device
-
-
-def build_adam(params, args):
-    return torch.optim.Adam(params, **args)
-
-
-__build_optim_dict = {
-    "adam": build_adam
-}
-
-# custom weights initialization called on netG and netD
-def weights_init(m):
-    classname = m.__class__.__name__
-    if classname.find('Conv') != -1:
-        pass
-        #m.weight.data.normal_(0.0, 0.01)
-    elif classname.find('BatchNorm') != -1:
-        m.weight.data.normal_(1.0, 0.02)
-        m.bias.data.fill_(0)
-
-
-def build(hparams, is_training):
-    if isinstance(hparams, str):
-        hparams = JsonConfig(hparams)
-    # get graph and criterions from build function
-    graph, optim, lrschedule, criterion_dict = None, None, None, None  # init with None
-    graph_prior = None
-    cpu, devices = "cpu", None
-    get_loss = None
-    # 1. build graph and criterion_dict, (on cpu)
-    # build and append `device attr` to graph
-    if hparams.Mixture.naive:
-        graph_dict = {"old": GenMM(hparams), "new": GenMM(hparams)}
-        # for component in graph.flow_list:
-        #     component.apply(weights_init)
-    else:
-        graph_dict = {"old": LatMM(hparams), "new": LatMM(hparams)}
-        #graph.get_component().apply(weights_init)
-    
-    for key, graph in graph_dict.items():
-        graph.device = hparams.Device.glow
-        if graph is not None:
-            # get device
-            devices = get_proper_device(graph.device)
-            graph.device = devices
-            graph.to(cpu)
-        # 2. get optim (on cpu)
-        try:
-            if graph is not None and is_training and key is "new":
-                optim_name = hparams.Optim.name
-                optim = __build_optim_dict[optim_name](graph.parameters(), hparams.Optim.args.to_dict())
-                print("[Builder]: Using optimizer `{}`, with args:{}".format(optim_name, hparams.Optim.args))
-                # get lrschedule
-                schedule_name = "default"
-                schedule_args = {}
-                if "Schedule" in hparams.Optim:
-                    schedule_name = hparams.Optim.Schedule.name
-                    schedule_args = hparams.Optim.Schedule.args.to_dict()
[…]

### 2019-03-14 code/glow/models.py

@@ -1,521 +1,0 @@
-import torch
-import torch.nn as nn
-import torch.nn.functional as F
-import numpy as np
-from tqdm import tqdm
-from . import thops
-from . import modules
-from . import utils
-from scipy import stats
-
-def prior_sampler(pk, size=64):
-        pk = tuple(pk.numpy())
-        xk = np.arange(len(pk))
-        custm = stats.rv_discrete(name='custm', values=(xk, pk))
-        prior_samples  = custm.rvs(size=size)
-        prior_samples = torch.from_numpy(prior_samples)
-        return prior_samples
-    
-
-
-def f(in_channels, out_channels, hidden_channels):
-    return nn.Sequential(
-        modules.Conv2d(in_channels, hidden_channels), nn.ReLU(inplace=False),
-        modules.Conv2d(hidden_channels, hidden_channels, kernel_size=[1, 1]), nn.ReLU(inplace=False),
-        modules.Conv2dZeros(hidden_channels, out_channels))
-
-
-class FlowStep(nn.Module):
-    FlowCoupling = ["additive", "affine"]
-    FlowPermutation = {
-        "reverse": lambda obj, z, logdet, rev: (obj.reverse(z, rev), logdet),
-        "shuffle": lambda obj, z, logdet, rev: (obj.shuffle(z, rev), logdet),
-        "invconv": lambda obj, z, logdet, rev: obj.invconv(z, logdet, rev)
-    }
-
-    def __init__(self, in_channels, hidden_channels,
-                 actnorm_scale=1.0,
-                 flow_permutation="invconv",
-                 flow_coupling="additive",
-                 LU_decomposed=False):
-        # check configures
-        assert flow_coupling in FlowStep.FlowCoupling,\
-            "flow_coupling should be in `{}`".format(FlowStep.FlowCoupling)
-        assert flow_permutation in FlowStep.FlowPermutation,\
-            "float_permutation should be in `{}`".format(
-                FlowStep.FlowPermutation.keys())
-        super().__init__()
-        self.flow_permutation = flow_permutation
-        self.flow_coupling = flow_coupling
-        # 1. actnorm
-        self.actnorm = modules.ActNorm2d(in_channels, actnorm_scale)
-        # 2. permute
-        if flow_permutation == "invconv":
-            self.invconv = modules.InvertibleConv1x1(
-                in_channels, LU_decomposed=LU_decomposed)
-        elif flow_permutation == "shuffle":
-            self.shuffle = modules.Permute2d(in_channels, shuffle=True)
-        else:
[…]

### 2019-03-14 code/glow/modules.py

@@ -1,372 +1,0 @@
-import torch
-import torch.nn as nn
-import torch.nn.functional as F
-import numpy as np
-import scipy.linalg
-from . import thops
-
-
-class _ActNorm(nn.Module):
-    """
-    Activation Normalization
-    Initialize the bias and scale with a given minibatch,
-    so that the output per-channel have zero mean and unit variance for that.
-
-    After initialization, `bias` and `logs` will be trained as parameters.
-    """
-
-    def __init__(self, num_features, scale=1.):
-        super().__init__()
-        # register mean and scale
-        size = [1, num_features, 1, 1]
-        self.register_parameter("bias", nn.Parameter(torch.zeros(*size)))
-        self.register_parameter("logs", nn.Parameter(torch.zeros(*size)))
-        self.num_features = num_features
-        self.scale = float(scale)
-        self.inited = False
-
-    def _check_input_dim(self, input):
-        return NotImplemented
-
-    def initialize_parameters(self, input):
-        self._check_input_dim(input)
-        if not self.training:
-            return
-        assert input.device == self.bias.device
-        with torch.no_grad():
-            bias = thops.mean(input.clone(), dim=[0, 2, 3], keepdim=True) * -1.0
-            vars = thops.mean((input.clone() + bias) ** 2, dim=[0, 2, 3], keepdim=True)
-            logs = torch.log(self.scale/(torch.sqrt(vars)+1e-6))
-            self.bias.data.copy_(bias.data)
-            self.logs.data.copy_(logs.data)
-            self.inited = True
-
-    def _center(self, input, reverse=False):
-        if not reverse:
-            return input + self.bias
-        else:
-            return input - self.bias
-
-    def _scale(self, input, logdet=None, reverse=False):
-        logs = self.logs
-        if not reverse:
-            input = input * torch.exp(logs)
-        else:
-            input = input * torch.exp(-logs)
-        if logdet is not None:
-            """
-            logs is log_std of `mean of channels`
-            so we need to multiply pixels
-            """
-            dlogdet = thops.sum(logs) * thops.pixels(input)
-            if reverse:
-                dlogdet *= -1
-            logdet = logdet + dlogdet
-        return input, logdet
-
-    def forward(self, input, logdet=None, reverse=False):
-        if not self.inited:
[…]

### 2019-03-14 code/glow/utils.py

@@ -1,243 +1,0 @@
-import os
-import re
-import copy
-import torch
-import numpy as np
-import matplotlib
-matplotlib.use("Agg")
-import matplotlib.pyplot as plt
-from shutil import copyfile
-
-def save_image(true_dist, fake_dist, img_dir, step):
-    """
-    Generates and saves a plot of the true distribution, the generator, and the
-    critic.
-    """
-    true_dist = true_dist.squeeze()
-    fake_dist = fake_dist.squeeze()
-    fake_dist = fake_dist.cpu().data.numpy()
-    true_dist = true_dist.cpu().data.numpy()
-    plt.clf()
-    plt.scatter(true_dist[:, 0], true_dist[:, 1], c='orange', marker='+')
-    plt.scatter(fake_dist[:, 0], fake_dist[:, 1], c='green', marker='+')
-    plt.savefig(img_dir + '/frame' + '_' + str(step) + '.jpg')
-
-
-def get_proper_cuda_device(device, verbose=True):
-    if not isinstance(device, list):
-        device = [device]
-    count = torch.cuda.device_count()
-    if verbose:
-        print("[Builder]: Found {} gpu".format(count))
-    for i in range(len(device)):
-        d = device[i]
-        did = None
-        if isinstance(d, str):
-            if re.search("cuda:[\d]+", d):
-                did = int(d[5:])
-        elif isinstance(d, int):
-            did = d
-        if did is None:
-            raise ValueError("[Builder]: Wrong cuda id {}".format(d))
-        if did < 0 or did >= count:
-            if verbose:
-                print("[Builder]: {} is not found, ignore.".format(d))
-            device[i] = None
-        else:
-            device[i] = did
-    device = [d for d in device if d is not None]
-    return device
-
-
-def get_proper_device(devices, verbose=True):
-    origin = copy.copy(devices)
-    devices = copy.copy(devices)
-    if not isinstance(devices, list):
-        devices = [devices]
-    use_cpu = any([d.find("cpu")>=0 for d in devices])
-    use_gpu = any([(d.find("cuda")>=0 or isinstance(d, int)) for d in devices])
-    assert not (use_cpu and use_gpu), "{} contains cpu and cuda device.".format(devices)
-    if use_gpu:
-        devices = get_proper_cuda_device(devices, verbose)
-        if len(devices) == 0:
-            if verbose:
[…]

### 2019-03-14 code/hmmlearn/doc/source/conf.py

@@ -1,164 +1,0 @@
-# -*- coding: utf-8 -*-
-
-from __future__ import unicode_literals
-
-
-extensions = [
-    'sphinx.ext.autodoc',
-    'sphinx.ext.autosummary',
-    'sphinx.ext.mathjax',
-    'sphinx.ext.napoleon',
-    'matplotlib.sphinxext.plot_directive',
-    'sphinx_gallery.gen_gallery',
-]
-
-autodoc_default_flags = ['members', 'inherited-members']
-
-sphinx_gallery_conf = {
-    'examples_dirs': '../../examples',
-    'gallery_dirs': 'auto_examples'
-}
-
-# Add any paths that contain templates here, relative to this directory.
-templates_path = ['_templates']
-
-# generate autosummary even if no references
-# autosummary_generate = False
-
-# The suffix of source filenames.
-source_suffix = '.rst'
-
-# The encoding of source files.
-source_encoding = 'utf-8'
-
-# Generate the plots for the gallery
-#plot_gallery = True
-
-# The master toctree document.
-master_doc = 'index'
-
-# General information about the project.
-project = 'hmmlearn'
-copyright = '2010-present, hmmlearn developers (BSD License)'
-
-# The version info for the project you're documenting, acts as replacement for
-# |version| and |release|, also used in various other places throughout the
-# built documents.
-#
-# The short X.Y version.
-# The full version, including alpha/beta/rc tags.
-import hmmlearn
-version = release = hmmlearn.__version__
-
-language = 'en'
-
-exclude_trees = ['_build', 'includes']
-
-# The reST default role (used for this markup: `text`) to use for all
-# documents.
-#default_role = None
-
-# If true, '()' will be appended to :func: etc. cross-reference text.
-add_function_parentheses = False
-
-# If true, the current module name will be prepended to all description
-# unit titles (such as .. function::).
-#add_module_names = True
-
-# If true, sectionauthor and moduleauthor directives will be shown in the
-# output. They are ignored by default.
-#show_authors = False
-
-# The name of the Pygments (syntax highlighting) style to use.
-pygments_style = 'sphinx'
-
-# A list of ignored prefixes for module index sorting.
-#modindex_common_prefix = []
-
-# -- Options for HTML output -------------------------------------------------
-
-## Read the docs style:
-try:
-    import sphinx_rtd_theme
-except ImportError:
-    html_theme = 'classic'
-else:
[…]

### 2019-03-14 code/hmmlearn/examples/plot_hmm_sampling.py

@@ -1,61 +1,0 @@
-"""
-Sampling from HMM
------------------
-
-This script shows how to sample points from a Hiden Markov Model (HMM):
-we use a 4-components with specified mean and covariance.
-
-The plot show the sequence of observations generated with the transitions
-between them. We can see that, as specified by our transition matrix,
-there are no transition between component 1 and 3.
-"""
-print(__doc__)
-
-import numpy as np
-import matplotlib.pyplot as plt
-
-from hmmlearn import hmm
-
-##############################################################
-# Prepare parameters for a 4-components HMM
-# Initial population probability
-startprob = np.array([0.6, 0.3, 0.1, 0.0])
-# The transition matrix, note that there are no transitions possible
-# between component 1 and 3
-transmat = np.array([[0.7, 0.2, 0.0, 0.1],
-                     [0.3, 0.5, 0.2, 0.0],
-                     [0.0, 0.3, 0.5, 0.2],
-                     [0.2, 0.0, 0.2, 0.6]])
-# The means of each component
-means = np.array([[0.0,  0.0],
-                  [0.0, 11.0],
-                  [9.0, 10.0],
-                  [11.0, -1.0]])
-# The covariance of each component
-covars = .5 * np.tile(np.identity(2), (4, 1, 1))
-
-# Build an HMM instance and set parameters
-model = hmm.GaussianHMM(n_components=4, covariance_type="full")
-
-# Instead of fitting it from the data, we directly set the estimated
-# parameters, the means and covariance of the components
-model.startprob_ = startprob
-model.transmat_ = transmat
-model.means_ = means
-model.covars_ = covars
-###############################################################
-
-# Generate samples
-X, Z = model.sample(500)
-
-# Plot the sampled data
-plt.plot(X[:, 0], X[:, 1], ".-", label="observations", ms=6,
-         mfc="orange", alpha=0.7)
-
-# Indicate the component numbers
-for i, m in enumerate(means):
-    plt.text(m[0], m[1], 'Component %i' % (i + 1),
-             size=17, horizontalalignment='center',
-             bbox=dict(alpha=.7, facecolor='w'))
-plt.legend(loc='best')
-plt.show()

### 2019-03-14 code/hmmlearn/lib/hmmlearn/__init__.py

@@ -1,18 +1,0 @@
-"""
-hmmlearn
-========
-
-``hmmlearn`` is a set of algorithms for learning and inference of
-Hiden Markov Models.
-"""
-
-try:
-    import setuptools_scm
-    __version__ = setuptools_scm.get_version(  # xref setup.py
-        root="../..", relative_to=__file__,
-        version_scheme="post-release", local_scheme="node-and-date")
-except (ImportError, LookupError):
-    try:
-        from ._version import version as __version__
-    except ImportError:
-        pass

### 2019-03-14 code/hmmlearn/lib/hmmlearn/_hmmc.pyx

@@ -1,157 +1,0 @@
-# cython: language_level=3, boundscheck=False, wraparound=False
-
-from cython cimport view
-from numpy.math cimport expl, logl, log1pl, isinf, fabsl, INFINITY
-
-import numpy as np
-
-ctypedef double dtype_t
-
-
-cdef inline int _argmax(dtype_t[:] X) nogil:
-    cdef dtype_t X_max = -INFINITY
-    cdef int pos = 0
-    cdef int i
-    for i in range(X.shape[0]):
-        if X[i] > X_max:
-            X_max = X[i]
-            pos = i
-    return pos
-
-
-cdef inline dtype_t _max(dtype_t[:] X) nogil:
-    return X[_argmax(X)]
-
-
-cdef inline dtype_t _logsumexp(dtype_t[:] X) nogil:
-    cdef dtype_t X_max = _max(X)
-    if isinf(X_max):
-        return -INFINITY
-
-    cdef dtype_t acc = 0
-    for i in range(X.shape[0]):
-        acc += expl(X[i] - X_max)
-
-    return logl(acc) + X_max
-
-
-cdef inline dtype_t _logaddexp(dtype_t a, dtype_t b) nogil:
-    if isinf(a) and a < 0:
-        return b
-    elif isinf(b) and b < 0:
-        return a
-    else:
-        return max(a, b) + log1pl(expl(-fabsl(a - b)))
-
-
-def _forward(int n_samples, int n_components,
-             dtype_t[:] log_startprob,
-             dtype_t[:, :] log_transmat,
-             dtype_t[:, :] framelogprob,
-             dtype_t[:, :] fwdlattice):
-
-    cdef int t, i, j
-    cdef dtype_t[::view.contiguous] work_buffer = np.zeros(n_components)
-
-    with nogil:
-        for i in range(n_components):
-            fwdlattice[0, i] = log_startprob[i] + framelogprob[0, i]
-
-        for t in range(1, n_samples):
-            for j in range(n_components):
-                for i in range(n_components):
-                    work_buffer[i] = fwdlattice[t - 1, i] + log_transmat[i, j]
-
-                fwdlattice[t, j] = _logsumexp(work_buffer) + framelogprob[t, j]
-
-
-def _backward(int n_samples, int n_components,
-              dtype_t[:] log_startprob,
-              dtype_t[:, :] log_transmat,
-              dtype_t[:, :] framelogprob,
[…]

### 2019-03-14 code/hmmlearn/lib/hmmlearn/_utils.py

@@ -1,61 +1,0 @@
-"""Private utilities."""
-
-import numpy as np
-
-
-# Copied from scikit-learn 0.19.
-def _validate_covars(covars, covariance_type, n_components):
-    """Do basic checks on matrix covariance sizes and values."""
-    from scipy import linalg
-    if covariance_type == 'spherical':
-        if len(covars) != n_components:
-            raise ValueError("'spherical' covars have length n_components")
-        elif np.any(covars <= 0):
-            raise ValueError("'spherical' covars must be non-negative")
-    elif covariance_type == 'tied':
-        if covars.shape[0] != covars.shape[1]:
-            raise ValueError("'tied' covars must have shape (n_dim, n_dim)")
-        elif (not np.allclose(covars, covars.T)
-              or np.any(linalg.eigvalsh(covars) <= 0)):
-            raise ValueError("'tied' covars must be symmetric, "
-                             "positive-definite")
-    elif covariance_type == 'diag':
-        if len(covars.shape) != 2:
-            raise ValueError("'diag' covars must have shape "
-                             "(n_components, n_dim)")
-        elif np.any(covars <= 0):
-            raise ValueError("'diag' covars must be non-negative")
-    elif covariance_type == 'full':
-        if len(covars.shape) != 3:
-            raise ValueError("'full' covars must have shape "
-                             "(n_components, n_dim, n_dim)")
-        elif covars.shape[1] != covars.shape[2]:
-            raise ValueError("'full' covars must have shape "
-                             "(n_components, n_dim, n_dim)")
-        for n, cv in enumerate(covars):
-            if (not np.allclose(cv, cv.T)
-                    or np.any(linalg.eigvalsh(cv) <= 0)):
-                raise ValueError("component %d of 'full' covars must be "
-                                 "symmetric, positive-definite" % n)
-    else:
-        raise ValueError("covariance_type must be one of " +
-                         "'spherical', 'tied', 'diag', 'full'")
-
-
-# Copied from scikit-learn 0.19.
-def distribute_covar_matrix_to_match_covariance_type(
-        tied_cv, covariance_type, n_components):
-    """Create all the covariance matrices from a given template."""
-    if covariance_type == 'spherical':
-        cv = np.tile(tied_cv.mean() * np.ones(tied_cv.shape[1]),
-                     (n_components, 1))
-    elif covariance_type == 'tied':
-        cv = tied_cv
[…]

### 2019-03-14 code/hmmlearn/lib/hmmlearn/base.py

@@ -1,681 +1,0 @@
-from __future__ import print_function
-
-import string
-import sys
-from collections import deque
-
-import numpy as np
-from scipy.special import logsumexp
-from sklearn.base import BaseEstimator, _pprint
-from sklearn.utils import check_array, check_random_state
-from sklearn.utils.validation import check_is_fitted
-
-from . import _hmmc
-from .utils import normalize, log_normalize, iter_from_X_lengths, log_mask_zero
-
-
-#: Supported decoder algorithms.
-DECODER_ALGORITHMS = frozenset(("viterbi", "map"))
-
-
-class ConvergenceMonitor(object):
-    """Monitors and reports convergence to :data:`sys.stderr`.
-
-    Parameters
-    ----------
-    tol : double
-        Convergence threshold. EM has converged either if the maximum
-        number of iterations is reached or the log probability
-        improvement between the two consecutive iterations is less
-        than threshold.
-
-    n_iter : int
-        Maximum number of iterations to perform.
-
-    verbose : bool
-        If ``True`` then per-iteration convergence reports are printed,
-        otherwise the monitor is mute.
-
-    Attributes
-    ----------
-    history : deque
-        The log probability of the data for the last two training
-        iterations. If the values are not strictly increasing, the
-        model did not converge.
-
-    iter : int
-        Number of iterations performed while training the model.
-
-    Examples
-    --------
-    Use custom convergence criteria by subclassing ``ConvergenceMonitor``
-    and redefining the ``converged`` method. The resulting subclass can
-    be used by creating an instance and pointing a model's ``monitor_``
-    attribute to it prior to fitting.
-
-    >>> from hmmlearn.base import ConvergenceMonitor
-    >>> from hmmlearn import hmm
-    >>>
-    >>> class ThresholdMonitor(ConvergenceMonitor):
-    ...     @property
-    ...     def converged(self):
-    ...         return (self.iter == self.n_iter or
-    ...                 self.history[-1] >= self.tol)
-    >>>
-    >>> model = hmm.GaussianHMM(n_components=2, tol=5, verbose=True)
-    >>> model.monitor_ = ThresholdMonitor(model.monitor_.tol,
-    ...                                   model.monitor_.n_iter,
-    ...                                   model.monitor_.verbose)
-    """
-    _template = "{iter:>10d} {logprob:>16.4f} {delta:>+16.4f}"
-
-    def __init__(self, tol, n_iter, verbose):
[…]

### 2019-03-14 code/hmmlearn/lib/hmmlearn/hmm.py

@@ -1,1206 +1,0 @@
-# Hidden Markov Models
-#
-# Author: Ron Weiss <ronweiss@gmail.com>
-#         Shiqiao Du <lucidfrontier.45@gmail.com>
-# API changes: Jaques Grobler <jaquesgrobler@gmail.com>
-# Modifications to create of the HMMLearn module: Gael Varoquaux
-# More API changes: Sergei Lebedev <superbobry@gmail.com>
-
-"""
-The :mod:`hmmlearn.hmm` module implements hidden Markov models.
-"""
-
-import numpy as np
-from scipy.special import logsumexp
-from sklearn import cluster
-from sklearn.utils import check_random_state
-
-from . import _utils
-from .stats import log_multivariate_normal_density
-from .base import _BaseHMM
-from .utils import iter_from_X_lengths, normalize, fill_covars
-
-__all__ = ["GMMHMM", "GaussianHMM", "MultinomialHMM", "GenHMM"]
-
-COVARIANCE_TYPES = frozenset(("spherical", "diag", "full", "tied"))
-
-
-class GaussianHMM(_BaseHMM):
-    r"""Hidden Markov Model with Gaussian emissions.
-
-    Parameters
-    ----------
-    n_components : int
-        Number of states.
-
-    covariance_type : string, optional
-        String describing the type of covariance parameters to
-        use.  Must be one of
-
-        * "spherical" --- each state uses a single variance value that
-          applies to all features.
-        * "diag" --- each state uses a diagonal covariance matrix.
-        * "full" --- each state uses a full (i.e. unrestricted)
-          covariance matrix.
-        * "tied" --- all states use **the same** full covariance matrix.
-
-        Defaults to "diag".
-
-    min_covar : float, optional
-        Floor on the diagonal of the covariance matrix to prevent
-        overfitting. Defaults to 1e-3.
-
-    startprob_prior : array, shape (n_components, ), optional
-        Parameters of the Dirichlet prior distribution for
-        :attr:`startprob_`.
-
-    transmat_prior : array, shape (n_components, n_components), optional
-        Parameters of the Dirichlet prior distribution for each row
-        of the transition probabilities :attr:`transmat_`.
-
-    means_prior, means_weight : array, shape (n_components, ), optional
-        Mean and precision of the Normal prior distribtion for
-        :attr:`means_`.
-
-    covars_prior, covars_weight : array, shape (n_components, ), optional
[…]

### 2019-03-14 code/hmmlearn/lib/hmmlearn/stats.py

@@ -1,90 +1,0 @@
-import numpy as np
-from scipy import linalg
-
-
-def log_multivariate_normal_density(X, means, covars, covariance_type='diag'):
-    """Compute the log probability under a multivariate Gaussian distribution.
-    Parameters
-    ----------
-    X : array_like, shape (n_samples, n_features)
-        List of n_features-dimensional data points. Each row corresponds to a
-        single data point.
-    means : array_like, shape (n_components, n_features)
-        List of n_features-dimensional mean vectors for n_components Gaussians.
-        Each row corresponds to a single mean vector.
-    covars : array_like
-        List of n_components covariance parameters for each Gaussian. The shape
-        depends on `covariance_type`:
-            (n_components, n_features)      if 'spherical',
-            (n_features, n_features)    if 'tied',
-            (n_components, n_features)    if 'diag',
-            (n_components, n_features, n_features) if 'full'
-    covariance_type : string
-        Type of the covariance parameters.  Must be one of
-        'spherical', 'tied', 'diag', 'full'.  Defaults to 'diag'.
-    Returns
-    -------
-    lpr : array_like, shape (n_samples, n_components)
-        Array containing the log probabilities of each data point in
-        X under each of the n_components multivariate Gaussian distributions.
-    """
-    log_multivariate_normal_density_dict = {
-        'spherical': _log_multivariate_normal_density_spherical,
-        'tied': _log_multivariate_normal_density_tied,
-        'diag': _log_multivariate_normal_density_diag,
-        'full': _log_multivariate_normal_density_full}
-    return log_multivariate_normal_density_dict[covariance_type](
-        X, means, covars
-    )
-
-
-def _log_multivariate_normal_density_diag(X, means, covars):
-    """Compute Gaussian log-density at X for a diagonal model."""
-    n_samples, n_dim = X.shape
-    lpr = -0.5 * (n_dim * np.log(2 * np.pi) + np.sum(np.log(covars), 1)
-                  + np.sum((means ** 2) / covars, 1)
-                  - 2 * np.dot(X, (means / covars).T)
-                  + np.dot(X ** 2, (1.0 / covars).T))
-    return lpr
-
-
-def _log_multivariate_normal_density_spherical(X, means, covars):
-    """Compute Gaussian log-density at X for a spherical model."""
-    cv = covars.copy()
-    if covars.ndim == 1:
[…]

### 2019-03-14 code/hmmlearn/lib/hmmlearn/tests/__init__.py

@@ -1,48 +1,0 @@
-import numpy as np
-from sklearn.datasets.samples_generator import make_spd_matrix
-from sklearn.utils import check_random_state
-
-from hmmlearn.utils import normalize
-
-# Make NumPy complain about underflows/overflows etc.
-np.seterr(all="warn")
-
-
-def make_covar_matrix(covariance_type, n_components, n_features,
-                      random_state=None):
-    mincv = 0.1
-    prng = check_random_state(random_state)
-    if covariance_type == 'spherical':
-        return (mincv + mincv * prng.random_sample((n_components,))) ** 2
-    elif covariance_type == 'tied':
-        return (make_spd_matrix(n_features)
-                + mincv * np.eye(n_features))
-    elif covariance_type == 'diag':
-        return (mincv + mincv *
-                prng.random_sample((n_components, n_features))) ** 2
-    elif covariance_type == 'full':
-        return np.array([
-            (make_spd_matrix(n_features, random_state=prng)
-             + mincv * np.eye(n_features))
-            for x in range(n_components)
-        ])
-
-
-def normalized(X, axis=None):
-    X_copy = X.copy()
-    normalize(X_copy, axis=axis)
-    return X_copy
-
-
-def log_likelihood_increasing(h, X, lengths, n_iter):
-    h.n_iter = 1        # make sure we do a single iteration at a time
-    h.init_params = ''  # and don't re-init params
-    log_likelihoods = np.empty(n_iter, dtype=float)
-    for i in range(n_iter):
-        h.fit(X, lengths=lengths)
-        log_likelihoods[i] = h.score(X, lengths=lengths)
-
-    # XXX the rounding is necessary because LL can oscillate in the
-    #     fractional part, failing the tests.
-    diff = np.round(np.diff(log_likelihoods), 10) >= 0
-    return diff.all()

### 2019-03-14 code/hmmlearn/lib/hmmlearn/tests/conftest.py

@@ -1,12 +1,0 @@
-import os
-import random
-
-import numpy as np
-
-
-def pytest_configure(config):
-    _random_seed = int(os.environ.get("HMMLEARN_SEED",
-                                      np.random.uniform() * (2**31 - 1)))
-    print("set RNG seed to {}".format(_random_seed))
-    np.random.seed(_random_seed)
-    random.seed(_random_seed)

### 2019-03-14 code/hmmlearn/lib/hmmlearn/tests/test_base.py

@@ -1,198 +1,0 @@
-import numpy as np
-import pytest
-
-from hmmlearn.base import _BaseHMM, ConvergenceMonitor
-from hmmlearn.utils import logsumexp
-
-
-class TestMonitor(object):
-    def test_converged_by_iterations(self):
-        m = ConvergenceMonitor(tol=1e-3, n_iter=2, verbose=False)
-        assert not m.converged
-        m.report(-0.01)
-        assert not m.converged
-        m.report(-0.1)
-        assert m.converged
-
-    def test_converged_by_logprob(self):
-        m = ConvergenceMonitor(tol=1e-3, n_iter=10, verbose=False)
-        for logprob in [-0.03, -0.02, -0.01]:
-            m.report(logprob)
-            assert not m.converged
-
-        m.report(-0.0101)
-        assert m.converged
-
-    def test_reset(self):
-        m = ConvergenceMonitor(tol=1e-3, n_iter=10, verbose=False)
-        m.iter = 1
-        m.history.append(-0.01)
-        m._reset()
-        assert m.iter == 0
-        assert not m.history
-
-    def test_report_first_iteration(self, capsys):
-        m = ConvergenceMonitor(tol=1e-3, n_iter=10, verbose=True)
-        m.report(-0.01)
-        out, err = capsys.readouterr()
-        assert not out
-        expected = m._template.format(iter=1, logprob=-0.01, delta=np.nan)
-        assert err.splitlines() == [expected]
-
-    def test_report(self, capsys):
-        n_iter = 10
-        m = ConvergenceMonitor(tol=1e-3, n_iter=n_iter, verbose=True)
-        for i in reversed(range(n_iter)):
-            m.report(-0.01 * i)
-
-        out, err = capsys.readouterr()
-        assert not out
-        assert len(err.splitlines()) == n_iter
-
-
-class StubHMM(_BaseHMM):
-    """An HMM with hardcoded observation probabilities."""
-    def _compute_log_likelihood(self, X):
-        return self.framelogprob
-
-
-class TestBaseAgainstWikipedia(object):
-    def setup_method(self, method):
-        # Example from http://en.wikipedia.org/wiki/Forward-backward_algorithm
-        self.framelogprob = np.log([[0.9, 0.2],
-                                    [0.9, 0.2],
-                                    [0.1, 0.8],
-                                    [0.9, 0.2],
-                                    [0.9, 0.2]])
-
-        h = StubHMM(2)
-        h.transmat_ = [[0.7, 0.3], [0.3, 0.7]]
[…]

### 2019-03-14 code/hmmlearn/lib/hmmlearn/tests/test_gaussian_hmm.py

@@ -1,235 +1,0 @@
-from __future__ import absolute_import
-
-from unittest import TestCase
-
-import numpy as np
-import pytest
-
-from hmmlearn import hmm
-
-from . import log_likelihood_increasing, make_covar_matrix, normalized
-
-
-class GaussianHMMTestMixin(object):
-    covariance_type = None  # set by subclasses
-
-    def setUp(self):
-        self.prng = prng = np.random.RandomState(10)
-        self.n_components = n_components = 3
-        self.n_features = n_features = 3
-        self.startprob = prng.rand(n_components)
-        self.startprob = self.startprob / self.startprob.sum()
-        self.transmat = prng.rand(n_components, n_components)
-        self.transmat /= np.tile(self.transmat.sum(axis=1)[:, np.newaxis],
-                                 (1, n_components))
-        self.means = prng.randint(-20, 20, (n_components, n_features))
-        self.covars = make_covar_matrix(
-            self.covariance_type, n_components, n_features, random_state=prng
-        )
-
-    def test_bad_covariance_type(self):
-        with pytest.raises(ValueError):
-            h = hmm.GaussianHMM(20, covariance_type='badcovariance_type')
-            h.means_ = self.means
-            h.covars_ = []
-            h.startprob_ = self.startprob
-            h.transmat_ = self.transmat
-            h._check()
-
-    def test_score_samples_and_decode(self):
-        h = hmm.GaussianHMM(self.n_components, self.covariance_type,
-                            init_params="st")
-        h.means_ = self.means
-        h.covars_ = self.covars
-
-        # Make sure the means are far apart so posteriors.argmax()
-        # picks the actual component used to generate the observations.
-        h.means_ = 20 * h.means_
-
-        gaussidx = np.repeat(np.arange(self.n_components), 5)
-        n_samples = len(gaussidx)
-        X = self.prng.randn(n_samples, self.n_features) + h.means_[gaussidx]
-        h._init(X)
-        ll, posteriors = h.score_samples(X)
-
-        self.assertEqual(posteriors.shape, (n_samples, self.n_components))
-        assert np.allclose(posteriors.sum(axis=1), np.ones(n_samples))
-
-        viterbi_ll, stateseq = h.decode(X)
-        assert np.allclose(stateseq, gaussidx)
-
-    def test_sample(self, n=1000):
[…]

### 2019-03-14 code/hmmlearn/lib/hmmlearn/tests/test_gmm_hmm.py

@@ -1,116 +1,0 @@
-from __future__ import absolute_import
-
-import numpy as np
-import pytest
-from sklearn.mixture import GaussianMixture
-from sklearn.utils import check_random_state
-
-from hmmlearn import hmm
-
-from . import log_likelihood_increasing, make_covar_matrix, normalized
-
-pytestmark = pytest.mark.xfail()
-
-
-def create_random_gmm(n_mix, n_features, covariance_type, prng=0):
-    prng = check_random_state(prng)
-    g = GaussianMixture(n_mix, covariance_type=covariance_type)
-    g.means_ = prng.randint(-20, 20, (n_mix, n_features))
-    g.covars_ = make_covar_matrix(covariance_type, n_mix, n_features)
-    g.weights_ = normalized(prng.rand(n_mix))
-    return g
-
-
-class GMMHMMTestMixin(object):
-    def setup_method(self, method):
-        self.prng = np.random.RandomState(9)
-        self.n_components = 3
-        self.n_mix = 2
-        self.n_features = 2
-        self.startprob = self.prng.rand(self.n_components)
-        self.startprob = self.startprob / self.startprob.sum()
-        self.transmat = normalized(
-            self.prng.rand(self.n_components, self.n_components),
-            axis=1)
-
-        self.gmms = []
-        for state in range(self.n_components):
-            self.gmms.append(create_random_gmm(
-                self.n_mix, self.n_features, self.covariance_type,
-                prng=self.prng))
-
-    def test_score_samples_and_decode(self):
-        h = hmm.GMMHMM(self.n_components, covariance_type=self.covariance_type)
-        h.startprob_ = self.startprob
-        h.transmat_ = self.transmat
-        h.gmms_ = self.gmms
-
-        # Make sure the means are far apart so posteriors.argmax()
-        # picks the actual component used to generate the observations.
-        for g in h.gmms_:
-            g.means_ *= 20
-
-        refstateseq = np.repeat(np.arange(self.n_components), 5)
-        n_samples = len(refstateseq)
-        X = [h.gmms_[x].sample(1).flatten() for x in refstateseq]
-
-        _ll, posteriors = h.score_samples(X)
-
-        assert posteriors.shape == (n_samples, self.n_components)
-        assert np.allclose(posteriors.sum(axis=1), np.ones(n_samples))
-
-        _logprob, stateseq = h.decode(X)
[…]

### 2019-03-14 code/hmmlearn/lib/hmmlearn/tests/test_gmm_hmm_new.py

@@ -1,178 +1,0 @@
-import numpy as np
-import pytest
-
-from . import log_likelihood_increasing
-from . import normalized
-from ..hmm import GMMHMM
-
-
-def sample_from_parallelepiped(low, high, n_samples, random_state):
-    (n_features,) = low.shape
-    X = np.zeros((n_samples, n_features))
-    for i in range(n_features):
-        X[:, i] = random_state.uniform(low[i], high[i], n_samples)
-    return X
-
-
-def prep_params(n_comps, n_mix, n_features, covar_type,
-                low, high, random_state):
-    # the idea is to generate ``n_comps`` bounding boxes and then
-    # generate ``n_mix`` mixture means in each of them
-
-    dim_lims = np.zeros((n_comps + 1, n_features))
-
-    # this generates a sequence of coordinates, which are then used as
-    # vertices of bounding boxes for mixtures
-    dim_lims[1:] = np.cumsum(
-        random_state.uniform(low, high, (n_comps, n_features)), axis=0
-    )
-
-    means = np.zeros((n_comps, n_mix, n_features))
-    for i, (left, right) in enumerate(zip(dim_lims, dim_lims[1:])):
-        means[i] = sample_from_parallelepiped(left, right, n_mix,
-                                              random_state)
-
-    startprob = np.zeros(n_comps)
-    startprob[0] = 1
-
-    transmat = normalized(random_state.uniform(size=(n_comps, n_comps)),
-                          axis=1)
-
-    if covar_type == "spherical":
-        covs = random_state.uniform(0.1, 5, size=(n_comps, n_mix))
-    elif covar_type == "diag":
-        covs = random_state.uniform(0.1, 5, size=(n_comps, n_mix, n_features))
-    elif covar_type == "tied":
-        covs = np.zeros((n_comps, n_features, n_features))
-        for i in range(n_comps):
-            low = random_state.uniform(-2, 2, (n_features, n_features))
-            covs[i] = np.dot(low.T, low)
-    elif covar_type == "full":
-        covs = np.zeros((n_comps, n_mix, n_features, n_features))
-        for i in range(n_comps):
-            for j in range(n_mix):
-                low = random_state.uniform(-2, 2,
-                                           size=(n_features, n_features))
-                covs[i, j] = np.dot(low.T, low)
-
-    weights = normalized(random_state.uniform(size=(n_comps, n_mix)),
[…]

### 2019-03-14 code/hmmlearn/lib/hmmlearn/tests/test_multinomial_hmm.py

@@ -1,132 +1,0 @@
-from __future__ import absolute_import
-
-import numpy as np
-import pytest
-
-from hmmlearn import hmm
-
-from . import log_likelihood_increasing, normalized
-
-
-class TestMultinomialAgainstWikipedia(object):
-    """Using examples from Wikipedia
-
-    - http://en.wikipedia.org/wiki/Hidden_Markov_model
-    - http://en.wikipedia.org/wiki/Viterbi_algorithm
-    """
-    def setup_method(self, method):
-        n_components = 2   # ['Rainy', 'Sunny']
-        n_features = 3     # ['walk', 'shop', 'clean']
-        self.h = hmm.MultinomialHMM(n_components)
-        self.h.n_features = n_features
-        self.h.startprob_ = np.array([0.6, 0.4])
-        self.h.transmat_ = np.array([[0.7, 0.3], [0.4, 0.6]])
-        self.h.emissionprob_ = np.array([[0.1, 0.4, 0.5],
-                                         [0.6, 0.3, 0.1]])
-
-    def test_decode_viterbi(self):
-        # From http://en.wikipedia.org/wiki/Viterbi_algorithm:
-        # "This reveals that the observations ['walk', 'shop', 'clean']
-        #  were most likely generated by states ['Sunny', 'Rainy', 'Rainy'],
-        #  with probability 0.01344."
-        X = [[0], [1], [2]]
-        logprob, state_sequence = self.h.decode(X, algorithm="viterbi")
-        assert round(np.exp(logprob), 5) == 0.01344
-        assert np.allclose(state_sequence, [1, 0, 0])
-
-    def test_decode_map(self):
-        X = [[0], [1], [2]]
-        _logprob, state_sequence = self.h.decode(X, algorithm="map")
-        assert np.allclose(state_sequence, [1, 0, 0])
-
-    def test_predict(self):
-        X = [[0], [1], [2]]
-        state_sequence = self.h.predict(X)
-        posteriors = self.h.predict_proba(X)
-        assert np.allclose(state_sequence, [1, 0, 0])
-        assert np.allclose(posteriors, [
-            [0.23170303, 0.76829697],
-            [0.62406281, 0.37593719],
-            [0.86397706, 0.13602294],
-        ])
-
-
-class TestMultinomailHMM(object):
-    def setup_method(self, method):
-        self.n_components = 2
-        self.n_features = 3
-        self.h = hmm.MultinomialHMM(self.n_components)
-        self.h.startprob_ = np.array([0.6, 0.4])
[…]

### 2019-03-14 code/hmmlearn/lib/hmmlearn/tests/test_utils.py

@@ -1,44 +1,0 @@
-import numpy as np
-
-from hmmlearn.utils import normalize, fill_covars
-
-
-def test_normalize():
-    A = np.random.normal(42., size=128)
-    A[np.random.choice(len(A), size=16)] = 0.0
-    assert (A == 0.0).any()
-    normalize(A)
-    assert np.allclose(A.sum(), 1.)
-
-
-def test_normalize_along_axis():
-    A = np.random.normal(42., size=(128, 4))
-    for axis in range(A.ndim):
-        A[np.random.choice(len(A), size=16), axis] = 0.0
-        assert (A[:, axis] == 0.0).any()
-        normalize(A, axis=axis)
-        assert np.allclose(A.sum(axis=axis), 1.)
-
-
-def test_fill_covars():
-    full = np.arange(12).reshape(3, 2, 2) + 1
-    np.testing.assert_equal(fill_covars(full, 'full', 3, 2), full)
-
-    diag = np.arange(6).reshape(3, 2) + 1
-    expected = np.array([[[1, 0], [0, 2]],
-                         [[3, 0], [0, 4]],
-                         [[5, 0], [0, 6]]])
-    np.testing.assert_equal(fill_covars(diag, 'diag', 3, 2), expected)
-
-    tied = np.arange(4).reshape(2, 2) + 1
-    expected = np.array([[[1, 2], [3, 4]],
-                         [[1, 2], [3, 4]],
-                         [[1, 2], [3, 4]]])
-    np.testing.assert_equal(fill_covars(tied, 'tied', 3, 2), expected)
-
-    spherical = np.array([1, 2, 3])
-    expected = np.array([[[1, 0], [0, 1]],
-                         [[2, 0], [0, 2]],
-                         [[3, 0], [0, 3]]])
-    np.testing.assert_equal(
-        fill_covars(spherical, 'spherical', 3, 2), expected)

### 2019-03-14 code/hmmlearn/lib/hmmlearn/utils.py

@@ -1,92 +1,0 @@
-import numpy as np
-from scipy.special import logsumexp
-
-
-def normalize(a, axis=None):
-    """Normalizes the input array so that it sums to 1.
-
-    Parameters
-    ----------
-    a : array
-        Non-normalized input data.
-
-    axis : int
-        Dimension along which normalization is performed.
-
-    Notes
-    -----
-    Modifies the input **inplace**.
-    """
-    a_sum = a.sum(axis)
-    if axis and a.ndim > 1:
-        # Make sure we don't divide by zero.
-        a_sum[a_sum == 0] = 1
-        shape = list(a.shape)
-        shape[axis] = 1
-        a_sum.shape = shape
-
-    a /= a_sum
-
-
-def log_normalize(a, axis=None):
-    """Normalizes the input array so that the exponent of the sum is 1.
-
-    Parameters
-    ----------
-    a : array
-        Non-normalized input data.
-
-    axis : int
-        Dimension along which normalization is performed.
-
-    Notes
-    -----
-    Modifies the input **inplace**.
-    """
-    with np.errstate(under="ignore"):
-        a_lse = logsumexp(a, axis)
-    a -= a_lse[:, np.newaxis]
-
-
-def iter_from_X_lengths(X, lengths):
-    if lengths is None:
-        yield 0, len(X)
-    else:
-        n_samples = X.shape[0]
-        end = np.cumsum(lengths).astype(np.int32)
-        start = end - lengths
-        if end[-1] > n_samples:
-            raise ValueError("more than {:d} samples in lengths array {!s}"
-                             .format(n_samples, lengths))
-
-        for i in range(len(lengths)):
-            yield start[i], end[i]
-
-
-def log_mask_zero(a):
-    """Computes the log of input probabilities masking divide by zero in log.
-
-    Notes
-    -----
-    During the M-step of EM-algorithm, very small intermediate start
-    or transition probabilities could be normalized to zero, causing a
-    *RuntimeWarning: divide by zero encountered in log*.
-
-    This function masks this unharmful warning.
-    """
-    a = np.asarray(a)
-    with np.errstate(divide="ignore"):
-        return np.log(a)
-
-
-def fill_covars(covars, covariance_type='full', n_components=1, n_features=1):
-    if covariance_type == 'full':
-        return covars
-    elif covariance_type == 'diag':
-        return np.array(list(map(np.diag, covars)))
-    elif covariance_type == 'tied':
[…]

### 2019-03-14 code/hmmlearn/setup.py

@@ -1,101 +1,0 @@
-# Copyright (C) 2007-2009 Cournapeau David <cournape@gmail.com>
-#               2010 Fabian Pedregosa <fabian.pedregosa@inria.fr>
-#               2014 Gael Varoquaux
-#               2014-2016 Sergei Lebedev <superbobry@gmail.com>
-#               2018- Antony Lee
-
-
-from distutils.version import LooseVersion
-from io import open
-
-import setuptools
-from setuptools import Extension, find_packages, setup
-from setuptools.command.build_ext import build_ext
-
-
-# Added support for environment markers in install_requires.
-if LooseVersion(setuptools.__version__) < "36.2":
-    raise ImportError("setuptools>=36.2 is required")
-
-
-class build_ext(build_ext, object):
-
-    def finalize_options(self):
-        # The key point: here, Cython and numpy will have been installed by
-        # pip.
-        from Cython.Build import cythonize
-        import numpy as np
-        import numpy.distutils
-
-        self.distribution.ext_modules[:] = cythonize("**/*.pyx")
-        # Sadly, this part needs to be done manually.
-        for ext in self.distribution.ext_modules:
-            for k, v in np.distutils.misc_util.get_info("npymath").items():
-                setattr(ext, k, v)
-            ext.include_dirs = [np.get_include()]
-
-        super(build_ext, self).finalize_options()
-
-    def build_extensions(self):
-        try:
-            self.compiler.compiler_so.remove("-Wstrict-prototypes")
-        except (AttributeError, ValueError):
-            pass
-        super(build_ext, self).build_extensions()
-
-
-setup(
-    name="hmmlearn",
-    description="",
-    long_description=open("README.rst", encoding="utf-8").read(),
-    maintainer="Antony Lee",
-    maintainer_email="",
-    url="https://github.com/hmmlearn/hmmlearn",
-    license="new BSD",
-    classifiers=[
-        "Development Status :: 3 - Alpha",
-        "License :: OSI Approved",
-        "Intended Audience :: Developers",
-        "Intended Audience :: Science/Research",
-        "Topic :: Software Development",
-        "Topic :: Scientific/Engineering",
-        "Programming Language :: Cython",
-        "Programming Language :: Python",
-        "Programming Language :: Python :: 2",
-        "Programming Language :: Python :: 2.7",
[…]

### 2019-03-14 code/infer_mix.py

@@ -1,234 +1,0 @@
-"""Train script.
-Usage:
-    infer.py <hparams> <dataset_root> <mode>
-"""
-import os
-import cv2
-import random
-import torch
-import torchvision.datasets as dset
-import torch
-import numpy as np
-from docopt import docopt
-from torchvision import transforms
-from glow.builder import build
-from glow.config import JsonConfig
-from glow import modules
-import torchvision.utils as vutils
-from scipy import stats
-
-
-def prior_sampler(pk, size=64):
-    """ Define the sample generator for mixture distribution
-    xk = np.arrange(number of components)
-    pk = (0.1, 0.2, 0.3, 0.1, 0.1, 0.0, 0.2), tuple
-    R = custm.rvs(size=100)
-    """
-    xk = np.arange(len(pk))
-    custm = stats.rv_discrete(name='custm', values=(xk, pk))
-    prior_samples  = custm.rvs(size=size)
-    return prior_samples
-
-def GenMMmap_encoding(graph, x):
-    z, nll = graph(x=x, y_onehot=None, reverse=False)
-       
-    nlog_joint_prob =nll - torch.log(pk.unsqueeze(1).expand_as(nll)+1e-6).cuda()
-    tmp_sum = torch.log( torch.sum( torch.exp(-nlog_joint_prob), dim=[0]) ).cuda
-    nlog_gamma = nlog_joint_prob + tmp_sum().expand_as(nlog_joint_prob)
-    _, imin = nlog_gamma.min(dim=0)
-    r_c_index = torch.stack([imin.data.cpu().type(torch.int64), torch.arange(0,64).type(torch.int64)])
-    
-    return z[r_c_index[0], r_c_index[1]], r_c_index
-
-def LatMMmap_encoding(graph, x):
-    graph = graph.get_component()
-    z, _, _ = graph(x=x, y_onehot=None, reverse=False)
-       
-    return z
-
-def interpolation(start, end, points=9):
-    weight = np.linspace(0.1, 0.9, points)
-    inter = []
-    for i in weight:
-        inter.append(torch.lerp(start, end, i))
-    return inter
-
-if __name__ == "__main__":
-    args = docopt(__doc__)
-    hparams = args["<hparams>"]
-    dataset_root = args["<dataset_root>"]
-    mode = args["<mode>"]
-    assert mode in ["Generating", "Interpolation"]
-
-    assert os.path.exists(dataset_root), (
[…]

### 2019-03-14 code/kaldi-io-for-python/build/lib/kaldi_io/__init__.py

@@ -1,16 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-# Copyright 2019  Brno University of Technology (author: Karel Vesely)
-# Licensed under the Apache License, Version 2.0 (the "License")
-
-# we will be importing from '../'
-from __future__ import absolute_import
-
-# just import the main module,
-from kaldi_io.kaldi_io import *
-
-# We might later subdivide the project into more modules,
-# and the interface will stay the same, if all modules
-# are listed in this '__init__.py'.
-

### 2019-03-14 code/kaldi-io-for-python/kaldi_io/kaldi_io.py

@@ -1,670 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-# Copyright 2014-2019 Brno University of Technology (author: Karel Vesely)
-# Licensed under the Apache License, Version 2.0 (the "License")
-
-from __future__ import print_function
-from __future__ import division
-
-import numpy as np
-import sys, os, re, gzip, struct
-
-#################################################
-# Adding kaldi tools to shell path,
-
-# Select kaldi,
-if not 'KALDI_ROOT' in os.environ:
-    # Default! To change run python with 'export KALDI_ROOT=/some_dir python'
-    os.environ['KALDI_ROOT']='/mnt/matylda5/iveselyk/Tools/kaldi-trunk'
-
-# Add kaldi tools to path,
-os.environ['PATH'] = os.popen('echo $KALDI_ROOT/src/bin:$KALDI_ROOT/tools/openfst/bin:$KALDI_ROOT/src/fstbin/:$KALDI_ROOT/src/gmmbin/:$KALDI_ROOT/src/featbin/:$KALDI_ROOT/src/lm/:$KALDI_ROOT/src/sgmmbin/:$KALDI_ROOT/src/sgmm2bin/:$KALDI_ROOT/src/fgmmbin/:$KALDI_ROOT/src/latbin/:$KALDI_ROOT/src/nnetbin:$KALDI_ROOT/src/nnet2bin:$KALDI_ROOT/src/nnet3bin:$KALDI_ROOT/src/online2bin/:$KALDI_ROOT/src/ivectorbin/:$KALDI_ROOT/src/lmbin/').readline().strip() + ':' + os.environ['PATH']
-
-
-#################################################
-# Define all custom exceptions,
-class UnsupportedDataType(Exception): pass
-class UnknownVectorHeader(Exception): pass
-class UnknownMatrixHeader(Exception): pass
-
-class BadSampleSize(Exception): pass
-class BadInputFormat(Exception): pass
-
-class SubprocessFailed(Exception): pass
-
-#################################################
-# Data-type independent helper functions,
-
-def open_or_fd(file, mode='rb'):
-    """ fd = open_or_fd(file)
-     Open file, gzipped file, pipe, or forward the file-descriptor.
-     Eventually seeks in the 'file' argument contains ':offset' suffix.
-    """
-    offset = None
-    try:
-        # strip 'ark:' prefix from r{x,w}filename (optional),
[…]

### 2019-03-14 code/kaldi-io-for-python/make_package.sh

@@ -1,42 +1,0 @@
-#!/bin/bash
-
-exit 0
-
-# Step 1: udpate version in 'setup.py'
-
-# Step 2: make packages,
-# python3 -m pip install --user --upgrade setuptools wheel twine
-rm dist/*
-python2 setup.py bdist_wheel
-python3 setup.py sdist bdist_wheel
-#$ ls dist/
-#    kaldi_io-vesis84-0.9.0.tar.gz
-#    kaldi_io_vesis84-0.9.0-py2-none-any.whl
-#    kaldi_io_vesis84-0.9.0-py3-none-any.whl
-
-# Hint: skip to 'Step 8' to skip sandboxing on 'test.pypi.org'
-
-{ # TEST_DEPLOYMENT_ON test.pypi.org,
-  # Step 3: upload the packages (test site),
-  python3 -m twine upload --verbose --repository-url https://test.pypi.org/legacy/ dist/*
-
-  # Step 4: see webpage,
-  # https://test.pypi.org/project/kaldi_io_vesis84
-
-  # Step 5: try installing it locally,
-  python3 -m pip install --user --index-url https://test.pypi.org/simple/ kaldi_io_vesis84
-
-  # Stepy 6: try to install it,
-  python3
-  <import kaldi_io
-  <print(kaldi_io)
-
-  # Step 7: remove the package,
-  python3 -m pip uninstall kaldi_io_vesis84
-}
-
-# Step 8: Put the packages to 'production' pypi,
-python3 -m twine upload --verbose dist/*
-python3 -m pip install --user kaldi_io
-python3 -m pip uninstall kaldi_io
-

### 2019-03-14 code/kaldi-io-for-python/run_tests.sh

@@ -1,7 +1,0 @@
-#!/bin/bash
-
-# test code for python 2,
-python2 -m unittest discover -s tests
-
-# test code for python 3,
-python3 -m unittest discover -s tests

### 2019-03-14 code/kaldi-io-for-python/setup.py

@@ -1,31 +1,0 @@
-#!/usr/bin/env python
-# -*- coding: utf-8 -*-
-
-# Copyright 2019  Oplatai.com (author: Ondrej Platek)
-# Copyright 2019  Brno University of Technology (author: Karel Vesely)
-
-# Licensed under the Apache License, Version 2.0 (the "License")
-
-import setuptools
-
-with open("README.md","r") as fh:
-    long_description = fh.read()
-
-setuptools.setup(
-    name='kaldi_io',
-    version='0.9.0',
-    author='Karel Vesely',
-    description='Glue code connecting Kaldi data and Python.',
-    long_description=long_description,
-    long_description_content_type="text/markdown",
-    url='https://github.com/vesis84/kaldi-io-for-python',
-    install_requires=[ 'numpy>=1.15.3', ],
-    packages=setuptools.find_packages(),
-    classifiers=[
-        "Programming Language :: Python :: 3",
-        "Programming Language :: Python :: 2",
-        "License :: OSI Approved :: Apache Software License",
-        "Operating System :: POSIX :: Linux",
-        "Topic :: Multimedia :: Sound/Audio :: Speech",
-    ],
-)

### 2019-03-14 code/kaldi-io-for-python/tests/test_cnet_posteriors.py

@@ -1,18 +1,0 @@
-#!/usr/bin/env python
-
-import unittest
-import sys
-
-import kaldi_io
-
-class CnetReadTest(unittest.TestCase):
-    def testCnetRead(self):
-        """ Test reading a 'confusion network', represented by same Kaldi type as Posterior, """
-
-        cnet_f='tests/data/1.cnet'
-        cntime_f='tests/data/1.cntime'
-
-        # just read, no regression test,
-        cnet = [ (k,v) for k,v in kaldi_io.read_cnet_ark(cnet_f) ]
-        cntime = [ (k,v) for k,v in kaldi_io.read_cntime_ark(cntime_f) ]
-

### 2019-03-14 code/kaldi-io-for-python/tests/test_compressed_feats.py

@@ -1,27 +1,0 @@
-#!/usr/bin/env python
-
-import unittest
-import sys
-
-import numpy as np
-import kaldi_io
-import timeit
-
-class CompressedFeatsTest(unittest.TestCase):
-    def testReading(self):
-        t_beg = timeit.default_timer()
-        orig = {k:m for k,m in kaldi_io.read_mat_ark('tests/data/feats.ark')}
-        t_read_not_compressed = timeit.default_timer() - t_beg
-
-        t_beg = timeit.default_timer()
-        comp = {k:m for k,m in kaldi_io.read_mat_ark('tests/data/feats_compressed.ark')}
-        t_read_compressed = timeit.default_timer() - t_beg
-
-        # reading the compressed data should be <5x slower,
-        self.assertLess(t_read_compressed, 5.*t_read_not_compressed)
-
-        # check that the values are similar
-        # (these are not identical due to discretization in compression),
-        for key in orig.keys():
-            self.assertGreater(1e-4, np.max(np.abs(comp[key]-orig[key])))
-

### 2019-03-14 code/kaldi-io-for-python/tests/test_kaldi_io.py

@@ -1,83 +1,0 @@
-#!/usr/bin/env python
-
-import unittest
-import sys
-
-import numpy as np
-
-import kaldi_io
-print("### KALDI_IO_PATH: %s ###" % str(kaldi_io))
-
-class KaldiIoTest(unittest.TestCase):
-    def testInt32VectorReadWrite(self):
-        """
-        Test read/write for int32 vectors.
-        """
-        # read,
-        i32_vec = { k:v for k,v in kaldi_io.read_vec_int_ark('tests/data/ali.ark') } # binary,
-        i32_vec2 = { k:v for k,v in kaldi_io.read_vec_int_ark('tests/data/ali_ascii.ark') } # ascii,
-        # re-save the data,
-        with kaldi_io.open_or_fd('tests/data_re-saved/ali.ark','wb') as f:
-            for k,v in i32_vec.items(): kaldi_io.write_vec_int(f, v, k)
-        # read and make sure it is the same,
-        for k,v in kaldi_io.read_vec_int_ark('tests/data_re-saved/ali.ark'):
-            self.assertTrue(np.array_equal(v,i32_vec[k]), msg="int32 vector same after re-saving")
-
-    def testFloatVectorReadWrite(self):
-        """
-        Test read/write for float vectors.
-        """
-        # read,
-        flt_vec = { k:v for k,v in kaldi_io.read_vec_flt_scp('tests/data/conf.scp') } # scp,
-        return
-
-        flt_vec2 = { k:v for k,v in kaldi_io.read_vec_flt_ark('tests/data/conf.ark') } # binary-ark,
-        flt_vec3 = { k:v for k,v in kaldi_io.read_vec_flt_ark('tests/data/conf_ascii.ark') } # ascii-ark,
-        # store,
-        with kaldi_io.open_or_fd('tests/data_re-saved/conf.ark','wb') as f:
-            for k,v in flt_vec.items(): kaldi_io.write_vec_flt(f, v, k)
-        # read and compare,
-        for k,v in kaldi_io.read_vec_flt_ark('tests/data_re-saved/conf.ark'):
-            self.assertTrue(np.array_equal(v,flt_vec[k]), msg="flt. vector same after re-saving")
-
-    def testMatrixReadWrite(self):
-        """
-        Test read/write for float matrices.
-        """
-        # read,
-        flt_mat = { k:m for k,m in kaldi_io.read_mat_scp('tests/data/feats_ascii.scp') } # ascii-scp,
[…]

### 2019-03-14 code/timit-preprocessor/clean.sh

@@ -1,3 +1,0 @@
-#!/usr/bin/env bash
-
-rm -r data/*

### 2019-03-14 code/timit-preprocessor/make_dataset.py

@@ -1,98 +1,0 @@
-import kaldi_io
-import sys
-import numpy as np
-import glob
-import os
-import pickle as pkl
-
-if __name__ == "__main__":
-    file = sys.argv[1]
-    out_file = sys.argv[2]
-
-    timit_folder = "/home/antoine/Documents/projects/asr/gm-hmm/data/timit"
-    fname_cb = "TIMIT-61.codebook"
-
-    out = kaldi_io.read_mat_scp(file)
-
-    n = 0
-    for _, mat in out:
-        n += mat.shape[0]
-
-    d = mat.shape[1]
-
-    X = np.zeros((n, d), dtype=np.float32)
-    lengths = []
-    keys = []
-    seq_start = 0
-    out = kaldi_io.read_mat_scp(file)
-    for key, mat in out:
-        l = mat.shape[0]
-        lengths.append(l)
-        keys.append(key)
-        X[seq_start:seq_start+l] = mat
-        seq_start += l
-
-    fe = 16e3
-    wsize_ms = 10
-    nsamples = len(lengths)
-    PHN = [0 for _ in range(nsamples)]
-    DATA = [0 for _ in range(nsamples)]
-
-    seq_start = 0
-    DATA_TYPE = "TEST" if "test" in out_file else "TRAIN"
-
-    for isample in range(nsamples):
-
-        # Get phone labels
-        fname = timit_folder + "/"+DATA_TYPE+"/" + keys[isample].replace("-", "/") + ".PHN"
-        #print(isample, "/", nsamples, ":", fname)
-        with open(fname, "r") as f:
-            lines = f.read().split("\n")[:-1]
-
-        PHN_file_data = np.array(list(map(lambda x: x.split(" "), lines)))
-        loc = PHN_file_data[:, :2]
-        y = PHN_file_data[:, -1]
-        del PHN_file_data
-        loc = loc.astype(np.int64)
-        x = X[seq_start:seq_start + lengths[isample]]
-
-        feats = np.arange(x.shape[0] * wsize_ms).reshape(-1, wsize_ms)
-
-
-        loc_ms = loc / fe * 1000
-        phones = np.copy(feats).reshape(-1)
-        # Clean loc_ms
-        if loc_ms[-1][1] < phones[-1]:
-            loc_ms[-1][1] = phones[-1]
-
-        if loc_ms[0][0] > 0:
-            loc_ms[0][0] = 0
-
-        for iy, irange in enumerate(loc_ms.tolist()):
[…]

### 2019-03-14 code/timit-preprocessor/parsing.py

@@ -1,44 +1,0 @@
-#!/usr/bin/env python
-
-import os
-import sys
-import argparse
-
-def main():
-    target_folder = os.path.join(args.datapath, args.folder)
-    target = args.folder
-
-    features = '{}/{}.wav.scp'.format(MATERIAL_PATH, target)
-    labels = '{}/{}.lbl'.format(MATERIAL_PATH, target)
-
-    with open(features, "w+") as f, open(labels, "w+") as l:
-        for person in os.listdir(target_folder):
-            person_dir = os.path.join(target_folder, person)
-            for task in os.listdir(person_dir):
-                task_dir = os.path.join(person_dir, task)
-                raw_sentences = [x.split('.')[0] for x in os.listdir(task_dir)]
-                sentences = list(set(raw_sentences))
-                for sentence in sentences:
-                    wav_file = os.path.join(task_dir, sentence+'.wav')
-                    f.write('{}-{}-{} {}\n'.format(person, task, sentence, os.path.abspath(wav_file)))
-
-                    phone_seq_file = os.path.join(task_dir, sentence+'.phn')
-                    phone_seq = []
-                    with open(phone_seq_file, "r+") as pf:
-                        for line in pf:
-                            phone_seq.append(line.split(' ')[-1].strip())
-                    l.write('{}-{}-{} {}\n'.format(person, task, sentence, ','.join(phone_seq)))
-            
-
-
-if __name__ == "__main__":
-    parser = argparse.ArgumentParser(description="parsing .wav.scp files for advanced use\ne.g. python3 parsing.py ~/Workspace/data/timit train", formatter_class=argparse.RawTextHelpFormatter)
-    parser.add_argument('datapath', metavar="<timit relative path>", type=str)
-    parser.add_argument('folder', metavar="<train|test>", type=str)
-    args = parser.parse_args()
-
-    MATERIAL_PATH = '.data/material'
-    if not os.path.exists(MATERIAL_PATH):
-        os.makedirs(MATERIAL_PATH)
-
-    main()

### 2019-03-14 code/train.py

@@ -1,47 +1,0 @@
-"""Train script.
-
-Usage:
-    train.py <hparams> <dataset> <dataset_root>
-"""
-import os
-import torchvision.datasets as dset
-from docopt import docopt
-from torchvision import transforms
-from glow.builder import build
-from glow.trainer import Trainer
-from glow.config import JsonConfig
-
-
-if __name__ == "__main__":
-    args = docopt(__doc__)
-    hparams = args["<hparams>"]
-    dataset = args["<dataset>"]
-
-    assert os.path.exists(hparams), (
-        "Failed to find hparams josn `{}`".format(hparams))
-
-    dataset_root = args["<dataset_root>"]
-    assert os.path.exists(dataset_root), (
-        "Failed to find root dir `{}` of dataset.".format(dataset_root))
-    hparams = JsonConfig(hparams)
-
-    if dataset == "mnist":
-        transform = transforms.Compose([
-            transforms.Resize(hparams.Data.resize),
-            transforms.ToTensor(),
-            transforms.Normalize((0.5,0.5,0.5),(0.5,0.5,0.5))
-        ])
-        dataset = dset.MNIST(root=dataset_root,
-                         download=True,
-                         transform=transform)
-    elif dataset == "fashion-mnist":
-        transform = transforms.Compose([
-            transforms.Resize(hparams.Data.resize),
-            transforms.ToTensor(),transforms.Normalize((0.5,0.5,0.5),(0.5,0.5,0.5))])
-        dataset = dset.FashionMNIST(root=dataset_root,
-                         download=True,
-                         transform=transform)
-    # build graph and dataset
-    built = build(hparams, True)
-    trainer = Trainer(**built, dataset=dataset, hparams=hparams)
-    trainer.train()

### 2019-03-14 src/genHMM.py
move files: Add new files in src/

@@ -1,0 +1,205 @@
+# file genHMM.py
+
+import numpy as np
+from hmmlearn.utils import normalize
+from src.glow.models import Glow, FlowNet
+from src.glow.config import JsonConfig
+from hmmlearn.base import _BaseHMM
+from functools import partial
+
+
+class GenHMM(_BaseHMM):
+    def __init__(self, n_components=None, n_prob_components=None, hparams=None,\
+            algorithm="viterbi", random_state=None, n_iter=10, tol=1e-2, verbose=False, params="st", init_params="st",\
+                 image_shape=(28, 28, 3)):
+
+        _BaseHMM.__init__(self, n_components,
+                          algorithm=algorithm, random_state=random_state, n_iter=n_iter,
+                          tol=tol, params=params, verbose=verbose, init_params=init_params)
+        self.image_shape=image_shape
+        self.n_components = n_components
+        self.n_prob_components = n_prob_components
+        self.hparams = JsonConfig(hparams)
+
+        self.model_params = ["transmat_", "startprob_", "networks", "pi"]
+
+        self.init_transmat()
+        self.init_startprob()
+        self.init_gen()
+
+        self.init_future()
+
+        print("here")
+
+    def init_startprob(self):
+        """
+        Initialize HMM initial coefficients.
+        """
+        init = 1. / self.n_components
+        self.startprob_ = np.full(self.n_components, init)
+        return self
+
+    def init_transmat(self):
+        """
+        Initialize HMM transition matrix.
+        """
+        init = 1/self.n_components
+        self.transmat_ = np.full((self.n_components, self.n_components),
+                                 init)
+        return self
+
+    def init_gen(self):
+        """
+        Initialize HMM probabilistic model.
+        """
+
+        # Init latent selector variable
+        self.pi = np.random.rand(self.n_components, self.n_prob_components)
+        normalize(self.pi, axis=1)
+
+        # Init networks
+
+        self.networks = [Glow(self.hparams) for _ in range(self.n_prob_components*self.n_components)]
+        #self.networks = [FlowNet(image_shape=self.image_shape,hidden_channels=5, L=2, K=2,\
+        #                         flow_coupling="additive", flow_permutation="shuffle")\
+        #                 for _ in range(self.n_prob_components * self.n_components)]
+
+        self.networks = np.array(self.networks).reshape(self.n_components, self.n_prob_components)
+        return self
+
+    def llh(self, X):
+        return self._compute_log_likelihood(X)
+
+
[…]

### 2019-03-14 src/glow/builder.py

@@ -1,0 +1,137 @@
+import re, os
+import copy
+import torch
+from collections import defaultdict
+from . import learning_rate_schedule
+from .config import JsonConfig
+from .models import LatMM, GenMM
+from .utils import load, save, get_proper_device
+
+
+def build_adam(params, args):
+    return torch.optim.Adam(params, **args)
+
+
+__build_optim_dict = {
+    "adam": build_adam
+}
+
+# custom weights initialization called on netG and netD
+def weights_init(m):
+    classname = m.__class__.__name__
+    if classname.find('Conv') != -1:
+        pass
+        #m.weight.data.normal_(0.0, 0.01)
+    elif classname.find('BatchNorm') != -1:
+        m.weight.data.normal_(1.0, 0.02)
+        m.bias.data.fill_(0)
+
+
+def build(hparams, is_training):
+    if isinstance(hparams, str):
+        hparams = JsonConfig(hparams)
+    # get graph and criterions from build function
+    graph, optim, lrschedule, criterion_dict = None, None, None, None  # init with None
+    graph_prior = None
+    cpu, devices = "cpu", None
+    get_loss = None
+    # 1. build graph and criterion_dict, (on cpu)
+    # build and append `device attr` to graph
+    if hparams.Mixture.naive:
+        graph_dict = {"old": GenMM(hparams), "new": GenMM(hparams)}
+        # for component in graph.flow_list:
+        #     component.apply(weights_init)
+    else:
+        graph_dict = {"old": LatMM(hparams), "new": LatMM(hparams)}
+        #graph.get_component().apply(weights_init)
+    
+    for key, graph in graph_dict.items():
+        graph.device = hparams.Device.glow
+        if graph is not None:
+            # get device
+            devices = get_proper_device(graph.device)
+            graph.device = devices
+            graph.to(cpu)
+        # 2. get optim (on cpu)
+        try:
+            if graph is not None and is_training and key is "new":
+                optim_name = hparams.Optim.name
+                optim = __build_optim_dict[optim_name](graph.parameters(), hparams.Optim.args.to_dict())
+                print("[Builder]: Using optimizer `{}`, with args:{}".format(optim_name, hparams.Optim.args))
+                # get lrschedule
+                schedule_name = "default"
+                schedule_args = {}
+                if "Schedule" in hparams.Optim:
+                    schedule_name = hparams.Optim.Schedule.name
+                    schedule_args = hparams.Optim.Schedule.args.to_dict()
[…]

### 2019-03-14 src/glow/models.py

@@ -1,0 +1,524 @@
+import torch
+import torch.nn as nn
+import torch.nn.functional as F
+import numpy as np
+from tqdm import tqdm
+from . import thops
+from . import modules
+from . import utils
+from scipy import stats
+
+
+def prior_sampler(pk, size=64):
+    pk = tuple(pk.numpy())
+    xk = np.arange(len(pk))
+    custm = stats.rv_discrete(name='custm', values=(xk, pk))
+    prior_samples = custm.rvs(size=size)
+    prior_samples = torch.from_numpy(prior_samples)
+    return prior_samples
+
+
+def f(in_channels, out_channels, hidden_channels):
+    return nn.Sequential(
+        modules.Conv2d(in_channels, hidden_channels), nn.ReLU(inplace=False),
+        modules.Conv2d(hidden_channels, hidden_channels, kernel_size=[1, 1]), nn.ReLU(inplace=False),
+        modules.Conv2dZeros(hidden_channels, out_channels))
+
+
+class FlowStep(nn.Module):
+    FlowCoupling = ["additive", "affine"]
+    FlowPermutation = {
+        "reverse": lambda obj, z, logdet, rev: (obj.reverse(z, rev), logdet),
+        "shuffle": lambda obj, z, logdet, rev: (obj.shuffle(z, rev), logdet),
+        "invconv": lambda obj, z, logdet, rev: obj.invconv(z, logdet, rev)
+    }
+
+    def __init__(self, in_channels, hidden_channels,
+                 actnorm_scale=1.0,
+                 flow_permutation="invconv",
+                 flow_coupling="additive",
+                 LU_decomposed=False):
+        # check configures
+        assert flow_coupling in FlowStep.FlowCoupling, \
+            "flow_coupling should be in `{}`".format(FlowStep.FlowCoupling)
+        assert flow_permutation in FlowStep.FlowPermutation, \
+            "float_permutation should be in `{}`".format(
+                FlowStep.FlowPermutation.keys())
+        super().__init__()
+        self.flow_permutation = flow_permutation
+        self.flow_coupling = flow_coupling
+        # 1. actnorm
+        self.actnorm = modules.ActNorm2d(in_channels, actnorm_scale)
+        # 2. permute
+        if flow_permutation == "invconv":
+            self.invconv = modules.InvertibleConv1x1(
+                in_channels, LU_decomposed=LU_decomposed)
+        elif flow_permutation == "shuffle":
[…]

### 2019-03-14 src/glow/modules.py

@@ -1,0 +1,375 @@
+import torch
+import torch.nn as nn
+import torch.nn.functional as F
+import numpy as np
+import scipy.linalg
+from . import thops
+
+
+class _ActNorm(nn.Module):
+    """
+    Activation Normalization
+    Initialize the bias and scale with a given minibatch,
+    so that the output per-channel have zero mean and unit variance for that.
+
+    After initialization, `bias` and `logs` will be trained as parameters.
+    """
+
+    def __init__(self, num_features, scale=1.):
+        super().__init__()
+        # register mean and scale
+        size = [1, num_features, 1, 1]
+        self.register_parameter("bias", nn.Parameter(torch.zeros(*size)))
+        self.register_parameter("logs", nn.Parameter(torch.zeros(*size)))
+        self.num_features = num_features
+        self.scale = float(scale)
+        self.inited = False
+
+    def _check_input_dim(self, input):
+        return NotImplemented
+
+    def initialize_parameters(self, input):
+        self._check_input_dim(input)
+        if not self.training:
+            return
+        assert input.device == self.bias.device
+        with torch.no_grad():
+            bias = thops.mean(input.clone(), dim=[0, 2, 3], keepdim=True) * -1.0
+            vars = thops.mean((input.clone() + bias) ** 2, dim=[0, 2, 3], keepdim=True)
+            logs = torch.log(self.scale / (torch.sqrt(vars) + 1e-6))
+            self.bias.data.copy_(bias.data)
+            self.logs.data.copy_(logs.data)
+            self.inited = True
+
+    def _center(self, input, reverse=False):
+        if not reverse:
+            return input + self.bias
+        else:
+            return input - self.bias
+
+    def _scale(self, input, logdet=None, reverse=False):
+        logs = self.logs
+        if not reverse:
+            input = input * torch.exp(logs)
+        else:
+            input = input * torch.exp(-logs)
+        if logdet is not None:
+            """
+            logs is log_std of `mean of channels`
+            so we need to multiply pixels
+            """
+            dlogdet = thops.sum(logs) * thops.pixels(input)
+            if reverse:
+                dlogdet *= -1
+            logdet = logdet + dlogdet
+        return input, logdet
+
+    def forward(self, input, logdet=None, reverse=False):
+        if not self.inited:
[…]

### 2019-03-14 src/glow/trainer.py

@@ -1,0 +1,271 @@
+import re
+import os
+import torch
+import torch.nn.functional as F
+
+import datetime
+import numpy as np
+from tqdm import tqdm
+from tensorboardX import SummaryWriter
+from torch.utils.data import DataLoader
+from .utils import save, load, plot_prob, new2old
+from .config import JsonConfig
+from .models import Glow
+from . import thops
+
+
+class Trainer(object):
+    def __init__(self, graph, graph_prior, optim, lrschedule, loaded_step,
+                 devices, data_device,
+                 dataset, hparams):
+        if isinstance(hparams, str):
+            hparams = JsonConfig(hparams)
+        # set members
+        # append date info
+        date = str(datetime.datetime.now())
+        date = date[:date.rfind(":")].replace("-", "")\
+                                     .replace(":", "")\
+                                     .replace(" ", "_")
+        self.log_dir = os.path.join(hparams.Dir.log_root, "log_" + date)
+        
+        self.checkpoints_dir = os.path.join(self.log_dir, "checkpoints")
+        if not os.path.exists(self.log_dir):
+            os.makedirs(self.log_dir)
+        # write hparams
+        hparams.dump(self.log_dir)
+        if not os.path.exists(self.checkpoints_dir):
+            os.makedirs(self.checkpoints_dir)
+        self.checkpoints_gap = hparams.Train.checkpoints_gap
+        self.prior_gap = hparams.Train.prior_gap
+        self.em_gap = hparams.Train.em_gap
+        self.max_checkpoints = hparams.Train.max_checkpoints
+        # set the mixture prior
+        self.num_component = hparams.Mixture.num_component
+        self.mix_prior = torch.ones(self.num_component)
+        # model relative
+        self.graph = graph
+
+        self.optim = optim
+        self.weight_y = hparams.Train.weight_y
+        # grad operation
+        self.max_grad_clip = hparams.Train.max_grad_clip
+        self.max_grad_norm = hparams.Train.max_grad_norm
+        # copy devices from built graph
+        self.devices = devices
+        self.data_device = data_device
+        # number of training batches
+        self.batch_size = hparams.Train.batch_size
+        self.data_loader = DataLoader(dataset,
+                                      batch_size=self.batch_size,
+                                    #   num_workers=8,
+                                      shuffle=True,
+                                      drop_last=True)
+        self.n_epoches = (hparams.Train.num_batches+len(self.data_loader)-1)
[…]

### 2019-03-14 src/timit-preprocessor/extract_mfcc.sh

@@ -1,0 +1,36 @@
+#!/usr/bin/env bash
+
+if [ ${#} != 3 ];then
+    echo "Usage: ./mfcc_extract.sh [Kaldi's relative path] [TIMIT relative path] [output .scp filename]"
+    exit -1
+fi
+
+kaldi=${1}
+timit=${2}
+target_file=${3}
+
+path=.data/processed
+options="--use-energy=false"
+
+mkdir -p $path
+
+
+fname=$(basename "$target_file")
+fname_trunk="${fname%.*}"
+target=(${fname_trunk//./ })
+
+echo $fname
+
+log=$path/$target.extract.log
+if [[ $fname == *".13."* ]]; then
+	${kaldi}/src/featbin/compute-mfcc-feats --verbose=2 $options scp:.data/material/$target.wav.scp ark,t,scp:$path/$target.13.ark,$target_file 2> $log
+fi
+
+if [[ $fname == *".39."* ]]; then
+	${kaldi}/src/featbin/compute-mfcc-feats --verbose=2 $options scp:.data/material/$target.wav.scp ark,t,scp:$path/$target.13.ark,$target_file 2> $log
+	${kaldi}/src/featbin/add-deltas ark:$path/$target.13.ark ark,t:$path/$target.13_deltas.ark
+	${kaldi}/src/featbin/compute-cmvn-stats ark:$path/$target.13_deltas.ark ark,t:$path/$target.cmvn_results.ark
+	${kaldi}/src/featbin/apply-cmvn ark:$path/$target.cmvn_results.ark ark:$path/$target.13_deltas.ark ark,t,scp:$path/$target.39.cmvn.ark,$target_file
+fi
+
+exit 0

### 2019-03-14 src/timit-preprocessor/make_dataset.py

@@ -1,0 +1,98 @@
+import kaldi_io
+import sys
+import numpy as np
+import glob
+import os
+import pickle as pkl
+
+if __name__ == "__main__":
+    file = sys.argv[1]
+    out_file = sys.argv[2]
+
+    timit_folder = "/home/antoine/Documents/projects/asr/gm-hmm/data/timit"
+    fname_cb = "TIMIT-61.codebook"
+
+    out = kaldi_io.read_mat_scp(file)
+
+    n = 0
+    for _, mat in out:
+        n += mat.shape[0]
+
+    d = mat.shape[1]
+
+    X = np.zeros((n, d), dtype=np.float32)
+    lengths = []
+    keys = []
+    seq_start = 0
+    out = kaldi_io.read_mat_scp(file)
+    for key, mat in out:
+        l = mat.shape[0]
+        lengths.append(l)
+        keys.append(key)
+        X[seq_start:seq_start+l] = mat
+        seq_start += l
+
+    fe = 16e3
+    wsize_ms = 10
+    nsamples = len(lengths)
+    PHN = [0 for _ in range(nsamples)]
+    DATA = [0 for _ in range(nsamples)]
+
+    seq_start = 0
+    DATA_TYPE = "TEST" if "test" in out_file else "TRAIN"
+
+    for isample in range(nsamples):
+
+        # Get phone labels
+        fname = timit_folder + "/"+DATA_TYPE+"/" + keys[isample].replace("-", "/") + ".PHN"
+        #print(isample, "/", nsamples, ":", fname)
+        with open(fname, "r") as f:
+            lines = f.read().split("\n")[:-1]
+
+        PHN_file_data = np.array(list(map(lambda x: x.split(" "), lines)))
+        loc = PHN_file_data[:, :2]
+        y = PHN_file_data[:, -1]
+        del PHN_file_data
+        loc = loc.astype(np.int64)
+        x = X[seq_start:seq_start + lengths[isample]]
+
+        feats = np.arange(x.shape[0] * wsize_ms).reshape(-1, wsize_ms)
+
+
+        loc_ms = loc / fe * 1000
+        phones = np.copy(feats).reshape(-1)
+        # Clean loc_ms
+        if loc_ms[-1][1] < phones[-1]:
+            loc_ms[-1][1] = phones[-1]
+
+        if loc_ms[0][0] > 0:
+            loc_ms[0][0] = 0
+
+        for iy, irange in enumerate(loc_ms.tolist()):
[…]

### 2019-03-14 src/timit-preprocessor/parsing.py

@@ -1,0 +1,44 @@
+#!/usr/bin/env python
+
+import os
+import sys
+import argparse
+
+def main():
+    target_folder = os.path.join(args.datapath, args.folder)
+    target = args.folder
+
+    features = '{}/{}.wav.scp'.format(MATERIAL_PATH, target)
+    labels = '{}/{}.lbl'.format(MATERIAL_PATH, target)
+
+    with open(features, "w+") as f, open(labels, "w+") as l:
+        for person in os.listdir(target_folder):
+            person_dir = os.path.join(target_folder, person)
+            for task in os.listdir(person_dir):
+                task_dir = os.path.join(person_dir, task)
+                raw_sentences = [x.split('.')[0] for x in os.listdir(task_dir)]
+                sentences = list(set(raw_sentences))
+                for sentence in sentences:
+                    wav_file = os.path.join(task_dir, sentence+'.wav')
+                    f.write('{}-{}-{} {}\n'.format(person, task, sentence, os.path.abspath(wav_file)))
+
+                    phone_seq_file = os.path.join(task_dir, sentence+'.phn')
+                    phone_seq = []
+                    with open(phone_seq_file, "r+") as pf:
+                        for line in pf:
+                            phone_seq.append(line.split(' ')[-1].strip())
+                    l.write('{}-{}-{} {}\n'.format(person, task, sentence, ','.join(phone_seq)))
+            
+
+
+if __name__ == "__main__":
+    parser = argparse.ArgumentParser(description="parsing .wav.scp files for advanced use\ne.g. python3 parsing.py ~/Workspace/data/timit train", formatter_class=argparse.RawTextHelpFormatter)
+    parser.add_argument('datapath', metavar="<timit relative path>", type=str)
+    parser.add_argument('folder', metavar="<train|test>", type=str)
+    args = parser.parse_args()
+
+    MATERIAL_PATH = '.data/material'
+    if not os.path.exists(MATERIAL_PATH):
+        os.makedirs(MATERIAL_PATH)
+
+    main()

### 2019-03-14 src/utils.py

@@ -1,0 +1,23 @@
+
+import numpy as np
+
+
+def norm_prob(x,axis=None):
+    coef_ = x.sum(axis)
+    if axis==0:
+        coef_ = coef_.reshape(1,-1)
+    elif axis==1:
+        coef_ = coef_.reshape(-1, 1)
+
+    return x / np.repeat(coef_, x.shape[axis], axis=axis)
+
+def test_norm_prob():
+    x = np.array([[1, 2], [4, 5]])
+    xn = norm_prob(x, axis=1)
+    assert(np.all(xn == np.array([[1/(1+2), 2/(1+2)], [4/(4+5), 5/(4+5)]])))
+    xn = norm_prob(x, axis=0)
+    assert (np.all(xn == np.array([[1 / (1 + 4), 2 / (5 + 2)], [4 / (4 + 1), 5 / (2 + 5)]])))
+
+
+if __name__ == "__main__":
+    test_norm_prob()

### 2019-03-14 bin/test.py
Add test bin.

@@ -1,0 +1,68 @@
+# file test.py
+import pickle as pkl
+import sys
+import os
+import numpy as np
+
+from src.genHMM import GenHMM
+
+def prepare_data(data_folder=None, fname_dtest=None, fname_dtrain=None, n_phn=None, verbose=False):
+    te_DATA, te_keys, te_lengths, codebook, te_PHN = pkl.load(open(fname_dtest, "rb"))
+    tr_DATA, tr_keys, tr_lengths, tr_PHN = pkl.load(open(fname_dtrain, "rb"))
+    if verbose:
+        print("Data loaded from files.")
+
+    # Pick n random phoneme
+    iphn = np.random.randint(len(codebook), size=n_phn)
+
+    # Prepare training data
+    X = np.concatenate(tuple(tr_DATA), axis=0)
+    d = find_phones(X, iphn)
+    ytrain = d[:, 0]
+    # Keep label as a dummy value to have even dimensions
+    xtrain = d[:, :]
+    del d
+
+    # Prepare test data
+    X = np.concatenate(tuple(te_DATA), axis=0)
+    d = find_phones(X, iphn)
+    ytest = d[:, 0]
+    # Keep label as a dummy value to have even dimensions
+    xtest = d[:, :]
+    del d
+
+    return xtrain, ytrain, xtest, ytest, iphn
+    pass
+
+
+def find_phones(X=None, iphn=None):
+    return X[np.concatenate(tuple([(X[:, 0] == iphn[i]).reshape(-1, 1) for i in range(iphn.shape[0])]), axis=1).sum(1) == 1, :]
+
+
+if __name__ == "__main__":
+    
+    usage = "Usage: python bin/test.py [data folder] [test pkl file] [train pkl file] [params .json file]\n\n\
+            Example: python bin/test.py data/ test13.pkl train13.pkl hparams/test.json\n"
+
+    if len(sys.argv) != 5:
+        print(usage)
+        sys.exit(1)
+
+    data_folder = sys.argv[1]
+    fname_dtest = os.path.join(data_folder, sys.argv[2])
+    fname_dtrain = os.path.join(data_folder, sys.argv[3])
+    hparams = sys.argv[4]
+
+    n_phn = 2
+    xtrain, ytrain, xtest, ytest, IPHN = prepare_data(data_folder=data_folder, \
+                                                fname_dtest=fname_dtest, fname_dtrain=fname_dtrain,\
[…]

### 2019-03-14 bin/test.py
Add comments.

@@ -6,37 +6,37 @@ import numpy as np
 
 from src.genHMM import GenHMM
 
+
+def getsubset(DATA,iphn):
+    # concat data
+    X = np.concatenate(tuple(DATA), axis=0)
+    # find subset
+    d = find_phones(X, iphn)
+    # Split in feats/target
+    # Note: Keep label in features as a dummy value to have even dimensions
+    return d[:, :], d[:, 0]
+
+
 def prepare_data(data_folder=None, fname_dtest=None, fname_dtrain=None, n_phn=None, verbose=False):
+    # read the datafiles
     te_DATA, te_keys, te_lengths, codebook, te_PHN = pkl.load(open(fname_dtest, "rb"))
     tr_DATA, tr_keys, tr_lengths, tr_PHN = pkl.load(open(fname_dtrain, "rb"))
     if verbose:
         print("Data loaded from files.")
 
-    # Pick n random phoneme
+    # Pick n random phonemes
     iphn = np.random.randint(len(codebook), size=n_phn)
 
-    # Prepare training data
-    X = np.concatenate(tuple(tr_DATA), axis=0)
-    d = find_phones(X, iphn)
-    ytrain = d[:, 0]
-    # Keep label as a dummy value to have even dimensions
-    xtrain = d[:, :]
-    del d
-
-    # Prepare test data
-    X = np.concatenate(tuple(te_DATA), axis=0)
-    d = find_phones(X, iphn)
-    ytest = d[:, 0]
-    # Keep label as a dummy value to have even dimensions
-    xtest = d[:, :]
-    del d
-
+    xtrain, ytrain = getsubset(tr_DATA, iphn)
+    xtest, ytest = getsubset(te_DATA, iphn)
     return xtrain, ytrain, xtest, ytest, iphn
-    pass
 
 
 def find_phones(X=None, iphn=None):
-    return X[np.concatenate(tuple([(X[:, 0] == iphn[i]).reshape(-1, 1) for i in range(iphn.shape[0])]), axis=1).sum(1) == 1, :]
+    # For each requested phonements, find their location in the data array
+    phonemes_location = [(X[:, 0] == iphn[i]).reshape(-1, 1) for i in range(iphn.shape[0])]
+
+    return X[np.concatenate(tuple(phonemes_location), axis=1).sum(1) == 1, :]
 
 
 if __name__ == "__main__":
@@ -53,16 +53,16 @@ if __name__ == "__main__":
[…]

### 2019-07-12 bin/train_class.py
get run on multiple gpus

@@ -5,12 +5,12 @@ sys.path.append("..")
 from parse import parse
 import pickle as pkl
 from src.genHMM import GenHMM, save_model, load_model
-from src.utils import pad_data, TheDataset
+from src.utils import pad_data, TheDataset, get_freer_gpu
 import torch
 from torch.utils.data import DataLoader
 import json
 import numpy as np
-
+import time
 if __name__ == "__main__":
     usage = "python bin/train_class.py data/train13.pkl models/epoch1_class1.mdlc param.json"
     if len(sys.argv) < 3 or sys.argv[1] == "-h" or sys.argv[1] == "--help":
@@ -60,18 +60,29 @@ if __name__ == "__main__":
 
     mdl.device = 'cpu'
     if torch.cuda.is_available():
-        device = torch.device('cuda')
-        mdl.device = device
-    
-    print("epoch:{}\tclass:{}\tPush model to {}...".format(epoch_str,iclass_str, mdl.device), file=sys.stdout)
-    mdl.pushto(mdl.device)   
-    
+        if not options["Mul_gpu"]:
+            # default case, only one gpu
+            device = torch.device('cuda:0')
+        else:
+            for i in range(4):
+                try:
+                    time.sleep(np.random.randint(20))
+                    device = torch.device('cuda:{}'.format(int(get_freer_gpu()) ))
+                    print("Try to push to device: {}".format(device))
+                    mdl.device = device
+                    mdl.pushto(mdl.device)   
+                    break
+                except:
+                    # if push error (maybe memory overflow, try again)
+                    print("Push to device cuda:{} fail, try again ...")
+                    continue
+    print("epoch:{}\tclass:{}\tPush model to {}. Done.".format(epoch_str,iclass_str, mdl.device), file=sys.stdout)
     
     # zero pad data for batch training
     max_len_ = max([x.shape[0] for x in xtrain])
     xtrain_padded = pad_data(xtrain, max_len_)
+
     traindata = DataLoader(dataset=TheDataset(xtrain_padded, lengths=l, device=mdl.device), batch_size=options["Train"]["batch_size"], shuffle=True)
-    
 
     # niter counts the number of em steps before saving a model checkpoint
     niter = options["Train"]["niter"]

### 2019-07-12 src/genHMM.py

@@ -71,8 +71,8 @@ class GenHMM(torch.nn.Module):
         self.em_skip = em_skip
         
         # Initialize HMM parameters
-        self.init_transmat()
-        self.init_startprob()
+        self.init_transmat(transmat_type)
+        self.init_startprob(startprob_type)
         
         # Initialize generative model networks
         self.init_gen(H=net_H, D=net_D, nchain=net_nchain, mask_type=mask_type)
@@ -83,7 +83,7 @@ class GenHMM(torch.nn.Module):
         self.global_step = 0
 
         
-    def init_startprob(self, startprob_type="first"):
+    def init_startprob(self, startprob_type="random"):
         """
         Initialize HMM initial coefficients.
         """
@@ -97,9 +97,11 @@ class GenHMM(torch.nn.Module):
             init[0] = 1
             self.startprob_ = init
 
-        else:
-            init = 1. / self.n_states
-            self.startprob_ = torch.ones(self.n_states) * init
+        elif "uniform" in startprob_type:
+            self.startprob_ = torch.ones(self.n_states)
+            
+            
+        normalize(self.startprob_, axis=0)
         
             
         return self
@@ -110,17 +112,21 @@ class GenHMM(torch.nn.Module):
         """
         if "random" in transmat_type:
             self.transmat_ = (torch.randn(self.n_states, self.n_states)).abs()
-            
-            
-        else:
+        elif "uniform" in transmat_type:
             init = 1/self.n_states
             self.transmat_ = torch.ones(self.n_states, self.n_states) * init
-        if "triangular" in transmat_type:
+        elif "triangular" in transmat_type:
             # use upper tra matrix
+            self.transmat_ = (torch.randn(self.n_states, self.n_states)).abs()
             for i in range(self.n_states):
                 for j in range(self.n_states):
                     if i > j:
                         self.transmat_[i,j] = 0
+
+        elif "ergodic" in transmat_type:
+            self.transmat_ = torch.ones(self.n_states, self.n_states) \
+                             + torch.randn(self.n_states, self.n_states) * 0.01 
+
         normalize(self.transmat_, axis=1)
         return self
 
@@ -196,6 +202,7 @@ class GenHMM(torch.nn.Module):
                 
         self.startprob_ = self.startprob_.to(device)
         self.transmat_ = self.transmat_.to(device)
+        self.pi = self.pi.to(device)
         self.logPIk_s = self.logPIk_s.to(device)
 
         self.device = device

### 2019-07-12 src/utils.py

@@ -1,4 +1,5 @@
 import numpy as np
+import os
 import torch
 from torch.utils.data import Dataset
 
@@ -48,6 +49,7 @@ def test_norm_prob():
     xn = norm_prob(x, axis=0)
     assert (np.all(xn == np.array([[1 / (1 + 4), 2 / (5 + 2)], [4 / (4 + 1), 5 / (2 + 5)]])))
 
+
 def step_learning_rate_decay(init_lr, global_step, minimum,
                              anneal_rate=0.98,
                              anneal_interval=1):
@@ -55,6 +57,12 @@ def step_learning_rate_decay(init_lr, global_step, minimum,
     if rate < minimum:
         rate = minimum
     return rate
+
+def get_freer_gpu():
+    os.system('nvidia-smi -q -d Memory |grep -A4 GPU|grep Free >tmp')
+    memory_available = [int(x.split()[2]) for x in open('tmp', 'r').readlines()]
+    return np.argmax(memory_available)
+
 
 if __name__ == "__main__":
     test_norm_prob()

### 2019-07-12 src/genHMM.py
correct statics accumulate for pi

@@ -296,25 +296,33 @@ class GenHMM(torch.nn.Module):
 
  #       print(loglh_sk.shape, self.n_states, self.n_prob_components)
         # max_loglh = torch.max(torch.max(loglh_sk, dim=1)[0],dim=1)[0]
-        ### dong: have not verify this computation
-        local_loglh_sk = loglh_sk.reshape(batch_size, self.n_states, self.n_prob_components, n_samples)
-        max_loglh = torch.max(torch.max(local_loglh_sk, dim=3)[0], dim=2)[0]
-#        print(max_loglh.shape)
+        ### dong: original update here was not correct
 
-        gamma_ = torch.zeros(batch_size, self.n_states, self.n_prob_components, n_samples, device=self.device)
+        local_loglh_sk = loglh_sk
+        max_loglh = torch.max(local_loglh_sk, dim=3, keepdim=True)[0]
+        # should be careful to use the minus max trick here
+        gamma_ = self.pi.reshape(1, 1, self.n_states, self.n_prob_components) * \
+                 (local_loglh_sk - max_loglh).exp()
+        gamma_ /= gamma_.sum(3, keepdim=True)
+        # # set the elements corresponding to padded values to be zeros, this is done by zeroes in posteriors
+        gamma = posteriors.unsqueeze(dim=3) * gamma_
 
-        for i in range(self.n_states):
-            for t in range(n_samples):
-                for m in range(self.n_prob_components):
-                    gamma_[:,i, m, t] = self.pi[i, m] * (local_loglh_sk - max_loglh[:,i].reshape(-1,1,1,1))[:, i, m, t].exp()
-                
- #               print(gamma_[:, i, :, t].shape,  gamma_[:,i, :, t].sum(1).shape) 
-                gamma_[:, i, :, t] /= (gamma_[:,i, :, t].sum(1).reshape(-1, 1) + 1e-6)   #.reshape(self.n_states,1,n_samples)
-#                print(gamma_[:, i, :, t].shape, posteriors[:, t, i].shape)
- 
-                gamma_[:, i, :, t] *= posteriors[:, t, i].reshape(-1,1)
 
-        self.stats["mixture"] += gamma_.sum(3).sum(0)
+        # In- line test for gamma computation, set the if condition to be true to compare gamm and statcs_prob_components
+        if False:
[…]

### 2019-07-12 manuscript/main.tex
In E-step, take the summation out
correct typos
formulate solution to update to pi

@@
   \begin{equation}\label{eq:em-q-funciton}
-    \Qq(\bm{H}; \bm{H}^{\mathrm{old}}) = \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}\left[ \sum_{r=1}^{R}\log\,p(\ubmr{x}{r}, \ubmr{s}{r}; \bm{H}) \right]
+    \Qq(\bm{H}; \bm{H}^{\mathrm{old}}) = \ \sum_{r=1}^{R}E_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}\left[\log\,p(\ubmr{x}{r}, \ubmr{s}{r}; \bm{H}) \right]
   \end{equation}
@@
 \begin{algorithm}[H]
-  \caption{Meta algorithm for HMM powered by Generative Models}
+  \caption{Meta algorithm for HMM powered by Generative Models (GMHMM)}
   \begin{algorithmic}[1]
@@
 \end{figure}
-
-Base on the generative model assumption:
-
+In this subsection, we specify the GenHMM model and give the learning algorithm of this model.
+Base on the generative model assumption, for each state $s$, the probabilistic generative model is
 \begin{equation}
@@
 \end{equation}
-where $\bm{\Phi}_s = \left\{ \bm{\Phi}_{s, \kappa}| \kappa = 1, 2, \cdots, K. \right\}$, and 
+where $\bm{\Phi}_s = \left\{ \pi_{s, \kappa}, \bm{\Phi}_{s, \kappa}| \kappa = 1, 2, \cdots, K. \right\}$, $\pi_{s, \kappa} = p(\kappa | \bm{x}, s; \bm{\Phi}_{s, \kappa})$ and 
 \begin{equation}
@@
 \begin{align}\label{eq:obj-q-gen-mix}
-  \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) & = \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|} \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})  \log\,p(\smtr{\kappa}{t}{r}, \bmtr{x}{t}{r}| \smtr{s}{t}{r}; \bm{\Phi})\nonumber \\
+  \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) & = \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|} \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})  \log\,p(\smtr{\kappa}{t}{r}, \bmtr{x}{t}{r}| \smtr{s}{t}{r}; \bm{\Phi}_{s})\nonumber \\
   =& \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})p(\smtr{\kappa}{t}{r}|\smtr{s}{t}{r}, \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \bigg[\log\,\pi_{\smtr{s}{t}{r},\smtr{\kappa}{t}{r}} \nonumber \\
-                                        & + \log\, p(\bmtr{x}{t}{r} | \smtr{s}{t}{r}, \smtr{\kappa}{t}{r}; \bm{\Phi}) \bigg]
+                                        & + \log\, p(\bmtr{x}{t}{r} | \smtr{s}{t}{r}, \smtr{\kappa}{t}{r}; \bm{\Phi}_s) \bigg]
 \end{align}
@@
 The latent prior for mixture of each source $s$ is obtained by solving the following problem:
-\begin{align}
-  \pi_{s, \kappa} & = \uargmax{pi_{s, \kappa}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) \\ \nonumber
-                  & s.t. \, \sum_{\kappa=1}^{K} \pi_{s, \kappa}= 1
+\begin{align}\label{opm:pi}
+  \pi_{s, \kappa} & = \uargmax{\pi_{s, \kappa}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) \\ \nonumber
+                  & s.t. \, \sum_{\kappa=1}^{K} \pi_{s, \kappa}= 1, \forall s = 1, 2, \cdots, |\Ss|. 
 \end{align}
-which gives the solution:
+
+To solve problem \autoref{opm:pi}, we formulate its Lagrange function as
+\begin{equation}
+  \Ll = \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) + \sum_{\lambda_s} \lambda_s\left( 1-  \sum_{\kappa=1}^{K}\pi_{s, \kappa}  \right).
+\end{equation}
+Solving $\pd{\Ll}{\pi_{s, \kappa}} = 0$ gives
+\begin{equation}
+  \pi_{s,\kappa} = \frac{1}{\lambda_s}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}} p(\smtr{s}{t}{r}=s, \smtr{\kappa}{t}{r} =\kappa| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})
+\end{equation}
+With condition $\sum_{\kappa=1}^{K} \pi_{s, \kappa}= 1, \forall s = 1, 2, \cdots, |\Ss|$, we have
+\begin{equation}
+  \lambda_s = \sum_{\kappa=1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}} p(\smtr{s}{t}{r}=s, \smtr{\kappa}{t}{r} =\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}})
+\end{equation}
+Then the solution to \autoref{opm:pi} is
 \begin{equation}\label{eq:mix-latent-parameter-solution}
-  \pi_{s, \kappa} = \frac{\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }{\sum_{k =1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=k | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }
+  \pi_{s, \kappa} = \frac{\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }{\sum_{k =1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=k | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }
 \end{equation}
 where $p(s, \kappa | \ubm{x}; \bm{H}^{\mathrm{old}}) = p(s| \ubm{x}; \bm{H}^{\mathrm{old}}) p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$ that can be computed by results of forward/backward and \autoref{eq:kappa-posterior}.
+
+\textcolor{red}{I am going to put specific update formula for $\pi$ here....}
 

### 2019-07-13 src/genHMM.py
keep testing update for pi, avoid 0 update for pi

@@ -303,7 +303,7 @@ class GenHMM(torch.nn.Module):
         # should be careful to use the minus max trick here
         gamma_ = self.pi.reshape(1, 1, self.n_states, self.n_prob_components) * \
                  (local_loglh_sk - max_loglh).exp()
-        gamma_ /= gamma_.sum(3, keepdim=True)
+        gamma_ = gamma_ / (gamma_.sum(3, keepdim=True) +  self.EPS)
         # # set the elements corresponding to padded values to be zeros, this is done by zeroes in posteriors
         gamma = posteriors.unsqueeze(dim=3) * gamma_
 
@@ -314,13 +314,15 @@ class GenHMM(torch.nn.Module):
     #        print(max_loglh.shape)
 
             gamma_ = torch.zeros(batch_size, n_samples, self.n_states, self.n_prob_components, device=self.device)
-            for i in range(self.n_states):
-                for k in range(self.n_prob_components):
-                    gamma_[:,:,i, k] = self.pi[i, k] * local_loglh_sk[:, :, i, k].exp()
+            for b in range(batch_size):
+                for n in range(n_samples):
+                    for i in range(self.n_states):
+                        for k in range(self.n_prob_components):
+                            gamma_[b,n,i, k] = self.pi[i, k] * (local_loglh_sk[b, n, i, k]-local_loglh_sk[b,n,i,:].max()).exp()
 
-                gamma_[:,:,i,:] = gamma_[:, :, i, :] / (gamma_[:, :, i, :].sum(dim=2, keepdim=True) + self.EPS)
+                        gamma_[b,n,i,:] = gamma_[b, n, i, :] / (gamma_[b, n, i, :].sum() + self.EPS)
 
-                statcs_prob_components[:,:,i,:] = posteriors[:, :, i].reshape(batch_size, n_samples, 1) * gamma_[:, :, i, :]
+                        statcs_prob_components[b,n,i,:] = posteriors[b, n, i] * gamma_[b, n, i, :]
 
         self.stats["mixture"] += gamma.sum(1).sum(0)
 
@@ -486,7 +488,7 @@ class GenHMM(torch.nn.Module):
         # get the adaptive learning rate
         ada_lr = step_learning_rate_decay(init_lr=self.lr,
                                           global_step=self.global_step,
-                                          minimum=4e-4,
+                                          minimum=1e-4,
                                           anneal_rate=0.98)
         optimizer = torch.optim.Adam(
[…]

### 2019-07-21 src/timit-preprocessor/make_dataset.py
Major change in labeling of mfccs.

@@ -13,7 +13,7 @@ def test_make_phone_index_sequence():
     expected[0:20] = 0
     expected[20:36] = 1
     expected[36:45] = 1
-    assert((make_phone_index_sequence(a) == expected).all())
+    assert((make_phone_index_sequence(a,45) == expected).all())
 
 
 def make_phone_index_sequence(loc_ms, utt_end):
@@ -59,13 +59,14 @@ if __name__ == "__main__":
     timit_folder = "/home/antoine/Documents/projects/asr/gm-hmm/data/timit"
     fname_cb = "TIMIT-61.codebook"
 
-    seq_start = 0
-
     # Now we must have the phonemes corresponding the the MFCC for each utterance
     fe = 16e3
 
-    # That's the step size used to compute the MFCC
-    wsize_ms = 10
+    # That's the step and window size used to compute the MFCC
+    wsize_ms = 25
+    wsize_samp = int(wsize_ms / 1000 * fe)
+    step_ms = 10
+    step_samp = int(step_ms / 1000 * fe)
 
     lengths = [] # Size of all sequnces
     keys = [] # Name of the utterance
@@ -95,26 +96,26 @@ if __name__ == "__main__":
         PHN_file_data = np.array(list(map(lambda x: x.split(" "), lines)))
 
         # Range of the phoneme
-        label_location_sample = PHN_file_data[:, :2]
+        label_location_samp = PHN_file_data[:, :2]
 
         # Actual Phoneme name
         label_name = PHN_file_data[:, -1]
 
-        label_location_sample = label_location_sample.astype(np.int64)
-        utt_end = l * wsize_ms
-        label_location_ms = label_location_sample / fe * 1000
-        label_location_ms = make_range_match(label_location_ms, utt_end)
+        label_location_samp = label_location_samp.astype(int)
+        utt_end = int((l-1) * step_samp + wsize_samp)
+        label_location_samp = make_range_match(label_location_samp, utt_end)
 
-        # label_seq_ms [ 0 0 1 1 1 1 1 1 1 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 2 3 3 3 3 3 3 3 3 3 3 3 3]
-        label_seq_ms = make_phone_index_sequence(label_location_ms,utt_end)
[…]

### 2019-07-21 manuscript/main.tex
add shared structure

@@
 \end{equation}
-where $p(s, \kappa | \ubm{x}; \bm{H}^{\mathrm{old}}) = p(s| \ubm{x}; \bm{H}^{\mathrm{old}}) p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$ that can be computed by results of forward/backward and \autoref{eq:kappa-posterior}.
-
-\textcolor{red}{I am going to put specific update formula for $\pi$ here....}
+where $p(s, \kappa | \ubm{x}; \bm{H}^{\mathrm{old}}) = p(s| \ubm{x}; \bm{H}^{\mathrm{old}}) p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$.
+$p(s| \ubm{x}; \bm{H}^{\mathrm{old}})$ that can be computed by results of forward/backward and \autoref{eq:kappa-posterior}.
+The posterior for $\kappa$ can be obtained by
+\begin{equation}
+  p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}}) = \frac{p(\kappa, \ubm{x} | s; \bm{H}^{\mathrm{old}})}{\sum_{\kappa}{p(\kappa, \ubm{x} | s; \bm{H}^{\mathrm{old}})}} = \frac{\pi_{s, \kappa} p( \bm{x} | s; \bm{H}^{\mathrm{old}})}{\sum_{\kappa}{\pi_{s, \kappa} p(\bm{x} | s; \bm{H}^{\mathrm{old}})}}
+  \end{equation}
 
@@
 \end{algorithm}
+
+\subsubsection{Generator-tied HMM (GtHMM)}
+\begin{figure}[!h]
+  \centering
+  \begin{tikzpicture}
+    \tikzstyle{enode} = [thick, draw=blue, ellipse, inner sep = 1pt,  align=center]
+                                           %                                            \tikzstyle{nnode} = [thick, rectangle, rounded corners = 1pt,minimum size = 1.2cm,draw,inner sep = 2pt]
+    \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]
+    \node[enode] (z1) at (0,1.8) {$\bm{y}\sim p_1(\bm{y})$};
+    \node[enode] (z2) at (0,0.5){$\bm{y}\sim p_{2}(\bm{y})$};
+    \node[enode] (zK) at (0,-1.8) {$\bm{y}\sim p_{|\Ss|}(\bm{y})$}; 
+    \node[enode] (x) at (5.8,1){$\ubm{x}\sim p(\ubm{x};\bm{H})$};
+                                           %                                            \node at (3,-1){$p(\bm{z}) = \sum_k\pi_kp_k(\bm{z})$};
+    \node[nnode] (g) at (3.5,1) {$\bm{g}$};
+
+    \node[nnode] (m) at (3.5,-2) {Memory};
+    \node[nnode] (a) at (3.5,-0.5) {${A}$};
+
+    \draw[dotted,line width=2pt] (0,-0.3) -- (0,-1.2);
+    \filldraw[->] (1.8, 0.5)circle (2pt) --  (g) ;
+    \draw[->] (z1) -- (1.6, 1.8);
+    \draw[->] (z2) -- (1.6, 0.5);
+                                           %                                            \draw[->] (1,-0.5) -- (1.8, -0.5);
+    \draw[->] (zK) -- (1.6, -1.8);
+                                           %                                            \draw[->] (z1) [in=180,out=0]  to node[right]{$\pi_1$} (g);
+                                           %                                            \draw[->] (z2) [in=180,out=0]  to node[above]{$\pi_2$} (g);
+                                           %                                            \draw[->,dashed] (zK) [in=180,out=0]  to node[right]{$\pi_K$} (g);
+    \draw[->] (g) to (x);
+    \begin{scope}[ thick, every node/.style={sloped,allow upside down}]
+      \draw (2.1,0.9)-- (2.2, 0.);
+      \draw (2.2,0.)-- node {\midarrow} (2.2,-2);
+      \draw (2.2,-2)-- (m);
+      \draw (m)-- (5, -2);
+      \draw (5, -2)-- node {\midarrow} (5 ,-0.5);
+      \draw (5, -0.5) -- (a);
+      \draw (a)-- node {\midarrow} (2.2, -0.5);
+    \end{scope}
+    \node at (4.8, -1) {$s_{t}$};
+    \node at (2.56, -0.25) {$s_{t+1}$};
+
+  \end{tikzpicture}
+  \caption{LatM-HMM Model defined by $\bm{H} = \{\Ss, \bm{q}, A, p({\bm{x}}|{\bm{s}}; \bm{\Phi}) \}$}
+  \vspace{0.1cm}
+\end{figure}
+
+\textcolor{blue}{to be continued...}
+
+Alternatively, we can use a latent-source mixed HMM (LatM-HMM) where different latent source share the same generator functioning as feature mapping. Then the generator of the LatM-HMM is defined as
+\begin{equation}
+  \left\{\bm{g}| \bm{g}: \bm{z}\rightarrow \bm{x}, s \in \Ss, \bm{z}\sim p_{s}(\bm{z})\right\}.
+\end{equation}
+We use $\bm{f} = \bm{g}^{-1}$ to denote inverse of $\bm{g}$ and use $\bm{\theta}$ to denote the parameter set of $\bm{g}$. Then the conditional probability for LatM-HMM is modeled as
+\begin{align}
+  p(\bm{x} | s; \bm{\Phi})
+  &= p_s(\bm{z}) \bigg| \det \left( \pd{\bm{z}}{\bm{x}} \right) \bigg| \nonumber \\
+  &= p_s(\bm{f}(\bm{x})) \bigg| \det \left( \pd{\bm{f}(\bm{x})}{\bm{x}} \right) \bigg|
+\end{align}
+
+The parameter set for this model to be decide is $\bm{\Phi} = \left\{ \bm{\theta},  \bm{\omega}_s, \forall s\in \Ss\right\}$.
+Then the problem in \autoref{eq:sub-gm} can be reformulated as:
+\begin{align}
+  &\umax{\bm{\Phi}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) \nonumber \\
+  =&\umax{ \bm{\theta}, \bm{\omega}_s, \forall s\in \Ss} \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \left[\log\,p_{\smtr{s}{t}{r}}(\bm{f}(\bmtr{x}{t}{r})) + \log\,\bigg| \det\, \left( \pd{\bm{f}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg| \right].
+\end{align}
 
@@
 
-\subsubsection{Generator-shared HMM (GSHMM)...to be continued}
-\begin{figure}[!h]
-  \centering
-  \begin{tikzpicture}
-    \tikzstyle{enode} = [thick, draw=blue, ellipse, inner sep = 1pt,  align=center]
-                                           %                                            \tikzstyle{nnode} = [thick, rectangle, rounded corners = 1pt,minimum size = 1.2cm,draw,inner sep = 2pt]
-    \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]
-    \node[enode] (z1) at (0,1.8) {$\bm{z}\sim p_1(\bm{z})$};
-    \node[enode] (z2) at (0,0.5){$\bm{z}\sim p_{2}(\bm{z})$};
-    \node[enode] (zK) at (0,-1.8) {$\bm{z}\sim p_{|\Ss|}(\bm{z})$}; 
-    \node[enode] (x) at (5.8,1){$\ubm{x}\sim p(\ubm{x};\bm{H})$};
-                                           %                                            \node at (3,-1){$p(\bm{z}) = \sum_k\pi_kp_k(\bm{z})$};
-    \node[nnode] (g) at (3.5,1) {$\bm{g}$};
-
-    \node[nnode] (m) at (3.5,-2) {Memory};
-    \node[nnode] (a) at (3.5,-0.5) {${A}$};
-
-    \draw[dotted,line width=2pt] (0,-0.3) -- (0,-1.2);
-    \filldraw[->] (1.8, 0.5)circle (2pt) --  (g) ;
-    \draw[->] (z1) -- (1.6, 1.8);
-    \draw[->] (z2) -- (1.6, 0.5);
-                                           %                                            \draw[->] (1,-0.5) -- (1.8, -0.5);
-    \draw[->] (zK) -- (1.6, -1.8);
-                                           %                                            \draw[->] (z1) [in=180,out=0]  to node[right]{$\pi_1$} (g);
-                                           %                                            \draw[->] (z2) [in=180,out=0]  to node[above]{$\pi_2$} (g);
-                                           %                                            \draw[->,dashed] (zK) [in=180,out=0]  to node[right]{$\pi_K$} (g);
-    \draw[->] (g) to (x);
-    \begin{scope}[ thick, every node/.style={sloped,allow upside down}]
-      \draw (2.1,0.9)-- (2.2, 0.);
-      \draw (2.2,0.)-- node {\midarrow} (2.2,-2);
-      \draw (2.2,-2)-- (m);
-      \draw (m)-- (5, -2);
-      \draw (5, -2)-- node {\midarrow} (5 ,-0.5);
-      \draw (5, -0.5) -- (a);
-      \draw (a)-- node {\midarrow} (2.2, -0.5);
-    \end{scope}
-    \node at (4.8, -1) {$s_{t}$};
-    \node at (2.56, -0.25) {$s_{t+1}$};
-
-  \end{tikzpicture}
-  \caption{LatM-HMM Model defined by $\bm{H} = \{\Ss, \bm{q}, A, p({\bm{x}}|{\bm{s}}; \bm{\Phi}) \}$}
-  \vspace{0.1cm}
-\end{figure}
-
-\textcolor{blue}{to be continued...}
-
-Alternatively, we can use a latent-source mixed HMM (LatM-HMM) where different latent source share the same generator functioning as feature mapping. Then the generator of the LatM-HMM is defined as
-\begin{equation}
-  \left\{\bm{g}| \bm{g}: \bm{z}\rightarrow \bm{x}, s \in \Ss, \bm{z}\sim p_{s}(\bm{z})\right\}.
-\end{equation}
-We use $\bm{f} = \bm{g}^{-1}$ to denote inverse of $\bm{g}$ and use $\bm{\theta}$ to denote the parameter set of $\bm{g}$. Then the conditional probability for LatM-HMM is modeled as
-\begin{align}
-  p(\bm{x} | s; \bm{\Phi})
-  &= p_s(\bm{z}) \bigg| \det \left( \pd{\bm{z}}{\bm{x}} \right) \bigg| \nonumber \\
-  &= p_s(\bm{f}(\bm{x})) \bigg| \det \left( \pd{\bm{f}(\bm{x})}{\bm{x}} \right) \bigg|
-\end{align}
-
-The parameter set for this model to be decide is $\bm{\Phi} = \left\{ \bm{\theta},  \bm{\omega}_s, \forall s\in \Ss\right\}$.
-Then the problem in \autoref{eq:sub-gm} can be reformulated as:
-\begin{align}
-  &\umax{\bm{\Phi}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) \nonumber \\
-  =&\umax{ \bm{\theta}, \bm{\omega}_s, \forall s\in \Ss} \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \left[\log\,p_{\smtr{s}{t}{r}}(\bm{f}(\bmtr{x}{t}{r})) + \log\,\bigg| \det\, \left( \pd{\bm{f}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg| \right].
-\end{align}
-
-\subsubsection{Generator Shared HMM (GSHMM)}
-
-\textbf{To be continued}...
-
 \section{On Implementation of acoustic signal}

### 2019-08-09 bin/compute_accuracy.py
stop tracking codes at the branch

@@ -1,112 +1,0 @@
-import sys
-sys.path.append("..")
-
-import os
-from src.genHMM import load_model
-from functools import partial
-import pickle as pkl
-import numpy as np
-from parse import parse
-from src.utils import pad_data, TheDataset
-
-import torch
-from torch.utils.data import DataLoader
-
-
-def accuracy_fun(data_file, mdl=None):
-    X = pkl.load(open(data_file, "rb"))
-    # Get the length of all the sequences
-    l = [xx.shape[0] for xx in X]
-    # zero pad data for batch training
-    max_len_ = max([xx.shape[0] for xx in X])
-    x_padded = pad_data(X, max_len_)
-    batchdata = DataLoader(dataset=TheDataset(x_padded,
-                                              lengths=l,
-                                              device=mdl.hmms[0].device),
-                           batch_size=512, shuffle=True)
-    
-    true_class = parse("{}_{}.pkl", os.path.basename(data_file))[1]
-    out_list = [mdl.forward(x) for x in batchdata]
-    out = np.concatenate(out_list, axis=1)
-
-    # the out here should be the shape: data_size * nclasses
-    class_hat = np.argmax(out, axis=0) + 1
-    istrue = class_hat == int(true_class)
-    return "{}/{}".format(str(istrue.sum()), str(istrue.shape[0]))
-
-
-def append_class(data_file, iclass):
-    return data_file.replace(".pkl", "_" + str(iclass)+".pkl")
-
-
-def divide(res_int):
-    return res_int[0] / res_int[1]
-
-
-def parse_(res_str):
-    res_str_split = res_str.split("/")
-    res_int = [int(x) for x in res_str_split]
-    return res_int
-
-
-def print_results(mdl_file, data_files, results):
-    epoch=parse("epoch{}.mdl",os.path.basename(mdl_file))[0]
-    # Print class by class accuracy
-    for res, data_f in zip(results, data_files):
-        true_class = parse("{}_{}.pkl", os.path.basename(data_f[0]))[1]
-        
-        print("epoch:",epoch, "class:",true_class, mdl_file, data_f[0].astype("<U"), res[0], divide(parse_(res[0].astype("<U"))), sep='\t', file=sys.stdout)
[…]

### 2019-08-09 bin/test.py

@@ -1,122 +1,0 @@
-# file test.py
-import sys
-sys.path.append('/home/dong/Documents/phd_research/my_papers/gm-hmm')
-
-import pickle as pkl
-import sys
-import os
-import numpy as np
-
-from src.genHMM import GenHMM
-
-
-
-
-def getsubset(data, label, iphn):
-    # concat data
-    # find subset
-    idx = np.in1d(label, iphn)
-    return data[idx], label[idx]
-
-
-def to_phoneme_level(DATA):
-    n_sequences = len(DATA)
-
-    seq_train = [0 for _ in range(n_sequences)]
-    targets_train = [0 for _ in range(n_sequences)]
-    data_tr = []
-    labels_tr = []
-
-    # For all sentences
-    for i, x in enumerate(DATA):
-        # Note: first column contains labels. Find when the label changes, it indicates a change of phoneme in the sentence
-        dx = np.diff(x[:, 0])
-
-        # Make a clean vector to delimit phonemes
-        change_locations = np.array([0] + (1 + np.argwhere(dx != 0)).reshape(-1).tolist() + [x.shape[0]-1])
-
-        # Make an array of size n_phoneme_in_sentence x 2, containing begining and end of each phoneme in a sentence
-        seq_train[i] = np.array([[change_locations[i-1], change_locations[i]]\
-                                 for i in range(1, change_locations.shape[0]) ])#np.diff(change_locations)
-
-        # Save an instance of labels
-        targets_train[i] = x[change_locations[:-1], 0]
-
-        # For each phoneme found in the sentence, get the sequence of MFCCs and the label
-
-        for j in range(seq_train[i].shape[0]):
-            data_tr += [x[seq_train[i][j][0]:seq_train[i][j][1]]]
-            labels_tr += [targets_train[i][j]]
-
-    # Return an array of arrays for the data, and an array of float for the labels
-    return np.array(data_tr), np.array(labels_tr)
-
-
-def prepare_data(data_folder=None, fname_dtest=None, fname_dtrain=None, n_phn=None, verbose=False):
-    # Read the datafiles
-    te_DATA, te_keys, te_lengths, codebook, te_PHN = pkl.load(open(fname_dtest, "rb"))
-    tr_DATA, tr_keys, tr_lengths, tr_PHN = pkl.load(open(fname_dtrain, "rb"))
-    if verbose:
[…]

### 2019-08-09 bin/train.py

@@ -1,196 +1,0 @@
-# file test.py
-import sys
-sys.path.append('.')
-
-import collections
-import pickle as pkl
-
-import sys
-import os
-import numpy as np
-
-from src.genHMM import GenHMM
-
-
-
-
-def getsubset(data, label, iphn):
-    # concat data
-    # find subset
-    idx = np.in1d(label, iphn)
-    return data[idx], label[idx]
-
-
-def to_phoneme_level(DATA):
-    n_sequences = len(DATA)
-
-    seq_train = [0 for _ in range(n_sequences)]
-    targets_train = [0 for _ in range(n_sequences)]
-    data_tr = []
-    labels_tr = []
-
-    # For all sentences
-    for i, x in enumerate(DATA):
-        # Note: first column contains labels. Find when the label changes, it indicates a change of phoneme in the sentence
-        dx = np.diff(x[:, 0])
-
-        # Make a clean vector to delimit phonemes
-        change_locations = np.array([0] + (1 + np.argwhere(dx != 0)).reshape(-1).tolist() + [x.shape[0]-1])
-
-        # Make an array of size n_phoneme_in_sentence x 2, containing begining and end of each phoneme in a sentence
-        seq_train[i] = np.array([[change_locations[i-1], change_locations[i]]\
-                                 for i in range(1, change_locations.shape[0]) ])#np.diff(change_locations)
-
-        # Save an instance of labels
-        targets_train[i] = x[change_locations[:-1], 0]
-
-        # Delete label from data
-        x[:, 0] = 0
-
-        # For each phoneme found in the sentence, get the sequence of MFCCs and the label
-
-        for j in range(seq_train[i].shape[0]):
-            data_tr += [x[seq_train[i][j][0]:seq_train[i][j][1]]]
-            labels_tr += [targets_train[i][j]]
-
-    # Return an array of arrays for the data, and an array of float for the labels
-    return np.array(data_tr), np.array(labels_tr)
-
-
-def prepare_data(data_folder=None, fname_dtest=None, fname_dtrain=None, n_phn=None, verbose=False):
-    # Read the datafiles
-    te_DATA, te_keys, te_lengths, codebook, te_PHN = pkl.load(open(fname_dtest, "rb"))
-    tr_DATA, tr_keys, tr_lengths, tr_PHN = pkl.load(open(fname_dtrain, "rb"))
-    if verbose:
-        print("Data loaded from files.")
-
[…]

### 2019-08-09 bin/train_class.py

@@ -1,91 +1,0 @@
-import os
-import sys
-sys.path.append("..")
-
-from parse import parse
-import pickle as pkl
-from src.genHMM import GenHMM, save_model, load_model
-from src.utils import pad_data, TheDataset
-import torch
-from torch.utils.data import DataLoader
-import json
-
-
-if __name__ == "__main__":
-    usage = "python bin/train_class.py data/train13.pkl models/epoch1_class1.mdlc param.json"
-    if len(sys.argv) < 3 or sys.argv[1] == "-h" or sys.argv[1] == "--help":
-        print(usage)
-        sys.exit(1)
-
-    # Parse
-    train_inputfile = sys.argv[1]
-    out_mdl = sys.argv[2]
-
-    # Test for a third parameter
-    try:
-        param_file = sys.argv[3]
-    except IndexError:
-        param_file = "default.json"
-
-
-    epoch_str, iclass_str = parse('epoch{}_class{}.mdlc',os.path.basename(out_mdl))
-    train_class_inputfile = train_inputfile.replace(".pkl", "_{}.pkl".format(iclass_str))
-
-    #  Load data
-    
-    xtrain = pkl.load(open(train_class_inputfile, "rb"))
-    #xtrain = xtrain[:100]
-    # Get the length of all the sequences
-    l = [x.shape[0] for x in xtrain]
-
-    #  Load or create model
-    if epoch_str == '1':
-        with open(param_file) as f_in:
-            options = json.load(f_in)
-
-        mdl = GenHMM(**options)
-
-    else:
-        # Load previous model
-        mdl = load_model(out_mdl.replace("epoch" + epoch_str, "epoch" + str(int(epoch_str)-1)))
-
-    mdl.iepoch = epoch_str
-    mdl.iclass = iclass_str
-
-
-    mdl.device = 'cpu'
-    if torch.cuda.is_available():
-        device = torch.device('cuda')
-        mdl.device = device
-    
-    print("epoch:{}\tclass:{}\tPush model to {}...".format(epoch_str,iclass_str, mdl.device), file=sys.stdout)
-    mdl.pushto(mdl.device)   
-    
-    
-    # zero pad data for batch training
-    max_len_ = max([x.shape[0] for x in xtrain])
-    xtrain_padded = pad_data(xtrain, max_len_)
-    traindata = DataLoader(dataset=TheDataset(xtrain_padded, lengths=l, device=mdl.device), batch_size=1024, shuffle=True)
-    
-
[…]

### 2019-08-09 practice/realnvp.py

@@ -1,75 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-#% matplotlib inline
-
-from pylab import rcParams
-
-rcParams['figure.figsize'] = 10, 8
-rcParams['figure.dpi'] = 300
-
-import torch
-from torch import nn
-from torch import distributions
-from torch.nn.parameter import Parameter
-
-from sklearn import cluster, datasets, mixture
-from sklearn.preprocessing import StandardScaler
-from src.realnvp import RealNVP
-
-if __name__ == "__main__":
-    H = 28
-    d = 2
-    nchain = 3
-    niter = 5000
-    n_samples_train = 100
-    n_samples_test = 1000
-
-    nets = lambda: nn.Sequential(nn.Linear(d, H), nn.LeakyReLU(), nn.Linear(H, H), nn.LeakyReLU(), nn.Linear(H, d),
-                                 nn.Tanh())
-    nett = lambda: nn.Sequential(nn.Linear(d, H), nn.LeakyReLU(), nn.Linear(H, H), nn.LeakyReLU(), nn.Linear(H, d))
-
-    masks = torch.from_numpy(np.array([[0, 1], [1, 0]] * nchain).astype(np.float32))
-    prior = distributions.MultivariateNormal(torch.zeros(d), torch.eye(d))
-    flow = RealNVP(nets, nett, masks, prior)
-
-    optimizer = torch.optim.Adam([p for p in flow.parameters() if p.requires_grad == True], lr=1e-4)
-
-    for t in range(niter):
-        noisy_moons = datasets.make_moons(n_samples=n_samples_train, noise=.05)[0].astype(np.float32)
-        x = torch.from_numpy(noisy_moons).reshape(n_samples_train, 1, d)
-        mask = torch.ones((x.shape[0], x.shape[1]), dtype=torch.uint8)
-        loss = -flow.log_prob(x, mask).mean()
-
-        optimizer.zero_grad()
-        loss.backward(retain_graph=True)
-        optimizer.step()
-
-        if t % 500 == 0:
-            print('iter %s:' % t, 'loss = %.3f' % loss)
-
-
-    noisy_moons = datasets.make_moons(n_samples=n_samples_test, noise=.05)[0].astype(np.float32)
[…]

### 2019-08-09 src/glow/models.py

@@ -1,524 +1,0 @@
-import torch
-import torch.nn as nn
-import torch.nn.functional as F
-import numpy as np
-from tqdm import tqdm
-from . import thops
-from . import modules
-from . import utils
-from scipy import stats
-
-
-def prior_sampler(pk, size=64):
-    pk = tuple(pk.numpy())
-    xk = np.arange(len(pk))
-    custm = stats.rv_discrete(name='custm', values=(xk, pk))
-    prior_samples = custm.rvs(size=size)
-    prior_samples = torch.from_numpy(prior_samples)
-    return prior_samples
-
-
-def f(in_channels, out_channels, hidden_channels):
-    return nn.Sequential(
-        modules.Conv2d(in_channels, hidden_channels), nn.ReLU(inplace=False),
-        modules.Conv2d(hidden_channels, hidden_channels, kernel_size=[1, 1]), nn.ReLU(inplace=False),
-        modules.Conv2dZeros(hidden_channels, out_channels))
-
-
-class FlowStep(nn.Module):
-    FlowCoupling = ["additive", "affine"]
-    FlowPermutation = {
-        "reverse": lambda obj, z, logdet, rev: (obj.reverse(z, rev), logdet),
-        "shuffle": lambda obj, z, logdet, rev: (obj.shuffle(z, rev), logdet),
-        "invconv": lambda obj, z, logdet, rev: obj.invconv(z, logdet, rev)
-    }
-
-    def __init__(self, in_channels, hidden_channels,
-                 actnorm_scale=1.0,
-                 flow_permutation="invconv",
-                 flow_coupling="additive",
-                 LU_decomposed=False):
-        # check configures
-        assert flow_coupling in FlowStep.FlowCoupling, \
-            "flow_coupling should be in `{}`".format(FlowStep.FlowCoupling)
-        assert flow_permutation in FlowStep.FlowPermutation, \
-            "float_permutation should be in `{}`".format(
-                FlowStep.FlowPermutation.keys())
-        super().__init__()
-        self.flow_permutation = flow_permutation
-        self.flow_coupling = flow_coupling
-        # 1. actnorm
-        self.actnorm = modules.ActNorm2d(in_channels, actnorm_scale)
-        # 2. permute
-        if flow_permutation == "invconv":
-            self.invconv = modules.InvertibleConv1x1(
-                in_channels, LU_decomposed=LU_decomposed)
-        elif flow_permutation == "shuffle":
[…]

### 2019-08-09 src/glow/trainer.py

@@ -1,271 +1,0 @@
-import re
-import os
-import torch
-import torch.nn.functional as F
-
-import datetime
-import numpy as np
-from tqdm import tqdm
-from tensorboardX import SummaryWriter
-from torch.utils.data import DataLoader
-from .utils import save, load, plot_prob, new2old
-from .config import JsonConfig
-from .models import Glow
-from . import thops
-
-
-class Trainer(object):
-    def __init__(self, graph, graph_prior, optim, lrschedule, loaded_step,
-                 devices, data_device,
-                 dataset, hparams):
-        if isinstance(hparams, str):
-            hparams = JsonConfig(hparams)
-        # set members
-        # append date info
-        date = str(datetime.datetime.now())
-        date = date[:date.rfind(":")].replace("-", "")\
-                                     .replace(":", "")\
-                                     .replace(" ", "_")
-        self.log_dir = os.path.join(hparams.Dir.log_root, "log_" + date)
-        
-        self.checkpoints_dir = os.path.join(self.log_dir, "checkpoints")
-        if not os.path.exists(self.log_dir):
-            os.makedirs(self.log_dir)
-        # write hparams
-        hparams.dump(self.log_dir)
-        if not os.path.exists(self.checkpoints_dir):
-            os.makedirs(self.checkpoints_dir)
-        self.checkpoints_gap = hparams.Train.checkpoints_gap
-        self.prior_gap = hparams.Train.prior_gap
-        self.em_gap = hparams.Train.em_gap
-        self.max_checkpoints = hparams.Train.max_checkpoints
-        # set the mixture prior
-        self.num_component = hparams.Mixture.num_component
-        self.mix_prior = torch.ones(self.num_component)
-        # model relative
-        self.graph = graph
-
-        self.optim = optim
-        self.weight_y = hparams.Train.weight_y
-        # grad operation
-        self.max_grad_clip = hparams.Train.max_grad_clip
-        self.max_grad_norm = hparams.Train.max_grad_norm
-        # copy devices from built graph
-        self.devices = devices
-        self.data_device = data_device
-        # number of training batches
-        self.batch_size = hparams.Train.batch_size
-        self.data_loader = DataLoader(dataset,
-                                      batch_size=self.batch_size,
-                                    #   num_workers=8,
-                                      shuffle=True,
-                                      drop_last=True)
-        self.n_epoches = (hparams.Train.num_batches+len(self.data_loader)-1)
[…]

### 2019-08-09 src/test.py

@@ -1,141 +1,0 @@
-# test the batch forward, backward algorithm, and _compute_log_xi_sum
-import os
-import sys
-sys.path.append("..")
-from parse import parse
-from scipy.special import logsumexp
-import pickle as pkl
-import numpy as np
-import torch
-from torch.utils.data import Dataset, DataLoader
-from hmmlearn import _hmmc
-from _torch_hmmc import _forward, _backward, _compute_log_xi_sum
-
-
-
-def batch_data(n_components):
-    """ Generate a batch of framelogprob with zero paddings`
-    Input:
-    n_components: number of hmm states 
-
-    OUTPUT: 
-
-    data: Batch framelogprob (randomly generated) with zero paddings
-          Shape: batch_size * max_sequence_length * feature vector length
-    
-    mask: The mask indicating which element of data is actually framelogprob, which element is padding
-          Shape: batch_size * max_sequence_length
-    """
-    x = []
-    # generate framelogprob 1, assume number of states is 4
-    x.append(- np.abs(np.random.randn(3, n_components) ) )
-    x.append(- np.abs(np.random.randn(5, n_components) ) )
-
-    # assume do padding to 8 samples for every sequence
-
-    length = 8
-    d = x[0].shape[1]
-    # generate xx here
-    # mask should be included here
-    data = np.array([np.concatenate((xx,np.zeros((length - xx.shape[0] + 1,d)))) for xx in x])
-    mask = np.array( [xx.sum(axis=1) < 0 for xx in data ], dtype=np.uint8)
-    
-    return (data, mask)
-
-# 1. batch data load
-n_components = 4
-startprob = np.abs(np.random.randn(n_components))
-startprob = startprob/startprob.sum()
-print("hmm start probability: {}".format(startprob) )
-
-transmat = np.abs(np.random.randn(n_components, n_components))
-transmat = transmat/transmat.sum(axis=0)
-print("hmm transition probability: \n{}".format(transmat) )
-
-data, mask = batch_data(n_components)
-print("batch_framelogprob is: \n{}".format(data))
-print("batch_mask is: \n{}".format(mask))
-
-# 2. forward comparison
-
-# 2.1 forward by hmmlearn method:
-hmmlearn_fwdlattice = []
-hmmlearn_logprob = []
-for idx, framelogprob in enumerate(data):
-    fwdlattice = np.zeros((mask[idx].sum(), n_components))
-    _hmmc._forward(int(mask[idx].sum()), \
-                   int(n_components), \
-                   np.log(startprob), \
[…]

### 2019-08-09 src/timit-preprocessor/extract_mfcc.sh

@@ -1,36 +1,0 @@
-#!/usr/bin/env bash
-
-if [ ${#} != 3 ];then
-    echo "Usage: ./mfcc_extract.sh [Kaldi's relative path] [TIMIT relative path] [output .scp filename]"
-    exit -1
-fi
-
-kaldi=${1}
-timit=${2}
-target_file=${3}
-
-path=.data/processed
-options="--use-energy=false"
-
-mkdir -p $path
-
-
-fname=$(basename "$target_file")
-fname_trunk="${fname%.*}"
-target=(${fname_trunk//./ })
-
-echo $fname
-
-log=$path/$target.extract.log
-if [[ $fname == *".13."* ]]; then
-	${kaldi}/src/featbin/compute-mfcc-feats --verbose=2 $options scp:.data/material/$target.wav.scp ark,t,scp:$path/$target.13.ark,$target_file 2> $log
-fi
-
-if [[ $fname == *".39."* ]]; then
-	${kaldi}/src/featbin/compute-mfcc-feats --verbose=2 $options scp:.data/material/$target.wav.scp ark,t,scp:$path/$target.13.ark,$target_file 2> $log
-	${kaldi}/src/featbin/add-deltas ark:$path/$target.13.ark ark,t:$path/$target.13_deltas.ark
-	${kaldi}/src/featbin/compute-cmvn-stats ark:$path/$target.13_deltas.ark ark,t:$path/$target.cmvn_results.ark
-	${kaldi}/src/featbin/apply-cmvn ark:$path/$target.cmvn_results.ark ark:$path/$target.13_deltas.ark ark,t,scp:$path/$target.39.cmvn.ark,$target_file
-fi
-
-exit 0

### 2019-08-10 manuscript/main.tex
adjust sections

@@
-\documentclass{article}
+%\def\year{2020}\relax
+%File: formatting-instruction.tex
+\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
+\usepackage{aaai20}  % DO NOT CHANGE THIS
+\usepackage{times}  % DO NOT CHANGE THIS
+\usepackage{helvet} % DO NOT CHANGE THIS
+\usepackage{courier}  % DO NOT CHANGE THIS
+\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
+\usepackage{graphicx} % DO NOT CHANGE THIS
+\urlstyle{rm} % DO NOT CHANGE THIS
+\def\UrlFont{\rm}  % DO NOT CHANGE THIS
+\usepackage{graphicx}  % DO NOT CHANGE THIS
+\frenchspacing  % DO NOT CHANGE THIS
+\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
+\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
+%\nocopyright
+%PDF Info Is REQUIRED.
+% For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
+% For /Title, add Title in Mixed Case. No accents or commands. Retain the parentheses.
+ \pdfinfo{/Title gm_hmm
+} %Leave this	
+\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}}
+\newcommand{\citep}{\cite}
+\newcommand{\citealp}[1]{\citeauthor{#1} \citeyear{#1}}
+%%%%%%%%%%%%%%%%%%%%%%%%
 
-\usepackage{hyperref} %BUG if put after
 \usepackage{tikz}
@@
 
-\usepackage{url}
-\usepackage{fullpage}
-\usepackage{cite}
 \usepackage{caption}
@@
 
-\hypersetup{  
-  bookmarks=true,
-  backref=true,
-  pagebackref=false,
-  colorlinks=true,
-  linkcolor=blue,
-  citecolor=red,
-  urlcolor=blue,
-  pdftitle={Generative Model HMM},
-  pdfauthor={Dong Liu},
-  pdfsubject={}
-}
+% \hypersetup{  
+%   bookmarks=true,
+%   backref=true,
+%   pagebackref=false,
+%   colorlinks=true,
+%   linkcolor=blue,
+%   citecolor=red,
+%   urlcolor=blue,
+%   pdftitle={Generative Model HMM},
+%   pdfauthor={Dong Liu},
+%   pdfsubject={}
+% }
+
+\setcounter{secnumdepth}{2} %May be changed to 1 or 2 if section numbers are desired.
+
+% The file aaai20.sty is the style file for AAAI Press 
+% proceedings, working notes, and technical reports.
+%
+\setlength\titlebox{2.5in} % If your paper contains an overfull \vbox too high warning at the beginning of the document, use this
+% command to correct it. You may not alter the value below 2.5 in
 
@@
 \author{
-  Firsthand Scientists
-}
+  Dong Liu, Antoine Honoré, Saikat Chatterjee, and Lars~K. Rasmussen\\
+  KTH Royal Institute of Technology, Stockholm, Sweden \\
+  E-mail: \{doli, honore, sach, lkra\}@kth.se}
 
@@
 \maketitle
-\section{Notation}
-Time is indexed by subscript and sequence is denoted by underline. $\bm{x}_t$ is signal at time $t$. The sequential time is denoted by $\ubar{\bm{x}} = \left[ \bm{x}_1,
-  \cdots, \bm{x}_T\right]^{\intercal}$, where $[\cdot]^{\intercal}$ means transpose and $T$ is the length of the sequence.
-Sequential signal or clip uses underline notation and is indexed by superscript, for instance $\ubar{\bm{x}}^{(r)}$ means the $r$-th sequential signal, where $r = 1, 2, \cdots, R.$, and $\ubmr{x}{r} = \left[ \bmtr{x}{1}{r}, \bmtr{x}{2}{r}, \cdots, \bmtr{x}{T^{(r)}}{r} \right]$ with length ${T^{(r)}}$.
-Note different sequential signal $\ubmr{x}{r}$ could have different lengths.
+\begin{abstract}
+to be written...
+\end{abstract}
 
-The hypothesis of Hidden Markov Model (HMM): $\Hh := \{\bm{H} | \{\Ss, \bm{q}, A, p(\bm{x}|{s}; \bm{\Phi})\}$,
+\section{introduction}
+
+\section{Preliminaries}
+\begin{figure}[!h]
+  \centering
+  \begin{tikzpicture}
+    \tikzstyle{enode} = [thick, draw=black, ellipse, inner sep = 1pt,  align=center]
+    \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]
+    \node[enode] (g1) at (-0.5,1.8) {$p(\bm{x}| s=1; \bm{\Phi}_{1})$};
+    \node[enode] (g2) at (-0.5,0.5) {$p(\bm{x}| s=2; \bm{\Phi}_{2})$};
+    \node[enode] (gs) at (-0.5, -1.8) {$p(\bm{x}| s=|\Ss|; \bm{\Phi}_{|\Ss|})$};
+    \node[enode] (x) at (4.5,1.5){$\ubm{x}\sim p(\ubm{x};\bm{H})$};
+
+    \draw[dotted,line width=2pt] (0,-0.3) -- (0,-1.2);
+    \filldraw[->] (1.9, 0.5)circle (2pt) --  (x) ;
+    \draw[->] (g1) -- (1.8, 1.8);
+    \draw[->] (g2) -- (1.8, 0.5);
+    \draw[->] (gs) -- (1.8, -1.8);
+
+    \begin{scope}[xshift=0.5cm, thick, every node/.style={sloped,allow upside down}]
+      \node[nnode] (m) at (3.5,-2) {Memory};
+      \node[nnode] (a) at (3.5,-0.5) {$\bm{A}$};
+
+      \draw (2.1,0.9)-- (2.2, 0.);
+      \draw (2.2,0.)-- node {\midarrow} (2.2,-2);
+      \draw (2.2,-2)-- (m);
+      \draw (m)-- (5, -2);
+      \draw (5, -2)-- node {\midarrow} (5 ,-0.5);
+      \draw (5, -0.5) -- (a);
+      \draw (a)-- node {\midarrow} (2.2, -0.5);
+      \node at (4.8, -1) {$s_{t}$};
+      \node at (2.56, -0.25) {$s_{t+1}$};
+    \end{scope}
+  \end{tikzpicture}
+  \caption{HMM model illustration.}\label{fig:hmm}
+  \vspace{0.1cm}
+\end{figure}
+
+Our framework is an Hidden Markov Model (HMM).
+An HMM model $\bm{H}$ defined in a hypothesis space $\Hh$, i.e.
+$\bm{H} \in \Hh$, is capable to model time-span signal $\ubar{\bm{x}} = \left[ \bm{x}_1, \cdots, \bm{x}_T\right]^{\intercal}$, where $\bm{x}_t\in \RR^{N}$ is the signal at time $t$, $[\cdot]^{\intercal}$ means transpose, and $T$ denote the time length.
+Using HMM for signal representation or generating is illustrated in \ref{fig:hmm}.
+The assumption is that different instant signal of $\ubar{\bm{x}}$ is generated by different source, while the source selection is done by a discrete hidden markov process.
+We define the hypothesis set of HMM as $\Hh := \{\bm{H} | \{\Ss, \bm{q}, \bm{A}, p(\bm{x}|{s}; \bm{\Phi}_{s})\}$, where:
 \begin{itemize}
@@
 \item $\bm{q} = \left[ q_1, q_2, \cdots, q_{|\Ss|}\right]^\intercal$ initial distribution of HMM $\bm{H}$ with $|\Ss|$ is cardinality of $\Ss$, $q_k = p(s=k)$ for random state variable $s$.
-\item $A$ is the transition matrix for the HMM $\bm{H}$ of size $|\Ss| \times |\Ss|$.
-\item Observable signal density $p({\bm{x}}|{s};\bm{\Phi})$ given hidden state sequence, where $\bm{\Phi}$ is the parameter set that defines this conditional probabilistic model.
+In the following of this paper, we use $s_t$ to denote the state $s$ at time $t$.
+\item $\bm{A}$ is the transition matrix for the HMM $\bm{H}$ of size $|\Ss| \times |\Ss|$, $\bm{A}_{i,j} = p(s_{t+1}=j|s_{t}=i)$.
+\item For given hidden state $s$, the observable signal density is $p({\bm{x}}|{s};\bm{\Phi}_{s})$, where $\bm{\Phi}_{s}$ is the parameter set that defines this conditional probabilistic model.
 \end{itemize}
 
-\section{Problem Statement}
+In the framework of HMM, at each time instance, signal $\bm{x}_t$ is generated by density $p(\bm{x}_t| s_t; \bm{\Phi}_{s_t})$, and $s_t$ is decided by the hidden markov process.
+This process gives us the probabilistic model $p(\ubm{x};\bm{H})$
+\section{Generator-mixed HMM}
+
+\begin{figure}[!th]
+  \centering
+  \begin{tikzpicture}
+    \tikzstyle{enode} = [thick, draw=black, ellipse, inner sep = 2pt,  align=center]
+    \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]
+    \node[enode] (z1) at (-1.2,1.8) {$\bm{z}\sim p_{s,1}(\bm{z})$};
+    \node[nnode] (g1) at (1,1.8) {$\bm{g}_{s,1}$};
+    \node[enode] (z2) at (-1.2,0.5){$\bm{z}\sim p_{s,2}(\bm{z})$};
+    \node[nnode] (g2) at (1,0.5) {$\bm{g}_{s,2}$};
+    \node[enode] (zK) at (-1.2,-1.8) {$\bm{z}\sim p_{s,K}(\bm{z})$};
+    \node[nnode] (gs) at (1, -1.8) {$\bm{g}_{s,K}$};
+    \node[enode] (x) at (4.5,0){$\bm{x}\sim p(\bm{x}| s; \bm{\Phi}_{s})$};
+
+    \draw[dotted,line width=2pt] (0,-0.3) -- (0,-1.2);
+    \filldraw[->] (1.9, 0.5)circle (2pt) --  node[above=0.2]{${\kappa}\sim \bm{\pi}_{s}$} (x)  ;
+    \draw[->] (z1) -- (g1);
+    \draw[->] (g1) -- (1.8, 1.8);
+
+    \draw[->] (z2) -- (g2);
+    \draw[->] (g2) -- (1.8, 0.5);
+
+    \draw[->] (zK) -- (gs);
+    \draw[->] (gs) -- (1.8, -1.8);
+  \end{tikzpicture}
+  \caption{Source $s$ powered by generator mixed generative model (GenHMM).}
+  \vspace{0.1cm}
+\end{figure}
+
+In this section, we introduce a neural network based generator-mixed HMM model.
+We term it as GenHMM.
+GenHMM has a hidden markov process as typical HMM, while the probabilistic model of GenHMM for each hidden state is induced by mixture of neural network based generators.
+For $s\in \Ss$, we define the state's probabilistic model in GenHMM as
+\begin{equation}
+  p(\bm{x}| s; \bm{\Phi}_{s}) = \sum_{\kappa=1}^{K}\pi_{s, \kappa} p(\bm{x}| s, \kappa; \bm{\Phi}_{s, \kappa})
+\end{equation}
+where $\kappa$ is a random variable following a categorical distribution, and $\pi_{s, \kappa} = p(\kappa | \bm{x}, s; \bm{\Phi}_{s, \kappa})$.
+Naturally $\sum_{\kappa = 1}^{K} \pi_{s, \kappa}= 1$.
+Denote $\bm{\pi}_{s} = [\pi_{s,1}, \pi_{s,2}, \cdots, \pi_{s,K}]$,
+$\bm{\Phi}_s = \left\{ \pi_{s, \kappa}, \bm{\Phi}_{s, \kappa}| \kappa = 1, 2, \cdots, K \right\}$,
+
+\subsubsection{Generator Mixed HMM (GenHMM)}\label{subsubsec:GenHMM}
+In this subsection, we specify the GenHMM model and give the learning algorithm of this model.
+Base on the generative model assumption, for each state $s$, the probabilistic generative model is
+
+\subsubsection{Flow structure}
+
+\section{Learning Algorithm of GenHMM}
+\subsection{Problem Statement}
 Given a empirical distribution $\hat{p}(\ubm{x}) = \frac{1}{R}\sum_{r=1}^{R} \delta_{\ubmr{x}{r}}(\ubm{x})$. We want to find a probabilistic model such that:
@@
 
-\section{Proposal}
+\subsection{Proposal}
 
-\begin{figure}[!h]
-  \centering
-  \begin{tikzpicture}
-    \tikzstyle{enode} = [thick, draw=blue, ellipse, inner sep = 1pt,  align=center]
-    % \tikzstyle{nnode} = [thick, rectangle, rounded corners = 1pt,minimum size = 1.2cm,draw,inner sep = 2pt]
-    \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]
-    % \node[enode] (z1) at (-1.2,1.8) {Source 1}; 
-    \node[nnode] (g1) at (-0.5,1.8) {Source 1};
-
-    % \node[enode] (z2) at (-1.2,0.5){$\bm{z}\sim p_{2}(\bm{z})$};
-    \node[nnode] (g2) at (-0.5,0.5) {Source 2};
-
-    % \node[enode] (zK) at (-1.2,-1.8) {$\bm{z}\sim p_{|\Ss|}(\bm{z})$};
-    \node[nnode] (gs) at (-0.5, -1.8) {Source $|\Ss|$};
-
-    \node[enode] (x) at (4.5,1.5){$\ubm{x}\sim p(\ubm{x};\bm{H})$};
-    % \node at (3,-1){$p(\bm{z}) = \sum_k\pi_kp_k(\bm{z})$};
-
-    \draw[dotted,line width=2pt] (0,-0.3) -- (0,-1.2);
-    \filldraw[->] (1.9, 0.5)circle (2pt) --  (x) ;
-
-    \draw[->] (g1) -- (1.8, 1.8);
-
-    \draw[->] (g2) -- (1.8, 0.5);
-
-    % \draw[->] (1,-0.5) -- (1.8, -0.5);
-
-    \draw[->] (gs) -- (1.8, -1.8);
-
-    % \draw[->] (z1) [in=180,out=0]  to node[right]{$\pi_1$} (g);
-    % \draw[->] (z2) [in=180,out=0]  to node[above]{$\pi_2$} (g);
-    % \draw[->,dashed] (zK) [in=180,out=0]  to node[right]{$\pi_K$} (g);
-    % \draw[->] (g) to (x);
-    \begin{scope}[xshift=0.5cm, thick, every node/.style={sloped,allow upside down}]
-      \node[nnode] (m) at (3.5,-2) {Memory};
-      \node[nnode] (a) at (3.5,-0.5) {$\bm{A}$};
-
-      \draw (2.1,0.9)-- (2.2, 0.);
-      \draw (2.2,0.)-- node {\midarrow} (2.2,-2);
-      \draw (2.2,-2)-- (m);
-      \draw (m)-- (5, -2);
-      \draw (5, -2)-- node {\midarrow} (5 ,-0.5);
-      \draw (5, -0.5) -- (a);
-      \draw (a)-- node {\midarrow} (2.2, -0.5);
-      \node at (4.8, -1) {$s_{t}$};
-      \node at (2.56, -0.25) {$s_{t+1}$};
-    \end{scope}
-
-  \end{tikzpicture}
-  \caption{HMM Model defined by $\bm{H} = \{\Ss, \bm{q}, \bm{A}, p({\bm{x}}|{\bm{s}}; \bm{\Phi}) \}$}
-  \vspace{0.1cm}
-\end{figure}
-
-Since model $\bm{H}$ contains hidden sequential variable $\ubm{s}$, we can not directly solve the maximum likelihood problem in \autoref{eq:ml-of-hmm}.
+Since model $\bm{H}$ contains hidden sequential variable $\ubm{s}$, we can not directly solve the maximum likelihood problem in \ref{eq:ml-of-hmm}.
 We use expectation maximization (EM) to address the hidden variable problem by
@@
 
-The \autoref{eq:em-m-opt} can be reformulated as:
+The \eqref{eq:em-m-opt} can be reformulated as:
 \begin{equation}\label{eq:m-step-subs}
@@
     \STATE Compute $p(\smtr{s}{t}{r} | \ubmr{x}{r}; \bm{H}^{\mathrm{old}})$, $p(\smtr{s}{t}{r}, \smtr{s}{t+1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})$ by forward/backward algorithm;
-    \STATE $\bm{q} \gets \uargmin{\bm{q}}\, \Qq(\bm{q}; \bm{H}^{\mathrm{old}})$ by \autoref{eq:update-initial-state-prob};
-    \STATE $\bm{A} \gets \uargmin{\bm{A}}\Qq(\bm{A}; \bm{H}^{\mathrm{old}})$ by \autoref{eq:update-transition-prob};
-    \STATE $\bm{\Phi} \gets \uargmin{\bm{\Phi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})$ by calling Algorithm \autoref{algo:opt-Phi-GenHMM} or \autoref{algo:opt-Phi-LatHMM};
+    \STATE $\bm{q} \gets \uargmin{\bm{q}}\, \Qq(\bm{q}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:update-initial-state-prob};
+    \STATE $\bm{A} \gets \uargmin{\bm{A}}\Qq(\bm{A}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:update-transition-prob};
+    \STATE $\bm{\Phi} \gets \uargmin{\bm{\Phi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})$ by calling Algorithm~\ref{algo:opt-Phi-GenHMM} or \ref{algo:opt-Phi-LatHMM};
     \STATE $\bm{H}^{\mathrm{old}} \gets \bm{H}$
@@
 \subsection{Initial Probability Update}
-\autoref{eq:init-distribution-update} can be written as:
+\eqref{eq:init-distribution-update} can be written as:
 \begin{align}
@@
 \subsection{Transition Probability Update}
-\autoref{eq:transition-update} can be written as
+\eqref{eq:transition-update} can be written as
 \begin{align}
@@
 \subsection{Generative Model Update}
-\autoref{eq:generative-model-update} can be rewritten as
+\eqref{eq:generative-model-update} can be rewritten as
 \begin{align}\label{eq:gm-update}
@@
 
-Then the third subproblem of \autoref{eq:m-step-subs} becomes:
+Then the third subproblem of \eqref{eq:m-step-subs} becomes:
 \begin{align}\label{eq:sub-gm}
@@
 
-It could be seen from \autoref{eq:gm-update} that the key to update generate  model is to evaluate $p(\bm{x}|{s}; \bm{\Phi})$ for all $s \in \Ss$.
+It could be seen from \eqref{eq:gm-update} that the key to update generate  model is to evaluate $p(\bm{x}|{s}; \bm{\Phi})$ for all $s \in \Ss$.
 In Forward/Backward algorithm, evaluation of $p(\bm{x}|{s}; \bm{\Phi})$ is also all what is needed to compute $p(s|\bm{x};\bm{\Phi})$.
 In the following two subsections, we will provide two neural network based generative models that fulfill this requirement and also have high capability for complex signal modeling.
-
-\subsubsection{Generator Mixed HMM (GenHMM)}\label{subsubsec:GenHMM}
-\begin{figure}[!h]
-  \centering
-  \begin{tikzpicture}
-    % \tikzstyle{enode} = [thick, draw=blue, circle, inner sep = 3pt,
-    % align=center]
-    \tikzstyle{enode} = [thick, draw=blue, ellipse, inner sep = 2pt,  align=center]
-    \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]
-
-    \node[enode] (z1) at (-1.2,1.8) {$\bm{z}\sim p_{s,1}(\bm{z})$};
-    \node[nnode] (g1) at (1,1.8) {$\bm{g}_{s,1}$};
-
-    \node[enode] (z2) at (-1.2,0.5){$\bm{z}\sim p_{s,2}(\bm{z})$};
-    \node[nnode] (g2) at (1,0.5) {$\bm{g}_{s,2}$};
-
-    \node[enode] (zK) at (-1.2,-1.8) {$\bm{z}\sim p_{s,K}(\bm{z})$};
-    \node[nnode] (gs) at (1, -1.8) {$\bm{g}_{s,K}$};
-
-    \node[enode] (x) at (4.5,0){$\bm{x}\sim p(\bm{x}| s; \bm{\Phi}_{s})$};
-    % \node at (3,-1){$p(\bm{z}) = \sum_k\pi_kp_k(\bm{z})$};
-
-    \draw[dotted,line width=2pt] (0,-0.3) -- (0,-1.2);
-    \filldraw[->] (1.9, 0.5)circle (2pt) --  node[above=0.2]{${\kappa}\sim \bm{\pi}_{s}$} (x)  ;
-    \draw[->] (z1) -- (g1);
-    \draw[->] (g1) -- (1.8, 1.8);
-
-    \draw[->] (z2) -- (g2);
-    \draw[->] (g2) -- (1.8, 0.5);
-
-    % \draw[->] (1,-0.5) -- (1.8, -0.5);
-    \draw[->] (zK) -- (gs);
-    \draw[->] (gs) -- (1.8, -1.8);
-
-    % \node[enode] (z) at (0,0) {$\bm{z}\sim p(\bm{z})$};
-    % \node[enode] (x) at (5.5,0){$\bm{x}\sim p(\bm{x}; \bm{\Phi})$};
-    %               % \node at (5.2,-1) {$p(\bm{x};\bm{\Phi}) = \textstyle\sum_{k=1}^K \pi_k  p_k(\bm{x})$};
-    % \node[nnode] (g1) at (2.6,1.8) {$\bm{g}_1$};
-    % \node[nnode] (g2) at (2.6,0.5) {$\bm{g}_2$};
-    % \node[nnode] (gk) at (2.6,-1.8) {$\bm{g}_K$};
-    % \draw[dotted,line width=2pt] (2.6,-0.3) -- (2.6,-1.2);
-    % \draw[->] (z) [in= 180, out =0] to (g1);
-    % \draw[->] (z) [in= 180, out =0] to (g2);
-    % \draw[->] (z) [in= 180, out =0] to (gk);
-    % \filldraw[->] (3.7, 0.5)circle (2pt) -- node[above=0.2]{$\bm{s}\sim \bm{\pi}$} (x) ;
-    %               % \draw[->] (3,-0.8) -- (3.5, -0.8);
-    % \draw[->] (g1) -- (3.5,1.8);
-    % \draw[->] (g2) -- (3.5, 0.5);
-    % \draw[->] (gk) -- (3.5, -1.8);
-  \end{tikzpicture}
-  \caption{Source $s$ powered by generator mixed generative model (GenHMM).}
-  \vspace{0.1cm}
-\end{figure}
-In this subsection, we specify the GenHMM model and give the learning algorithm of this model.
-Base on the generative model assumption, for each state $s$, the probabilistic generative model is
-\begin{equation}
-  p(\bm{x}| s; \bm{\Phi}_{s}) = \sum_{\kappa=1}^{K}\pi_{s, \kappa} p(\bm{x}| s, \kappa; \bm{\Phi}_{s, \kappa})
-\end{equation}
-where $\bm{\Phi}_s = \left\{ \pi_{s, \kappa}, \bm{\Phi}_{s, \kappa}| \kappa = 1, 2, \cdots, K. \right\}$, $\pi_{s, \kappa} = p(\kappa | \bm{x}, s; \bm{\Phi}_{s, \kappa})$ and 
-\begin{equation}
-  \sum_{\kappa = 1}^{K} \pi_{s, \kappa}= 1
-\end{equation}
 
@@
                                           %   \end{align}
-In \autoref{eq:obj-q-gen-mix}, $p(s_t| \ubm{x}, \bm{H}^{\mathrm{old}})$ is computed by forward/backward algorithm. The posterior of $\kappa$ is:
+In \eqref{eq:obj-q-gen-mix}, $p(s_t| \ubm{x}, \bm{H}^{\mathrm{old}})$ is computed by forward/backward algorithm. The posterior of $\kappa$ is:
 \begin{align}\label{eq:kappa-posterior}
@@
 
-To solve problem \autoref{opm:pi}, we formulate its Lagrange function as
+To solve problem \eqref{opm:pi}, we formulate its Lagrange function as
 \begin{equation}
@@
 \end{equation}
-Then the solution to \autoref{opm:pi} is
+Then the solution to \eqref{opm:pi} is
 \begin{equation}\label{eq:mix-latent-parameter-solution}
@@
 where $p(s, \kappa | \ubm{x}; \bm{H}^{\mathrm{old}}) = p(s| \ubm{x}; \bm{H}^{\mathrm{old}}) p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$.
-$p(s| \ubm{x}; \bm{H}^{\mathrm{old}})$ that can be computed by results of forward/backward and \autoref{eq:kappa-posterior}.
+$p(s| \ubm{x}; \bm{H}^{\mathrm{old}})$ that can be computed by results of forward/backward and \eqref{eq:kappa-posterior}.
 The posterior for $\kappa$ can be obtained by
@@
 
-    \STATE Compute $p(\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})$  by Forward/backward and \autoref{eq:kappa-posterior}.
+    \STATE Compute $p(\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})$  by Forward/backward and \eqref{eq:kappa-posterior}.
     %     = \frac{\pi_k \Nn\left(f(X); \bm{\mu}_k^{t},\bm{\sigma}_k^{t}\right)}{\sum_{k=1}^K\; \pi_k\Nn\left(f(X); \bm{\mu}_k^{t},\bm{\sigma}_k^{t}\right) }, \forall \left\{ x^{(i)}\right\}_{i=1}^{n_b}$
     % \Comment{Calculate posterior \autoref{eq-posterior-gamma}}
-    \STATE Compute loss ${\Qq}\left({\bm{\Phi}}, {\bm{H}}^{\mathrm{old}}\right)$ in \autoref{eq:obj-q-gen-mix}
+    \STATE Compute loss ${\Qq}\left({\bm{\Phi}}, {\bm{H}}^{\mathrm{old}}\right)$ in \eqref{eq:obj-q-gen-mix}
 
@@
 
-    \STATE Update $\pi_{s, \kappa}, \forall, s\in \Ss, \kappa= 1, 2, \cdots, K$, according to \autoref{eq:mix-latent-parameter-solution}
+    \STATE Update $\pi_{s, \kappa}, \forall, s\in \Ss, \kappa= 1, 2, \cdots, K$, according to \eqref{eq:mix-latent-parameter-solution}
 
@@
 The parameter set for this model to be decide is $\bm{\Phi} = \left\{ \bm{\theta},  \bm{\omega}_s, \forall s\in \Ss\right\}$.
-Then the problem in \autoref{eq:sub-gm} can be reformulated as:
+Then the problem in \eqref{eq:sub-gm} can be reformulated as:
 \begin{align}
@@
 
-\subsubsection{Latent-source Mixed HMM (LatMM)}
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}
-    \tikzstyle{enode} = [thick, draw=blue, ellipse, inner sep = 1pt,  align=center]
-    % \tikzstyle{nnode} = [thick, rectangle, rounded corners = 1pt,minimum size = 1.2cm,draw,inner sep = 2pt]
-    \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]
-    \node[enode] (z1) at (0,1.8) {$\bm{z}\sim p_{s,1}(\bm{z})$};
-    \node[enode] (z2) at (0,0.5){$\bm{z}\sim p_{s,2}(\bm{z})$};
-    \node[enode] (zK) at (0,-1.8) {$\bm{z}\sim p_{s, K}(\bm{z})$}; 
-    \node[enode] (x) at (5.5,0){$\bm{x}\sim p(\bm{x}|s;{\bm{\Phi}}_{s})$};
-    % \node at (3,-1){$p(\bm{z}) = \sum_k\pi_kp_k(\bm{z})$};
-    \node[nnode] (g) at (3.2,0) {$\bm{g}_s$};
-    \draw[dotted,line width=2pt] (0,-0.3) -- (0,-1.2);
-    \filldraw[->] (1.8, 0.5)circle (2pt) -- node[above=0.2]{${\kappa}\sim \bm{\pi}_s$} (g) ;
-    \draw[->] (z1) -- (1.6, 1.8);
-    \draw[->] (z2) -- (1.6, 0.5);
-    % \draw[->] (1,-0.5) -- (1.8, -0.5);
-    \draw[->] (zK) -- (1.6, -1.8);
-    % \draw[->] (z1) [in=180,out=0]  to node[right]{$\pi_1$} (g);
-    % \draw[->] (z2) [in=180,out=0]  to node[above]{$\pi_2$} (g);
-    % \draw[->,dashed] (zK) [in=180,out=0]  to node[right]{$\pi_K$} (g);
-    \draw[->] (g) to (x);
-  \end{tikzpicture}
-  \caption{Source $s$ powered by Latent mixed generative model (LatHMM).}
-  \vspace{0.1cm}
-\end{figure}
-
-LatHMM, (\textbf{benching marking}), the latent source mixed emission model is built by allowing different latent source sharing the same generator $\bm{g}_s$:
-\begin{align}
-  & p(\bm{x}| s, \kappa; \bm{\Phi}_{s}) \nonumber \\
-  =&p_{s, \kappa}(\bm{z}) \bigg| \det \left( \pd{\bm{z}}{\bm{x}} \right) \bigg| \nonumber \\
-  =& \Nn\left( \bm{f}_{{s}}(\bm{x}); \bm{\mu}_{{s},{\kappa}}, \bm{C}_{{s},{\kappa}} \right) \bigg| \det\, \left( \pd{\bm{f}_{{s}}(\bm{x})}{\bm{x}} \right) \bigg|
-\end{align}
-where $\bm{f}_{{s}} = \bm{g^{-1}}_{s}$ is defined by $\bm{\theta}$, and $\bm{\Phi}_s = \left\{ \bm{\theta}_s, \bm{\mu}_{{s},{\kappa}}, \bm{C}_{{s},{\kappa}}| \kappa = 1, 2, \cdots, K \right\}$.
-
-This emission generator sharing makes the posterior computation w.r.t. of $\kappa$ easier:
-
-\begin{equation}\label{eq:kappa-posterior-LatHMM}
-  p(\kappa| s, \ubm{x}; \bm{H}^{\mathrm{old}})
-  = \frac{\pi_{s, \kappa} p(\bm{x}| s, \kappa, \bm{H}^{\mathrm{old}})}{\sum_{\kappa=1}^{K}  \pi_{s, \kappa} p(\bm{x}| s, \kappa,\bm{H}^{\mathrm{old}})}
-  = \frac{\pi_{s, \kappa} \Nn\left( \bm{z}; \bm{\mu}_{s, \kappa}, \bm{C}_{s, \kappa} \right)}{\sum_{\kappa=1}^{K}  \pi_{s, \kappa} \Nn\left( \bm{z}; \bm{\mu}_{s, \kappa}, \bm{C}_{s, \kappa}\right)} \bigg|_{\bm{z} = \bm{f}_s(\bm{x})}.
-  % & = \frac{\pi_{s_t, \kappa_t} \Nn\left( \bm{f}_{{s}_{t},{\kappa}_{t}}(\bm{x}); \bm{\mu}_{{s}_{t},{\kappa}_{t}}, \bm{C}_{{s}_{t},{\kappa}_{t}} \right) \bigg| \det\, \left( \pd{\bm{f}_{{s}_{t},{\kappa}_{t}}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg| }{\sum_{\kappa_t=1}^{K}\pi_{s_t, \kappa_t} \Nn\left( \bm{f}_{{s}_{t},{\kappa}_{t}}(\bm{x}); \bm{\mu}_{{s}_{t},{\kappa}_{t}}, \bm{C}_{{s}_{t},{\kappa}_{t}} \right) \bigg| \det\, \left( \pd{\bm{f}_{{s}_{t},{\kappa}_{t}}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg|} \nonumber \\
-\end{equation}
-Computation of $p(s_t| \ubm{x}, \bm{H}^{\mathrm{old}})$ remains the same, relying on forward/backward algorithm.
-
-In implementation, $\bm{C}_{s, \kappa} = diag\left( \bm{\sigma}_{s, \kappa} \right)$ for simplicity.
-To avoid the singularity problem of Gaussian mixture, we put a Gamma distribution as prior for $\bm{\sigma}_{s, \kappa}$, i.e.
-$\Gamma(\bm{\sigma}_{s, \kappa}^{-1}, a, b)$ where $a$ and $b$ are hyperparameter for Gamma distribution.
-Then the problem can be reformulated as:
-\begin{equation}\label{eq:loss-on-Phi-LatHMM}
-  \uargmax{\bm{\Phi}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) + \frac{1}{K}\log\prod_{k=1}^K\Gamma(\bm{\sigma}_{s, k}^{-1}; a, b)
-\end{equation}
-
-Apart from the emission prboability model difference, rest computation is the same as \autoref{subsubsec:GenHMM}. We summarize the algorithm as:
-
-\begin{algorithm}[H]
-  \caption{M-step w.r.t. $\bm{\Phi}$ powered by LatMM}\label{algo:opt-Phi-LatHMM}
-  \begin{algorithmic}[1]
-    \STATE {\bfseries Input:}
-    Latent mixture distribution: $\sum_{k=1}^{K}\pi_{s, \kappa} \Nn\left(\bm{z}; \bm{\mu}_{s, \kappa}, \mathrm{diag}(\bm{\sigma}_{s, \kappa}^2)\right)$\\
-    Empirical distribution $\hat{p}(\bm{x})$ of dataset; \\
-    \STATE Set a total number of epochs $T$ of training as stop criterion. A learning rate $\eta$. Set hyperparameter $a$ , $b$ for prior of $\bm{\sigma}_{s, \kappa}^{-1}, \forall k$.
-    \FOR {epoch $t < T$}
-    \STATE Sample a batch of data $\left\{ \ubmr{x}{r} \right\}_{r=1}^{R_b}$ from dataset $P_d(\ubm{x})$
-    \STATE Compute $p(\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})$  by Forward/backward and \autoref{eq:kappa-posterior-LatHMM}.
-%= \frac{\pi_k \Nn\left(f(X); \bm{\mu}_k^{t},\bm{\sigma}_k^{t}\right)}{\sum_{k=1}^K\; \pi_k\Nn\left(f(X); \bm{\mu}_k^{t},\bm{\sigma}_k^{t}\right) }, \forall \left\{ x^{(i)}\right\}_{i=1}^{n_b}$
-    % \Comment{Calculate posterior \autoref{eq-posterior-gamma}}
-    \STATE Compute loss in \autoref{eq:loss-on-Phi-LatHMM}%\Comment{Expectation}
-
-    \STATE $\partial{\bm{\th}_s}, \partial{\bm{\mu}_{s, \kappa}}, \partial{\bm{\sigma}_{s, \kappa}}\gets 
-    \nabla_{\bm{\theta}, \bm{\mu}_k, \bm{\sigma}_k} -
-    \frac{1}{R_{b}}{\Qq}\left({\bm{\Phi}},\ubar{\bm{H}}^{\mathrm{old}}\right)
-    -\frac{1}{K}\sum_{k=1}^K \log \, \Gamma(\bm{\sigma}_{s, \kappa}^{-1}; a, b)$ %\Comment{E-step}
-    % \STATE $\partial{g_k} \gets \nabla_{\th_k}\widetilde{\Qq}\left(\left\{ \th_k
-    %   \right\}_{k=1}^{K};\left\{ \th_k^t\right\}_{k=1}^{K}\right)$,
-    % $\forall \th_k \in\left\{ \th_k\right\}_{k=1}^{K}$
-    \STATE $\bm{\th}_s \gets \bm{\th}_s - \eta \cdot \partial{\bm{\theta}_s}$, $\forall s \in \Ss$
-    \STATE $\bm{\mu}_{s, \kappa} \gets \bm{\mu}_{s, \kappa} - \eta \cdot \partial{\bm{\mu}_{s, \kappa}}, \forall \kappa, s$
-    \STATE $\bm{\sigma}_{s, \kappa} \gets \bm{\sigma}_{s, \kappa} - \eta \cdot \partial{\bm{\sigma}_{s, \kappa}}, \forall \kappa, s$
-    \ENDFOR
-
-    \STATE Update $\pi_k$ according to \autoref{eq:mix-latent-parameter-solution}
-
-    \STATE Assemble $\bm{\Phi} = \left\{  \bm{\Phi}_s| s \in \Ss\right\}$ for $\bm{\Phi}_s = \left\{ \bm{\theta}_s, \bm{\mu}_{{s},{\kappa}}, \bm{C}_{{s},{\kappa}}| \kappa = 1, 2, \cdots, K \right\}$.
-
-  \end{algorithmic}
-\end{algorithm}
-
-    %     This mixture emission probability model introduces anther latent variable $\kappa$, indicating the actual component of the mixture probability model that corresponding to the signal $\bm{x}$.
-
-    % %     For this proposal, we seek to use a generator mixed HMM scheme, termed as GenM-HMM. We define a set of generators for GenM-HMM:
-    % %     \begin{equation}
-    % %       \{\bm{g}_{s}| s \in \Ss, \bm{g}_s: \bm{z}\rightarrow \bm{x}, \bm{z}\sim p_{s}(\bm{z})\}.
-    % %     \end{equation}
-    %     Thus there are total $|\Ss|$ generators.
-$p(\bm{x}|{s}; \bm{\Phi})$ is induced as $\bm{g}_s(\bm{z}) \sim p(\bm{x}|{s}; \bm{\Phi})$ where $\bm{z} \sim p_s(\bm{z})$ for $s \in \Ss$.
-We have the $s$-th component of the GenM-HMM model as
-    %     \begin{align}
-    %     p(\bm{x} | s; \bm{\Phi})
-    %                                    &= p_s(\bm{z}) \bigg| \det \left( \pd{\bm{z}}{\bm{x}} \right) \bigg| \nonumber \\
-    %                                    &= p_s(\bm{f}_s(\bm{x})) \bigg| \det \left( \pd{\bm{f}_s(\bm{x})}{\bm{x}} \right) \bigg|
-                                           %   \end{align}
-                                           %                                            where $p_s(\bm{z})$ is the latent source distribution for $s=1, 2, \cdots, |\Ss|$.
-
-                                           %                                            Let us denote the parameter set that defines latent distribution $p_s(z)$ by $\bm{\omega}_s$ and the parameter set that defines generator $\bm{g}_s$ by $\bm{\theta}_s$.
-Then $\bm{\Phi} = \left\{ \bm{\theta}_s,
-                                           %                                            \bm{\omega}_s, \forall s\in \Ss\right\}$. The problem in \autoref{eq:sub-gm} can be reformulated as:
+\section{Convergence Analysis}
 
 \section{On Implementation of acoustic signal}
-Found a HMM python lib that basics provide needed API for us, see \href{https://hmmlearn.readthedocs.io/en/stable/index.html}{hmmlearn}. Saikat also has suggestion.
+test citing \cite{2018arXiv180703039K}
 
-For problem \autoref{eq:sub-gm} we are going to use our generative models to solve. I have the following consideration to revised our LatMM and GenMM for this application:
+For problem \eqref{eq:sub-gm} we are going to use our generative models to solve. I have the following consideration to revised our LatMM and GenMM for this application:
 \begin{itemize}
@@
 
-\section{On Implementation of Planning}
-Refer to \cite{kurutach2018learning_plan} and its \href{https://github.com/thanard/causal-infogan}{experiments}.
-
-\bibliographystyle{plain}
+\bibliography{myref}
+\bibliographystyle{aaai}
 
      %      \bibliography{bibliography}
-\bibliography{myref}
 

### 2019-08-12 manuscript/main.tex
structure orginized

@@
 \newcommand{\ubm}[1]{\ubar{\bm{#1}}}
-\newcommand{\ubmr}[2]{\ubar{\bm{#1}}^{(#2)}}
+\newcommand{\ubmr}[2]{\ubar{\bm{#1}}^{#2}}
 
@@
 
-\newcommand{\bmtr}[3]{\bm{#1}^{(#3)}_{#2}}
-\newcommand{\smtr}[3]{{#1}^{(#3)}_{#2}}
+\newcommand{\bmtr}[3]{\bm{#1}^{#3}_{#2}}
+\newcommand{\smtr}[3]{{#1}^{#3}_{#2}}
 
@@
 
-\section{Preliminaries}
+\section{Generator-mixed HMM}
+
 \begin{figure}[!h]
@@
 This process gives us the probabilistic model $p(\ubm{x};\bm{H})$
-\section{Generator-mixed HMM}
+\subsection{Emission Model of GenHMM}
 
@@
   \end{tikzpicture}
-  \caption{Source $s$ powered by generator mixed generative model (GenHMM).}
+  \caption{Source of state $s$ in GenHMM.}
+  \label{fig:gen-mix}
   \vspace{0.1cm}
@@
 GenHMM has a hidden markov process as typical HMM, while the probabilistic model of GenHMM for each hidden state is induced by mixture of neural network based generators.
-For $s\in \Ss$, we define the state's probabilistic model in GenHMM as
+For given state $s\in \Ss$, we define its probabilistic model in GenHMM as
 \begin{equation}
-  p(\bm{x}| s; \bm{\Phi}_{s}) = \sum_{\kappa=1}^{K}\pi_{s, \kappa} p(\bm{x}| s, \kappa; \bm{\Phi}_{s, \kappa})
+  p(\bm{x}| s; \bm{\Phi}_{s}) = \sum_{\kappa=1}^{K}\pi_{s, \kappa} p(\bm{x}| s, \kappa; \bm{\theta}_{s, \kappa}),
 \end{equation}
-where $\kappa$ is a random variable following a categorical distribution, and $\pi_{s, \kappa} = p(\kappa | \bm{x}, s; \bm{\Phi}_{s, \kappa})$.
+where $\kappa$ is a random variable following a categorical distribution, and $\pi_{s, \kappa} = p(\kappa | s; \bm{\Phi}_{s, \kappa})$.
 Naturally $\sum_{\kappa = 1}^{K} \pi_{s, \kappa}= 1$.
 Denote $\bm{\pi}_{s} = [\pi_{s,1}, \pi_{s,2}, \cdots, \pi_{s,K}]$,
-$\bm{\Phi}_s = \left\{ \pi_{s, \kappa}, \bm{\Phi}_{s, \kappa}| \kappa = 1, 2, \cdots, K \right\}$,
+$\bm{\theta}_s = \left\{ \bm{\theta}_{s, \kappa}| \kappa = 1, 2, \cdots, K \right\}$. 
+We define $p(\bm{x}| s, \kappa; \bm{\theta}_{s, \kappa})$ as induced distribution by a generator $\bm{g}_{s,\kappa}: \RR^{N}\rightarrow\RR^{N}$, such that $\bm{x}=\bm{g}_{s, \kappa}(\bm{z})$, where $\bm{z}\sim p_{s,\kappa}(\bm{z})$.
+By change of variable, we have
+\begin{equation}\label{eq:changel-variable}
+  p(\bm{x}| s, \kappa; \bm{\theta}_{s, \kappa}) = p_{s,\kappa}(\bm{z})\bigg| \det\left( \pd{\bm{g}_{s,\kappa}(\bm{z})}{\bm{z}} \right)\bigg|^{-1}.
+\end{equation}
 
-\subsubsection{Generator Mixed HMM (GenHMM)}\label{subsubsec:GenHMM}
-In this subsection, we specify the GenHMM model and give the learning algorithm of this model.
-Base on the generative model assumption, for each state $s$, the probabilistic generative model is
+For a state $s$ of GenHMM, the induced probability distribution is illustrated in Figure~\ref{fig:gen-mix}.
 
-\subsubsection{Flow structure}
+\subsection{Problem Statement}
 
-\section{Learning Algorithm of GenHMM}
-\subsection{Problem Statement}
-Given a empirical distribution $\hat{p}(\ubm{x}) = \frac{1}{R}\sum_{r=1}^{R} \delta_{\ubmr{x}{r}}(\ubm{x})$. We want to find a probabilistic model such that:
+For a given dataset, we denote its empirical distribution by $\hat{p}(\ubm{x}) = \frac{1}{R}\sum_{r=1}^{R} \delta_{\ubmr{x}{r}}(\ubm{x})$, where superscipt $(\cdot)^{r}$ denotes the index of sequential signal.
+The natural question is how to use GenHMM to represent the dataset.
+Alternative, we are looking for the answer to question:
 \begin{equation}
-  \min KL(\hat{p}(\ubm{x})\| p(\ubm{x}))
+  \umin{\bm{H}\in \Hh} KL(\hat{p}(\ubm{x})\| p(\ubm{x};\bm{H}))
 \end{equation}
-where $KL(\cdot\|\cdot)$ denotes the Kullback-Leibler divergence.
+where $KL(\cdot\|\cdot)$ denotes the Kullback-Leibler divergence. The KL divergence problem can be boiled down to a maximizing a loglikelihood problem:
+\begin{equation}\label{eq:ml-of-hmm}
+  \uargmax{\bm{H} \in \Hh} \sum_{r=1}^{R}\log\,p(\ubmr{x}{r}; \bm{H}).
+\end{equation}
 
-When we use HMM to model the empirical distribution and approach the unknown true distribution, the problem boils down to:
-\begin{equation}
-  \uargmax{\bm{H} \in \Hh} p(\ubm{X}; \bm{H})
-\end{equation}
-where $\ubm{X} = \left[ \ubmr{x}{1}, \ubmr{x}{2}, \cdots, \ubmr{x}{R} \right]$
+With such a model GenHMM at hand, the remaining problems are:
+\begin{itemize}
+\item How to realize GenHMM by neural networks to have high modeling capacity?
+\item How to train GenHMM to sovle problem in \eqref{eq:ml-of-hmm} using practical algorithm?
+\item Would the training of GenHMM converges?
+\end{itemize}
+We would answer these questions in the following section.
 
-The problem can be reformulated as
-\begin{equation}\label{eq:ml-of-hmm}
-  \uargmax{\bm{H} \in \Hh} \sum_{r=1}^{R}\log\,p(\ubmr{x}{r}; \bm{H})
-\end{equation}
-for independent identical distributed assumption of $\ubm{x}$.
+\section{Practical Solution for GenHMM}
 
-\subsection{Proposal}
+In this section, we provide answers to the questions at the end of last section.
 
-Since model $\bm{H}$ contains hidden sequential variable $\ubm{s}$, we can not directly solve the maximum likelihood problem in \ref{eq:ml-of-hmm}.
+Since model $\bm{H}$ contains hidden sequential variable $\ubm{s}$ and $\ubm{\kappa}$, we can not directly solve the maximum likelihood problem in \ref{eq:ml-of-hmm}.
 We use expectation maximization (EM) to address the hidden variable problem by
@@
   \begin{equation}\label{eq:em-q-funciton}
-    \Qq(\bm{H}; \bm{H}^{\mathrm{old}}) = \ \sum_{r=1}^{R}E_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}\left[\log\,p(\ubmr{x}{r}, \ubmr{s}{r}; \bm{H}) \right]
+    \Qq(\bm{H}; \bm{H}^{\mathrm{old}}) = \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H})\right],
   \end{equation}
+  where $\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \cdot\right]$ denotes the expectation operator by distribution $\hat{p}(\ubm{x})$ and $p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})$.
 \item M-step: the optimization step:
@@
 The \eqref{eq:em-m-opt} can be reformulated as:
-\begin{equation}\label{eq:m-step-subs}
-  \umax{\bm{H}} \Qq(\bm{H}; \bm{H}^{\mathrm{old}}) = \umax{\bm{q}}\Qq(\bm{q}; \bm{H}^{\mathrm{old}}) + \umax{{A}}\Qq({A}; \bm{H}^{\mathrm{old}}) + \umax{\bm{\Phi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})
-\end{equation}
+\begin{align}\label{eq:m-step-subs}
+  &\umax{\bm{H}} \Qq(\bm{H}; \bm{H}^{\mathrm{old}}) \nonumber \\
+  =&\umax{\bm{q}}\Qq(\bm{q}; \bm{H}^{\mathrm{old}}) + \umax{{A}}\Qq({A}; \bm{H}^{\mathrm{old}}) 
+  + \umax{\bm{\Phi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})
+\end{align}
 where
 \begin{align}
-  \Qq(\bm{q}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,p(\smtr{s}{1}{r}; \bm{q})\right] \label{eq:init-distribution-update}\\
-  \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})\right] \label{eq:transition-update}\\
-  \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,p(\ubmr{x}{r} | \ubmr{s}{r}; \bm{\Phi})\right]\label{eq:generative-model-update}
+  \Qq(\bm{q}; \bm{H}^{\mathrm{old}}) 
+    &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p({s}_{1})  \right] \nonumber\\
+    &= \EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p({s}_{1})  \right]\label{eq:init-distribution-update}\\
+  \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \sum_{t}\log\,p({s}_{t+1}|{s}_{t}; \bm{H}) \right] \label{eq:transition-update}\\
+  \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) &= \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p(\ubm{x}, \ubm{\kappa}| \ubm{s}; \bm{H}) \right]\label{eq:generative-model-update}
 \end{align}
@@
 The marginal $p(s_t| \ubm{x}; \bm{H})$ is also efficiently computed as the joint posterior.
+
+\subsection{Modeling $\bm{g}_{s,\kappa}$ by a flow model }
+We use feed-forward neural network to implement every generator $\bm{g}_{s,\kappa}$.
+For notation simplicity, denote $\bm{g}_{s,\kappa}=\bm{g}_{s,\kappa}$.
+Define  a feed-forward neural network $\bm{g}_{s,\kappa}$ that has multiple hidden layers
+$\bm{g}_{s,\kappa}=\bm{g}_{s,\kappa}^{[L]}\circ \bm{g}_{s,\kappa}^{[L-1]}\circ \cdots
+\circ \bm{g}_{s,\kappa}^{[1]}$ and is invertible $\bm{f}_{s,\kappa}=\bm{g}_{s,\kappa}^{-1}$, where $\circ$ denotes mapping concatenation. Then the signal flow can be depicted as
+\begin{equation*}
+  \vspace{-8pt}
+  \centering
+  \begin{tikzpicture}
+    \node (z) at (0,0) {};
+    \node at ($(z)-(0.5,0)$){$\bm{z}=\bm{h}_0$};
+    \node (xi1) at (1.5,0) {$\bm{h}_1$};
+    \node (xi2) at (3,0) {};
+    \node (xi3) at (4.5,0){};
+    \node (x) at (6,0) {};
+    \node at ($(x)+(0.5,0)$){${\bm{x}} = \bm{h}_L$};
+    \draw[->] ($(z) + (0.3,0.1)$) -- node[above]{$\bm{g}_{s,\kappa}^{[1]}$} ($(xi1)+(-0.3,0.1)$); 
+    \draw[->] ($(xi1)-(0.3,0.1)$) -- node[below]{${\bm{f}}_{s,\kappa}^{[1]}$}($(z) - (-0.3,0.1)$);
+    \draw[->] ($(xi1) + (0.3,0.1)$) -- node[above]{$\bm{g}_{s,\kappa}^{[2]}$} ($(xi2)+(-0.3,0.1)$); 
+    \draw[->] ($(xi2)-(0.3,0.1)$) -- node[below]{${\bm{f}}_{s,\kappa}^{[2]}$}($(xi1) - (-0.3,0.1)$);
+    \draw[->] ($(xi3) + (0.3,0.1)$) -- node[above]{$\bm{g}_{s,\kappa}^{[L]}$} ($(x)+(-0.3,0.1)$); 
+    \draw[->] ($(x)-(0.3,0.1)$) -- node[below]{${\bm{f}}_{s,\kappa}^{[L]}$}($(xi3) - (-0.3,0.1)$);
+    \draw[dotted,line width = 0.3 mm] (xi2) -- (xi3);
+  \end{tikzpicture},
+\end{equation*}
+where $\bm{g}_{s,\kappa}^{[l]}$ and $\bm{f}_{s,\kappa}^{[l]}$ are the $l$-th layer of $\bm{g}_{s,\kappa}$ and $\bm{f}_{s,\kappa}$, respectively.
+In a feed-forward neural network, if every layer is invertible,
+the full feed-forward neural network is invertible. The inverse
+function is given by $\bm{z}=\bm{f}_{s,\kappa}(\bm{x})$. Flow-based
+network, proposed in \cite{DBLP:journals/corr/DinhKB14}, is such a
+feed-forward neural network, which is further improved in subsequent works \cite{2016arXiv160508803D, 2018arXiv180703039K}.
+It also has additional advantages as efficient Jacobian computation and low computational complexity.
+%Computation of inverse mapping for flow-based network and determinant of Jacobin were addressed in the literature.
+
+For a flow-based neural network architecture, let us assume that
+the feature $\bm{h}_l$ at the $l$'th layer has two subparts as
+$\bm{h}_l = [\bm{h}_{l,a}^{T} \, , \, \bm{h}_{l,b}^{T}]^{T}$ where
+$(\cdot)^{T}$ denotes transpose operation. Then considering $\bm{h}_0 = \bm{z}$, we have the following forward and inverse relations between $(l-1)$'th and $l$'th layers:
+\begin{equation}\label{eq-gl}
+  \begin{array}{l}
+    \bm{h}_{l-1} =
+    \begin{bmatrix}
+      \bm{h}_{l-1,a}\\
+      \bm{h}_{l-1,b}
+    \end{bmatrix}
+    =
+    \begin{bmatrix}
+      \bm{h}_{l,a}\\
+      \bm{m}_a(\bm{h}_{l,a})\odot \bm{h}_{l,b} + \bm{m}_b(\bm{h}_{l,a})
+    \end{bmatrix},\vspace{10pt}\\
+    \bm{h}_{l} =
+    \begin{bmatrix}
+      \bm{h}_{l,a}\\
+      \bm{h}_{l,b}
+    \end{bmatrix}
+    =
+    \begin{bmatrix}
+      \bm{h}_{l-1,a}\\
+      \left(  \bm{h}_{l-1,b} - \bm{m}_b(\bm{h}_{l-1,a}) \right)\oslash \bm{m}_a(\bm{h}_{l-1,a}) 
+    \end{bmatrix}, \\
+  \end{array}  
+\end{equation}
+where $\odot$ denotes element-wise product, $\oslash$ denotes
+element-wise division, and $\bm{m}_a(\cdot), \bm{m}_b(\cdot)$ can be
+complex non-linear mappings (implemented by neural networks).
+For the flow-based neural network, the determinant of Jacobian matrix is
+\begin{equation}\label{eq:cat-jacobian}
+  \begin{array}{rl}
+    \mathrm{det}(\nabla{\bm{f}_{s,\kappa}}) & = \prod_{l=1}^L \det (\nabla{\bm{f}_{s,\kappa}^{[l]}}),
+  \end{array}
+\end{equation}
+where $\nabla{f_{s,\kappa}^{[l]}}$ is the Jacobian of the transformation from the $l$-th layer to the $(l-1)$-th layer, i.e., the inverse transformation.
+We compute the determinate of the Jacobian matrix as
+\begin{align}\label{eq-hl-determinate}
+  \det (\nabla{f_{s,\kappa}^{[l]}})& = \det \left[  \pd{\bm{h}_{l-1}}{\bm{h}_l} \right] \nonumber\\
+                               & = \det
+                                 \begin{bmatrix}
+                                   \bm{I}_a & \mathbf{0} \nonumber\\
+                                   \pd{\bm{h}_{l-1,b}}{\bm{h}_{l,a}} & \mathrm{diag}(\bm{m}_a(\bm{h}_{l,a}))
+                                 \end{bmatrix}\nonumber\\
+                               &= \det \left( \mathrm{diag}(\bm{m}_a(\bm{h}_{l,a})) \right),
+\end{align}
+where $\bm{I}_a$ is identity matrix and $\mathrm{diag}(\cdot)$ returns a square matrix with the elements of $(\cdot)$ on the main diagnal.
+\textcolor{red}{do I still need this?}
+Then the pdf is
+\begin{align}
+  p(\bm{x}) & =  p(\bm{z}) \big| \mathrm{det}(\bm{J}) |_{\bm{z}={\bm{f}}({\bm{x}})}\big| \nonumber\\
+            &  = p(\bm{z}) \prod_{l=1}^L \abs{\det\left( \mathrm{diag}(\bm{m}_a(\bm{h}_{l,a}))]  \right)}.
+\end{align}
+[] describes a \textit{coupling} mapping between layers.
+Since the coupling has a partial identity mapping, direct concatenation of multiple such coupling mappings would result in a partial identity mapping of the whole neural network $\tilde{\bm{g}}$.
+Alternating the positions of identity mapping \cite{2016arXiv160508803D} or using $1\times1$ convolution operations \cite{2018arXiv180703039K} before each coupling mapping is used to treat the issue.
+              %               \textcolor{blue}{Some reading problem in this sentence: Since the coupling has a partial identity mapping, alternating the positions of identity mapping \cite{2016arXiv160508803D} or using $1\times1$ convolution operations \cite{2018arXiv180703039K} when
+              %               concatenating multiple layers together were proposed to avoid the situation that a part of the output is equal to a part of the input.} %In addition, applying affine mapping does not affect its advantage of trivial inverse and determinate computation.
+Furthermore, \cite{2016arXiv160508803D}\cite{2018arXiv180703039K} split some hidden
+layer signal $\bm{h}$ and model a part of it directly as standard Gaussian to reduce computation and memory burden.
+% % \textcolor{blue}{Therefore the dimension
+              %               of latent signal $\bm{z}$ to model can be smaller than $N$.}
+
+\subsection{Learning of GenHMM}
+
+\subsubsection{Generative Model Update}
+We use notation $\bm{\Pi} = \left\{  \bm{\pi}_{s}| s\in \Ss \right\}$, $\bm{\Theta}=\left\{ \bm{\theta}_s| s\in \Ss \right\}$.
+Then the third subproblem, \eqref{eq:m-step-subs}, becomes:
+\begin{align}\label{eq:sub-gm}
+  &\umax{\bm{\Phi}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) = \umax{\bm{\Pi}} \Qq(\bm{\Pi}; \bm{H}^{\mathrm{old}}) + \umax{\bm{\Theta}} \Qq(\bm{\Theta}; \bm{H}^{\mathrm{old}}),
+\end{align}
+where
+\begin{align}
+  &\Qq(\bm{\Pi}; \bm{H}^{\mathrm{old}})  =\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[  \log\,p(\ubm{\kappa}| \ubm{s}; \bm{H})\right] \nonumber \\
+  &\Qq(\bm{\Theta}; \bm{H}^{\mathrm{old}}) =\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[  \log\,p(\ubm{x}| \ubm{s},\ubm{\kappa}; \bm{H})\right] 
+\end{align}
+
+The neural network loss can be thus formulated as
+\begin{align}\label{eq:obj-q-gen-mix}
+  &\Qq(\bm{\Theta}; \bm{H}^{\mathrm{old}}) \nonumber \\
+  = &\frac{1}{R}\sum_{r=1}^{R}\sum_{\ubmr{s}{r}}\sum_{\ubmr{\kappa}{r}}{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \sum_{t=1}^{{T}^{r}}\log\,p(\bm{x}^{(r)} | \smtr{s}{t}{r}, \smtr{\kappa}{t}{r}; \bm{H}) \nonumber \\
+  =& \frac{1}{R}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})p(\smtr{\kappa}{t}{r}|\smtr{s}{t}{r}, \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \nonumber\\
+&  \log\, p(\bmtr{x}{t}{r} | \smtr{s}{t}{r}, \smtr{\kappa}{t}{r}; \bm{H}) 
+\end{align}
+
+Here, $p(s_t| \ubm{x}, \bm{H}^{\mathrm{old}})$ is computed by forward/backward algorithm. The posterior of $\kappa$ is:
+\begin{align}\label{eq:kappa-posterior}
+  p(\kappa| s, \ubm{x}; \bm{H}^{\mathrm{old}})
+  &=  \frac{p(\kappa, \ubm{x}| s; \bm{H}^{\mathrm{old}})}{p(\ubm{x}| s,\bm{H}^{\mathrm{old}})} \nonumber \\
+  & = \frac{\pi_{s, \kappa}^{\mathrm{old}} p(\bm{x}| s, \kappa, \bm{H}^{\mathrm{old}})}{\sum_{\kappa=1}^{K}  \pi_{s, \kappa}^{\mathrm{old}} p(\bm{x}| s, \kappa,\bm{H}^{\mathrm{old}})} 
+\end{align}
+where the last equation is due to the fact that only $\bm{x}_t$ among sequence $\ubm{x}$ depends on $s_t, \kappa_t$.
+
+\textcolor{red}{expand to neural network here}By substituting \eqref{eq:changel-variable} and \eqref{eq:cat-jacobian} into \eqref{eq:obj-q-gen-mix}, we have loss function for neural networks as
+\begin{align}\label{eq:obj-q-gen-mix}
+  &\Qq(\bm{\Theta}; \bm{H}^{\mathrm{old}}) \nonumber \\
+  =& \frac{1}{R}\hspace{-3pt}\sum_{r=1}^{R}\hspace{-3pt} \sum_{t=1}^{{T}^{r}-1}\hspace{-3pt} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|} \hspace{-3pt} \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})p(\smtr{\kappa}{t}{r}|\smtr{s}{t}{r}, \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \nonumber\\
+&\left[ \log\, p_{\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}}(\bm{f}_{\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}}(\bmtr{x}{t}{r})) + \sum_{l=1}^{L}\log\,| \det (\nabla{\bm{f}_{s,\kappa}^{[l]}})|\right]
+\end{align}
+
+In implementation, the generators of GenHMM uses standard Gaussian distribution for variables $\bm{z} = \bm{f}_{s,\kappa}(\bm{x})$, i.e. $p_{s,\kappa}(\bm{z})$ used standard Gaussian density function.
+
+------------------------------
+
+The latent prior for mixture of each source $s$ is obtained by solving the following problem:
+\begin{align}\label{opm:pi}
+  \pi_{s, \kappa} & = \uargmax{\pi_{s, \kappa}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) \\ \nonumber
+                  & s.t. \, \sum_{\kappa=1}^{K} \pi_{s, \kappa}= 1, \forall s = 1, 2, \cdots, |\Ss|. 
+\end{align}
+
+To solve problem \eqref{opm:pi}, we formulate its Lagrange function as
+\begin{equation}
+  \Ll = \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) + \sum_{\lambda_s} \lambda_s\left( 1-  \sum_{\kappa=1}^{K}\pi_{s, \kappa}  \right).
+\end{equation}
+Solving $\pd{\Ll}{\pi_{s, \kappa}} = 0$ gives
+\begin{equation}
+  \pi_{s,\kappa} = \frac{1}{\lambda_s}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}} p(\smtr{s}{t}{r}=s, \smtr{\kappa}{t}{r} =\kappa| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})
+\end{equation}
+With condition $\sum_{\kappa=1}^{K} \pi_{s, \kappa}= 1, \forall s = 1, 2, \cdots, |\Ss|$, we have
+\begin{equation}
+  \lambda_s = \sum_{\kappa=1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}} p(\smtr{s}{t}{r}=s, \smtr{\kappa}{t}{r} =\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}})
+\end{equation}
+Then the solution to \eqref{opm:pi} is
+\begin{equation}\label{eq:mix-latent-parameter-solution}
+  \pi_{s, \kappa} = \frac{\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }{\sum_{k =1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}-1} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=k | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }
+\end{equation}
+where $p(s, \kappa | \ubm{x}; \bm{H}^{\mathrm{old}}) = p(s| \ubm{x}; \bm{H}^{\mathrm{old}}) p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$.
+$p(s| \ubm{x}; \bm{H}^{\mathrm{old}})$ that can be computed by results of forward/backward in Viterbi algorithm, while $p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$.
+$p(s| \ubm{x}; \bm{H}^{\mathrm{old}})$ is given by \eqref{eq:kappa-posterior}.
+
+-----------------------
+
+\subsubsection{Initial Probability Update}
+\eqref{eq:init-distribution-update} can be written as:
+\begin{align}
+  &\Qq(\bm{q}; \bm{H}^{\mathrm{old}}) \nonumber\\
+  &=\frac{1}{R} \sum_{r=1}^{R}\sum_{\ubmr{s}{r}} {p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r}) \nonumber \\
+% & = \frac{1}{R}\sum_{r=1}^{R}\sum_{\smtr{s}{1}{r}=1}^{|\Ss|}\sum_{\smtr{s}{2}{r}=1}^{|\Ss|}\cdots \sum_{\smtr{s}{T^{r}}{r}}^{{|\Ss|}} {p(\smtr{s}{1}{r}, \smtr{s}{2}{r}, \cdots, \smtr{s}{T^{r}}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r}) \\
+& = \frac{1}{R}\sum_{r=1}^{R}\sum_{\smtr{s}{1}{r}=1}^{|\Ss|}{p(\smtr{s}{1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r}) 
+\end{align}
+
+Since $p(\smtr{s}{1}{r})$ is the probability of initial state of GenHMM $\bm{H}$ for $r$-th sequential sample, actually $q_i = p({s}_{1} =i) $, $i= 1, 2, \cdots, |\Ss|$ for $\bm{H}$.
+Solution to problem:
+\begin{align}
+  \bm{q} &= \uargmax{\bm{q}} \Qq(\bm{q}; \bm{H}^{\mathrm{old}}), \nonumber \\
+                        &\mathrm{s.t.} \sum_{i=1}^{ |\Ss| }q_i = 1, q_i \geq 0, \forall s.
+\end{align}
+is
+\begin{equation}\label{eq:update-initial-state-prob}
+  q_i = \frac{1}{R} \sum_{r=1}^{R} p(\smtr{s}{1}{r}=i | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}), \forall\; i = 1, 2, \cdots, |\Ss|.
+\end{equation}
+
+\subsubsection{Transition Probability Update}
+\eqref{eq:transition-update} can be written as
+\begin{align}
+  &\Qq(\bm{A}; \bm{H}^{\mathrm{old}})\nonumber \\
+  % &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})\right] \nonumber\\
+  &= \sum_{r=1}^{R} \sum_{\ubmr{s}{r}}{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \sum_{t=1}^{T^{(r)}-1}\log\,p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A}) \nonumber \\
+  &= \sum_{r=1}^{R} \sum_{t=1}^{T^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}\sum_{\smtr{s}{t+1}{r}=1}^{|\Ss|}{p(\smtr{s}{t}{r}, \smtr{s}{t+1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})
+\end{align}
+
+Since $\bm{A}_{i, j}  = p(\smtr{s}{t+1}{r}=j|\smtr{s}{t}{r}=i; {A})$ where $A_{i, j}$ is the element of transition matrix $A$, the solution to problem:
+\begin{align}\label{eq:update-transition-prob}
+  \bm{A} = &\uargmax{\bm{A}} \Qq(\bm{A}; \bm{H}^{\mathrm{old}}), \nonumber \\
+  \mathrm{s.t.} &\hspace{0.2cm} \bm{A} \cdot \bm{1} = \bm{1} \nonumber \\
+                          &  \bm{A}^{\intercal} \cdot \bm{1} = \bm{1} \nonumber \\
+                          & \bm{A}_{i,j} \geq 0.
+\end{align}
+is
+\begin{equation}
+  \bm{A}_{i,j} = \frac{\bar{\xi}_{i,j}}{\sum_{k = 1}^{|\Ss|} \bar{\xi}_{i,k}},
+\end{equation}
+where
+\begin{equation}
+  \bar{\xi}_{i,j} = \sum_{r= 1}^{R} \sum_{t= 1}^{T^{(r)}-1}{p(\smtr{s}{t}{r}=i, \smtr{s}{t+1}{r}=j| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}
+\end{equation}
+
+\subsection{Algorithm of GenHMM}
+Algorithm here:
+things to talk:
+\begin{itemize}
+\item separate optimization of different sets of parameters
+\item batch optimization optimization
+\item not going to optimal for $\bm{\Phi}$, leaving to convergence analysis in next section
+\end{itemize}
 
@@
 
-\subsection{Initial Probability Update}
-\eqref{eq:init-distribution-update} can be written as:
-\begin{align}
-  \Qq(\bm{q}; \bm{H}^{\mathrm{old}}) &= \sum_{r=1}^{R}\sum_{\ubmr{s}{r}} {p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r}; \bm{q}) \nonumber \\
-                                     & = \sum_{r=1}^{R}\sum_{\smtr{s}{1}{r}=1}^{|\Ss|}\sum_{\smtr{s}{2}{r}=1}^{|\Ss|}\cdots \sum_{\smtr{s}{T^{r}}{r}}^{{|\Ss|}} {p(\smtr{s}{1}{r}, \smtr{s}{2}{r}, \cdots, \smtr{s}{T^{r}}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r}; \bm{q}) \\
-                                     & = \sum_{r=1}^{R}\sum_{\smtr{s}{1}{r}=1}^{|\Ss|}{p(\smtr{s}{1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r}; \bm{q}) 
-\end{align}
-
-Since $p(\smtr{s}{1}{r}; \bm{H})$ is the probability of initial state of HMM $\bm{H}$ for $r$-th sequential, actually $q_i = p(\smtr{s}{1}{r} =i;\bm{H} ) $ for $i= 1, 2, \cdots, |\Ss|$.
-Solution to problem:
-\begin{align}
-  \bm{q}^{\mathrm{new}} &= \uargmax{\bm{q}} \Qq(\bm{q}; \bm{H}^{\mathrm{old}}), \nonumber \\
-                        &\mathrm{s.t.} \sum_{i=1}^{ |\Ss| }q_i = 1 \nonumber\\
-                        &q_i \geq 0, \forall s.
-\end{align}
-is
-\begin{equation}\label{eq:update-initial-state-prob}
-  q_i = \frac{1}{R} \sum_{r=1}^{R} p(\smtr{s}{1}{r}=i | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}), \forall\; i = 1, 2, \cdots, |\Ss|.
-\end{equation}
-
-\subsection{Transition Probability Update}
-\eqref{eq:transition-update} can be written as
-\begin{align}
-  \Qq(\bm{A}; \bm{H}^{\mathrm{old}})
-  &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})\right] \nonumber\\
-  &= \sum_{r=1}^{R} \sum_{\ubmr{s}{r}}{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A}) \nonumber \\
-  &= \sum_{r=1}^{R} \sum_{t=1}^{T^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}\sum_{\smtr{s}{t+1}{r}=1}^{|\Ss|}{p(\smtr{s}{t}{r}, \smtr{s}{t+1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})
-\end{align}
-
-Since $\bm{A}_{i, j}  = p(\smtr{s}{t+1}{r}=j|\smtr{s}{t}{r}=i; {A})$ where $A_{i, j}$ is the element of transition matrix $A$, the solution to problem:
-\begin{align}\label{eq:update-transition-prob}
-  \bm{A}^{\mathrm{new}} = &\uargmax{\bm{A}} \Qq(\bm{A}; \bm{H}^{\mathrm{old}}), \nonumber \\
-  \mathrm{s.t.} &\hspace{0.2cm} \bm{A} \cdot \bm{1} = \bm{1} \nonumber \\
-                          &  \bm{A}^{\intercal} \cdot \bm{1} = \bm{1} \nonumber \\
-                          & \bm{A}_{i,j} \geq 0.
-\end{align}
-is
-\begin{equation}
-  \bm{A}_{i,j}^{\mathrm{new}} = \frac{\bar{\xi}_{i,j}}{\sum_{k = 1}^{|\Ss|} \bar{\xi}_{i,k}},
-\end{equation}
-where
-\begin{equation}
-  \bar{\xi}_{i,j} = \sum_{r= 1}^{R} \sum_{t= 1}^{T^{(r)}-1}{p(\smtr{s}{t}{r}=i, \smtr{s}{t+1}{r}=j| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}
-\end{equation}
-
-\subsection{Generative Model Update}
-\eqref{eq:generative-model-update} can be rewritten as
-\begin{align}\label{eq:gm-update}
-  \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})
-  &= \sum_{r=1}^{R}\sum_{\ubmr{s}{r}}{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\ubmr{x}{r} | \ubmr{s}{r}; \bm{\Phi}) \nonumber \\
-  &= \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \log\, p(\bmtr{x}{t}{r} | \smtr{s}{t}{r}; \bm{\Phi}).
-\end{align}
-
-Then the third subproblem of \eqref{eq:m-step-subs} becomes:
-\begin{align}\label{eq:sub-gm}
-  &\uargmax{\bm{\Phi}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}), \nonumber \\
-  \mathrm{s.t.} &\hspace{0.2cm} p(\bm{x} | {s}; \bm{\Phi}) \mathrm{~is~our~ general~model}
-\end{align}
-
-It could be seen from \eqref{eq:gm-update} that the key to update generate  model is to evaluate $p(\bm{x}|{s}; \bm{\Phi})$ for all $s \in \Ss$.
-In Forward/Backward algorithm, evaluation of $p(\bm{x}|{s}; \bm{\Phi})$ is also all what is needed to compute $p(s|\bm{x};\bm{\Phi})$.
-In the following two subsections, we will provide two neural network based generative models that fulfill this requirement and also have high capability for complex signal modeling.
-
-% Let us denote the inverse of $\bm{g}_s$ as $\bm{f}_s = \bm{g}_s^{-1}$.
-% For this case we assume the latent sources are Gaussian. We write the emission probability model as a $K$ generator mixed model:
-% \begin{align}
-    %     p(\bm{x} | s; \bm{\Phi}) &= \sum_{\kappa = 1}^{K} \pi_{s, \kappa} \Nn\left( \bm{f}_{s,\kappa}(\bm{x}); \bm{\mu}_{s,\kappa}, \bm{C}_{s,\kappa}  \right)\bigg|\det\,\left( \pd{\bm{f_{s,\kappa}(\bm{x})}}{\bm{x}} \right)\bigg| \nonumber \\
-    %     \sum_{\kappa = 1}^{K} \pi_{s, \kappa}&= 1
-                                                 %   \end{align}
-
-The help function should be revised to deal with the new latent variable $\kappa$ into:
-\begin{align}\label{eq:obj-q-gen-mix}
-  \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) & = \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|} \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})  \log\,p(\smtr{\kappa}{t}{r}, \bmtr{x}{t}{r}| \smtr{s}{t}{r}; \bm{\Phi}_{s})\nonumber \\
-  =& \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})p(\smtr{\kappa}{t}{r}|\smtr{s}{t}{r}, \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \bigg[\log\,\pi_{\smtr{s}{t}{r},\smtr{\kappa}{t}{r}} \nonumber \\
-                                        & + \log\, p(\bmtr{x}{t}{r} | \smtr{s}{t}{r}, \smtr{\kappa}{t}{r}; \bm{\Phi}_s) \bigg]
-\end{align}
-                                          %                                           \Nn\left( \bm{f}_{\smtr{s}{t}{r},\smtr{\kappa}{t}{r}}(\bm{x}); \bm{\mu}_{\smtr{s}{t}{r},\smtr{\kappa}{t}{r}}, \bm{C}_{\smtr{s}{t}{r},\smtr{\kappa}{t}{r}} \right) \log\,\bigg| \det\, \left( \pd{\bm{f}_{\smtr{s}{t}{r},\smtr{\kappa}{t}{r}}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg| \bigg].
-                                          %   \end{align}
-In \eqref{eq:obj-q-gen-mix}, $p(s_t| \ubm{x}, \bm{H}^{\mathrm{old}})$ is computed by forward/backward algorithm. The posterior of $\kappa$ is:
-\begin{align}\label{eq:kappa-posterior}
-  p(\kappa| s, \ubm{x}; \bm{H}^{\mathrm{old}})
-  &=  \frac{p(\kappa, \ubm{x}| s; \bm{H}^{\mathrm{old}})}{p(\ubm{x}| s,\bm{H}^{\mathrm{old}})} \nonumber \\
-  & = \frac{\pi_{s, \kappa} p(\ubm{x}| s, \kappa, \bm{H}^{\mathrm{old}})}{\sum_{\kappa=1}^{K} \pi_{s, \kappa} p(\ubm{x}| s, \kappa,\bm{H}^{\mathrm{old}})} \nonumber \\
-  & = \frac{\pi_{s, \kappa} p(\bm{x}| s, \kappa, \bm{H}^{\mathrm{old}})}{\sum_{\kappa=1}^{K}  \pi_{s, \kappa} p(\bm{x}| s, \kappa,\bm{H}^{\mathrm{old}})} 
-  % & = \frac{\pi_{s_t, \kappa_t} \Nn\left( \bm{f}_{{s}_{t},{\kappa}_{t}}(\bm{x}); \bm{\mu}_{{s}_{t},{\kappa}_{t}}, \bm{C}_{{s}_{t},{\kappa}_{t}} \right) \bigg| \det\, \left( \pd{\bm{f}_{{s}_{t},{\kappa}_{t}}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg| }{\sum_{\kappa_t=1}^{K}\pi_{s_t, \kappa_t} \Nn\left( \bm{f}_{{s}_{t},{\kappa}_{t}}(\bm{x}); \bm{\mu}_{{s}_{t},{\kappa}_{t}}, \bm{C}_{{s}_{t},{\kappa}_{t}} \right) \bigg| \det\, \left( \pd{\bm{f}_{{s}_{t},{\kappa}_{t}}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg|} \nonumber \\
-\end{align}
-where the last equation is due to the fact that only $\bm{x}_t$ among sequence $\ubm{x}$ depends on $s_t, \kappa_t$.
-
-    %     \begin{align}
-
-    %   \end{align}
-
-The latent prior for mixture of each source $s$ is obtained by solving the following problem:
-\begin{align}\label{opm:pi}
-  \pi_{s, \kappa} & = \uargmax{\pi_{s, \kappa}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) \\ \nonumber
-                  & s.t. \, \sum_{\kappa=1}^{K} \pi_{s, \kappa}= 1, \forall s = 1, 2, \cdots, |\Ss|. 
-\end{align}
-
-To solve problem \eqref{opm:pi}, we formulate its Lagrange function as
-\begin{equation}
-  \Ll = \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) + \sum_{\lambda_s} \lambda_s\left( 1-  \sum_{\kappa=1}^{K}\pi_{s, \kappa}  \right).
-\end{equation}
-Solving $\pd{\Ll}{\pi_{s, \kappa}} = 0$ gives
-\begin{equation}
-  \pi_{s,\kappa} = \frac{1}{\lambda_s}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}} p(\smtr{s}{t}{r}=s, \smtr{\kappa}{t}{r} =\kappa| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})
-\end{equation}
-With condition $\sum_{\kappa=1}^{K} \pi_{s, \kappa}= 1, \forall s = 1, 2, \cdots, |\Ss|$, we have
-\begin{equation}
-  \lambda_s = \sum_{\kappa=1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}} p(\smtr{s}{t}{r}=s, \smtr{\kappa}{t}{r} =\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}})
-\end{equation}
-Then the solution to \eqref{opm:pi} is
-\begin{equation}\label{eq:mix-latent-parameter-solution}
-  \pi_{s, \kappa} = \frac{\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }{\sum_{k =1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=k | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }
-\end{equation}
-where $p(s, \kappa | \ubm{x}; \bm{H}^{\mathrm{old}}) = p(s| \ubm{x}; \bm{H}^{\mathrm{old}}) p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$.
-$p(s| \ubm{x}; \bm{H}^{\mathrm{old}})$ that can be computed by results of forward/backward and \eqref{eq:kappa-posterior}.
-The posterior for $\kappa$ can be obtained by
-\begin{equation}
-  p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}}) = \frac{p(\kappa, \ubm{x} | s; \bm{H}^{\mathrm{old}})}{\sum_{\kappa}{p(\kappa, \ubm{x} | s; \bm{H}^{\mathrm{old}})}} = \frac{\pi_{s, \kappa} p( \bm{x} | s; \bm{H}^{\mathrm{old}})}{\sum_{\kappa}{\pi_{s, \kappa} p(\bm{x} | s; \bm{H}^{\mathrm{old}})}}
-  \end{equation}
-
-In implementation, GenHMM uses the generator mixed emission model with $\kappa$-th component as:
-
-\begin{align}
-  &p(\bm{x}| s, \kappa; \bm{\Phi}_{s, \kappa}) \nonumber \\
-  =&p_{s, \kappa}(\bm{z}) \bigg| \det \left( \pd{\bm{z}}{\bm{x}} \right) \bigg| \nonumber \\
-  = &\Nn\left( \bm{f}_{{s},{\kappa}}(\bm{x}); \bm{\mu}_{{s},{\kappa}}, \bm{C}_{{s},{\kappa}} \right) \bigg| \det\, \left( \pd{\bm{f}_{{s},{\kappa}}(\bm{x})}{\bm{x}} \right) \bigg|
-\end{align}
-where $\bm{f}_{{s},{\kappa}} = \bm{g^{-1}_{{s},{\kappa}}}$ is defined by parameter set $\bm{\theta}_{s, \kappa}$, and $\bm{\Phi}_{s, \kappa} = \left\{\bm{\mu}_{{s},{\kappa}}, \bm{C}_{{s},{\kappa}} , \bm{\theta}_{s, \kappa} \right\}, \kappa = 1, 2, \cdots, K$.
-We can start by setting $\bm{\mu}_{{s},{\kappa}} = \bm{0}$ and $\bm{C}_{{s},{\kappa}} = diag(\bm{1})$.
 \begin{algorithm}[H]
@@
 
-\subsubsection{Generator-tied HMM (GtHMM)}
-\begin{figure}[!h]
-  \centering
-  \begin{tikzpicture}
-    \tikzstyle{enode} = [thick, draw=blue, ellipse, inner sep = 1pt,  align=center]
-                                           %                                            \tikzstyle{nnode} = [thick, rectangle, rounded corners = 1pt,minimum size = 1.2cm,draw,inner sep = 2pt]
-    \tikzstyle{nnode} = [thick, rectangle, rounded corners = 2pt,minimum size = 0.8cm,draw,inner sep = 2pt]
-    \node[enode] (z1) at (0,1.8) {$\bm{y}\sim p_1(\bm{y})$};
-    \node[enode] (z2) at (0,0.5){$\bm{y}\sim p_{2}(\bm{y})$};
-    \node[enode] (zK) at (0,-1.8) {$\bm{y}\sim p_{|\Ss|}(\bm{y})$}; 
-    \node[enode] (x) at (5.8,1){$\ubm{x}\sim p(\ubm{x};\bm{H})$};
-                                           %                                            \node at (3,-1){$p(\bm{z}) = \sum_k\pi_kp_k(\bm{z})$};
-    \node[nnode] (g) at (3.5,1) {$\bm{g}$};
-
-    \node[nnode] (m) at (3.5,-2) {Memory};
-    \node[nnode] (a) at (3.5,-0.5) {${A}$};
-
-    \draw[dotted,line width=2pt] (0,-0.3) -- (0,-1.2);
-    \filldraw[->] (1.8, 0.5)circle (2pt) --  (g) ;
-    \draw[->] (z1) -- (1.6, 1.8);
-    \draw[->] (z2) -- (1.6, 0.5);
-                                           %                                            \draw[->] (1,-0.5) -- (1.8, -0.5);
-    \draw[->] (zK) -- (1.6, -1.8);
-                                           %                                            \draw[->] (z1) [in=180,out=0]  to node[right]{$\pi_1$} (g);
-                                           %                                            \draw[->] (z2) [in=180,out=0]  to node[above]{$\pi_2$} (g);
-                                           %                                            \draw[->,dashed] (zK) [in=180,out=0]  to node[right]{$\pi_K$} (g);
-    \draw[->] (g) to (x);
-    \begin{scope}[ thick, every node/.style={sloped,allow upside down}]
-      \draw (2.1,0.9)-- (2.2, 0.);
-      \draw (2.2,0.)-- node {\midarrow} (2.2,-2);
-      \draw (2.2,-2)-- (m);
-      \draw (m)-- (5, -2);
-      \draw (5, -2)-- node {\midarrow} (5 ,-0.5);
-      \draw (5, -0.5) -- (a);
-      \draw (a)-- node {\midarrow} (2.2, -0.5);
-    \end{scope}
-    \node at (4.8, -1) {$s_{t}$};
-    \node at (2.56, -0.25) {$s_{t+1}$};
-
-  \end{tikzpicture}
-  \caption{LatM-HMM Model defined by $\bm{H} = \{\Ss, \bm{q}, A, p({\bm{x}}|{\bm{s}}; \bm{\Phi}) \}$}
-  \vspace{0.1cm}
-\end{figure}
-
-\textcolor{blue}{to be continued...}
-
-Alternatively, we can use a latent-source mixed HMM (LatM-HMM) where different latent source share the same generator functioning as feature mapping. Then the generator of the LatM-HMM is defined as
-\begin{equation}
-  \left\{\bm{g}| \bm{g}: \bm{z}\rightarrow \bm{x}, s \in \Ss, \bm{z}\sim p_{s}(\bm{z})\right\}.
-\end{equation}
-We use $\bm{f} = \bm{g}^{-1}$ to denote inverse of $\bm{g}$ and use $\bm{\theta}$ to denote the parameter set of $\bm{g}$. Then the conditional probability for LatM-HMM is modeled as
-\begin{align}
-  p(\bm{x} | s; \bm{\Phi})
-  &= p_s(\bm{z}) \bigg| \det \left( \pd{\bm{z}}{\bm{x}} \right) \bigg| \nonumber \\
-  &= p_s(\bm{f}(\bm{x})) \bigg| \det \left( \pd{\bm{f}(\bm{x})}{\bm{x}} \right) \bigg|
-\end{align}
-
-The parameter set for this model to be decide is $\bm{\Phi} = \left\{ \bm{\theta},  \bm{\omega}_s, \forall s\in \Ss\right\}$.
-Then the problem in \eqref{eq:sub-gm} can be reformulated as:
-\begin{align}
-  &\umax{\bm{\Phi}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) \nonumber \\
-  =&\umax{ \bm{\theta}, \bm{\omega}_s, \forall s\in \Ss} \sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \left[\log\,p_{\smtr{s}{t}{r}}(\bm{f}(\bmtr{x}{t}{r})) + \log\,\bigg| \det\, \left( \pd{\bm{f}(\bmtr{x}{t}{r})}{\bmtr{x}{t}{r}} \right) \bigg| \right].
-\end{align}
-
 \section{Convergence Analysis}

### 2019-08-13 manuscript/main.tex
add convg

@@
-%\def\year{2020}\relax
-%File: formatting-instruction.tex
+% \def\year{2020}\relax
+% File: formatting-instruction.tex
 \documentclass[letterpaper]{article} % DO NOT CHANGE THIS
@@
 \setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
-%\nocopyright
-%PDF Info Is REQUIRED.
+% \nocopyright
+% PDF Info Is REQUIRED.
 % For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
@@
 
-\subsection{Problem Statement}
+\subsection{Learning in EM framework}
 
@@
 \end{equation}
-
-With such a model GenHMM at hand, the remaining problems are:
-\begin{itemize}
-\item How to realize GenHMM by neural networks to have high modeling capacity?
-\item How to train GenHMM to sovle problem in \eqref{eq:ml-of-hmm} using practical algorithm?
-\item Would the training of GenHMM converges?
-\end{itemize}
-We would answer these questions in the following section.
-
-\section{Practical Solution for GenHMM}
-
-In this section, we provide answers to the questions at the end of last section.
 
@@
 The marginal $p(s_t| \ubm{x}; \bm{H})$ is also efficiently computed as the joint posterior.
+
+With such a model GenHMM at hand, the remaining problems are:
+\begin{itemize}
+\item How to realize GenHMM by neural networks to have high modeling capacity?
+\item How to train GenHMM to sovle problem in \eqref{eq:ml-of-hmm} using practical algorithm?
+\item Would the training of GenHMM converges?
+\end{itemize}
+We would answer these questions in the following section.
+
+\section{Practical Solution for GenHMM}
+
+In this section, we provide answers to the questions at the end of last section.
 
@@
 
-\subsection{Learning of GenHMM}
+\subsection{Learning of GenHMM}\label{subsec:optmGenHMM}
 
@@
 
-\textcolor{red}{expand to neural network here}By substituting \eqref{eq:changel-variable} and \eqref{eq:cat-jacobian} into \eqref{eq:obj-q-gen-mix}, we have loss function for neural networks as
+By substituting \eqref{eq:changel-variable} and \eqref{eq:cat-jacobian} into \eqref{eq:obj-q-gen-mix}, we have loss function for neural networks as
 \begin{align}\label{eq:obj-q-gen-mix}
@@
 where
-\begin{equation}
+\begin{equation}\label{eq:update-transition-solt}
   \bar{\xi}_{i,j} = \sum_{r= 1}^{R} \sum_{t= 1}^{T^{(r)}-1}{p(\smtr{s}{t}{r}=i, \smtr{s}{t+1}{r}=j| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}
@@
 \subsection{Algorithm of GenHMM}
+We goal is to minimize the KL divergence $KL(\hat{p}(\ubm{x})\| p(\ubm{x};\bm{H}))$ or maximizing$\sum_{r=1}^{R}\log\,p(\ubmr{x}{r}; \bm{H})$. The 
 Algorithm here:
@@
 \begin{algorithm}[H]
-  \caption{Meta algorithm for HMM powered by Generative Models (GMHMM)}
+  \caption{Learning of GenHMM}
   \begin{algorithmic}[1]
@@
       Building $\bm{H}^{\mathrm{old}}, \bm{H} \in \Hh$ gives: \\
-      $\bm{H}^{\mathrm{old}} = \{\Ss, \bm{q}^{\mathrm{old}}, A^{\mathrm{old}}, p(\bm{x}|s; \bm{\Phi}^{\mathrm{old}})\}$, $\bm{H} = \{\Ss, \bm{q}, A, p(\bm{x}|s; \bm{\Phi})\}$
-    }, \\
+      $\bm{H}^{\mathrm{old}} = \{\Ss, \bm{q}^{\mathrm{old}}, A^{\mathrm{old}}, p(\bm{x}|s; \bm{\Phi}^{\mathrm{old}})\}$, $\bm{H} = \{\Ss, \bm{q}, A, p(\bm{x}|s; \bm{\Phi})\}$\\
+    Empirical distribution $\hat{p}(\bm{x})$ of dataset;} \\
     \STATE Initialize $\bm{H}$
     \STATE $\bm{H}^{\mathrm{old}} \gets \bm{H}$
+    \STATE Set learning rate $\eta$, neural network optimization epochs $T$ per EM step.
     \FOR { $\bm{H}$ not converge}
-    \STATE Sample a batch of data $\left\{ \ubmr{x}{r} \right\}_{r=1}^{R_b}$ from the dataset $\hat{p}(\ubm{x})$
-    \STATE Compute $p(\smtr{s}{t}{r} | \ubmr{x}{r}; \bm{H}^{\mathrm{old}})$, $p(\smtr{s}{t}{r}, \smtr{s}{t+1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})$ by forward/backward algorithm;
-    \STATE $\bm{q} \gets \uargmin{\bm{q}}\, \Qq(\bm{q}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:update-initial-state-prob};
-    \STATE $\bm{A} \gets \uargmin{\bm{A}}\Qq(\bm{A}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:update-transition-prob};
-    \STATE $\bm{\Phi} \gets \uargmin{\bm{\Phi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})$ by calling Algorithm~\ref{algo:opt-Phi-GenHMM} or \ref{algo:opt-Phi-LatHMM};
+        \FOR {epoch $t < T$}
+    \STATE Sample a batch of data $\left\{ \ubmr{x}{r} \right\}_{r=1}^{R_b}$ from dataset $\hat{p}(\ubm{x})$
+
+    \STATE Compute posterior $p(\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})$  
+    \STATE Formulate loss ${\Qq}\left({\bm{\Theta}}, {\bm{H}}^{\mathrm{old}}\right)$ in \eqref{eq:obj-q-gen-mix}
+
+    \STATE $\partial{\bm{\Theta}} \gets  \nabla_{\bm{\Theta}} {\Qq}\left({\bm{\Theta}},{\bm{H}}^{\mathrm{old}}\right)$
+    \STATE $\bm{\Theta} \gets \bm{\Theta} + \eta \cdot \partial{\bm{\theta}_s}$
+    \ENDFOR
+    \STATE $\bm{q} \gets \uargmax{\bm{q}}\, \Qq(\bm{q}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:update-initial-state-prob};
+    \STATE $\bm{A} \gets \uargmax{\bm{A}}\Qq(\bm{A}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:update-transition-prob};
+    \STATE $\bm{\Pi} \gets \uargmax{\bm{\Pi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:update-transition-solt};
     \STATE $\bm{H}^{\mathrm{old}} \gets \bm{H}$
@@
 
-\begin{algorithm}[H]
-  \caption{M-step w.r.t. $\bm{\Phi}$ powered by GenMM}\label{algo:opt-Phi-GenHMM}
-  \begin{algorithmic}[1]
-    \STATE {\bfseries Input:}
-    Latent mixture distribution: $\Nn\left(\bm{z}; \bm{0}, \mathrm{diag}(\bm{1})\right)$, $\forall \, s \in \Ss, \kappa = 1, 2, \cdots, K$;\\
-    Empirical distribution $\hat{p}(\bm{x})$ of dataset; \\
-    \STATE Set a total number of epochs $T$ of training as stop criterion. A learning rate $\eta$.
-    \FOR {epoch $t < T$}
-    \STATE Sample a batch of data $\left\{ \ubmr{x}{r} \right\}_{r=1}^{R_b}$ from dataset $P_d(\ubm{x})$
+\subsection{Convergence Analysis}
+\begin{prop}
+  Assume that parameters of GenHMM are in a compact set.
+Further assume that $\Qq(\bm{\Theta}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) \geq \Qq(\bm{\Theta}^{\mathrm{old}}; \bm{H}^{\mathrm{old}})$, then we always have $KL(\hat{p}(\ubm{x})\| p(\ubm{x};\bm{H}^{\mathrm{new}})) \geq KL(\hat{p}(\ubm{x})\| p(\ubm{x};\bm{H}^{\mathrm{old}}))$, and GenHMM converges.
+\end{prop}
 
-    \STATE Compute $p(\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})$  by Forward/backward and \eqref{eq:kappa-posterior}.
-    %     = \frac{\pi_k \Nn\left(f(X); \bm{\mu}_k^{t},\bm{\sigma}_k^{t}\right)}{\sum_{k=1}^K\; \pi_k\Nn\left(f(X); \bm{\mu}_k^{t},\bm{\sigma}_k^{t}\right) }, \forall \left\{ x^{(i)}\right\}_{i=1}^{n_b}$
-    % \Comment{Calculate posterior \autoref{eq-posterior-gamma}}
-    \STATE Compute loss ${\Qq}\left({\bm{\Phi}}, {\bm{H}}^{\mathrm{old}}\right)$ in \eqref{eq:obj-q-gen-mix}
+\begin{proof}
+  We begin with the comparison of log-likelihood evaluated under $\bm{H}^{\mathrm{old}}$ and $\bm{H}^{\mathrm{old}}$. The log-likelihood of dataset given by $\hat{p}(\ubm{x})$ can be reformulated as
+  \begin{align*}
+    &\EE_{\hat{p}(\ubm{x})}\left[ \log\,p(\ubm{x};\bm{H}^{\mathrm{new}}) \right] \nonumber \\
+    =& \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{new}})}{p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right] \nonumber \\
+    &+ \EE_{\hat{p}(\ubm{x})}\left[ KL(p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})\|p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{new}})) \right],
+  \end{align*}
+  where the first term on the right hand side of equality can be further written as
+  \begin{align*}
+    &\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{new}})}{p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right] \nonumber \\
+    = &\Qq(\bm{H}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) + \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})\right].
+  \end{align*}
+  According to subsection~\ref{subsec:optmGenHMM}, the subproblem solutions gives
+  \begin{align*}
+    \Qq(\bm{q}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) &\geq \Qq(\bm{q}^{\mathrm{old}}; \bm{H}^{\mathrm{old}}),\nonumber \\
+    \Qq(\bm{A}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) &\geq \Qq(\bm{A}^{\mathrm{old}}; \bm{H}^{\mathrm{old}}),\nonumber \\
+    \Qq(\bm{\Pi}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) &\geq \Qq(\bm{\Pi}^{\mathrm{old}}; \bm{H}^{\mathrm{old}}).
+  \end{align*}
+  For the learning w.r.t.
+neural network parameter set $\bm{\Theta}$, as long as the lose function does not decrease during EM iterations, i.e.
+$\Qq(\bm{\Theta}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) \geq \Qq(\bm{\Theta}^{\mathrm{old}}; \bm{H}^{\mathrm{old}})$, we would have
+  \begin{equation*}
+    \Qq(\bm{H}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) \geq \Qq(\bm{H}^{\mathrm{old}}; \bm{H}^{\mathrm{old}}),
+  \end{equation*}
+  since
+  \begin{align*}
+    \Qq(\bm{H}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) = &\Qq(\bm{q}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) + \Qq(\bm{A}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) \nonumber \\
+    &+ \Qq(\bm{\Pi}^{\mathrm{new}}; \bm{H}^{\mathrm{old}})
+    + \Qq(\bm{\Theta}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}).
+  \end{align*}
+  Therefore we have the inequality 
+  \begin{align*}
+    &\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{new}})}{p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right] \nonumber \\
+    \geq &\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{old}})}{p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right].
+  \end{align*}
+  Due to the fact that $KL(p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})\|p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}}))=0$, we have
+  \begin{align*}
+      &\EE_{\hat{p}(\ubm{x})}\left[ \log\,p(\ubm{x};\bm{H}^{\mathrm{new}}) \right] \nonumber \\
+    \geq & \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{old}})}{p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right] \nonumber \\
+    = & \EE_{\hat{p}(\ubm{x})}\left[ \log\,p(\ubm{x};\bm{H}^{\mathrm{old}}) \right].
+  \end{align*}
+  Therefore
+  \begin{align*}
+    KL(\hat{p}(\ubm{x})\|p(\ubm{x};\bm{H}^{\mathrm{new}})) = &\EE_{\hat{p}(\ubm{x})}\left[ \log\,\frac{\hat{p}(\ubm{x})}{p(\ubm{x};\bm{H}^{\mathrm{new}})} \right] \nonumber \\
+    \leq &\EE_{\hat{p}(\ubm{x})}\left[ \log\,\frac{\hat{p}(\ubm{x})}{p(\ubm{x};\bm{H}^{\mathrm{old}})} \right] \\
+    = & KL(\hat{p}(\ubm{x})\|p(\ubm{x};\bm{H}^{\mathrm{old}})).
+    \end{align*}
+  Since KL divergence is non-negative and thus lower bounded, GenHMM will converge.
+\end{proof}
 
-    \STATE $\partial{\bm{\th}_s} \gets  \nabla_{\bm{\theta}} -
-    \frac{1}{R_{b}}{\Qq}\left({\bm{\Phi}},{\bm{H}}^{\mathrm{old}}\right)$
-    % \Comment{E-step}
-    % \STATE $\partial{g_k} \gets \nabla_{\th_k}\widetilde{\Qq}\left(\left\{ \th_k
-    %   \right\}_{k=1}^{K};\left\{ \th_k^t\right\}_{k=1}^{K}\right)$,
-    % $\forall \th_k \in\left\{ \th_k\right\}_{k=1}^{K}$
-    \STATE $\bm{\th}_s \gets \bm{\th}_s - \eta \cdot \partial{\bm{\theta}_s}$, $\forall s \in \Ss$
-    \ENDFOR
-
-    \STATE Update $\pi_{s, \kappa}, \forall, s\in \Ss, \kappa= 1, 2, \cdots, K$, according to \eqref{eq:mix-latent-parameter-solution}
-
-    \STATE Assemble $\bm{\Phi} = \left\{  \bm{\Phi}_s| s \in \Ss\right\}$ for $\bm{\Phi}_s = \left\{ \bm{\theta}_{s, \kappa}, \bm{\mu}_{{s},{\kappa}}, \bm{C}_{{s},{\kappa}}| \kappa = 1, 2, \cdots, K \right\}$.
-
-  \end{algorithmic}
-\end{algorithm}
-
-\section{Convergence Analysis}
-
-\section{On Implementation of acoustic signal}
+\section{Experiments}
 test citing \cite{2018arXiv180703039K}

### 2019-08-15 manuscript/main.tex
update manuscript

@@
 \section{introduction}
+Sequential data modeling is a challenging topic in pattern recognition and machine learning, since for many applications, the assumption of independent and identically distributed data points is too strong to model data properly.
+Hidden Markov model (HMM) is classic way to model sequential data.
+HMM has been widely used in different practical problems, and especially is known for applications in reinforcement learning [], natural language modeling [, biological sequence analysis such as proteins and DNA [], etc.
+
+HMM provides an framework for sequential data modeling.
+But for ..., a probabilistic model is used to describe how each state of HMM should represent a part of sequential data.
+The typical way is to use a Gaussian mixture model (GMM) per state of HMM, where GMMs are used to represent the relationship between states of HMM and sequential data input.
+A large amount of research on GMM based HMM (GMMHMM) has been carried out[] in literature.
+
+Given the success of GMMHMM, it is not efficient for modeling data in nonlinear manifold.
+Research attempts of training HMM with neural networks have been done to boost the modeling capacity of HMM.
+A successful work is to bring restrictive Boltzmann machine (RBM) \cite{Hinton2012} into HMM learning \cite{hinton2012deepSpeech}.
+In RBM based HMM, the hierarchical learning scheme is used, where a stack of RBMs are trained in unsupervised fashion , and then the stack of RBMs with a final softmax layer are used to mapping sequential input signal to states of a HMM.
+...
+talk about some other DNNHMMs.
+
+Introduce our model...
+
+Contribution ...
 

### 2019-08-16 manuscript/main.tex
add intro section

--- 2019-08-13 manuscript/main.tex
@@
 \section{introduction}
+Sequential data modeling is a challenging topic in pattern recognition and machine learning, since for many applications, the assumption of independent and identically distributed data points is too strong to model data properly.
+Hidden Markov model (HMM) is classic way to model sequential data.
+HMM has been widely used in different practical problems, and especially is known for applications in reinforcement learning \cite{ding2018reinforcementhmm,levine2018reinforcementReview}, natural language modeling \cite{khan2016survey,Hariyanti_2019}, biological sequence analysis such as proteins \cite{ASHWIN20172} and DNA \cite{ren2015dna}, etc.
+
+HMM provides an framework for sequential data modeling.
+But for ..., a probabilistic model is used to describe how each state of HMM should represent a part of sequential data.
+The typical way is to use a Gaussian mixture model (GMM) per state of HMM, where GMMs are used to represent the relationship between states of HMM and sequential data input.
+A large amount of research on GMM based HMM (GMM-HMM) has been carried out[] in literature.
+
+Given the success of GMM-HMM, it is not efficient for modeling data in nonlinear manifold.
+Research attempts of training HMM with neural networks have been done to boost the modeling capacity of HMM.
+A successful work is to bring restrictive Boltzmann machine (RBM) \cite{Hinton2012} into HMM learning \cite{hinton2012deepSpeech,li2013hybrid,Miao2013ImprovingLC}.
+In RBM based HMM, the hierarchical learning scheme is used: i) a stack of RBMs are trained in unsupervised fashion, ii) then a final softmax layer is added to the stack of RBMs to represent the posterior of state of HMM given data input, iii) a discriminative training is followed to do the final tuning of the model.
+
+Another track of related work is hybrid method of temporal neural network models and HMM.
+In \cite{liu2019lstmHmmHyb,buys2018bridging,vik2016rnnHmm}, a long short-term memory (LSTM) model/recurrent neural network (RNN) is combined with HMM as hybridization.
+A hierarchical training is carried out by: i) training a HMM first, ii) and then doing modified training of LSTM using trained HMM.
+This training procedure is argued by the authors to use LSTM or RNN to fill in the gap where HMM can not learn.
+
+work of epfl fiting GMM with neural network...
+
+The above work help improve modeling capacity of HMM based models by bringing in neural networks.
+A softmax layer is usually used to represent probability whenever a conditional distribution is needed.
+These hierarchical schemes are built based on intuition of domain knowledge.
+Training of these hierarchical models usually need specific experience in relevant area to be able to the hierarchical procedure of training.
+
+In this work, we propose a generator-mixed HMM model, termed as GenHMM, where generators are realized by neural networks to help the model gain high modeling capacity.
+Our proposed model, GenHMM:
+\begin{itemize}
+\item has advantage of easy to do training, since GenHMM training employs expectation maximization (EM) framework.
+Training a GenHMM is as easy as training a GMM-HMM model, but configure of GenHMM is flexible.
+\item has high modeling capability of sequence data, though its training and using are handy
+\item is able to compute loglikelihood exactly and efficiently.
+\end{itemize}
+Instead of using softmax for probability representation, our GenHMM is able to compute exact loglikelihood of given sequence data based on change of variable formula.
+To make the loglikelihood computation efficient, neural network based generators of GenHMM are implemented as flow models.
+
+Our contributions in the paper are as follows:
+\begin{itemize}
+\item Proposing a neural network based HMM for sequential data modeling, i.e. GenHMM. The likelihood computation of sequential data under GenHMM can be computed exactly and efficiently.
+\item Designing practical algorithm for training GenHMM under EM framework. Stochastic gradient  search in batch fashion is embedded in this algorithm.
+\item Convergence analysis for GenHMM with the learning algorithm with gradient based optimization.
+\end{itemize}
 
@@
 network, proposed in \cite{DBLP:journals/corr/DinhKB14}, is such a
-feed-forward neural network, which is further improved in subsequent works \cite{2016arXiv160508803D, 2018arXiv180703039K}.
+feed-forward neural network, which is further improved in subsequent works \cite{2016arXiv160508803D,kingma2018glow}.
 It also has additional advantages as efficient Jacobian computation and low computational complexity.
@@
 Since the coupling has a partial identity mapping, direct concatenation of multiple such coupling mappings would result in a partial identity mapping of the whole neural network $\tilde{\bm{g}}$.
-Alternating the positions of identity mapping \cite{2016arXiv160508803D} or using $1\times1$ convolution operations \cite{2018arXiv180703039K} before each coupling mapping is used to treat the issue.
+Alternating the positions of identity mapping \cite{2016arXiv160508803D} or using $1\times1$ convolution operations \cite{kingma2018glow} before each coupling mapping is used to treat the issue.
 % \textcolor{blue}{Some reading problem in this sentence: Since the coupling has a partial identity mapping, alternating the positions of identity mapping \cite{2016arXiv160508803D} or using $1\times1$ convolution operations \cite{2018arXiv180703039K} when
 % concatenating multiple layers together were proposed to avoid the situation that a part of the output is equal to a part of the input.} %In addition, applying affine mapping does not affect its advantage of trivial inverse and determinate computation.
-Furthermore, \cite{2016arXiv160508803D}\cite{2018arXiv180703039K} split some hidden
+Furthermore, \cite{2016arXiv160508803D}\cite{kingma2018glow} split some hidden
 layer signal $\bm{h}$ and model a part of it directly as standard Gaussian to reduce computation and memory burden.
@@
 \section{Experiments}
-test citing \cite{2018arXiv180703039K}
 
@@
 
+\newpage
+
 \bibliography{myref}
 \bibliographystyle{aaai}
-
-% \bibliography{bibliography}
-
-% \appendix
-
-% \input{section/sec-appendixA}
 \end{document}

### 2019-08-16 manuscript/main.tex
add exp section

@@
 HMM provides an framework for sequential data modeling.
-But for ..., a probabilistic model is used to describe how each state of HMM should represent a part of sequential data.
-The typical way is to use a Gaussian mixture model (GMM) per state of HMM, where GMMs are used to represent the relationship between states of HMM and sequential data input.
-A large amount of research on GMM based HMM (GMM-HMM) has been carried out[] in literature.
+To connect states of HMM with input data , a probabilistic model is used to describe how each state of HMM should represent a part of sequential data.
+The typical way is to use a Gaussian mixture model (GMM) per state of HMM\cite{juang1986maximum}, where GMMs are used to represent the relationship between states of HMM and sequential data input.
+GMM based HMM (GMM-HMM) has becomes a standard model for sequential data modeling, and employed widely for practical applications, especially in speech recognition \cite{gales2008application}.
 
@@
 \section{Experiments}
+To show the validity of our model, we implement our model in PyTorch and test it with sequential data.
+The code is available at github repository\footnote{\textcolor{blue}{link}}.
+In the next subsection, we first discuss the experimental setups and the show the experimental results.
 
-For problem \eqref{eq:sub-gm} we are going to use our generative models to solve. I have the following consideration to revised our LatMM and GenMM for this application:
-\begin{itemize}
-\item Use factorized model instead of additive mixture model, to make likelihood computation logarithm domain compatible; 
-\item Use full EM fashion instead of mini-batch fashion for training: store generative model as old for EM, there are always two neural networks working, one old for probability evaluation and one new for optimization.
-\end{itemize}
+\subsection{Experimental Setups}
+The dataset used for sequential data modeling and classification is TIMIT dataset where the speech is sampled at $16$kHz.
+For extraction of feature vectors, we use $25$ms frame length and $10$ms frame shift to convert sound track into standard Mel-frequency cepstral coefficients (MFCCs) features.
+The deltas and delta-deltas features are also concatenated and used in our experiments.
+
+The TIMIT dataset consists of $5300$ phoneme-labeled speech utterances which are partitioned into two sets: \textcolor{blue}{a train set consists of $4620$ utterance, and a test set consists of $1680$ utterances.} There are totally $61$ different types of phones in TIMIT labels.
+Our experiments in this section is carried out in two different cases: i) full $61$-phoneme classification case, where both training and testing use the $61$-phoneme data; ii) $39$-phonme classification case, where $61$ phonemes are folded into $39$ phonemes as described in \cite{Perdigao11}.
+
+In all the following experiments, all generators $\left\{\bm{g}_{s,\kappa}|s\in \Ss, \kappa=1, 2, \cdots, K \right\}$ of GenHMM are implemented as flow models.
+To be more specific, our generator structure follows the structure of flow model in RealNVP \cite{2016arXiv160508803D}.
+The coupling layer shown as \eqref{eq-gl} maps a part of its input signal identically.
+The following coupling layer would alternate the input signal older such that after two consequential coupling layers, there is no signal remains the same any more.
+We term such two consequential coupling layers as a \textit{flow block}.
+In our experiments, a generator $\bm{g}_{s,\kappa}$ consists of four flow blocks.
+The latent source signal is assumed as standard Gaussian, i.e.
+$p_{s,\kappa}(\bm{z})$ is density function of standard Gaussian.
+
+In learning of GenHMM, the number of states is set as $|\Ss| \in \left\{ 3,4,5 \right\}$.
+The exact number of states for a phoneme is decided by computing the average length of MFCC frames per phone in training dataset, and clipping the average length into $\left\{ 3,4,5 \right\}$.
+Transition matrix $\bm{A}$ is initialized as upper triangular matrix for each GenHMM.
+
+\subsection{Experimental Results}
 

### 2019-08-18 manuscript/main.tex
write expmt section

@@
 
-\usepackage{caption}
 \usepackage{bm}
@@
 Our experiments in this section is carried out in two different cases: i) full $61$-phoneme classification case, where both training and testing use the $61$-phoneme data; ii) $39$-phonme classification case, where $61$ phonemes are folded into $39$ phonemes as described in \cite{Perdigao11}.
+\textcolor{blue}{The mapping table for the folding process is attached in appendix.}
 
@@
 
+\begin{table}
+  \caption{Test accuracy table for $39$ dimensional features }\label{tab:acc-classification39f}
+  \vspace{-0.3cm}
+  \centering
+  \subcaption{Folded $39$ Phonemes}\label{tab:acc-classification39f_a}
+  \centering  
+  \begin{tabular}{l|c|c|c}
+    \toprule
+    {Model} &  K=1 &  K=3 &  K=5  \\  \midrule
+    GMM-HMM & $62.30\%$ &  $68.06\%$ &  $68.73\%$  \\ \midrule
+    GenHMM & $76.78\%$ & $77.68\%$ &   \\ 
+    \bottomrule                                                                  
+  \end{tabular}
+  \bigskip
+  \subcaption{$61$ Phonemes}\label{tab:acc-classification39f_b}
+  % \centering  
+  \begin{tabular}{l|c|c|c} \toprule
+    {Model} &  K=1 &  K=3 &  K=5
+    \\ \midrule
+    GMM-HMM & $53.64\%$ &  $59.65\%$ & $61.98\%$  \\ \midrule
+    GenHMM & $69.51\%$ & $70.67\%$ & $70.78\%$   \\ 
+    \bottomrule
+  \end{tabular}
+  \vspace{0.3cm}
+\end{table}
+
+We firstly show the phoneme classification using 39 dimensional MFCC features (MFCC coefficients, deltas, and delta-deltas), to validate one possible usage of our proposed model.
+GMM-HMM is trained and tested as a reference model in our experiments.
+Training and testing of GMM-HMM is in the same condition as GenHMMs are trained and tested.
+Dataset usage for GenMM and GMM-HMM is the same, and number of states for GMM-HMM is the same as that for GenHMM in modeling each phoneme.
+Apart from setting the reference model, we also run the experiment comparisons with different total number of mixture components ($K$).
+
+Table \ref{tab:acc-classification39f} shows the results for this experiments, in which we tested both the folded $39$-phoneme classification case (the conventional way) in table \ref{tab:acc-classification39f_a} and also the $61$-phoneme classification case in table \ref{tab:acc-classification39f_b}.
+As shown in both $61$-phoneme and $39$-phoneme cases, GenHMM gets significant higher accuracy than GMM-HMM for the same number of mixture components.
+As our expectation, GenHMM should have better modeling capacity of sequential data since we bring in the neural network based generators into GenHMM, which should be able to represent complex relationship between states of HMM and sequential data.
+This is some increase of accuracy as total number of mixture components in GenHMM is increased from $K=1$ to $K=5$.
+The sequential dependency of data is modeled by HMM itself, while each state of HMM can have better representation using a mixture probabilistic model if data represented by the state is multi-mode.
+Comparing the results in $39$-phoneme and $61$-phoneme cases, GenHMM gets higher accuracy for $39$-phoneme classification than it does for $61$-phoneme classification.
+The total training dataset size remains the same as $61$ phonemes are folded into $39$ phonemes.
+There are less training data available and more classes to be recognized in the $61$-phoneme case, which makes the take more challenging.
+
+\begin{table}
+  \caption{Test accuracy table for $13$ dimensional features }
+  \label{tab:acc-classification13f}
+  \vspace{-0.3cm}
+  \centering
+  \subcaption{Folded $39$ Phonemes}
+  \centering  
+  \begin{tabular}{l|c|c|c}
+    \toprule
+    {Model} &  K=1 &  K=3 &  K=5  \\  \midrule
+    GMM-HMM & $48.59\%$ &  $51.26\%$ &  $52.40\%$  \\ \midrule
+    GenHMM & $\%$ & $\%$ &   \\ 
+    \bottomrule                                                                  
+  \end{tabular}
+  \bigskip
+  \subcaption{$61$ Phonemes}
+  \begin{tabular}{l|c|c|c}
+    \toprule
+    {Model} &  K=1 &  K=3 &  K=5
+    \\ \midrule
+    GMM-HMM & $37.15\%$ &  $40.65\%$ & $42.27\%$  \\ \midrule
+    GenHMM & $\%$ & $\%$ & $\%$   \\ 
+    \bottomrule
+  \end{tabular}
+  \vspace{0.3cm}
+\end{table}
+
+Similar experiments to that in table~\ref{tab:acc-classification13f} are carried out by using only the MFCC coefficients as feature input (excluding deltas and delta-deltas).
+The results are shown in table~\ref{tab:acc-classification13f}.
+The superior performance of GenHMM remains compared with reference model GMM-HMM.
+The gain by using mixture generators are also present in this set of experiments while the difference between $61$-phoneme and $39$-phoneme cases is also similar to the set of experiments in table~\ref{tab:acc-classification39f}.
+
+\section{Conclusion}
+
 \newpage

### 2019-08-18 bin/compute_accuracy_class.py
add noisy acc test

@@ -11,8 +11,8 @@ import time
 import numpy as np
 
 if __name__ == "__main__":
-    usage = "bin/compute_accuracy_class.py exp/gaus/13feats/models/epoch2_class1.mdlc exp/gaus/13feats/data/train.13.pkl exp/gaus/13feats/data/test.13.pkl"
-    if len(sys.argv) != 4 or sys.argv[1] == "-h" or sys.argv[1] == "--help":
+    usage = "bin/compute_accuracy_class.py exp/gaus/13feats/models/epoch2_class1.mdlc exp/gaus/13feats/data/train.13.pkl exp/gaus/13feats/data/test.13.pkl noise_std"
+    if len(sys.argv) < 4 or sys.argv[1] == "-h" or sys.argv[1] == "--help":
         print(usage, file=sys.stderr)
         sys.exit(1)
 
@@ -28,18 +28,28 @@ if __name__ == "__main__":
     
     training_data_file = sys.argv[2]
     testing_data_file = sys.argv[3]
+    # Test for a forth parameter
+    try:
+        noise_std = float(sys.argv[4])
+    except IndexError:
+        noise_std = 0
+
 
     mdl_file = os.path.join(models_dir, "epoch{}.mdl".format(epoch))
     # Builds an array of string containing the train and test data sets for each class
     # size: nclass x 2 (train, test)
-    data_files = [append_class(training_data_file, iclass), append_class(testing_data_file, iclass)]
+    # if noise_std ==0, do standard testing of training and testing. If noise_std > 0, noise testing case
+    if noise_std==0:
+        data_files = [append_class(training_data_file, iclass), append_class(testing_data_file, iclass)]
+    elif noise_std > 0:
+        data_files = [append_class(testing_data_file, iclass)]
     # Load Model
     device = 'cpu'
     if model_type == 'gaus':
         with open(mdl_file, "rb") as handle:
             mdl = pkl.load(handle)
         mdl.device = 'cpu'
-        f = lambda x: accuracy_fun(x, mdl=mdl)
+        f = lambda x: accuracy_fun(x, mdl=mdl, noise_std=noise_std)
     elif model_type == 'gen':
         mdl = load_model(mdl_file)
         if torch.cuda.is_available():
@@ -64,7 +74,7 @@ if __name__ == "__main__":
         # set model into eval mode
         mdl.eval()
         
[…]

### 2019-08-18 src/utils.py

@@ -7,15 +7,15 @@ import sys
 import pickle as pkl
 from parse import parse
 
-
-def accuracy_fun(data_file, mdl=None):
+def accuracy_fun(data_file, mdl=None, noise_std=0):
     X = pkl.load(open(data_file, "rb"))
     # Get the length of all the sequences
     l = [xx.shape[0] for xx in X]
     # zero pad data for batch training
 
     true_class = parse("{}_{}.pkl", os.path.basename(data_file))[1]
-    out_list = [mdl.forward(x_i[:,1:]) for x_i in X]
+    out_list = [mdl.forward(x_i[:,1:] + np.random.randn(*x_i[:,1:].shape) * noise_std ) for x_i in X]
+    
     out = np.array(out_list).transpose()
 
     # the out here should be the shape: data_size * nclasses
@@ -24,8 +24,16 @@ def accuracy_fun(data_file, mdl=None):
     print(data_file, "Done ...", "{}/{}".format(str(istrue.sum()), str(istrue.shape[0])), file=sys.stderr)
     return "{}/{}".format(str(istrue.sum()), str(istrue.shape[0]))
 
-def accuracy_fun_torch(data_file, mdl=None, batch_size_=128):
+def accuracy_fun_torch(data_file, mdl=None, batch_size_=128, noise_std=0):
     X = pkl.load(open(data_file, "rb"))
+    # if noise testing case, manipulate data input with noise
+    if noise_std>0:
+        X = []
+        for xx in X:
+            xx += np.random.randn(*xx.shape) * noise_std
+            xx[:, 0] = 0
+            X.append(xx)
+        X = np.array(X)
     # Get the length of all the sequences
     l = [xx.shape[0] for xx in X]
     # zero pad data for batch training

### 2019-08-18 src/utils.py
one line fix

@@ -28,12 +28,12 @@ def accuracy_fun_torch(data_file, mdl=None, batch_size_=128, noise_std=0):
     X = pkl.load(open(data_file, "rb"))
     # if noise testing case, manipulate data input with noise
     if noise_std>0:
-        X = []
+        Xlist = []
         for xx in X:
             xx += np.random.randn(*xx.shape) * noise_std
             xx[:, 0] = 0
-            X.append(xx)
-        X = np.array(X)
+            Xlist.append(xx)
+        X = np.array(Xlist)
     # Get the length of all the sequences
     l = [xx.shape[0] for xx in X]
     # zero pad data for batch training

### 2019-08-20 manuscript/main.tex
draft v1

@@
 
-\title{Powering Hidden Markov Model by Generative Models}
+\title{Powering Hidden Markov Model by Neural Network based Generative Models}
 \author{
-  Dong Liu, Antoine Honoré, Saikat Chatterjee, and Lars~K. Rasmussen\\
-  KTH Royal Institute of Technology, Stockholm, Sweden \\
+  Dong Liu$^{\dagger}$, Antoine Honoré$^{\ast \dagger}$, Saikat Chatterjee$^{\dagger}$, and Lars~K. Rasmussen$^{\dagger}$\\
+  $^{\dagger}$KTH Royal Institute of Technology, Stockholm, Sweden \\
+  $^{\ast}$Karolinska Institute, Stockholm, Sweden \\
   E-mail: \{doli, honore, sach, lkra\}@kth.se}
@@
 \begin{abstract}
-  to be written...
+  Hidden Markov model (HMM) has been successfully used for sequential data modeling problems.
+In this work, we propose to power the modeling capacity of HMM by bringing in neural network based generative models.
+The proposed model is termed as GenHMM.
+In GenHMM, the hidden states of HMM and input data are related by neural network based generative model that provides exact likelihood computation.
+This is realized by a flow model.
+Apart from modeling itself, we also give learning algorithm for GenHMM in expectation-maximization framework.
+The convergence of learning GenHMM is analyzed.
+We demonstrate the efficiency of GenHMM by classification tasks on practical sequential data.
 \end{abstract}
@@
 Sequential data modeling is a challenging topic in pattern recognition and machine learning, since for many applications, the assumption of independent and identically distributed data points is too strong to model data properly.
-Hidden Markov model (HMM) is classic way to model sequential data.
+Hidden Markov model (HMM) is a classic way to model sequential data.
 HMM has been widely used in different practical problems, and especially is known for applications in reinforcement learning \cite{ding2018reinforcementhmm,levine2018reinforcementReview}, natural language modeling \cite{khan2016survey,Hariyanti_2019}, biological sequence analysis such as proteins \cite{ASHWIN20172} and DNA \cite{ren2015dna}, etc.
@@
 HMM provides an framework for sequential data modeling.
-To connect states of HMM with input data , a probabilistic model is used to describe how each state of HMM should represent a part of sequential data.
-The typical way is to use a Gaussian mixture model (GMM) per state of HMM\cite{juang1986maximum}, where GMMs are used to represent the relationship between states of HMM and sequential data input.
-GMM based HMM (GMM-HMM) has becomes a standard model for sequential data modeling, and employed widely for practical applications, especially in speech recognition \cite{gales2008application}.
+To connect states of HMM with input data, a probabilistic model is used to describe how each state of HMM should represent a pattern of sequential data.
+The typical way is to use a Gaussian mixture model (GMM) per state of HMM \cite{juang1986maximum}, where GMMs are used to represent the relationship between states of HMM and sequential data input.
+GMM based HMM (GMM-HMM) has becomes a standard model for sequential data modeling, and employed widely for practical applications, especially in speech recognition \cite{gales2008application,chatterjee2011auditory}.
 
 Given the success of GMM-HMM, it is not efficient for modeling data in nonlinear manifold.
-Research attempts of training HMM with neural networks have been done to boost the modeling capacity of HMM.
-A successful work is to bring restrictive Boltzmann machine (RBM) \cite{Hinton2012} into HMM learning \cite{hinton2012deepSpeech,li2013hybrid,Miao2013ImprovingLC}.
-In RBM based HMM, the hierarchical learning scheme is used: i) a stack of RBMs are trained in unsupervised fashion, ii) then a final softmax layer is added to the stack of RBMs to represent the posterior of state of HMM given data input, iii) a discriminative training is followed to do the final tuning of the model.
+Research attempts of training HMM with neural networks have been made to boost the modeling capacity of HMM.
+A successful work is to bring restrictive Boltzmann machine (RBM) \cite{Hinton2012} into HMM based models \cite{hinton2012deepSpeech,li2013hybrid,Miao2013ImprovingLC}.
+In RBM based HMM, a hierarchical learning scheme is used: i) a stack of RBMs are trained one after another in unsupervised fashion, ii) then a final softmax layer is added to the stack of RBMs to represent the posterior of state of HMM given data input, iii) a discriminative training is followed to do the final tuning of the model.
 
@@
 
-work of epfl fiting GMM with neural network...
+% work of epfl fiting GMM with neural network...
 
@@
 These hierarchical schemes are built based on intuition of domain knowledge.
-Training of these hierarchical models usually need specific experience in relevant area to be able to the hierarchical procedure of training.
+Training of these hierarchical models usually need specific experience in relevant area to be able to proceed with the hierarchical procedure of training.
 
@@
 \begin{itemize}
+  \item firstly has high modeling capacity of sequential data, due to the neural network based generators;
 \item has advantage of easy to do training, since GenHMM training employs expectation maximization (EM) framework.
-Training a GenHMM is as easy as training a GMM-HMM model, but configure of GenHMM is flexible.
-\item has high modeling capability of sequence data, though its training and using are handy
+Training a GenHMM is as easy as training a GMM-HMM model, while configuration of GenHMM is flexible.
+% \item has high modeling capability of sequence data, though its training and using are handy
 \item is able to compute loglikelihood exactly and efficiently.
@@
 \item Designing practical algorithm for training GenHMM under EM framework. Stochastic gradient  search in batch fashion is embedded in this algorithm.
-\item Convergence analysis for GenHMM with the learning algorithm with gradient based optimization.
+\item Convergence analysis for GenHMM under the proposed learning algorithm.
 \end{itemize}
@@
 
-Our framework is an Hidden Markov Model (HMM).
-An HMM model $\bm{H}$ defined in a hypothesis space $\Hh$, i.e.
+Our framework is a Hidden Markov Model (HMM).
+A HMM model $\bm{H}$ defined in a hypothesis space $\Hh$, i.e.
 $\bm{H} \in \Hh$, is capable to model time-span signal $\ubar{\bm{x}} = \left[ \bm{x}_1, \cdots, \bm{x}_T\right]^{\intercal}$, where $\bm{x}_t\in \RR^{N}$ is the signal at time $t$, $[\cdot]^{\intercal}$ means transpose, and $T$ denote the time length.
@@
 \item $\Ss$ is the set of states of HMM $\bm{H}$;
-\item $\bm{q} = \left[ q_1, q_2, \cdots, q_{|\Ss|}\right]^\intercal$ initial distribution of HMM $\bm{H}$ with $|\Ss|$ is cardinality of $\Ss$, $q_k = p(s=k)$ for random state variable $s$.
+\item $\bm{q} = \left[ q_1, q_2, \cdots, q_{|\Ss|}\right]^\intercal$ initial distribution of HMM $\bm{H}$ with $|\Ss|$ is cardinality of $\Ss$, $q_s = p(s;\bm{H})$ for random state variable $s$.
 In the following of this paper, we use $s_t$ to denote the state $s$ at time $t$.
-\item $\bm{A}$ is the transition matrix for the HMM $\bm{H}$ of size $|\Ss| \times |\Ss|$, $\bm{A}_{i,j} = p(s_{t+1}=j|s_{t}=i)$.
-\item For given hidden state $s$, the observable signal density is $p({\bm{x}}|{s};\bm{\Phi}_{s})$, where $\bm{\Phi}_{s}$ is the parameter set that defines this conditional probabilistic model.
+\item $\bm{A}$ is the transition matrix for the HMM $\bm{H}$ of size $|\Ss| \times |\Ss|$, $\bm{A}_{i,j} = p(s_{t+1}=j|s_{t}=i; \bm{H})$.
+\item For given hidden state $s$, the observable signal density function is $p({\bm{x}}|{s};\bm{\Phi}_{s})$, where $\bm{\Phi}_{s}$ is the parameter set that defines this conditional probabilistic model.
 \end{itemize}
@@
 In the framework of HMM, at each time instance, signal $\bm{x}_t$ is generated by density $p(\bm{x}_t| s_t; \bm{\Phi}_{s_t})$, and $s_t$ is decided by the hidden markov process.
-This process gives us the probabilistic model $p(\ubm{x};\bm{H})$
-\subsection{Emission Model of GenHMM}
+This process gives us the probabilistic model $p(\ubm{x};\bm{H})$.
+\subsection{Generative Model of GenHMM}
 
@@
 \end{equation}
-where $\kappa$ is a random variable following a categorical distribution, and $\pi_{s, \kappa} = p(\kappa | s; \bm{\Phi}_{s, \kappa})$.
+where $\kappa$ is a random variable following a categorical distribution, and $\pi_{s, \kappa} = p(\kappa | s; \bm{H})$.
 Naturally $\sum_{\kappa = 1}^{K} \pi_{s, \kappa}= 1$.
@@
 $\bm{\theta}_s = \left\{ \bm{\theta}_{s, \kappa}| \kappa = 1, 2, \cdots, K \right\}$. 
-We define $p(\bm{x}| s, \kappa; \bm{\theta}_{s, \kappa})$ as induced distribution by a generator $\bm{g}_{s,\kappa}: \RR^{N}\rightarrow\RR^{N}$, such that $\bm{x}=\bm{g}_{s, \kappa}(\bm{z})$, where $\bm{z}\sim p_{s,\kappa}(\bm{z})$.
+We define $p(\bm{x}| s, \kappa; \bm{\theta}_{s, \kappa})$ as induced distribution by a generator $\bm{g}_{s,\kappa}: \RR^{N}\rightarrow\RR^{N}$, such that $\bm{x}=\bm{g}_{s, \kappa}(\bm{z})$, where $\bm{z}$ is a latent variable following a distribution with density function $p_{s,\kappa}(\bm{z})$.
 By change of variable, we have
@@
 \begin{equation}
-  \umin{\bm{H}\in \Hh} KL(\hat{p}(\ubm{x})\| p(\ubm{x};\bm{H}))
+  \umin{\bm{H}\in \Hh} KL(\hat{p}(\ubm{x})\| p(\ubm{x};\bm{H})),
 \end{equation}
@@
 
-Since model $\bm{H}$ contains hidden sequential variable $\ubm{s}$ and $\ubm{\kappa}$, we can not directly solve the maximum likelihood problem in \ref{eq:ml-of-hmm}.
-We use expectation maximization (EM) to address the hidden variable problem by
+The first problem that we need to address is to deal with the hidden sequential variable $\ubm{s}$ and $\ubm{\kappa}$  in model $\bm{H}$.
+For a sequential observable variable $\ubm{x}$, $\ubm{s}$ is the hidden state sequence corresponding to $\ubm{x}$, and $\ubm{\kappa}$ is the hidden variable sequence representing the generator identity sequence that actually generates $\ubm{x}$.
+
+Since directly maximizing likelihood is not an option for our problem in \ref{eq:ml-of-hmm}, we address this problem in expectation maximization (EM) framework.
+This divides our problem into two steps: i) using the joint posterior of hidden variable sequences $\ubm{s}$ and $\ubm{\kappa}$ to obtain a ``expected likelihood'' of observable variable sequence $\ubm{x}$, i.e.
+the E-step; ii) maximizing the expected likelihood with regarding to (w.r.t.) the model $\bm{H}$, i.e.
+the M-step.
+We formulate these two steps as follows:
 \begin{itemize}
@@
   % \end{equation}
-  The ``expected likelihood'' function:
+  the expected likelihood function:
   \begin{equation}\label{eq:em-q-funciton}
@@
   where $\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \cdot\right]$ denotes the expectation operator by distribution $\hat{p}(\ubm{x})$ and $p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})$.
-\item M-step: the optimization step:
+\item M-step: the maximization step:
   \begin{equation}\label{eq:em-m-opt}
-    \umax{\bm{H}} \Qq(\bm{H}; \bm{H}^{\mathrm{old}})
+    \umax{\bm{H}} \Qq(\bm{H}; \bm{H}^{\mathrm{old}}).
   \end{equation}
@@
 
-The \eqref{eq:em-m-opt} can be reformulated as:
+The problem \eqref{eq:em-m-opt} can be reformulated as:
 \begin{align}\label{eq:m-step-subs}
   &\umax{\bm{H}} \Qq(\bm{H}; \bm{H}^{\mathrm{old}}) \nonumber \\
-  =&\umax{\bm{q}}\Qq(\bm{q}; \bm{H}^{\mathrm{old}}) + \umax{{A}}\Qq({A}; \bm{H}^{\mathrm{old}}) 
+  =&\umax{\bm{q}}\Qq(\bm{q}; \bm{H}^{\mathrm{old}}) + \umax{\bm{A}}\Qq(\bm{A}; \bm{H}^{\mathrm{old}}) 
      + \umax{\bm{\Phi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})
@@
   \Qq(\bm{q}; \bm{H}^{\mathrm{old}}) 
-  &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p({s}_{1})  \right] \nonumber\\
-  &= \EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p({s}_{1})  \right]\label{eq:init-distribution-update}\\
-  \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \sum_{t}\log\,p({s}_{t+1}|{s}_{t}; \bm{H}) \right] \label{eq:transition-update}\\
-  \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) &= \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p(\ubm{x}, \ubm{\kappa}| \ubm{s}; \bm{H}) \right]\label{eq:generative-model-update}
+  % &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p({s}_{1})  \right] \nonumber\\
+  &= \EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p({s}_{1};\bm{H})  \right], \label{eq:init-distribution-update}\\
+  \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \sum_{t}\log\,p({s}_{t+1}|{s}_{t}; \bm{H}) \right], \label{eq:transition-update}\\
+  \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) &= \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p(\ubm{x}, \ubm{\kappa}| \ubm{s}; \bm{H}) \right].
+\label{eq:generative-model-update}
 \end{align}
 
-We can see that the solution of $H$ depends on the posterior probability $p(\ubm{s}| \ubm{x}; \bm{H})$.
-Though the evaluation of posterior according to Bayesian theorem is simple, the computation complexity of $p(\ubm{s}| \ubm{x}; \bm{H})$ grows exponentially with the length of $\ubm{s}$.
-Therefore, we would employ Forward/Backward algorithm \cite{} to do the posterior computation efficiently.
-The marginal $p(s_t| \ubm{x}; \bm{H})$ is also efficiently computed as the joint posterior.
+We can see that the solution of $\bm{H}$ depends on the posterior probability $p(\ubm{s}| \ubm{x}; \bm{H})$.
+Though the evaluation of posterior according to Bayesian theorem is straightforward, the computation complexity of $p(\ubm{s}| \ubm{x}; \bm{H})$ grows exponentially with the length of $\ubm{s}$.
+Therefore, we employ forward-backward algorithm \cite{Bishop:2006:PRM:1162264} to do the posterior computation efficiently.
+As we would detail in next section, it is actually $p(s| \ubm{x}; \bm{H})$ and $p(s, \kappa| \ubm{x}; \bm{H})$ that are needed to formulate the problem.
+For the joint posterior $p(s, \kappa| \ubm{x}; \bm{H})$, it can be computed by following the Bayesian rule when posterior of hidden state is available.
 
-With such a model GenHMM at hand, the remaining problems are:
+With such a solution framework ready for model GenHMM, there are still remaining problems to address before it can be employed for practical usage, including:
 \begin{itemize}
-\item How to realize GenHMM by neural networks to have high modeling capacity?
-\item How to train GenHMM to sovle problem in \eqref{eq:ml-of-hmm} using practical algorithm?
-\item Would the training of GenHMM converges?
+\item how to realize GenHMM by neural network based generators such that likelihoond of their induced distributions can be computed explicitly and exactly?
+\item how to train GenHMM to sovle problem in \eqref{eq:ml-of-hmm} using practical algorithm?
+\item would the training of GenHMM converges?
 \end{itemize}
-We would answer these questions in the following section.
+We would attempt these problems in the following section.
 
@@
 
-In this section, we provide answers to the questions at the end of last section.
+In this section, we detail the solution for realizing and learning GenHMM.
+% We begin with introducing flow model that we are going to use to model our generators $\{\bm{g}_{s, \kappa}| s\in\Ss, \kappa=1,2,\cdots, K\}$ for efficient likelihood computation.
+The learning method of HMM is followed then.
+The convergence analysis is discussed at the end of this section.
 
-\subsection{Modeling $\bm{g}_{s,\kappa}$ by a flow model }
-We use feed-forward neural network to implement every generator $\bm{g}_{s,\kappa}$.
-For notation simplicity, denote $\bm{g}_{s,\kappa}=\bm{g}_{s,\kappa}$.
-Define  a feed-forward neural network $\bm{g}_{s,\kappa}$ that has multiple hidden layers
+\subsection{Realizing $\bm{g}_{s,\kappa}$ by a Flow Model}
+We use a feed-forward neural network to implement each generator $\bm{g}_{s,\kappa}$.
+Define  $\bm{g}_{s,\kappa}$ as a $L$-layer neural network and we formulate its mapping by each layer's concatenation:
 $\bm{g}_{s,\kappa}=\bm{g}_{s,\kappa}^{[L]}\circ \bm{g}_{s,\kappa}^{[L-1]}\circ \cdots
-\circ \bm{g}_{s,\kappa}^{[1]}$ and is invertible $\bm{f}_{s,\kappa}=\bm{g}_{s,\kappa}^{-1}$, where $\circ$ denotes mapping concatenation. Then the signal flow can be depicted as
+\circ \bm{g}_{s,\kappa}^{[1]}$, where superscipt $[l]$ denotes the $l$'s layer index and $\circ$ denotes mapping concatenation.
+Assume $\bm{g}_{s,\kappa}$ is invertible and denotes its inverse mapping as $\bm{f}_{s,\kappa}=\bm{g}_{s,\kappa}^{-1}$.
+For a latent variable $\bm{z}$ with density function $p_{s,\kappa}(\bm{z})$, the generated signal $\bm{x}$ follows an induced distribution with density function \eqref{eq:changel-variable}.
+We illustrate the signal flow between latent variable $\bm{z}$ and observable variable $\bm{x}$ as follows:
 \begin{equation*}
@@
     \node (x) at (6,0) {};
-    \node at ($(x)+(0.5,0)$){${\bm{x}} = \bm{h}_L$};
+    \node at ($(x)+(0.5,0)$){${\bm{h}_L=\bm{x}}$};
     \draw[->] ($(z) + (0.3,0.1)$) -- node[above]{$\bm{g}_{s,\kappa}^{[1]}$} ($(xi1)+(-0.3,0.1)$); 
@@
 \end{equation*}
-where $\bm{g}_{s,\kappa}^{[l]}$ and $\bm{f}_{s,\kappa}^{[l]}$ are the $l$-th layer of $\bm{g}_{s,\kappa}$ and $\bm{f}_{s,\kappa}$, respectively.
-In a feed-forward neural network, if every layer is invertible,
-the full feed-forward neural network is invertible. The inverse
-function is given by $\bm{z}=\bm{f}_{s,\kappa}(\bm{x})$. Flow-based
-network, proposed in \cite{DBLP:journals/corr/DinhKB14}, is such a
-feed-forward neural network, which is further improved in subsequent works \cite{2016arXiv160508803D,kingma2018glow}.
-It also has additional advantages as efficient Jacobian computation and low computational complexity.
+where $\bm{f}_{s,\kappa}^{[l]}$ is the $l$-th layer of $\bm{f}_{s,\kappa}$. We have $\bm{z}=\bm{f}_{s,\kappa}(\bm{x})$. If every layer of $\bm{g}_{s,\kappa}$ is invertible,
+the full feed-forward neural network is invertible. Flow model, proposed in \cite{DBLP:journals/corr/DinhKB14} as an image generating model, is such a invertible
+feed-forward neural network.
+It is further improved in subsquential works \cite{2016arXiv160508803D,kingma2018glow} for high-fidelity and high-resolution image generating and representation.
+Flow model also has advantages as efficient Jacobian computation and low computational complexity.
 %Computation of inverse mapping for flow-based network and determinant of Jacobin were addressed in the literature.
 
-For a flow-based neural network architecture, let us assume that
+For a flow model, let us assume that
 the feature $\bm{h}_l$ at the $l$'th layer has two subparts as
-$\bm{h}_l = [\bm{h}_{l,a}^{T} \, , \, \bm{h}_{l,b}^{T}]^{T}$ where
-$(\cdot)^{T}$ denotes transpose operation. Then considering $\bm{h}_0 = \bm{z}$, we have the following forward and inverse relations between $(l-1)$'th and $l$'th layers:
+$\bm{h}_l = [\bm{h}_{l,a}^{\intercal} \, , \, \bm{h}_{l,b}^{\intercal}]^{\intercal}$.
+The efficient invertible mapping of flow model comes from following forward and inverse relations between $(l-1)$'th and $l$'th layers:
 \begin{equation}\label{eq-gl}
@@
 where $\bm{I}_a$ is identity matrix and $\mathrm{diag}(\cdot)$ returns a square matrix with the elements of $(\cdot)$ on the main diagnal.
-\textcolor{red}{do I still need this?}
-Then the pdf is
-\begin{align}
-  p(\bm{x}) & =  p(\bm{z}) \big| \mathrm{det}(\bm{J}) |_{\bm{z}={\bm{f}}({\bm{x}})}\big| \nonumber\\
-            &  = p(\bm{z}) \prod_{l=1}^L \abs{\det\left( \mathrm{diag}(\bm{m}_a(\bm{h}_{l,a}))]  \right)}.
-\end{align}
-[] describes a \textit{coupling} mapping between layers.
-Since the coupling has a partial identity mapping, direct concatenation of multiple such coupling mappings would result in a partial identity mapping of the whole neural network $\tilde{\bm{g}}$.
+% \textcolor{red}{do I still need this?}
+% Then the pdf is
+% \begin{align}
+%   p(\bm{x}) & =  p(\bm{z}) \big| \mathrm{det}(\bm{J}) |_{\bm{z}={\bm{f}}({\bm{x}})}\big| \nonumber\\
+%             &  = p(\bm{z}) \prod_{l=1}^L \abs{\det\left( \mathrm{diag}(\bm{m}_a(\bm{h}_{l,a}))]  \right)}.
+                %   \end{align}
+
+\eqref{eq-gl} describes a \textit{coupling} layer in a flow model.
+A flow model is basically a stack of multiple coupling layers.
+But the issue of direct concatenation of multiple such coupling mappings is partial identity mapping of the whole model.
+This issue can be addressed by alternating hidden signal order after each coupling layer.
+
+% Since the coupling has a partial identity mapping, direct concatenation of multiple such coupling mappings would result in a partial identity mapping of the whole neural network $\tilde{\bm{g}}$.
 Alternating the positions of identity mapping \cite{2016arXiv160508803D} or using $1\times1$ convolution operations \cite{kingma2018glow} before each coupling mapping is used to treat the issue.
@@
 % concatenating multiple layers together were proposed to avoid the situation that a part of the output is equal to a part of the input.} %In addition, applying affine mapping does not affect its advantage of trivial inverse and determinate computation.
-Furthermore, \cite{2016arXiv160508803D}\cite{kingma2018glow} split some hidden
-layer signal $\bm{h}$ and model a part of it directly as standard Gaussian to reduce computation and memory burden.
+% Furthermore, \cite{2016arXiv160508803D}\cite{kingma2018glow} split some hidden
+% layer signal $\bm{h}$ and model a part of it directly as standard Gaussian to reduce computation and memory burden.
 % % \textcolor{blue}{Therefore the dimension
@@
 \subsection{Learning of GenHMM}\label{subsec:optmGenHMM}
-
-\subsubsection{Generative Model Update}
-We use notation $\bm{\Pi} = \left\{  \bm{\pi}_{s}| s\in \Ss \right\}$, $\bm{\Theta}=\left\{ \bm{\theta}_s| s\in \Ss \right\}$.
-Then the third subproblem, \eqref{eq:m-step-subs}, becomes:
+In this subsection, we address the problem of learning GenHMM.
+\subsubsection{Generative Model Learning}
+The generative model learning is actually to solve the problem in \eqref{eq:generative-model-update}, which can be further divided into two subproblems: i) generator learning; ii) weights of generators learning.
+Let us define notations:
+$\bm{\Pi} = \left\{  \bm{\pi}_{s}| s\in \Ss \right\}$, $\bm{\Theta}=\left\{ \bm{\theta}_s| s\in \Ss \right\}$. 
+Then the problem in \eqref{eq:generative-model-update} becomes:
 \begin{align}\label{eq:sub-gm}
@@
 \begin{align}
-  &\Qq(\bm{\Pi}; \bm{H}^{\mathrm{old}})  =\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[  \log\,p(\ubm{\kappa}| \ubm{s}; \bm{H})\right] \nonumber \\
-  &\Qq(\bm{\Theta}; \bm{H}^{\mathrm{old}}) =\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[  \log\,p(\ubm{x}| \ubm{s},\ubm{\kappa}; \bm{H})\right] 
+  \Qq(\bm{\Pi}; \bm{H}^{\mathrm{old}})  &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[  \log\,p(\ubm{\kappa}| \ubm{s}; \bm{H})\right], \\
+  \Qq(\bm{\Theta}; \bm{H}^{\mathrm{old}}) &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[  \log\,p(\ubm{x}| \ubm{s},\ubm{\kappa}; \bm{H})\right].
 \end{align}
 
-The neural network loss can be thus formulated as
+We firstly address the generator learning problem, i.e.
+$\umax{\bm{\Theta}} \Qq(\bm{\Theta}; \bm{H}^{\mathrm{old}})$.
+This is boiled down to maximizing the cost function of neural networks that can be formulated as
 \begin{align}\label{eq:obj-q-gen-mix}
@@
   =& \frac{1}{R}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})p(\smtr{\kappa}{t}{r}|\smtr{s}{t}{r}, \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \nonumber\\
-  &  \log\, p(\bmtr{x}{t}{r} | \smtr{s}{t}{r}, \smtr{\kappa}{t}{r}; \bm{H}) 
+  &  \log\, p(\bmtr{x}{t}{r} | \smtr{s}{t}{r}, \smtr{\kappa}{t}{r}; \bm{H}). 
 \end{align}
-
-Here, $p(s_t| \ubm{x}, \bm{H}^{\mathrm{old}})$ is computed by forward/backward algorithm. The posterior of $\kappa$ is:
+In \eqref{eq:obj-q-gen-mix}, the state posterior $p(s_t| \ubm{x}, \bm{H}^{\mathrm{old}})$ is computed by forward-backward algorithm. The posterior of $\kappa$ is:
 \begin{align}\label{eq:kappa-posterior}
@@
   &=  \frac{p(\kappa, \ubm{x}| s; \bm{H}^{\mathrm{old}})}{p(\ubm{x}| s,\bm{H}^{\mathrm{old}})} \nonumber \\
-  & = \frac{\pi_{s, \kappa}^{\mathrm{old}} p(\bm{x}| s, \kappa, \bm{H}^{\mathrm{old}})}{\sum_{\kappa=1}^{K}  \pi_{s, \kappa}^{\mathrm{old}} p(\bm{x}| s, \kappa,\bm{H}^{\mathrm{old}})} 
+  & = \frac{\pi_{s, \kappa}^{\mathrm{old}} p(\bm{x}| s, \kappa, \bm{H}^{\mathrm{old}})}{\sum_{\kappa=1}^{K}  \pi_{s, \kappa}^{\mathrm{old}} p(\bm{x}| s, \kappa,\bm{H}^{\mathrm{old}})},
 \end{align}
-where the last equation is due to the fact that only $\bm{x}_t$ among sequence $\ubm{x}$ depends on $s_t, \kappa_t$.
+where the last equation is due to the fact that $\bm{x}_t$ among sequence $\ubm{x}$ depends on $s_t, \kappa_t$ only.
 
-By substituting \eqref{eq:changel-variable} and \eqref{eq:cat-jacobian} into \eqref{eq:obj-q-gen-mix}, we have loss function for neural networks as
+By substituting \eqref{eq:changel-variable} and \eqref{eq:cat-jacobian} into \eqref{eq:obj-q-gen-mix}, we have cost function for neural networks as
 \begin{align}\label{eq:obj-q-gen-mix}
@@
   =& \frac{1}{R}\hspace{-3pt}\sum_{r=1}^{R}\hspace{-3pt} \sum_{t=1}^{{T}^{r}-1}\hspace{-3pt} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|} \hspace{-3pt} \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})p(\smtr{\kappa}{t}{r}|\smtr{s}{t}{r}, \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \nonumber\\
-  &\left[ \log\, p_{\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}}(\bm{f}_{\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}}(\bmtr{x}{t}{r})) + \sum_{l=1}^{L}\log\,| \det (\nabla{\bm{f}_{s,\kappa}^{[l]}})|\right]
+  &\left[ \log\, p_{\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}}(\bm{f}_{\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}}(\bmtr{x}{t}{r})) + \sum_{l=1}^{L}\log\,| \det (\nabla{\bm{f}_{s,\kappa}^{[l]}})|\right].
 \end{align}
+The generators of GenHMM can simply use standard Gaussian distribution for latent variables $\bm{z} \sim p_{s,\kappa}(\bm{z})$.
+It is straightforward to employ batch-size stochastic gradient decent to maximize $\Qq(\bm{\Theta; \bm{H}^{\mathrm{old}}})$ w.r.t.
+parameters of generators, i.e.
+$\bm{\Phi}$, since the dataset represented by $\hat{p}(\ubm{x})$ can be tool large to do whole-dataset iterations.
 
-In implementation, the generators of GenHMM uses standard Gaussian distribution for variables $\bm{z} = \bm{f}_{s,\kappa}(\bm{x})$, i.e. $p_{s,\kappa}(\bm{z})$ used standard Gaussian density function.
-
-------------------------------
-
-The latent prior for mixture of each source $s$ is obtained by solving the following problem:
+In what follows we address the problem $\max_{\bm{\Pi}} \Qq(\bm{\Pi}; \bm{H}^{\mathrm{old}})$ in our generative model learning.
+The conditional distribution of hidden variable $\kappa$, $\pi_{s, \kappa} = p(\kappa | s; \bm{H})$, is obtained by solving the following problem:
 \begin{align}\label{opm:pi}
@@
 where $p(s, \kappa | \ubm{x}; \bm{H}^{\mathrm{old}}) = p(s| \ubm{x}; \bm{H}^{\mathrm{old}}) p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$.
-$p(s| \ubm{x}; \bm{H}^{\mathrm{old}})$ that can be computed by results of forward/backward in Viterbi algorithm, while $p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$.
-$p(s| \ubm{x}; \bm{H}^{\mathrm{old}})$ is given by \eqref{eq:kappa-posterior}.
+$p(s| \ubm{x}; \bm{H}^{\mathrm{old}})$ that can be computed by results of forward-backward algorithm, while $p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$ is given by \eqref{eq:kappa-posterior}.
 
------------------------
+With the generative modeling learning obtained, it remains to solve the initial distribution update and transition matrix update of HMM in GenHMM, i.e.
+the problem \eqref{eq:init-distribution-update} and \eqref{eq:transition-update}.
+These two problem are basically two constrained optimization problems.
+The solution to them is provided in \cite{Bishop:2006:PRM:1162264}.
+But to keep learning algorithm for GenHMM complete, we give the update rules for $\bm{q}$ and $\bm{A}$ as follows.
 
 \subsubsection{Initial Probability Update}
-\eqref{eq:init-distribution-update} can be written as:
+Problem in \eqref{eq:init-distribution-update} can be reformulated as:
 \begin{align}
-  &\Qq(\bm{q}; \bm{H}^{\mathrm{old}}) \nonumber\\
-  &=\frac{1}{R} \sum_{r=1}^{R}\sum_{\ubmr{s}{r}} {p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r}) \nonumber \\
+  \Qq(\bm{q}; \bm{H}^{\mathrm{old}})
+  &=\frac{1}{R} \sum_{r=1}^{R}\sum_{\ubmr{s}{r}} {p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r};\bm{H}) \nonumber \\
   % & = \frac{1}{R}\sum_{r=1}^{R}\sum_{\smtr{s}{1}{r}=1}^{|\Ss|}\sum_{\smtr{s}{2}{r}=1}^{|\Ss|}\cdots \sum_{\smtr{s}{T^{r}}{r}}^{{|\Ss|}} {p(\smtr{s}{1}{r}, \smtr{s}{2}{r}, \cdots, \smtr{s}{T^{r}}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r}) \\
-  & = \frac{1}{R}\sum_{r=1}^{R}\sum_{\smtr{s}{1}{r}=1}^{|\Ss|}{p(\smtr{s}{1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r}) 
+  & = \frac{1}{R}\sum_{r=1}^{R}\sum_{\smtr{s}{1}{r}=1}^{|\Ss|}{p(\smtr{s}{1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r};\bm{H}) 
 \end{align}
 
-Since $p(\smtr{s}{1}{r})$ is the probability of initial state of GenHMM $\bm{H}$ for $r$-th sequential sample, actually $q_i = p({s}_{1} =i) $, $i= 1, 2, \cdots, |\Ss|$ for $\bm{H}$.
-Solution to problem:
+$p(\smtr{s}{1}{r};\bm{H})$ is the probability of initial state of GenHMM for $r$-th sequential sample. Actually $q_i = p({s}_{1} =i) $, $i= 1, 2, \cdots, |\Ss|$ for $\bm{H}$. Solution to problem:
 \begin{align}
@@
 \subsubsection{Transition Probability Update}
-\eqref{eq:transition-update} can be written as
+Problem \eqref{eq:transition-update} can be reformulated as
 \begin{align}
@@
   % &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})\right] \nonumber\\
-  &= \sum_{r=1}^{R} \sum_{\ubmr{s}{r}}{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \sum_{t=1}^{T^{(r)}-1}\log\,p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A}) \nonumber \\
-  &= \sum_{r=1}^{R} \sum_{t=1}^{T^{(r)}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}\sum_{\smtr{s}{t+1}{r}=1}^{|\Ss|}{p(\smtr{s}{t}{r}, \smtr{s}{t+1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})
+  &= \sum_{r=1}^{R} \sum_{\ubmr{s}{r}}{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \sum_{t=1}^{T^{(r)}-1}\log\,p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; \bm{H}) \nonumber \\
+  &= \sum_{r=1}^{R} \hspace{-0.1cm}\sum_{t=1}^{T^{r}-1}\hspace{-0.1cm} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}\hspace{-0.05cm}\sum_{\smtr{s}{t+1}{r}=1}^{|\Ss|}\hspace{-0.2cm}{p(\smtr{s}{t}{r}, \smtr{s}{t+1}{r}| \ubmr{x}{r};\hspace{-0.05cm} \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; \bm{H})
 \end{align}
 
-Since $\bm{A}_{i, j}  = p(\smtr{s}{t+1}{r}=j|\smtr{s}{t}{r}=i; {A})$ where $A_{i, j}$ is the element of transition matrix $A$, the solution to problem:
+Since $\bm{A}_{i, j}  = p(\smtr{s}{t+1}{r}=j|\smtr{s}{t}{r}=i; \bm{H})$ is the element of transition matrix $\bm{A}$, the solution to problem:
 \begin{align}\label{eq:update-transition-prob}
@@
   \mathrm{s.t.} &\hspace{0.2cm} \bm{A} \cdot \bm{1} = \bm{1} \nonumber \\
-           &  \bm{A}^{\intercal} \cdot \bm{1} = \bm{1} \nonumber \\
            & \bm{A}_{i,j} \geq 0.
@@
 \begin{equation}\label{eq:update-transition-solt}
-  \bar{\xi}_{i,j} = \sum_{r= 1}^{R} \sum_{t= 1}^{T^{(r)}-1}{p(\smtr{s}{t}{r}=i, \smtr{s}{t+1}{r}=j| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}
+  \bar{\xi}_{i,j} = \sum_{r= 1}^{R} \sum_{t= 1}^{T^{(r)}-1}{p(\smtr{s}{t}{r}=i, \smtr{s}{t+1}{r}=j| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}.
 \end{equation}
 
-\subsection{Algorithm of GenHMM}
-We goal is to minimize the KL divergence $KL(\hat{p}(\ubm{x})\| p(\ubm{x};\bm{H}))$ or maximizing$\sum_{r=1}^{R}\log\,p(\ubmr{x}{r}; \bm{H})$. The 
-Algorithm here:
-things to talk:
-\begin{itemize}
-\item separate optimization of different sets of parameters
-\item batch optimization optimization
-\item not going to optimal for $\bm{\Phi}$, leaving to convergence analysis in next section
-\end{itemize}
+\subsection{Algorithm of GenHMM and Convergence Analysis}
 
-We summarize the optimization algorithm as:
+To summarize the learning solution in subsection~\ref{subsec:optmGenHMM}, we wrap our algorithm into pseudocode as shown in Algorithm~\ref{algo:genhmm}.
+We use Adam \cite{DBLP:journals/corr/KingmaB14} optimizer for optimization w.r.t.
+the parameters of generators in GenHMM.
+As is shown from line $6$ to $12$ in Algorithm~\ref{algo:genhmm}, the batch-size stochastic gradient decent can be naturally embedded into the learning algorithm of GenHMM.
+
+As described by the pseudocode in Algorithm~\ref{algo:genhmm}, the learning of GenHMM is divided into optimization w.r.t.
+to generators' parameters $\bm{\Theta}$, initial probability $\bm{q}$ of hidden state, transition matrix $\bm{A}$, and generator mixture weights $\bm{\Pi}$.
+Different from the optimization w.r.t.
+to $\bm{q}$, $\bm{A}$ and $\bm{\Pi}$, which have optimal solutions, generator learning usually cannot give optimal solution to problem $\max_{\bm{\Theta}} \Qq(\bm{\Theta}; \bm{H}^{\mathrm{old}})$.
+In fact, given that no optimal $\bm{\Theta}$ is obtained in each iteration in Algorithm~\ref{algo:genhmm}, learning of GenHMM can still converge.
+
+In pursuit of representing a dataset by GenHMM, we are actually minimize a KL divergence $KL(\hat{p}(\ubm{x})\| p(\ubm{x};\bm{H}))$.
+Our following analysis shows that our proposed algorithm for learning GenHMM can converge, without requirement for optimal $\bm{\Theta}$ in each iteration.
 
 \begin{algorithm}[H]
-  \caption{Learning of GenHMM}
+  \caption{Learning of GenHMM}\label{algo:genhmm}
   \begin{algorithmic}[1]
     \STATE {\bfseries Input:}{
-      Building $\bm{H}^{\mathrm{old}}, \bm{H} \in \Hh$ gives: \\
+      Initializing $\bm{H}^{\mathrm{old}}, \bm{H} \in \Hh$ gives: \\
       $\bm{H}^{\mathrm{old}} = \{\Ss, \bm{q}^{\mathrm{old}}, A^{\mathrm{old}}, p(\bm{x}|s; \bm{\Phi}^{\mathrm{old}})\}$, $\bm{H} = \{\Ss, \bm{q}, A, p(\bm{x}|s; \bm{\Phi})\}$\\
@@
     \STATE $\bm{H}^{\mathrm{old}} \gets \bm{H}$
-    \STATE Set learning rate $\eta$, neural network optimization epochs $T$ per EM step.
+    \STATE Set learning rate $\eta$, neural network optimization epochs $N$ per EM step.
     \FOR { $\bm{H}$ not converge}
-    \FOR {epoch $t < T$}
+    \FOR {epoch $n < N$}
     \STATE Sample a batch of data $\left\{ \ubmr{x}{r} \right\}_{r=1}^{R_b}$ from dataset $\hat{p}(\ubm{x})$
@@
     \STATE $\partial{\bm{\Theta}} \gets  \nabla_{\bm{\Theta}} {\Qq}\left({\bm{\Theta}},{\bm{H}}^{\mathrm{old}}\right)$
-    \STATE $\bm{\Theta} \gets \bm{\Theta} + \eta \cdot \partial{\bm{\theta}_s}$
+    \STATE $\bm{\Theta} \gets \bm{\Theta} + \eta \cdot \partial{\bm{\Theta}}$
     \ENDFOR
     \STATE $\bm{q} \gets \uargmax{\bm{q}}\, \Qq(\bm{q}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:update-initial-state-prob};
-    \STATE $\bm{A} \gets \uargmax{\bm{A}}\Qq(\bm{A}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:update-transition-prob};
-    \STATE $\bm{\Pi} \gets \uargmax{\bm{\Pi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:update-transition-solt};
+    \STATE $\bm{A} \gets \uargmax{\bm{A}}\Qq(\bm{A}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:update-transition-solt};
+    \STATE $\bm{\Pi} \gets \uargmax{\bm{\Pi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:mix-latent-parameter-solution};
     \STATE $\bm{H}^{\mathrm{old}} \gets \bm{H}$
-
     \ENDFOR
@@
 
-\subsection{Convergence Analysis}
 \begin{prop}
   Assume that parameters of GenHMM are in a compact set.
-Further assume that $\Qq(\bm{\Theta}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) \geq \Qq(\bm{\Theta}^{\mathrm{old}}; \bm{H}^{\mathrm{old}})$, then we always have $KL(\hat{p}(\ubm{x})\| p(\ubm{x};\bm{H}^{\mathrm{new}})) \geq KL(\hat{p}(\ubm{x})\| p(\ubm{x};\bm{H}^{\mathrm{old}}))$, and GenHMM converges.
+Further assume that $\Qq(\bm{\Theta}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) \geq \Qq(\bm{\Theta}^{\mathrm{old}}; \bm{H}^{\mathrm{old}})$, then we always have $KL(\hat{p}(\ubm{x})\| p(\ubm{x};\bm{H}^{\mathrm{new}})) \leq KL(\hat{p}(\ubm{x})\| p(\ubm{x};\bm{H}^{\mathrm{old}}))$, and GenHMM converges.
 \end{prop}
@@
     GMM-HMM & $62.30\%$ &  $68.06\%$ &  $68.73\%$  \\ \midrule
-    GenHMM & $76.78\%$ & $77.68\%$ &   \\ 
+    GenHMM & $76.78\%$ & $77.77\%$ &  $\%$ \\ 
     \bottomrule                                                                  
@@
 We firstly show the phoneme classification using 39 dimensional MFCC features (MFCC coefficients, deltas, and delta-deltas), to validate one possible usage of our proposed model.
-GMM-HMM is trained and tested as a reference model in our experiments.
+Since generative training is carried out in our experiments, GMM-HMM is trained and tested as a reference model in our experiments.
 Training and testing of GMM-HMM is in the same condition as GenHMMs are trained and tested.
@@
     GMM-HMM & $48.59\%$ &  $51.26\%$ &  $52.40\%$  \\ \midrule
-    GenHMM & $\%$ & $\%$ &   \\ 
+    GenHMM & $61.10\%$ & $\%$ &   \\ 
     \bottomrule                                                                  
@@
     GMM-HMM & $37.15\%$ &  $40.65\%$ & $42.27\%$  \\ \midrule
-    GenHMM & $\%$ & $\%$ & $\%$   \\ 
+    GenHMM & $50.33\%$ & $\%$ & $\%$   \\ 
     \bottomrule
@@
 \section{Conclusion}
-
-\newpage
+In this work, we proposed a generative model based HMM (GenHMM) whose generators are realized by neural networks.
+We provide the training method for GenHMM.
+The validity of GenHMM is demonstrated by the experiments of classification task on practical sequential dataset.
+The learning method in this work is based on generative training.
+Future work would consider discriminative training for classification tasks of sequential data.
 

### 2019-08-23 manuscript/main.tex
minor revision and complete tables

@@
 The proposed model is termed as GenHMM.
-In GenHMM, the hidden states of HMM and input data are related by neural network based generative model that provides exact likelihood computation.
+In GenHMM, the hidden states of HMM and input data are related by a neural network based generative model that has tractability of exact likelihood and provides efficient likelihood computation.
 This is realized by a flow model.
-Apart from modeling itself, we also give learning algorithm for GenHMM in expectation-maximization framework.
-The convergence of learning GenHMM is analyzed.
+A learning algorithm for GenHMM is proposed in expectation-maximization framework.
+The convergence of the learning GenHMM is analyzed.
 We demonstrate the efficiency of GenHMM by classification tasks on practical sequential data.
@@
 \section{introduction}
-Sequential data modeling is a challenging topic in pattern recognition and machine learning, since for many applications, the assumption of independent and identically distributed data points is too strong to model data properly.
+Sequential data modeling is a challenging topic in pattern recognition and machine learning.
+For many applications, the assumption of independent and identically distributed data points is too strong to model data properly.
 Hidden Markov model (HMM) is a classic way to model sequential data.
@@
 
-HMM provides an framework for sequential data modeling.
+HMM provides a framework for sequential data modeling.
 To connect states of HMM with input data, a probabilistic model is used to describe how each state of HMM should represent a pattern of sequential data.
@@
 A hierarchical training is carried out by: i) training a HMM first, ii) and then doing modified training of LSTM using trained HMM.
-This training procedure is argued by the authors to use LSTM or RNN to fill in the gap where HMM can not learn.
+This hierarchical training procedure is motivated by the intuition of using LSTM or RNN to fill in the gap where HMM can not learn.
 
@@
 
-The above work help improve modeling capacity of HMM based models by bringing in neural networks.
+The above works help improve modeling capacity of HMM based models by bringing in neural networks.
 A softmax layer is usually used to represent probability whenever a conditional distribution is needed.
 These hierarchical schemes are built based on intuition of domain knowledge.
-Training of these hierarchical models usually need specific experience in relevant area to be able to proceed with the hierarchical procedure of training.
+Training of these hierarchical models usually need specific expertise in relevant areas to be able to proceed with the hierarchical procedure of training and application usage.
 
@@
   \item firstly has high modeling capacity of sequential data, due to the neural network based generators;
-\item has advantage of easy to do training, since GenHMM training employs expectation maximization (EM) framework.
+\item has advantage of easy to do training.
+Training of GenHMM employs expectation maximization (EM) framework.
 Training a GenHMM is as easy as training a GMM-HMM model, while configuration of GenHMM is flexible.
@@
 \end{itemize}
-Instead of using softmax for probability representation, our GenHMM is able to compute exact loglikelihood of given sequence data based on change of variable formula.
+Instead of using softmax for probability representation, our GenHMM has tractability of exact loglikelihood of given sequential data, which is based on change of variable formula.
 To make the loglikelihood computation efficient, neural network based generators of GenHMM are implemented as flow models.
@@
 \item Designing practical algorithm for training GenHMM under EM framework. Stochastic gradient  search in batch fashion is embedded in this algorithm.
-\item Convergence analysis for GenHMM under the proposed learning algorithm.
+\item Giving convergence analysis for GenHMM under the proposed learning algorithm.
 \end{itemize}
@@
 A HMM model $\bm{H}$ defined in a hypothesis space $\Hh$, i.e.
-$\bm{H} \in \Hh$, is capable to model time-span signal $\ubar{\bm{x}} = \left[ \bm{x}_1, \cdots, \bm{x}_T\right]^{\intercal}$, where $\bm{x}_t\in \RR^{N}$ is the signal at time $t$, $[\cdot]^{\intercal}$ means transpose, and $T$ denote the time length.
-Using HMM for signal representation or generating is illustrated in \ref{fig:hmm}.
-The assumption is that different instant signal of $\ubar{\bm{x}}$ is generated by different source, while the source selection is done by a discrete hidden markov process.
+$\bm{H} \in \Hh$, is capable to model time-span signal $\ubar{\bm{x}} = \left[ \bm{x}_1, \cdots, \bm{x}_T\right]^{\intercal}$, where $\bm{x}_t\in \RR^{N}$ is the signal at time $t$, $[\cdot]^{\intercal}$ denotes transpose, and $T$ denotes the time length.
+Using HMM for signal representation is illustrated in Figure~\ref{fig:hmm}.
+The model assumption is that different instant signal of $\ubar{\bm{x}}$ is generated by a different signal source associated with a hidden state of HMM, while the source selection is done by a discrete hidden markov process.
 We define the hypothesis set of HMM as $\Hh := \{\bm{H} | \{\Ss, \bm{q}, \bm{A}, p(\bm{x}|{s}; \bm{\Phi}_{s})\}$, where:
 \begin{itemize}
-\item $\Ss$ is the set of states of HMM $\bm{H}$;
-\item $\bm{q} = \left[ q_1, q_2, \cdots, q_{|\Ss|}\right]^\intercal$ initial distribution of HMM $\bm{H}$ with $|\Ss|$ is cardinality of $\Ss$, $q_s = p(s;\bm{H})$ for random state variable $s$.
+\item $\Ss$ is the set of hidden states of HMM $\bm{H}$;
+\item $\bm{q} = \left[ q_1, q_2, \cdots, q_{|\Ss|}\right]^\intercal$ is initial distribution of HMM $\bm{H}$ with $|\Ss|$ as cardinality of $\Ss$.
+For a state $s$, $q_s = p(s;\bm{H})$.
 In the following of this paper, we use $s_t$ to denote the state $s$ at time $t$.
-\item $\bm{A}$ is the transition matrix for the HMM $\bm{H}$ of size $|\Ss| \times |\Ss|$, $\bm{A}_{i,j} = p(s_{t+1}=j|s_{t}=i; \bm{H})$.
-\item For given hidden state $s$, the observable signal density function is $p({\bm{x}}|{s};\bm{\Phi}_{s})$, where $\bm{\Phi}_{s}$ is the parameter set that defines this conditional probabilistic model.
+\item $\bm{A}$ is the transition matrix for the HMM $\bm{H}$ of size $|\Ss| \times |\Ss|$ and $\bm{A}_{i,j} = p(s_{t+1}=j|s_{t}=i; \bm{H})$.
+\item For given hidden state $s$, the density function of observable signal is $p({\bm{x}}|{s};\bm{\Phi}_{s})$, where $\bm{\Phi}_{s}$ is the parameter set that defines this probabilistic model.
 \end{itemize}
 
-In the framework of HMM, at each time instance, signal $\bm{x}_t$ is generated by density $p(\bm{x}_t| s_t; \bm{\Phi}_{s_t})$, and $s_t$ is decided by the hidden markov process.
-This process gives us the probabilistic model $p(\ubm{x};\bm{H})$.
+In the framework of HMM, at each time instance, signal $\bm{x}_t$ is assumed to be generated by a distribution with density function $p(\bm{x}_t| s_t; \bm{\Phi}_{s_t})$, and $s_t$ is decided by the hidden markov process.
+This gives us the probabilistic model $p(\ubm{x};\bm{H})$.
 \subsection{Generative Model of GenHMM}
@@
 We term it as GenHMM.
-GenHMM has a hidden markov process as typical HMM, while the probabilistic model of GenHMM for each hidden state is induced by mixture of neural network based generators.
+GenHMM has a hidden markov process as a typical HMM does, while the probabilistic model of GenHMM for each hidden state is induced by mixture of neural network based generators.
 For given state $s\in \Ss$, we define its probabilistic model in GenHMM as
@@
 \end{equation}
-where $\kappa$ is a random variable following a categorical distribution, and $\pi_{s, \kappa} = p(\kappa | s; \bm{H})$.
+where $\kappa$ is a random variable following a categorical distribution, with probability $\pi_{s, \kappa} = p(\kappa | s; \bm{H})$.
 Naturally $\sum_{\kappa = 1}^{K} \pi_{s, \kappa}= 1$.
-Denote $\bm{\pi}_{s} = [\pi_{s,1}, \pi_{s,2}, \cdots, \pi_{s,K}]$,
+Denote $\bm{\pi}_{s} = [\pi_{s,1}, \pi_{s,2}, \cdots, \pi_{s,K}]^{\intercal}$,
 $\bm{\theta}_s = \left\{ \bm{\theta}_{s, \kappa}| \kappa = 1, 2, \cdots, K \right\}$. 
@@
   \begin{array}{l}
+    \bm{h}_{l} =
+    \begin{bmatrix}
+      \bm{h}_{l,a}\\
+      \bm{h}_{l,b}
+    \end{bmatrix}
+    =
+    \begin{bmatrix}
+      \bm{h}_{l-1,a}\\
+      \left(  \bm{h}_{l-1,b} - \bm{m}_b(\bm{h}_{l-1,a}) \right)\oslash \bm{m}_a(\bm{h}_{l-1,a}) 
+    \end{bmatrix}, \vspace{10pt} \\
         \bm{h}_{l-1} =
@@
       \bm{m}_a(\bm{h}_{l,a})\odot \bm{h}_{l,b} + \bm{m}_b(\bm{h}_{l,a})
-    \end{bmatrix},\vspace{10pt}\\
-    \bm{h}_{l} =
-    \begin{bmatrix}
-      \bm{h}_{l,a}\\
-      \bm{h}_{l,b}
-    \end{bmatrix}
-    =
-    \begin{bmatrix}
-      \bm{h}_{l-1,a}\\
-      \left(  \bm{h}_{l-1,b} - \bm{m}_b(\bm{h}_{l-1,a}) \right)\oslash \bm{m}_a(\bm{h}_{l-1,a}) 
-    \end{bmatrix}, \\
+    \end{bmatrix},\\
   \end{array}  
@@
   \end{align*}
-  For the learning w.r.t.
-neural network parameter set $\bm{\Theta}$, as long as the lose function does not decrease during EM iterations, i.e.
-$\Qq(\bm{\Theta}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) \geq \Qq(\bm{\Theta}^{\mathrm{old}}; \bm{H}^{\mathrm{old}})$, we would have
+  For the learning w.r.t. neural network parameter set $\bm{\Theta}$, as long as the lose function does not decrease during EM iterations, we would have
   \begin{equation*}
@@
 To show the validity of our model, we implement our model in PyTorch and test it with sequential data.
-The code is available at github repository\footnote{\textcolor{blue}{link}}.
+The code is available at github repository\footnote{Due to double-blind review, the author's repository is not present here temporarily.}.
 In the next subsection, we first discuss the experimental setups and the show the experimental results.
@@
     {Model} &  K=1 &  K=3 &  K=5  \\  \midrule
-    GMM-HMM & $62.30\%$ &  $68.06\%$ &  $68.73\%$  \\ \midrule
-    GenHMM & $76.78\%$ & $77.77\%$ &  $\%$ \\ 
+    GMM-HMM & $62.3\%$ &  $68.0\%$ &  $68.7\%$  \\ \midrule
+    GenHMM & $76.7\%$ & $77.7\%$ &  $77.7\%$ \\ 
     \bottomrule                                                                  
@@
     \\ \midrule
-    GMM-HMM & $53.64\%$ &  $59.65\%$ & $61.98\%$  \\ \midrule
-    GenHMM & $69.51\%$ & $70.67\%$ & $70.78\%$   \\ 
+    GMM-HMM & $53.6\%$ &  $59.6\%$ & $61.9\%$  \\ \midrule
+    GenHMM & $69.5\%$ & $70.6\%$ & $70.7\%$   \\ 
     \bottomrule
@@
     {Model} &  K=1 &  K=3 &  K=5  \\  \midrule
-    GMM-HMM & $48.59\%$ &  $51.26\%$ &  $52.40\%$  \\ \midrule
-    GenHMM & $61.10\%$ & $\%$ &   \\ 
+    GMM-HMM & $48.5\%$ &  $51.2\%$ &  $52.4\%$  \\ \midrule
+    GenHMM & $61.1\%$ & $62.1\%$ & $62.1\%$   \\ 
     \bottomrule                                                                  
@@
     \\ \midrule
-    GMM-HMM & $37.15\%$ &  $40.65\%$ & $42.27\%$  \\ \midrule
-    GenHMM & $50.33\%$ & $\%$ & $\%$   \\ 
+    GMM-HMM & $37.1\%$ &  $40.6\%$ & $42.2\%$  \\ \midrule
+    GenHMM & $50.3\%$ & $51.0\%$ & $52.3\%$   \\ 
     \bottomrule

### 2019-08-26 manuscript/main.tex
revise with minh's comment

@@
 A HMM model $\bm{H}$ defined in a hypothesis space $\Hh$, i.e.
-$\bm{H} \in \Hh$, is capable to model time-span signal $\ubar{\bm{x}} = \left[ \bm{x}_1, \cdots, \bm{x}_T\right]^{\intercal}$, where $\bm{x}_t\in \RR^{N}$ is the signal at time $t$, $[\cdot]^{\intercal}$ denotes transpose, and $T$ denotes the time length.
+$\bm{H} \in \Hh$, is capable to model time-span signal $\ubar{\bm{x}} = \left[ \bm{x}_1, \cdots, \bm{x}_T\right]^{\intercal}$, where $\bm{x}_t\in \RR^{N}$ is the signal at time $t$, $[\cdot]^{\intercal}$ denotes transpose, and $T$ denotes the time length\footnote{The length for difference sequential data can be different.}.
 Using HMM for signal representation is illustrated in Figure~\ref{fig:hmm}.
@@
 
-For a given dataset, we denote its empirical distribution by $\hat{p}(\ubm{x}) = \frac{1}{R}\sum_{r=1}^{R} \delta_{\ubmr{x}{r}}(\ubm{x})$, where superscipt $(\cdot)^{r}$ denotes the index of sequential signal.
-The natural question is how to use GenHMM to represent the dataset.
-Alternative, we are looking for the answer to question:
+Assume the sequential signal $\ubm{x}$ follows an unknown distribution $p(\ubm{x})$.
+We would like to use GenHMM to model this distribution.
+Alternatively, we are looking for the answer to the question:
 \begin{equation}
-  \umin{\bm{H}\in \Hh} KL(\hat{p}(\ubm{x})\| p(\ubm{x};\bm{H})),
+  \umin{\bm{H}\in \Hh} KL({p}(\ubm{x})\| p(\ubm{x};\bm{H})),
 \end{equation}
-where $KL(\cdot\|\cdot)$ denotes the Kullback-Leibler divergence. The KL divergence problem can be boiled down to a maximizing a loglikelihood problem:
+where $KL(\cdot\|\cdot)$ denotes the Kullback-Leibler divergence.
+For practical consideration, we only have access to the samples of $p(\ubm{x})$, i.e.
+the dataset of this distribution.
+For the given dataset, we denote its empirical distribution by $\hat{p}(\ubm{x}) = \frac{1}{R}\sum_{r=1}^{R} \delta_{\ubmr{x}{r}}(\ubm{x})$, where superscipt $(\cdot)^{r}$ denotes the index of sequential signal.
+The KL divergence problem can be boiled down to a maximizing a loglikelihood problem:
 \begin{equation}\label{eq:ml-of-hmm}
-  \uargmax{\bm{H} \in \Hh} \sum_{r=1}^{R}\log\,p(\ubmr{x}{r}; \bm{H}).
+  \uargmax{\bm{H} \in \Hh} \frac{1}{R}\sum_{r=1}^{R}\log\,p(\ubmr{x}{r}; \bm{H}).
 \end{equation}
@@
   &= \EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p({s}_{1};\bm{H})  \right], \label{eq:init-distribution-update}\\
-  \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \sum_{t}\log\,p({s}_{t+1}|{s}_{t}; \bm{H}) \right], \label{eq:transition-update}\\
+  \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \sum_{t=1}^{T}\log\,p({s}_{t+1}|{s}_{t}; \bm{H}) \right], \label{eq:transition-update}\\
   \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) &= \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p(\ubm{x}, \ubm{\kappa}| \ubm{s}; \bm{H}) \right].
@@
   &\Qq(\bm{\Theta}; \bm{H}^{\mathrm{old}}) \nonumber \\
-  = &\frac{1}{R}\sum_{r=1}^{R}\sum_{\ubmr{s}{r}}\sum_{\ubmr{\kappa}{r}}{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \sum_{t=1}^{{T}^{r}}\log\,p(\bm{x}^{(r)} | \smtr{s}{t}{r}, \smtr{\kappa}{t}{r}; \bm{H}) \nonumber \\
-  =& \frac{1}{R}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}-1} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})p(\smtr{\kappa}{t}{r}|\smtr{s}{t}{r}, \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \nonumber\\
-  &  \log\, p(\bmtr{x}{t}{r} | \smtr{s}{t}{r}, \smtr{\kappa}{t}{r}; \bm{H}). 
+  = &\frac{1}{R}\sum_{r=1}^{R}\sum_{\ubmr{s}{r}}\sum_{\ubmr{\kappa}{r}}{p(\ubmr{s}{r}, \ubmr{\kappa}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \sum_{t=1}^{{T}^{r}}\log\,p(\bmtr{x}{t}{r} | \smtr{s}{t}{r}, \smtr{\kappa}{t}{r}; \bm{H}) \nonumber \\
+  =& \frac{1}{R}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}  \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})p(\smtr{\kappa}{t}{r}|\smtr{s}{t}{r}, \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \nonumber\\
+  &  \log\, p(\bmtr{x}{t}{r} | \smtr{s}{t}{r}, \smtr{\kappa}{t}{r}; \bm{H}), 
 \end{align}
-In \eqref{eq:obj-q-gen-mix}, the state posterior $p(s_t| \ubm{x}, \bm{H}^{\mathrm{old}})$ is computed by forward-backward algorithm. The posterior of $\kappa$ is:
+where $T^r$ is the length of the $r$-th sequential data.
+In \eqref{eq:obj-q-gen-mix}, the state posterior $p(s_t| \ubm{x}, \bm{H}^{\mathrm{old}})$ is computed by forward-backward algorithm.
+The posterior of $\kappa$ is:
 \begin{align}\label{eq:kappa-posterior}
@@
   &\Qq(\bm{\Theta}; \bm{H}^{\mathrm{old}}) \nonumber \\
-  =& \frac{1}{R}\hspace{-3pt}\sum_{r=1}^{R}\hspace{-3pt} \sum_{t=1}^{{T}^{r}-1}\hspace{-3pt} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|} \hspace{-3pt} \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})p(\smtr{\kappa}{t}{r}|\smtr{s}{t}{r}, \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \nonumber\\
+  =& \frac{1}{R}\hspace{-3pt}\sum_{r=1}^{R}\hspace{-3pt} \sum_{t=1}^{{T}^{r}}\hspace{-3pt} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|} \hspace{-3pt} \sum_{\smtr{\kappa}{t}{r}=1}^{K}p(\smtr{s}{t}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})p(\smtr{\kappa}{t}{r}|\smtr{s}{t}{r}, \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) \nonumber\\
   &\left[ \log\, p_{\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}}(\bm{f}_{\smtr{s}{t}{r}, \smtr{\kappa}{t}{r}}(\bmtr{x}{t}{r})) + \sum_{l=1}^{L}\log\,| \det (\nabla{\bm{f}_{s,\kappa}^{[l]}})|\right].
@@
 parameters of generators, i.e.
-$\bm{\Phi}$, since the dataset represented by $\hat{p}(\ubm{x})$ can be tool large to do whole-dataset iterations.
+$\bm{\Phi}$, since the dataset represented by $\hat{p}(\ubm{x})$ can be too large to do whole-dataset iterations.
 
@@
 \begin{align}\label{opm:pi}
-  \pi_{s, \kappa} & = \uargmax{\pi_{s, \kappa}} \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) \\ \nonumber
+  \pi_{s, \kappa} & = \uargmax{\pi_{s, \kappa}} \Qq(\bm{\Pi}; \bm{H}^{\mathrm{old}}) \\ \nonumber
                   & s.t. \, \sum_{\kappa=1}^{K} \pi_{s, \kappa}= 1, \forall s = 1, 2, \cdots, |\Ss|. 
@@
 \begin{equation}
-  \Ll = \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) + \sum_{\lambda_s} \lambda_s\left( 1-  \sum_{\kappa=1}^{K}\pi_{s, \kappa}  \right).
+  \Ll = \Qq(\bm{\Pi}; \bm{H}^{\mathrm{old}}) + \sum_{s=1}^{|\Ss|} \lambda_s\left( 1-  \sum_{\kappa=1}^{K}\pi_{s, \kappa}  \right).
 \end{equation}
@@
 \begin{equation}
-  \pi_{s,\kappa} = \frac{1}{\lambda_s}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}} p(\smtr{s}{t}{r}=s, \smtr{\kappa}{t}{r} =\kappa| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})
+  \pi_{s,\kappa} = \frac{1}{\lambda_s}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}} p(\smtr{s}{t}{r}=s, \smtr{\kappa}{t}{r} =\kappa| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})
 \end{equation}
@@
   % &= \sum_{r=1}^{R} \EE_{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \left[\log\,\sum_{t=1}^{T^{(r)}-1}p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; {A})\right] \nonumber\\
-  &= \sum_{r=1}^{R} \sum_{\ubmr{s}{r}}{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \sum_{t=1}^{T^{(r)}-1}\log\,p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; \bm{H}) \nonumber \\
+  &= \sum_{r=1}^{R} \sum_{\ubmr{s}{r}}{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \sum_{t=1}^{T^{r}-1}\log\,p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; \bm{H}) \nonumber \\
   &= \sum_{r=1}^{R} \hspace{-0.1cm}\sum_{t=1}^{T^{r}-1}\hspace{-0.1cm} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}\hspace{-0.05cm}\sum_{\smtr{s}{t+1}{r}=1}^{|\Ss|}\hspace{-0.2cm}{p(\smtr{s}{t}{r}, \smtr{s}{t+1}{r}| \ubmr{x}{r};\hspace{-0.05cm} \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; \bm{H})
@@
 \begin{equation}\label{eq:update-transition-solt}
-  \bar{\xi}_{i,j} = \sum_{r= 1}^{R} \sum_{t= 1}^{T^{(r)}-1}{p(\smtr{s}{t}{r}=i, \smtr{s}{t+1}{r}=j| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}.
+  \bar{\xi}_{i,j} = \sum_{r= 1}^{R} \sum_{t= 1}^{T^{r}-1}{p(\smtr{s}{t}{r}=i, \smtr{s}{t+1}{r}=j| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})}.
 \end{equation}
+\subsection{On Convergence of GenHMM}
+In pursuit of representing a dataset by GenHMM,  we are interested if the learning solution discussed in subsection~\ref{subsec:optmGenHMM} would convergence.
+The properties on GenHMM's convergence is analyzed as follows.
+
+\begin{prop}\label{proposition1}
+  Assume that parameter $\bm{\Theta} = \left\{ \bm{\theta}_{s,\kappa}| s\in \Ss, \kappa=1, 2, \cdots, K \right\}$ is in a compact set,  $\bm{f}_{s,\kappa}$ and  ${\nabla\bm{f}_{s,\kappa}}$ are continuous w.r.t.
+${\bm\theta}_{s,\kappa}$ in GenHMM.
+Then GenHMM converges.
+\end{prop}
+
+\begin{proof}
+  We begin with the comparison of log-likelihood evaluated under $\bm{H}^{\mathrm{old}}$ and $\bm{H}^{\mathrm{old}}$. The log-likelihood of dataset given by $\hat{p}(\ubm{x})$ can be reformulated as
+  \begin{align*}
+    &\EE_{\hat{p}(\ubm{x})}\left[ \log\,p(\ubm{x};\bm{H}^{\mathrm{new}}) \right] \nonumber \\
+    =& \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{new}})}{p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right] \nonumber \\
+    &+ \EE_{\hat{p}(\ubm{x})}\left[ KL(p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})\|p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{new}})) \right],
+  \end{align*}
+  where the first term on the right hand side of the above inequality can be further written as
+  \begin{align*}
+    &\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{new}})}{p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right] \nonumber \\
+    = &\Qq(\bm{H}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) + \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})\right].
+  \end{align*}
+  According to subsection~\ref{subsec:optmGenHMM}, the optimization problems give
+  \begin{align*}
+    \Qq(\bm{q}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) &\geq \Qq(\bm{q}^{\mathrm{old}}; \bm{H}^{\mathrm{old}}),\nonumber \\
+    \Qq(\bm{A}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) &\geq \Qq(\bm{A}^{\mathrm{old}}; \bm{H}^{\mathrm{old}}),\nonumber \\
+    \Qq(\bm{\Pi}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) &\geq \Qq(\bm{\Pi}^{\mathrm{old}}; \bm{H}^{\mathrm{old}}), \nonumber \\
+    \Qq(\bm{H}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) &\geq \Qq(\bm{H}^{\mathrm{old}}; \bm{H}^{\mathrm{old}})
+  \end{align*}
+  % For the learning w.r.t. neural network parameter set $\bm{\Theta}$, as long as the lose function does not decrease during EM iterations, we would have
+  Since
+  \begin{align*}
+    \Qq(\bm{H}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) = &\Qq(\bm{q}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) + \Qq(\bm{A}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) \nonumber \\
+                                                        &+ \Qq(\bm{\Pi}^{\mathrm{new}}; \bm{H}^{\mathrm{old}})
+                                                          + \Qq(\bm{\Theta}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}),
+  \end{align*}
+  it gives
+  \begin{equation}
+    \Qq(\bm{H}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) \geq \Qq(\bm{H}^{\mathrm{old}}; \bm{H}^{\mathrm{old}}).
+  \end{equation}
+  Therefore we have the inequality 
+  \begin{align*}
+    &\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{new}})}{p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right] \nonumber \\
+    \geq &\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{old}})}{p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right].
+  \end{align*}
+  Due to the fact that $KL(p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})\|p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}}))=0$, we have
+  \begin{align*}
+    &\EE_{\hat{p}(\ubm{x})}\left[ \log\,p(\ubm{x};\bm{H}^{\mathrm{new}}) \right] \nonumber \\
+    \geq & \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{old}})}{p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right] \nonumber \\
+    = & \EE_{\hat{p}(\ubm{x})}\left[ \log\,p(\ubm{x};\bm{H}^{\mathrm{old}}) \right].
+  \end{align*}
+  Since $\bm{f}_{s,\kappa}$ and  ${\nabla\bm{f}_{s,\kappa}}$ are continuous w.r.t.
+${\bm\theta}_{s,\kappa}$ in GenHMM, $\EE_{\hat{p}(\ubm{x})}\left[ \log\,p(\ubm{x};\bm{H}) \right]$ is bounded.
+The above inequality shows $\EE_{\hat{p}(\ubm{x})}\left[ \log\,p(\ubm{x};\bm{H}) \right]$ is non-decreasing in learning of GenHMM.
+Therefore, GenHMM will converge.
+
+  % Therefore
+  % \begin{align*}
+  %   KL(\hat{p}(\ubm{x})\|p(\ubm{x};\bm{H}^{\mathrm{new}})) = &\EE_{\hat{p}(\ubm{x})}\left[ \log\,\frac{\hat{p}(\ubm{x})}{p(\ubm{x};\bm{H}^{\mathrm{new}})} \right] \nonumber \\
+  %   \leq &\EE_{\hat{p}(\ubm{x})}\left[ \log\,\frac{\hat{p}(\ubm{x})}{p(\ubm{x};\bm{H}^{\mathrm{old}})} \right] \\
+  %   = & KL(\hat{p}(\ubm{x})\|p(\ubm{x};\bm{H}^{\mathrm{old}})).
+  % \end{align*}
+  % Since KL divergence is non-negative and thus lower bounded, GenHMM will converge.
+\end{proof}
 
@@
 to $\bm{q}$, $\bm{A}$ and $\bm{\Pi}$, which have optimal solutions, generator learning usually cannot give optimal solution to problem $\max_{\bm{\Theta}} \Qq(\bm{\Theta}; \bm{H}^{\mathrm{old}})$.
-In fact, given that no optimal $\bm{\Theta}$ is obtained in each iteration in Algorithm~\ref{algo:genhmm}, learning of GenHMM can still converge.
-
-In pursuit of representing a dataset by GenHMM, we are actually minimize a KL divergence $KL(\hat{p}(\ubm{x})\| p(\ubm{x};\bm{H}))$.
-Our following analysis shows that our proposed algorithm for learning GenHMM can converge, without requirement for optimal $\bm{\Theta}$ in each iteration.
+In fact, given that no optimal $\bm{\Theta}$ is obtained, learning of GenHMM can still converge as long as quantity $\Qq(\bm{\Theta}; \bm{H})$ are improving in iterations in Algorithm~\ref{algo:genhmm}, where the inequalities in Proposition~\ref{proposition1} still hold.
+Therefore optimal $\bm{\Theta}$ in each iteration is not required for convergence of GenHMM as long as the loss in $\eqref{eq:obj-q-gen-mix}$ is getting improved.
 
@@
     \FOR {epoch $n < N$}
-    \STATE Sample a batch of data $\left\{ \ubmr{x}{r} \right\}_{r=1}^{R_b}$ from dataset $\hat{p}(\ubm{x})$
+    \STATE Sample a batch of data $\left\{ \ubmr{x}{r} \right\}_{r=1}^{R_b}$ from dataset $\hat{p}(\ubm{x})$ with batch size $R_b$
 
@@
 \end{algorithm}
-
-\begin{prop}
-  Assume that parameters of GenHMM are in a compact set.
-Further assume that $\Qq(\bm{\Theta}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) \geq \Qq(\bm{\Theta}^{\mathrm{old}}; \bm{H}^{\mathrm{old}})$, then we always have $KL(\hat{p}(\ubm{x})\| p(\ubm{x};\bm{H}^{\mathrm{new}})) \leq KL(\hat{p}(\ubm{x})\| p(\ubm{x};\bm{H}^{\mathrm{old}}))$, and GenHMM converges.
-\end{prop}
-
-\begin{proof}
-  We begin with the comparison of log-likelihood evaluated under $\bm{H}^{\mathrm{old}}$ and $\bm{H}^{\mathrm{old}}$. The log-likelihood of dataset given by $\hat{p}(\ubm{x})$ can be reformulated as
-  \begin{align*}
-    &\EE_{\hat{p}(\ubm{x})}\left[ \log\,p(\ubm{x};\bm{H}^{\mathrm{new}}) \right] \nonumber \\
-    =& \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{new}})}{p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right] \nonumber \\
-    &+ \EE_{\hat{p}(\ubm{x})}\left[ KL(p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})\|p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{new}})) \right],
-  \end{align*}
-  where the first term on the right hand side of equality can be further written as
-  \begin{align*}
-    &\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{new}})}{p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right] \nonumber \\
-    = &\Qq(\bm{H}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) + \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})\right].
-  \end{align*}
-  According to subsection~\ref{subsec:optmGenHMM}, the subproblem solutions gives
-  \begin{align*}
-    \Qq(\bm{q}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) &\geq \Qq(\bm{q}^{\mathrm{old}}; \bm{H}^{\mathrm{old}}),\nonumber \\
-    \Qq(\bm{A}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) &\geq \Qq(\bm{A}^{\mathrm{old}}; \bm{H}^{\mathrm{old}}),\nonumber \\
-    \Qq(\bm{\Pi}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) &\geq \Qq(\bm{\Pi}^{\mathrm{old}}; \bm{H}^{\mathrm{old}}).
-  \end{align*}
-  For the learning w.r.t. neural network parameter set $\bm{\Theta}$, as long as the lose function does not decrease during EM iterations, we would have
-  \begin{equation*}
-    \Qq(\bm{H}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) \geq \Qq(\bm{H}^{\mathrm{old}}; \bm{H}^{\mathrm{old}}),
-  \end{equation*}
-  since
-  \begin{align*}
-    \Qq(\bm{H}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) = &\Qq(\bm{q}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) + \Qq(\bm{A}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) \nonumber \\
-                                                        &+ \Qq(\bm{\Pi}^{\mathrm{new}}; \bm{H}^{\mathrm{old}})
-                                                          + \Qq(\bm{\Theta}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}).
-  \end{align*}
-  Therefore we have the inequality 
-  \begin{align*}
-    &\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{new}})}{p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right] \nonumber \\
-    \geq &\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{old}})}{p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right].
-  \end{align*}
-  Due to the fact that $KL(p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})\|p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}}))=0$, we have
-  \begin{align*}
-    &\EE_{\hat{p}(\ubm{x})}\left[ \log\,p(\ubm{x};\bm{H}^{\mathrm{new}}) \right] \nonumber \\
-    \geq & \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{old}})}{p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right] \nonumber \\
-    = & \EE_{\hat{p}(\ubm{x})}\left[ \log\,p(\ubm{x};\bm{H}^{\mathrm{old}}) \right].
-  \end{align*}
-  Therefore
-  \begin{align*}
-    KL(\hat{p}(\ubm{x})\|p(\ubm{x};\bm{H}^{\mathrm{new}})) = &\EE_{\hat{p}(\ubm{x})}\left[ \log\,\frac{\hat{p}(\ubm{x})}{p(\ubm{x};\bm{H}^{\mathrm{new}})} \right] \nonumber \\
-    \leq &\EE_{\hat{p}(\ubm{x})}\left[ \log\,\frac{\hat{p}(\ubm{x})}{p(\ubm{x};\bm{H}^{\mathrm{old}})} \right] \\
-    = & KL(\hat{p}(\ubm{x})\|p(\ubm{x};\bm{H}^{\mathrm{old}})).
-  \end{align*}
-  Since KL divergence is non-negative and thus lower bounded, GenHMM will converge.
-\end{proof}
 

### 2019-08-27 manuscript/main.tex
update

@@
 Another track of related work is hybrid method of temporal neural network models and HMM.
-In \cite{liu2019lstmHmmHyb,buys2018bridging,vik2016rnnHmm}, a long short-term memory (LSTM) model/recurrent neural network (RNN) is combined with HMM as hybridization.
+In \cite{liu2019lstmHmmHyb,buys2018bridging,vik2016rnnHmm}, a long short-term memory (LSTM) model/recurrent neural network (RNN) is combined with HMM as hybrid.
 A hierarchical training is carried out by: i) training a HMM first, ii) and then doing modified training of LSTM using trained HMM.
@@
 
-In this work, we propose a generator-mixed HMM model, termed as GenHMM, where generators are realized by neural networks to help the model gain high modeling capacity.
+In this work, we propose a generative model based HMM.
+Specifically, we propose a generator-mixed HMM, termed as GenHMM, where generators are realized by neural networks to help the model gain high modeling capacity.
 Our proposed model, GenHMM:
@@
 Training of GenHMM employs expectation maximization (EM) framework.
-Training a GenHMM is as easy as training a GMM-HMM model, while configuration of GenHMM is flexible.
+Training a GenHMM is as easy as training a GMM-HMM model, while configuration of GenHMM is flexible;
 % \item has high modeling capability of sequence data, though its training and using are handy
@@
 \begin{itemize}
-\item Proposing a neural network based HMM for sequential data modeling, i.e. GenHMM. The likelihood computation of sequential data under GenHMM can be computed exactly and efficiently.
+\item Proposing a neural network based HMM for sequential data modeling, i.e. GenHMM. GenHMM has the tractability of exact likelihood.
 \item Designing practical algorithm for training GenHMM under EM framework. Stochastic gradient  search in batch fashion is embedded in this algorithm.
 \item Giving convergence analysis for GenHMM under the proposed learning algorithm.
+\item Verifying the efficiency of the proposed model on practical sequential data.
 \end{itemize}
 
-\section{Generator-mixed HMM}
+\section{Generator-mixed HMM (GenHMM)}
 
@@
 \begin{itemize}
-\item $\Ss$ is the set of hidden states of HMM $\bm{H}$;
+\item $\Ss$ is the set of hidden states of HMM $\bm{H}$.
 \item $\bm{q} = \left[ q_1, q_2, \cdots, q_{|\Ss|}\right]^\intercal$ is initial distribution of HMM $\bm{H}$ with $|\Ss|$ as cardinality of $\Ss$.
@@
 \item For given hidden state $s$, the density function of observable signal is $p({\bm{x}}|{s};\bm{\Phi}_{s})$, where $\bm{\Phi}_{s}$ is the parameter set that defines this probabilistic model.
+Denote $\bm{\Phi} = \left\{ \bm{\Phi}_{s}| s \in \Ss \right\}$.
 \end{itemize}
@@
 In the framework of HMM, at each time instance, signal $\bm{x}_t$ is assumed to be generated by a distribution with density function $p(\bm{x}_t| s_t; \bm{\Phi}_{s_t})$, and $s_t$ is decided by the hidden markov process.
-This gives us the probabilistic model $p(\ubm{x};\bm{H})$.
+Putting these together gives us the probabilistic model $p(\ubm{x};\bm{H})$.
+
 \subsection{Generative Model of GenHMM}
@@
 We define $p(\bm{x}| s, \kappa; \bm{\theta}_{s, \kappa})$ as induced distribution by a generator $\bm{g}_{s,\kappa}: \RR^{N}\rightarrow\RR^{N}$, such that $\bm{x}=\bm{g}_{s, \kappa}(\bm{z})$, where $\bm{z}$ is a latent variable following a distribution with density function $p_{s,\kappa}(\bm{z})$.
-By change of variable, we have
+Assuming $\bm{g}_{s, \kappa}$ is invertible, by change of variable, we have
 \begin{equation}\label{eq:changel-variable}
@@
 
-For a state $s$ of GenHMM, the induced probability distribution is illustrated in Figure~\ref{fig:gen-mix}.
+For a state $s$ of GenHMM, the induced probability distribution has the signal flow illustration as shown in Figure~\ref{fig:gen-mix}, in which the generator identity is up to the random variable $\kappa$.
 
@@
 
-The first problem that we need to address is to deal with the hidden sequential variable $\ubm{s}$ and $\ubm{\kappa}$  in model $\bm{H}$.
+The first problem that we need to address is to deal with the hidden sequential variable $\ubm{s}$ and $\ubm{\kappa}$  in model $\bm{H}$, where $\ubm{s}=[ \bm{s}_1, \bm{s}_2, \cdots, \bm{s}_T ]^{\intercal}$ and $\ubm{\kappa} = [\bm{\kappa}_1, \bm{\kappa}_2, \cdots, \bm{\kappa}_T]^{\intercal}$.
 For a sequential observable variable $\ubm{x}$, $\ubm{s}$ is the hidden state sequence corresponding to $\ubm{x}$, and $\ubm{\kappa}$ is the hidden variable sequence representing the generator identity sequence that actually generates $\ubm{x}$.
@@
 The learning method of HMM is followed then.
-The convergence analysis is discussed at the end of this section.
+The convergence of GenHMM is also discussed in this section.
 
@@
 We use a feed-forward neural network to implement each generator $\bm{g}_{s,\kappa}$.
-Define  $\bm{g}_{s,\kappa}$ as a $L$-layer neural network and we formulate its mapping by each layer's concatenation:
+Define  $\bm{g}_{s,\kappa}$ as a $L$-layer neural network and we formulate its mapping by layer-wise concatenation:
 $\bm{g}_{s,\kappa}=\bm{g}_{s,\kappa}^{[L]}\circ \bm{g}_{s,\kappa}^{[L-1]}\circ \cdots
@@
 This is boiled down to maximizing the cost function of neural networks that can be formulated as
-\begin{align}\label{eq:obj-q-gen-mix}
+\begin{align}\label{eq:obj-q-gen-mix-log}
   &\Qq(\bm{\Theta}; \bm{H}^{\mathrm{old}}) \nonumber \\
@@
 where $T^r$ is the length of the $r$-th sequential data.
-In \eqref{eq:obj-q-gen-mix}, the state posterior $p(s_t| \ubm{x}, \bm{H}^{\mathrm{old}})$ is computed by forward-backward algorithm.
+In \eqref{eq:obj-q-gen-mix-log}, the state posterior $p(s_t| \ubm{x}, \bm{H}^{\mathrm{old}})$ is computed by forward-backward algorithm.
 The posterior of $\kappa$ is:
@@
 
-By substituting \eqref{eq:changel-variable} and \eqref{eq:cat-jacobian} into \eqref{eq:obj-q-gen-mix}, we have cost function for neural networks as
+By substituting \eqref{eq:changel-variable} and \eqref{eq:cat-jacobian} into \eqref{eq:obj-q-gen-mix-log}, we have cost function for neural networks as
 \begin{align}\label{eq:obj-q-gen-mix}
@@
 
-$p(\smtr{s}{1}{r};\bm{H})$ is the probability of initial state of GenHMM for $r$-th sequential sample. Actually $q_i = p({s}_{1} =i) $, $i= 1, 2, \cdots, |\Ss|$ for $\bm{H}$. Solution to problem:
+$p(\smtr{s}{1}{r};\bm{H})$ is the probability of initial state of GenHMM for $r$-th sequential sample. Actually $q_i = p({s}_{1} =i;\bm{H}) $, $i= 1, 2, \cdots, |\Ss|$. Solution to problem:
 \begin{align}
@@
 
-\subsection{Algorithm of GenHMM and Convergence Analysis}
+\subsection{Algorithm of GenHMM}
 
@@
 \subsection{Experimental Setups}
-The dataset used for sequential data modeling and classification is TIMIT dataset where the speech is sampled at $16$kHz.
+The dataset used for sequential data modeling and classification is TIMIT dataset where the speech signal is sampled at $16$kHz.
 For extraction of feature vectors, we use $25$ms frame length and $10$ms frame shift to convert sound track into standard Mel-frequency cepstral coefficients (MFCCs) features.
@@
 The coupling layer shown as \eqref{eq-gl} maps a part of its input signal identically.
-The following coupling layer would alternate the input signal older such that after two consequential coupling layers, there is no signal remains the same any more.
+The following coupling layer would alternate the input signal order such that after two consequential coupling layers, there is no signal remains the same any more.
 We term such two consequential coupling layers as a \textit{flow block}.
@@
 The exact number of states for a phoneme is decided by computing the average length of MFCC frames per phone in training dataset, and clipping the average length into $\left\{ 3,4,5 \right\}$.
-Transition matrix $\bm{A}$ is initialized as upper triangular matrix for each GenHMM.
+Transition matrix $\bm{A}$ is initialized as upper triangular matrix for GenHMM.
 

### 2019-08-28 manuscript/main.tex
update

@@
 HMM provides a framework for sequential data modeling.
-To connect states of HMM with input data, a probabilistic model is used to describe how each state of HMM should represent a pattern of sequential data.
+To connect states of HMM with input data, a probabilistic model is used to describe how each state of HMM represents a pattern of sequential data.
 The typical way is to use a Gaussian mixture model (GMM) per state of HMM \cite{juang1986maximum}, where GMMs are used to represent the relationship between states of HMM and sequential data input.
@@
 In \cite{liu2019lstmHmmHyb,buys2018bridging,vik2016rnnHmm}, a long short-term memory (LSTM) model/recurrent neural network (RNN) is combined with HMM as hybrid.
-A hierarchical training is carried out by: i) training a HMM first, ii) and then doing modified training of LSTM using trained HMM.
+A hierarchical training is carried out by: i) training a HMM first, ii) then doing modified training of LSTM using trained HMM.
 This hierarchical training procedure is motivated by the intuition of using LSTM or RNN to fill in the gap where HMM can not learn.
@@
 \begin{itemize}
-  \item firstly has high modeling capacity of sequential data, due to the neural network based generators;
+  \item has high modeling capacity of sequential data, due to the neural network based generators;
 \item has advantage of easy to do training.
@@
 
-Our framework is a Hidden Markov Model (HMM).
-A HMM model $\bm{H}$ defined in a hypothesis space $\Hh$, i.e.
+Our framework is a HMM.
+A HMM $\bm{H}$ defined in a hypothesis space $\Hh$, i.e.
 $\bm{H} \in \Hh$, is capable to model time-span signal $\ubar{\bm{x}} = \left[ \bm{x}_1, \cdots, \bm{x}_T\right]^{\intercal}$, where $\bm{x}_t\in \RR^{N}$ is the signal at time $t$, $[\cdot]^{\intercal}$ denotes transpose, and $T$ denotes the time length\footnote{The length for difference sequential data can be different.}.
@@
 \item $\Ss$ is the set of hidden states of HMM $\bm{H}$.
-\item $\bm{q} = \left[ q_1, q_2, \cdots, q_{|\Ss|}\right]^\intercal$ is initial distribution of HMM $\bm{H}$ with $|\Ss|$ as cardinality of $\Ss$.
+\item $\bm{q} = \left[ q_1, q_2, \cdots, q_{|\Ss|}\right]^\intercal$ is the initial distribution of HMM $\bm{H}$ with $|\Ss|$ as cardinality of $\Ss$.
 For a state $s$, $q_s = p(s;\bm{H})$.
-In the following of this paper, we use $s_t$ to denote the state $s$ at time $t$.
+In the following, we use $s_t$ to denote the state $s$ at time $t$.
 \item $\bm{A}$ is the transition matrix for the HMM $\bm{H}$ of size $|\Ss| \times |\Ss|$ and $\bm{A}_{i,j} = p(s_{t+1}=j|s_{t}=i; \bm{H})$.
-\item For given hidden state $s$, the density function of observable signal is $p({\bm{x}}|{s};\bm{\Phi}_{s})$, where $\bm{\Phi}_{s}$ is the parameter set that defines this probabilistic model.
+\item For a given hidden state $s$, the density function of the observable signal is $p({\bm{x}}|{s};\bm{\Phi}_{s})$, where $\bm{\Phi}_{s}$ is the parameter set that defines this probabilistic model.
 Denote $\bm{\Phi} = \left\{ \bm{\Phi}_{s}| s \in \Ss \right\}$.

### 2019-08-28 Phoneme-Recognition-on-the-TIMIT-Database_IntechOpen_files/27.js
untrack not used files

@@ -1,1 +1,0 @@
-webpackJsonp([27],{601:function(t,e,i){!function(e,i){t.exports=i()}(0,function(){"use strict";function t(){var t=o.scrollY||o.pageYOffset,e=v.timeStamp,i="down"===d?Math.max:Math.min,n=u.offsetHeight-o.innerHeight;if(t=Math.max(0,t),t=Math.min(n,t),l.unshift({y:t,t:e}),l.pop(),t===i(m,t))return w=e,void(m=t);var b=e-s;if(b>w){m=t;for(var p=0;p<f&&(l[p]&&!(l[p].t<b));p+=1)m=i(m,l[p].y)}Math.abs(t-m)>c&&(m=t,w=e,d="down"===d?"up":"down",r.setAttribute(a,d))}function e(e){return v=e,o.requestAnimationFrame(t)}function i(t){return r=t&&t.el||n.el,o=t&&t.win||n.win,a=t&&t.attribute||n.attribute,d=t&&t.direction||n.dir,t&&!0===t.off?(r.setAttribute(a,"off"),o.removeEventListener("scroll",e)):(m=o.scrollY||o.pageYOffset,r.setAttribute(a,d),o.addEventListener("scroll",e))}var n={el:document.documentElement,win:window,attribute:"data-scrolldir",dir:"down"},r=void 0,o=void 0,a=void 0,d=void 0,u=document.body,f=32,s=512,c=64,l=Array(f),v=void 0,m=void 0,w=0;return i})}});

### 2019-08-28 Phoneme-Recognition-on-the-TIMIT-Database_IntechOpen_files/altmetric_badges-c09531d6f252ca72ffdadae2d9597c14.js

@@ -1,3 +1,0 @@
-(function(){window._altmetric||(window._altmetric={}),window._altmetric["export"]=function(e,t){return window._altmetric[e]=t},window._altmetric.exports=function(e){var t,n,i;n=[];for(t in e)i=e[t],n.push(window._altmetric[t]=i);return n}}).call(this),window._altmetric.api_uri="https://api.altmetric.com",window._altmetric.api_key="3c130976ca2b8f2e88f8377633751ba1",window._altmetric.api_version="v1",window._altmetric.details_uri="https://www.altmetric.com",function(){var e;e=function(e){var t,n;return t=document.createElement("div"),n=document.createTextNode(e),t.appendChild(n),t.innerHTML},_altmetric.exports({encodeHTML:e})}.call(this),function(e){"function"==typeof e.define&&(e._altmetric_define=e.define,e.define=void 0),"object"==typeof e.exports&&(e._altmetric_exports=e.exports,e.exports=void 0)}(window),function(e,t){"object"==typeof exports&&"object"==typeof module?module.exports=t():"function"==typeof define&&define.amd?define([],t):"object"==typeof exports?exports.Handlebars=t():e.Handlebars=t()}(this,function(){return function(e){function t(i){if(n[i])return n[i].exports;var r=n[i]={exports:{},id:i,loaded:!1};return e[i].call(r.exports,r,r.exports,t),r.loaded=!0,r.exports}var n={};return t.m=e,t.c=n,t.p="",t(0)}([function(e,t,n){"use strict";function i(){var e=new s.HandlebarsEnvironment;return h.extend(e,s),e.SafeString=u["default"],e.Exception=d["default"],e.Utils=h,e.escapeExpression=h.escapeExpression,e.VM=m,e.template=function(t){return m.template(t,e)},e}var r=n(1)["default"],o=n(2)["default"];t.__esModule=!0;var a=n(3),s=r(a),l=n(17),u=o(l),c=n(5),d=o(c),p=n(4),h=r(p),f=n(18),m=r(f),g=n(19),_=o(g),v=i();v.create=i,_["default"](v),v["default"]=v,t["default"]=v,e.exports=t["default"]},function(e,t){"use strict";t["default"]=function(e){if(e&&e.__esModule)return e;var t={};if(null!=e)for(var n in e)Object.prototype.h
[…]

### 2019-08-28 Phoneme-Recognition-on-the-TIMIT-Database_IntechOpen_files/api.js

@@ -1,1 +1,0 @@
-/* PLEASE DO NOT COPY AND PASTE THIS CODE. */(function(){var CFG='___grecaptcha_cfg';if(!window[CFG]){window[CFG]={};}var GR='grecaptcha';if(!window[GR]){window[GR]={};}window[GR].ready=window[GR].ready||function(f){(window[CFG]['fns']=window[CFG]['fns']||[]).push(f);};(window[CFG]['render']=window[CFG]['render']||[]).push('explicit');(window[CFG]['onload']=window[CFG]['onload']||[]).push('vueRecaptchaApiLoaded');window['__google_recaptcha_client']=true;var po=document.createElement('script');po.type='text/javascript';po.async=true;po.src='https://www.gstatic.com/recaptcha/api2/v1563777128698/recaptcha__en.js';var elem=document.querySelector('script[nonce]');var n=elem&&(elem['nonce']||elem.getAttribute('nonce'));if(n){po.setAttribute('nonce',n);}var s=document.getElementsByTagName('script')[0];s.parentNode.insertBefore(po, s);})();

### 2019-08-28 Phoneme-Recognition-on-the-TIMIT-Database_IntechOpen_files/badge.js

@@ -1,1 +1,0 @@
-!function(e){function n(i){if(t[i])return t[i].exports;var s=t[i]={i:i,l:!1,exports:{}};return e[i].call(s.exports,s,s.exports,n),s.l=!0,s.exports}var t={};n.m=e,n.c=t,n.d=function(e,t,i){n.o(e,t)||Object.defineProperty(e,t,{configurable:!1,enumerable:!0,get:i})},n.n=function(e){var t=e&&e.__esModule?function(){return e.default}:function(){return e};return n.d(t,"a",t),t},n.o=function(e,n){return Object.prototype.hasOwnProperty.call(e,n)},n.p="",n(n.s=1)}([function(e,n){function t(e){var n=i.call(e);return"[object Function]"===n||"function"==typeof e&&"[object RegExp]"!==n||"undefined"!=typeof window&&(e===window.setTimeout||e===window.alert||e===window.confirm||e===window.prompt)}e.exports=t;var i=Object.prototype.toString},function(e,n,t){"use strict";Object.defineProperty(n,"__esModule",{value:!0});var i=t(2),s=(t.n(i),t(3)),o=t(7),r=t(9);if(!window.__dimensions_embed){var a=function(){window.__dimensions_detail_url||(window.__dimensions_detail_url="https://badge.dimensions.ai"),window.__dimensions_metrics_api||(window.__dimensions_metrics_api="https://metrics-api.dimensions.ai"),window.__dimensions_badge_server||(window.__dimensions_badge_server="https://badge.dimensions.ai"),window.__dimensions_badge_url||(window.__dimensions_badge_url="https://badge.dimensions.ai")},c=function(){var e=document.querySelectorAll(".__dimensions_badge_embed__");if(e.length)for(var n=0;n<e.length;++n)!function(n){var t=new o.a(e[n].attributes,window.__dimensions_detail_url,window.__dimensions_metrics_api,window.__dimensions_badge_server);new r.a(t.metricsURL).fetch(function(i){new s.a(e[n],t,i).load()})}(n)},d=function(){var e=document.createElement("link");e.rel="stylesheet",e.href=window.__dimensions_badge_url+"/badge.css",document.getElementsByTagName("head")[0].appendChild(e)},l=function(){window.__dimensions_embed_installed__||(window.__dimensions_embed_installed__=!0,a(),d(),c())};(document.attachEvent?"complete"===document.readyState:"lo
[…]

### 2019-08-28 Phoneme-Recognition-on-the-TIMIT-Database_IntechOpen_files/embed.js

@@ -1,1 +1,0 @@
-!function(e,t,d){var n="createElement",c="getElementsByTagName",m="setAttribute",d=document.getElementById(e);return d&&d.parentNode&&d.parentNode.removeChild(d),d=document[n+"NS"]&&document.documentElement.namespaceURI,d=d?document[n+"NS"](d,"script"):document[n]("script"),d[m]("id",e),d[m]("src",t),(document[c]("head")[0]||document[c]("body")[0]).appendChild(d),d=new Image,void d[m]("src","https://d1uo4w7k31k5mn.cloudfront.net/donut/0.png")}("altmetric-embed-js","https://d1bxh8uas1mnw7.cloudfront.net/assets/altmetric_badges-c09531d6f252ca72ffdadae2d9597c14.js");

### 2019-08-28 src/glow/modules.py

@@ -1,375 +1,0 @@
-import torch
-import torch.nn as nn
-import torch.nn.functional as F
-import numpy as np
-import scipy.linalg
-from . import thops
-
-
-class _ActNorm(nn.Module):
-    """
-    Activation Normalization
-    Initialize the bias and scale with a given minibatch,
-    so that the output per-channel have zero mean and unit variance for that.
-
-    After initialization, `bias` and `logs` will be trained as parameters.
-    """
-
-    def __init__(self, num_features, scale=1.):
-        super().__init__()
-        # register mean and scale
-        size = [1, num_features, 1, 1]
-        self.register_parameter("bias", nn.Parameter(torch.zeros(*size)))
-        self.register_parameter("logs", nn.Parameter(torch.zeros(*size)))
-        self.num_features = num_features
-        self.scale = float(scale)
-        self.inited = False
-
-    def _check_input_dim(self, input):
-        return NotImplemented
-
-    def initialize_parameters(self, input):
-        self._check_input_dim(input)
-        if not self.training:
-            return
-        assert input.device == self.bias.device
-        with torch.no_grad():
-            bias = thops.mean(input.clone(), dim=[0, 2, 3], keepdim=True) * -1.0
-            vars = thops.mean((input.clone() + bias) ** 2, dim=[0, 2, 3], keepdim=True)
-            logs = torch.log(self.scale / (torch.sqrt(vars) + 1e-6))
-            self.bias.data.copy_(bias.data)
-            self.logs.data.copy_(logs.data)
-            self.inited = True
-
-    def _center(self, input, reverse=False):
-        if not reverse:
-            return input + self.bias
-        else:
-            return input - self.bias
-
-    def _scale(self, input, logdet=None, reverse=False):
-        logs = self.logs
-        if not reverse:
-            input = input * torch.exp(logs)
-        else:
-            input = input * torch.exp(-logs)
-        if logdet is not None:
-            """
-            logs is log_std of `mean of channels`
-            so we need to multiply pixels
-            """
-            dlogdet = thops.sum(logs) * thops.pixels(input)
-            if reverse:
-                dlogdet *= -1
-            logdet = logdet + dlogdet
-        return input, logdet
-
-    def forward(self, input, logdet=None, reverse=False):
-        if not self.inited:
[…]

### 2019-08-28 src/timit-preprocessor/add_noise.py
Noise computation script

@@ -1,0 +1,61 @@
+from functools import partial
+from multiprocessing.dummy import Pool
+from scipy.io import wavfile
+import argparse
+from glob import glob
+import os,sys
+import numpy as np
+
+
+def gen_noise(n, type, sigma):
+    """Generate noise of a certain type and std."""
+    if type == "white":
+        raw_noise = np.random.randn(n)
+    else:
+        print("Unknown {} noise".format(type), file=sys.stderr)
+        raw_noise = 0
+
+    return raw_noise * sigma
+
+
+def new_filename(file, type, snr):
+    """Append noise tyep and power at the end of wav filename."""
+    return file.replace(".WAV", ".WAV.{}.{}dB".format(type, snr))
+
+
+def corrupt_wav(file, type=None, snr=None):
+    """Corrupt a wav file with noise and write to a new file."""
+    rate, s = wavfile.read(file)
+    dtype = s.dtype
+
+    s_std = np.std(s)
+    n_std = 10 ** (- snr / 10) * s_std
+    n = gen_noise(s.shape[0], type, n_std)
+    sn = (s + n).astype(dtype)
+
+    wavfile.write(new_filename(file, type, snr), rate, sn)
+    return 0
+
+
+if __name__ == "__main__":
+    parser = argparse.ArgumentParser(description="Add a particular noise type to WAV files.")
+
+    parser.add_argument('-i', metavar="<Input folder>", type=str)
+    parser.add_argument('-snr', metavar="<Signal to Noise Ratio (dB)>", type=int)
+    parser.add_argument('-type', metavar="<Noise type>", type=str)
+    parser.add_argument('-j', metavar="<Number of jobs (default: numcpu)>",
+                        type=int, default=os.cpu_count())
+
+    args = parser.parse_args()
+
+    print(args.i, args.snr, args.type)
+
+    wavs = glob(os.path.join(args.i, "**" , "*.WAV"), recursive=True)
+    f = partial(corrupt_wav, type=args.type, snr=args.snr)
+
+    f(wavs[1])
+
+    with Pool(args.j) as pool:
+        pool.map(f, wavs)
+
+    sys.exit(0)

### 2019-08-28 src/timit-preprocessor/add_noise.py
Compatible with Noise files

@@ -3,14 +3,26 @@ from multiprocessing.dummy import Pool
 from scipy.io import wavfile
 import argparse
 from glob import glob
-import os,sys
+import os, sys
 import numpy as np
 
 
-def gen_noise(n, type, sigma):
+def gen_noise(n, type, sigma, noise_folder="../../data/NoiseDB/NoiseX_16kHz"):
     """Generate noise of a certain type and std."""
+
     if type == "white":
-        raw_noise = np.random.randn(n)
+        noise_filename = os.path.join(noise_folder, "{}_16kHz.wav".format(type))
+        _, loaded_noise = wavfile.read(noise_filename)
+        try:
+            assert(n < loaded_noise.shape[0])
+
+        except AssertionError as e:
+            print("Noise file: {} is too short.".format(noise_filename), file=sys.stderr)
+
+        # Find a random section in file.
+        istart = np.random.randint(loaded_noise.shape[0] - n)
+        raw_noise = loaded_noise[istart:istart+n]
+
     else:
         print("Unknown {} noise".format(type), file=sys.stderr)
         raw_noise = 0
@@ -23,16 +35,19 @@ def new_filename(file, type, snr):
     return file.replace(".WAV", ".WAV.{}.{}dB".format(type, snr))
 
 
-def corrupt_wav(file, type=None, snr=None):
-    """Corrupt a wav file with noise and write to a new file."""
-    rate, s = wavfile.read(file)
-    dtype = s.dtype
-
+def corrupt_data(s, type, snr):
+    """Corrupt a signal with a particular noise."""
     s_std = np.std(s)
     n_std = 10 ** (- snr / 10) * s_std
     n = gen_noise(s.shape[0], type, n_std)
-    sn = (s + n).astype(dtype)
+    sn = (s + n).astype(s.dtype)
+    return sn
 
+
+def corrupt_wav(file, type=None, snr=None):
+    """Corrupt a wav file with noise and write to a new file."""
+    rate, s = wavfile.read(file)
+    sn = corrupt_data(s, type, snr)
     wavfile.write(new_filename(file, type, snr), rate, sn)
     return 0
 
@@ -52,9 +67,7 @@ if __name__ == "__main__":
 
     wavs = glob(os.path.join(args.i, "**" , "*.WAV"), recursive=True)
     f = partial(corrupt_wav, type=args.type, snr=args.snr)
-
     f(wavs[1])
-
     with Pool(args.j) as pool:
         pool.map(f, wavs)
 

### 2019-08-28 manuscript/main.tex
adopt comment from xinyue

@@
 The typical way is to use a Gaussian mixture model (GMM) per state of HMM \cite{juang1986maximum}, where GMMs are used to represent the relationship between states of HMM and sequential data input.
-GMM based HMM (GMM-HMM) has becomes a standard model for sequential data modeling, and employed widely for practical applications, especially in speech recognition \cite{gales2008application,chatterjee2011auditory}.
+GMM based HMM (GMM-HMM) has become a standard model for sequential data modeling, and employed widely for practical applications, especially in speech recognition \cite{gales2008application,chatterjee2011auditory}.
 
@@
 The model assumption is that different instant signal of $\ubar{\bm{x}}$ is generated by a different signal source associated with a hidden state of HMM, while the source selection is done by a discrete hidden markov process.
-We define the hypothesis set of HMM as $\Hh := \{\bm{H} | \{\Ss, \bm{q}, \bm{A}, p(\bm{x}|{s}; \bm{\Phi}_{s})\}$, where:
+We define the hypothesis set of HMM as $\Hh := \{\bm{H} | \{\Ss, \bm{q}, \bm{A}, p(\bm{x}|{s}; \bm{\Phi}_{s})\}$, where
 \begin{itemize}
@@
 GenHMM has a hidden markov process as a typical HMM does, while the probabilistic model of GenHMM for each hidden state is induced by mixture of neural network based generators.
-For given state $s\in \Ss$, we define its probabilistic model in GenHMM as
+For a given state $s\in \Ss$, we define its probabilistic model in GenHMM as
 \begin{equation}
@@
 We would like to use GenHMM to model this distribution.
-Alternatively, we are looking for the answer to the question:
+Alternatively, we are looking for the answer to the question
 \begin{equation}
@@
 For the given dataset, we denote its empirical distribution by $\hat{p}(\ubm{x}) = \frac{1}{R}\sum_{r=1}^{R} \delta_{\ubmr{x}{r}}(\ubm{x})$, where superscipt $(\cdot)^{r}$ denotes the index of sequential signal.
-The KL divergence problem can be boiled down to a maximizing a loglikelihood problem:
+The KL divergence problem can be boiled down to a maximizing loglikelihood problem
 \begin{equation}\label{eq:ml-of-hmm}
@@
 Since directly maximizing likelihood is not an option for our problem in \ref{eq:ml-of-hmm}, we address this problem in expectation maximization (EM) framework.
-This divides our problem into two steps: i) using the joint posterior of hidden variable sequences $\ubm{s}$ and $\ubm{\kappa}$ to obtain a ``expected likelihood'' of observable variable sequence $\ubm{x}$, i.e.
-the E-step; ii) maximizing the expected likelihood with regarding to (w.r.t.) the model $\bm{H}$, i.e.
+This divides our problem into two steps: i) using the joint posterior of hidden variable sequences $\ubm{s}$ and $\ubm{\kappa}$ to obtain an ``expected likelihood'' of the observable variable sequence $\ubm{x}$, i.e.
+the E-step; ii) maximizing the expected likelihood with regard to (w.r.t.) the model $\bm{H}$, i.e.
 the M-step.
@@
   % \end{equation}
-  the expected likelihood function:
+  the expected likelihood function
   \begin{equation}\label{eq:em-q-funciton}
@@
   where $\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \cdot\right]$ denotes the expectation operator by distribution $\hat{p}(\ubm{x})$ and $p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})$.
-\item M-step: the maximization step:
+\item M-step: the maximization step
   \begin{equation}\label{eq:em-m-opt}
@@
 
-The problem \eqref{eq:em-m-opt} can be reformulated as:
+The problem \eqref{eq:em-m-opt} can be reformulated as
 \begin{align}\label{eq:m-step-subs}
@@
   =&\umax{\bm{q}}\Qq(\bm{q}; \bm{H}^{\mathrm{old}}) + \umax{\bm{A}}\Qq(\bm{A}; \bm{H}^{\mathrm{old}}) 
-     + \umax{\bm{\Phi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})
+     + \umax{\bm{\Phi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}),
 \end{align}
@@
 Therefore, we employ forward-backward algorithm \cite{Bishop:2006:PRM:1162264} to do the posterior computation efficiently.
-As we would detail in next section, it is actually $p(s| \ubm{x}; \bm{H})$ and $p(s, \kappa| \ubm{x}; \bm{H})$ that are needed to formulate the problem.
-For the joint posterior $p(s, \kappa| \ubm{x}; \bm{H})$, it can be computed by following the Bayesian rule when posterior of hidden state is available.
+As we would detail in next section, what are needed to formulate the problem, are actually the $p(s| \ubm{x}; \bm{H})$ and $p(s, \kappa| \ubm{x}; \bm{H})$.
+For the joint posterior $p(s, \kappa| \ubm{x}; \bm{H})$, it can be computed by the Bayesian rule when posterior of hidden state is available.
 
-With such a solution framework ready for model GenHMM, there are still remaining problems to address before it can be employed for practical usage, including:
+With such a solution framework ready for GenHMM, there are still remaining problems to address before it can be employed for practical usage, including
 \begin{itemize}
@@
 \item how to train GenHMM to sovle problem in \eqref{eq:ml-of-hmm} using practical algorithm?
-\item would the training of GenHMM converges?
+\item would the training of GenHMM converge?
 \end{itemize}
@@
 $\bm{g}_{s,\kappa}=\bm{g}_{s,\kappa}^{[L]}\circ \bm{g}_{s,\kappa}^{[L-1]}\circ \cdots
-\circ \bm{g}_{s,\kappa}^{[1]}$, where superscipt $[l]$ denotes the $l$'s layer index and $\circ$ denotes mapping concatenation.
-Assume $\bm{g}_{s,\kappa}$ is invertible and denotes its inverse mapping as $\bm{f}_{s,\kappa}=\bm{g}_{s,\kappa}^{-1}$.
+\circ \bm{g}_{s,\kappa}^{[1]}$, where superscript $[l]$ denotes the layer index and $\circ$ denotes mapping concatenation.
+Assume $\bm{g}_{s,\kappa}$ is invertible and denote its inverse mapping as $\bm{f}_{s,\kappa}=\bm{g}_{s,\kappa}^{-1}$.
 For a latent variable $\bm{z}$ with density function $p_{s,\kappa}(\bm{z})$, the generated signal $\bm{x}$ follows an induced distribution with density function \eqref{eq:changel-variable}.
-We illustrate the signal flow between latent variable $\bm{z}$ and observable variable $\bm{x}$ as follows:
+We illustrate the signal flow between latent variable $\bm{z}$ and observable variable $\bm{x}$ as follows
 \begin{equation*}
@@
     \draw[dotted,line width = 0.3 mm] (xi2) -- (xi3);
-  \end{tikzpicture},
+  \end{tikzpicture}
 \end{equation*}
 where $\bm{f}_{s,\kappa}^{[l]}$ is the $l$-th layer of $\bm{f}_{s,\kappa}$. We have $\bm{z}=\bm{f}_{s,\kappa}(\bm{x})$. If every layer of $\bm{g}_{s,\kappa}$ is invertible,
-the full feed-forward neural network is invertible. Flow model, proposed in \cite{DBLP:journals/corr/DinhKB14} as an image generating model, is such a invertible
+the full feed-forward neural network is invertible. Flow model, proposed in \cite{DBLP:journals/corr/DinhKB14} as an image generating model, is such an invertible
 feed-forward neural network.
@@
 $\bm{h}_l = [\bm{h}_{l,a}^{\intercal} \, , \, \bm{h}_{l,b}^{\intercal}]^{\intercal}$.
-The efficient invertible mapping of flow model comes from following forward and inverse relations between $(l-1)$'th and $l$'th layers:
+The efficient invertible mapping of flow model comes from following forward and inverse relations between $(l-1)$'th and $l$'th layers
 \begin{equation}\label{eq-gl}
@@
 where $\nabla{f_{s,\kappa}^{[l]}}$ is the Jacobian of the transformation from the $l$-th layer to the $(l-1)$-th layer, i.e., the inverse transformation.
-We compute the determinate of the Jacobian matrix as
+We compute the determinant of the Jacobian matrix as
 \begin{align}\label{eq-hl-determinate}
@@
 $\bm{\Pi} = \left\{  \bm{\pi}_{s}| s\in \Ss \right\}$, $\bm{\Theta}=\left\{ \bm{\theta}_s| s\in \Ss \right\}$. 
-Then the problem in \eqref{eq:generative-model-update} becomes:
+Then the problem in \eqref{eq:generative-model-update} becomes
 \begin{align}\label{eq:sub-gm}
@@
 $\umax{\bm{\Theta}} \Qq(\bm{\Theta}; \bm{H}^{\mathrm{old}})$.
-This is boiled down to maximizing the cost function of neural networks that can be formulated as
+This is boiled down to maximize the cost function of neural networks that can be formulated as
 \begin{align}\label{eq:obj-q-gen-mix-log}
@@
 In \eqref{eq:obj-q-gen-mix-log}, the state posterior $p(s_t| \ubm{x}, \bm{H}^{\mathrm{old}})$ is computed by forward-backward algorithm.
-The posterior of $\kappa$ is:
+The posterior of $\kappa$ is
 \begin{align}\label{eq:kappa-posterior}
@@
 \end{align}
-where the last equation is due to the fact that $\bm{x}_t$ among sequence $\ubm{x}$ depends on $s_t, \kappa_t$ only.
+where the last equation is due to the fact that $\bm{x}_t$ among sequence $\ubm{x}$ only depends on $s_t, \kappa_t$.
 
@@
 In what follows we address the problem $\max_{\bm{\Pi}} \Qq(\bm{\Pi}; \bm{H}^{\mathrm{old}})$ in our generative model learning.
-The conditional distribution of hidden variable $\kappa$, $\pi_{s, \kappa} = p(\kappa | s; \bm{H})$, is obtained by solving the following problem:
+The conditional distribution of hidden variable $\kappa$, $\pi_{s, \kappa} = p(\kappa | s; \bm{H})$, is obtained by solving the following problem
 \begin{align}\label{opm:pi}
@@
 \begin{equation}
-  \lambda_s = \sum_{\kappa=1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}} p(\smtr{s}{t}{r}=s, \smtr{\kappa}{t}{r} =\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}})
+  \lambda_s = \sum_{\kappa=1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}} p(\smtr{s}{t}{r}=s, \smtr{\kappa}{t}{r} =\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}).
 \end{equation}
@@
 \begin{equation}\label{eq:mix-latent-parameter-solution}
-  \pi_{s, \kappa} = \frac{\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }{\sum_{k =1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}-1} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=k | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }
+  \pi_{s, \kappa} = \frac{\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }{\sum_{k =1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}-1} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=k | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) },
 \end{equation}
@@
 
-With the generative modeling learning obtained, it remains to solve the initial distribution update and transition matrix update of HMM in GenHMM, i.e.
+With the generative model learning obtained, it remains to solve the initial distribution update and transition matrix update of HMM in GenHMM, i.e.
 the problem \eqref{eq:init-distribution-update} and \eqref{eq:transition-update}.
-These two problem are basically two constrained optimization problems.
-The solution to them is provided in \cite{Bishop:2006:PRM:1162264}.
+These two problems are basically two constrained optimization problems.
+The solutions to them are available in literature \cite{Bishop:2006:PRM:1162264}.
 But to keep learning algorithm for GenHMM complete, we give the update rules for $\bm{q}$ and $\bm{A}$ as follows.
@@
 \subsubsection{Initial Probability Update}
-Problem in \eqref{eq:init-distribution-update} can be reformulated as:
+The problem in \eqref{eq:init-distribution-update} can be reformulated as
 \begin{align}
@@
   % & = \frac{1}{R}\sum_{r=1}^{R}\sum_{\smtr{s}{1}{r}=1}^{|\Ss|}\sum_{\smtr{s}{2}{r}=1}^{|\Ss|}\cdots \sum_{\smtr{s}{T^{r}}{r}}^{{|\Ss|}} {p(\smtr{s}{1}{r}, \smtr{s}{2}{r}, \cdots, \smtr{s}{T^{r}}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r}) \\
-  & = \frac{1}{R}\sum_{r=1}^{R}\sum_{\smtr{s}{1}{r}=1}^{|\Ss|}{p(\smtr{s}{1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r};\bm{H}) 
+  & = \frac{1}{R}\sum_{r=1}^{R}\sum_{\smtr{s}{1}{r}=1}^{|\Ss|}{p(\smtr{s}{1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r};\bm{H}).
 \end{align}
 
-$p(\smtr{s}{1}{r};\bm{H})$ is the probability of initial state of GenHMM for $r$-th sequential sample. Actually $q_i = p({s}_{1} =i;\bm{H}) $, $i= 1, 2, \cdots, |\Ss|$. Solution to problem:
+$p(\smtr{s}{1}{r};\bm{H})$ is the probability of initial state of GenHMM for $r$-th sequential sample. Actually $q_i = p({s}_{1} =i;\bm{H}) $, $i= 1, 2, \cdots, |\Ss|$. Solution to the problem
 \begin{align}
-  \bm{q} &= \uargmax{\bm{q}} \Qq(\bm{q}; \bm{H}^{\mathrm{old}}), \nonumber \\
-         &\mathrm{s.t.} \sum_{i=1}^{ |\Ss| }q_i = 1, q_i \geq 0, \forall s.
+  \bm{q} &= \uargmax{\bm{q}} \Qq(\bm{q}; \bm{H}^{\mathrm{old}}) \nonumber \\
+         &\mathrm{s.t.} \sum_{i=1}^{ |\Ss| }q_i = 1, q_i \geq 0, \forall i.
 \end{align}
@@
 \subsubsection{Transition Probability Update}
-Problem \eqref{eq:transition-update} can be reformulated as
+The problem \eqref{eq:transition-update} can be reformulated as
 \begin{align}
@@
 
-Since $\bm{A}_{i, j}  = p(\smtr{s}{t+1}{r}=j|\smtr{s}{t}{r}=i; \bm{H})$ is the element of transition matrix $\bm{A}$, the solution to problem:
+Since $\bm{A}_{i, j}  = p(\smtr{s}{t+1}{r}=j|\smtr{s}{t}{r}=i; \bm{H})$ is the element of transition matrix $\bm{A}$, the solution to the problem
 \begin{align}\label{eq:update-transition-prob}
-  \bm{A} = &\uargmax{\bm{A}} \Qq(\bm{A}; \bm{H}^{\mathrm{old}}), \nonumber \\
+  \bm{A} = &\uargmax{\bm{A}} \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) \nonumber \\
   \mathrm{s.t.} &\hspace{0.2cm} \bm{A} \cdot \bm{1} = \bm{1} \nonumber \\
-           & \bm{A}_{i,j} \geq 0.
+           & \bm{A}_{i,j} \geq 0 \,\, \forall i,j,
 \end{align}
@@
 \begin{proof}
-  We begin with the comparison of log-likelihood evaluated under $\bm{H}^{\mathrm{old}}$ and $\bm{H}^{\mathrm{old}}$. The log-likelihood of dataset given by $\hat{p}(\ubm{x})$ can be reformulated as
+  We begin with the comparison of loglikelihood evaluated under $\bm{H}^{\mathrm{new}}$ and $\bm{H}^{\mathrm{old}}$. The loglikelihood of dataset given by $\hat{p}(\ubm{x})$ can be reformulated as
   \begin{align*}
@@
 the parameters of generators in GenHMM.
-As is shown from line $6$ to $12$ in Algorithm~\ref{algo:genhmm}, the batch-size stochastic gradient decent can be naturally embedded into the learning algorithm of GenHMM.
+As shown from line $6$ to $12$ in Algorithm~\ref{algo:genhmm}, the batch-size stochastic gradient decent can be naturally embedded into the learning algorithm of GenHMM.
 
-As described by the pseudocode in Algorithm~\ref{algo:genhmm}, the learning of GenHMM is divided into optimization w.r.t.
+As described by the pseudocode in Algorithm~\ref{algo:genhmm}, the learning of GenHMM is divided into optimizations w.r.t.
 to generators' parameters $\bm{\Theta}$, initial probability $\bm{q}$ of hidden state, transition matrix $\bm{A}$, and generator mixture weights $\bm{\Pi}$.
@@
       Initializing $\bm{H}^{\mathrm{old}}, \bm{H} \in \Hh$ gives: \\
-      $\bm{H}^{\mathrm{old}} = \{\Ss, \bm{q}^{\mathrm{old}}, A^{\mathrm{old}}, p(\bm{x}|s; \bm{\Phi}^{\mathrm{old}})\}$, $\bm{H} = \{\Ss, \bm{q}, A, p(\bm{x}|s; \bm{\Phi})\}$\\
-      Empirical distribution $\hat{p}(\bm{x})$ of dataset;} \\
+      $\bm{H}^{\mathrm{old}} = \{\Ss, \bm{q}^{\mathrm{old}}, A^{\mathrm{old}}, p(\bm{x}|s; \bm{\Phi}^{\mathrm{old}})\}$, $\bm{H} = \{\Ss, \bm{q}, A, p(\bm{x}|s; \bm{\Phi})\}$,\\
+      empirical distribution $\hat{p}(\bm{x})$ of dataset;} \\
     \STATE Initialize $\bm{H}$
     \STATE $\bm{H}^{\mathrm{old}} \gets \bm{H}$
-    \STATE Set learning rate $\eta$, neural network optimization epochs $N$ per EM step.
+    \STATE Set learning rate $\eta$, neural network optimization epochs $N$ per EM step
     \FOR { $\bm{H}$ not converge}
@@
     \ENDFOR
-    \STATE $\bm{q} \gets \uargmax{\bm{q}}\, \Qq(\bm{q}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:update-initial-state-prob};
-    \STATE $\bm{A} \gets \uargmax{\bm{A}}\Qq(\bm{A}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:update-transition-solt};
-    \STATE $\bm{\Pi} \gets \uargmax{\bm{\Pi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:mix-latent-parameter-solution};
+    \STATE $\bm{q} \gets \uargmax{\bm{q}}\, \Qq(\bm{q}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:update-initial-state-prob}
+    \STATE $\bm{A} \gets \uargmax{\bm{A}}\Qq(\bm{A}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:update-transition-solt}
+    \STATE $\bm{\Pi} \gets \uargmax{\bm{\Pi}}\Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}})$ by \eqref{eq:mix-latent-parameter-solution}
     \STATE $\bm{H}^{\mathrm{old}} \gets \bm{H}$
@@
 The code is available at github repository\footnote{Due to double-blind review, the author's repository is not present here temporarily.}.
-In the next subsection, we first discuss the experimental setups and the show the experimental results.
+In the next subsection, we first discuss the experimental setups and then show the experimental results.
 
@@
 The TIMIT dataset consists of $5300$ phoneme-labeled speech utterances which are partitioned into two sets: \textcolor{blue}{a train set consists of $4620$ utterance, and a test set consists of $1680$ utterances.} There are totally $61$ different types of phones in TIMIT labels.
-Our experiments in this section is carried out in two different cases: i) full $61$-phoneme classification case, where both training and testing use the $61$-phoneme data; ii) $39$-phonme classification case, where $61$ phonemes are folded into $39$ phonemes as described in \cite{Perdigao11}.
+Our experiments in this section are carried out in two different cases: i) full $61$-phoneme classification case, where both training and testing use the $61$-phoneme data; ii) $39$-phonme classification case, where $61$ phonemes are folded into $39$ phonemes as described in \cite{Perdigao11}.
 \textcolor{blue}{The mapping table for the folding process is attached in appendix.}
@@
 
-Table \ref{tab:acc-classification39f} shows the results for this experiments, in which we tested both the folded $39$-phoneme classification case (the conventional way) in table \ref{tab:acc-classification39f_a} and also the $61$-phoneme classification case in table \ref{tab:acc-classification39f_b}.
+Table \ref{tab:acc-classification39f} shows the results for this experiments, in which we test both the folded $39$-phoneme classification case (the conventional way) in table \ref{tab:acc-classification39f_a} and the $61$-phoneme classification case in table \ref{tab:acc-classification39f_b}.
 As shown in both $61$-phoneme and $39$-phoneme cases, GenHMM gets significant higher accuracy than GMM-HMM for the same number of mixture components.
@@
 The total training dataset size remains the same as $61$ phonemes are folded into $39$ phonemes.
-There are less training data available and more classes to be recognized in the $61$-phoneme case, which makes the take more challenging.
+There are less training data available and more classes to be recognized in the $61$-phoneme case, which makes the task more challenging.
 
@@
 
-Similar experiments to that in table~\ref{tab:acc-classification13f} are carried out by using only the MFCC coefficients as feature input (excluding deltas and delta-deltas).
+Similar experiments in table~\ref{tab:acc-classification13f} are carried out by using only the MFCC coefficients as feature input (excluding deltas and delta-deltas).
 The results are shown in table~\ref{tab:acc-classification13f}.
 The superior performance of GenHMM remains compared with reference model GMM-HMM.
-The gain by using mixture generators are also present in this set of experiments while the difference between $61$-phoneme and $39$-phoneme cases is also similar to the set of experiments in table~\ref{tab:acc-classification39f}.
+The gain by using mixture generators is also presented in this set of experiments while the difference between $61$-phoneme and $39$-phoneme cases is also similar to the set of experiments in table~\ref{tab:acc-classification39f}.
 
@@
 In this work, we proposed a generative model based HMM (GenHMM) whose generators are realized by neural networks.
-We provide the training method for GenHMM.
-The validity of GenHMM is demonstrated by the experiments of classification task on practical sequential dataset.
+We provided the training method for GenHMM.
+The validity of GenHMM was demonstrated by the experiments of classification tasks on practical sequential dataset.
 The learning method in this work is based on generative training.
-Future work would consider discriminative training for classification tasks of sequential data.
+For future work, we would consider discriminative training for classification tasks of sequential data.
 

### 2019-08-28 src/timit-preprocessor/add_noise.py
change the option for easier integration with make

@@ -3,6 +3,8 @@ from multiprocessing.dummy import Pool
 from scipy.io import wavfile
 import argparse
 from glob import glob
+from parse import parse
+
 import os, sys
 import numpy as np
 
@@ -13,9 +15,9 @@ def gen_noise(n, type, sigma, noise_folder="../../data/NoiseDB/NoiseX_16kHz"):
     if type == "white":
         noise_filename = os.path.join(noise_folder, "{}_16kHz.wav".format(type))
         _, loaded_noise = wavfile.read(noise_filename)
+
         try:
             assert(n < loaded_noise.shape[0])
-
         except AssertionError as e:
             print("Noise file: {} is too short.".format(noise_filename), file=sys.stderr)
 
@@ -30,45 +32,49 @@ def gen_noise(n, type, sigma, noise_folder="../../data/NoiseDB/NoiseX_16kHz"):
     return raw_noise * sigma
 
 
-def new_filename(file, type, snr):
+def new_filename(file, ntype, snr):
     """Append noise tyep and power at the end of wav filename."""
-    return file.replace(".WAV", ".WAV.{}.{}dB".format(type, snr))
+    return file.replace(".WAV", ".WAV.{}.{}dB".format(ntype, snr))
 
 
-def corrupt_data(s, type, snr):
+def corrupt_data(s, ntype, snr):
     """Corrupt a signal with a particular noise."""
     s_std = np.std(s)
     n_std = 10 ** (- snr / 10) * s_std
-    n = gen_noise(s.shape[0], type, n_std)
+    n = gen_noise(s.shape[0], ntype, n_std)
     sn = (s + n).astype(s.dtype)
     return sn
 
 
-def corrupt_wav(file, type=None, snr=None):
+def corrupt_wav(file, ntype=None, snr=None):
     """Corrupt a wav file with noise and write to a new file."""
     rate, s = wavfile.read(file)
-    sn = corrupt_data(s, type, snr)
-    wavfile.write(new_filename(file, type, snr), rate, sn)
+    sn = corrupt_data(s, ntype, snr)
+    wavfile.write(new_filename(file, ntype, snr), rate, sn)
     return 0
 
 
 if __name__ == "__main__":
     parser = argparse.ArgumentParser(description="Add a particular noise type to WAV files.")
 
-    parser.add_argument('-i', metavar="<Input folder>", type=str)
[…]

### 2019-08-28 src/timit-preprocessor/parsing.py
Make sure to refer to the noisy wav file

@@ -3,13 +3,27 @@
 import os
 import sys
 import argparse
+from parse import parse
 
-def main():
-    target_folder = os.path.join(args.datapath, args.folder.upper())
-    target = args.folder
 
-    features = '{}/{}.wav.scp'.format(MATERIAL_PATH, target)
-    labels = '{}/{}.lbl'.format(MATERIAL_PATH, target)
+def main(args=None):
+    try:
+        target, ntype, snr = parse("{}.{}.{:d}dB", args.folder)
+    except TypeError as e:
+        target = args.folder
+        ntype="clean"
+        snr=0
+
+    target_folder = os.path.join(args.datapath, target.upper())
+
+    #target = args.folder
+
+    features = '{}/{}.wav.scp'.format(MATERIAL_PATH, args.folder)
+    labels = '{}/{}.lbl'.format(MATERIAL_PATH, args.folder)
+    if ntype == "clean":
+        extension = ".WAV"
+    else:
+        extension = ".WAV.{}.{}dB".format(ntype, snr)
     
     with open(features, "w+") as f, open(labels, "w+") as l:
         for person in os.listdir(target_folder):
@@ -19,7 +33,7 @@ def main():
                 raw_sentences = [x.split('.')[0] for x in os.listdir(task_dir)]
                 sentences = list(set(raw_sentences))
                 for sentence in sentences:
-                    wav_file = os.path.join(task_dir, sentence + '.WAV')
+                    wav_file = os.path.join(task_dir, sentence + extension)
                     f.write('{}-{}-{} {}\n'.format(person, task, sentence, os.path.abspath(wav_file)))
 
                     phone_seq_file = os.path.join(task_dir, sentence+'.PHN')
@@ -35,10 +49,11 @@ if __name__ == "__main__":
     parser = argparse.ArgumentParser(description="parsing .wav.scp files for advanced use\ne.g. python3 parsing.py ~/Workspace/data/timit train", formatter_class=argparse.RawTextHelpFormatter)
     parser.add_argument('datapath', metavar="<timit relative path>", type=str)
     parser.add_argument('folder', metavar="<train|test>", type=str)
+
     args = parser.parse_args()
 
     MATERIAL_PATH = '.data/material'
     if not os.path.exists(MATERIAL_PATH):
         os.makedirs(MATERIAL_PATH)
 
-    main()
+    main(args)

### 2019-08-28 src/utils.py
Move some helper functions to utils

@@ -6,6 +6,139 @@ from torch.utils.data import DataLoader
 import sys
 import pickle as pkl
 from parse import parse
+import json
+from functools import partial
+
+
+def test_norm_minmax():
+    x = np.array([[1, 3], [2, 2]])
+    min_ = x.min(0)
+    max_ = x.max(0)
+    xn = norm_minmax(x, min_=min_, max_=max_)
+    assert ((xn == np.array([[0, 1], [1, 0]])).all())
+
+
+def norm_minmax(x, min_=None, max_=None):
+    return ((x - min_.reshape(1, -1)) / (max_.reshape(1, -1) - min_.reshape(1, -1)))
+
+
+def normalize(xtrain, xtest):
+    """Normalize training data set between 0 and 1. Perform the same scaling on the testing set."""
+    f_min = np.vectorize(lambda x : np.min(x, axis=0), signature="()->(k)")
+    f_max = np.vectorize(lambda x : np.max(x, axis=0), signature="()->(k)")
+    min_tr = np.min(f_min(xtrain), axis=0)
+    max_tr = np.max(f_max(xtrain), axis=0)
+
+    # The first component is zeros and can create division by 0
+    min_tr[0] = 0
+    max_tr[0] = 1
+    f_perform_normalize = np.vectorize(partial(norm_minmax, min_=min_tr, max_=max_tr), signature="()->()", otypes=[np.ndarray])
+    return f_perform_normalize(xtrain), f_perform_normalize(xtest)
+
+
+def getsubset(data, label, iphn):
+    # concat data
+    # find subset
+    idx = np.in1d(label, iphn)
+    return data[idx], label[idx]
+
+
+def find_change_loc(x):
+    dx = np.diff(x)
+    # Make a clean vector to delimit phonemes
+    change_locations = np.array([0] + (1 + np.argwhere(dx != 0)).reshape(-1).tolist() + [x.shape[0]])
+    # Make an array of size n_phoneme_in_sentence x 2, containing begining and end of each phoneme in a sentence
+
+    fmt_interv = np.array([[change_locations[i-1], change_locations[i]]\
+                                 for i in range(1, change_locations.shape[0]) ])
+    return fmt_interv, x[change_locations[:-1]]
+
+
+def test_find_change_loc():
+    l = np.array([1,1,1,1,1,1,1,0,0,0,0,0,0,2,2,2,2,2,2,3])
+    out, out2 = find_change_loc(l)
+    assert((out2 == np.array([1,0,2,3])).all())
[…]

### 2019-08-28 src/timit-preprocessor/add_noise.py
Error in the def of snr

@@ -29,7 +29,7 @@ def gen_noise(n, type, sigma, noise_folder="../../data/NoiseDB/NoiseX_16kHz"):
         print("Unknown {} noise".format(type), file=sys.stderr)
         raw_noise = 0
 
-    return raw_noise * sigma
+    return raw_noise / raw_noise.std() * sigma
 
 
 def new_filename(file, ntype, snr):
@@ -40,7 +40,7 @@ def new_filename(file, ntype, snr):
 def corrupt_data(s, ntype, snr):
     """Corrupt a signal with a particular noise."""
     s_std = np.std(s)
-    n_std = 10 ** (- snr / 10) * s_std
+    n_std = 10 ** (- snr / 20) * s_std
     n = gen_noise(s.shape[0], ntype, n_std)
     sn = (s + n).astype(s.dtype)
     return sn
@@ -69,11 +69,9 @@ if __name__ == "__main__":
         print("No noise to be added with option: {}.\nExit.".format(args.opt),file=sys.stderr)
         sys.exit(0)
 
-
     if dset == "test":
         wavs = glob(os.path.join(args.timit, "TEST", "**" , "*.WAV"), recursive=True)
         f = partial(corrupt_wav, ntype=ntype, snr=snr)
-
         with Pool(args.j) as pool:
             pool.map(f, wavs)
 

### 2019-08-28 src/timit-preprocessor/parsing.py
line break for clarity.

@@ -11,8 +11,8 @@ def main(args=None):
         target, ntype, snr = parse("{}.{}.{:d}dB", args.folder)
     except TypeError as e:
         target = args.folder
-        ntype="clean"
-        snr=0
+        ntype = "clean"
+        snr = 0
 
     target_folder = os.path.join(args.datapath, target.upper())
 
@@ -44,11 +44,14 @@ def main(args=None):
                     l.write('{}-{}-{} {}\n'.format(person, task, sentence, ','.join(phone_seq)))
             
 
-
 if __name__ == "__main__":
-    parser = argparse.ArgumentParser(description="parsing .wav.scp files for advanced use\ne.g. python3 parsing.py ~/Workspace/data/timit train", formatter_class=argparse.RawTextHelpFormatter)
-    parser.add_argument('datapath', metavar="<timit relative path>", type=str)
-    parser.add_argument('folder', metavar="<train|test>", type=str)
+    parser = argparse.ArgumentParser(
+        description="parsing .wav.scp files for advanced use\n"
+                    "e.g. python3 parsing.py ~/Workspace/data/timit train", formatter_class=argparse.RawTextHelpFormatter)
+    parser.add_argument('datapath',
+                        metavar="<timit relative path>", type=str)
+    parser.add_argument('folder',
+                        metavar="<train|test>", type=str)
 
     args = parser.parse_args()
 

### 2019-08-28 bin/split_dataset.py
Split a dataset according to an existing class_map.json.

@@ -1,0 +1,53 @@
+from parse import parse
+import argparse
+import os
+import pickle as pkl
+from gm_hmm.src.utils import read_classmap,to_phoneme_level,flip, phn61_to_phn39,remove_label, getsubset, normalize
+from functools import partial
+
+if __name__ == "__main__":
+    parser = argparse.ArgumentParser(description="Split a test dataset according to an existing class map.")
+    parser.add_argument('-input', metavar="<Input dataset to split>", type=str)
+    parser.add_argument('-classmap', metavar="<class_map.json file>", type=str)
+    parser.add_argument('-totclass', metavar="<Total number of classes>", type=int)
+    args = parser.parse_args()
+
+    fname_dtest = args.input
+    mode, nfeats, ntype, snr = parse("{}.{}.{}.{:d}db.pkl", os.path.basename(fname_dtest))
+    fname_dtrain = os.path.join(os.path.dirname(fname_dtest), "train.{}.pkl".format(nfeats))
+
+    cmap = read_classmap(os.path.dirname(args.classmap))
+    te_DATA, te_keys, te_lengths, phn2int_61, te_PHN = pkl.load(open(fname_dtest, "rb"))
+
+    tr_DATA, tr_keys, tr_lengths, tr_PHN = pkl.load(open(fname_dtrain, "rb"))
+
+    data_te, label_te = to_phoneme_level(te_DATA)
+    data_tr, label_tr = to_phoneme_level(tr_DATA)
+    phn2int = phn2int_61
+    if args.totclass == 39:
+        f = partial(phn61_to_phn39, int2phn_61=flip(phn2int_61), data_folder=os.path.dirname(fname_dtest))
+        label_tr, phn2int_39 = f(label_te)
+        label_te, _ = f(label_te, phn2int_39=phn2int_39)
+
+        data_te, label_te = remove_label(data_te, label_te, phn2int_39)
+        phn2int_39.pop('-', None)
+        phn2int = phn2int_39
+
+    iphn = [phn2int[v] for k, v in cmap.items()]
+
+    xtrain, ytrain = getsubset(data_tr, label_tr, iphn)
+    xtest, ytest = getsubset(data_te, label_te, iphn)
+    xtrain, xtest = normalize(xtrain, xtest)
+
+    # Assert length (If we add an already existing phoneme,
+    # the dictionary size will not be len(classmap) + len(class2phn)
+
[…]

### 2019-08-28 src/utils.py
Do not crash when class datafiles are missing. Skip the training instead.

@@ -159,7 +159,11 @@ def accuracy_fun(data_file, mdl=None):
 
 
 def accuracy_fun_torch(data_file, mdl=None, batch_size_=128):
-    X = pkl.load(open(data_file, "rb"))
+    try:
+        X = pkl.load(open(data_file, "rb"))
+    except:
+        return "0/1"
+
     # Get the length of all the sequences
     l = [xx.shape[0] for xx in X]
     # zero pad data for batch training
@@ -169,7 +173,7 @@ def accuracy_fun_torch(data_file, mdl=None, batch_size_=128):
                                               lengths=l,
                                               device=mdl.hmms[0].device),
                            batch_size=batch_size_, shuffle=True)
-    
+
     true_class = parse("{}_{}.pkl", os.path.basename(data_file))[1]
     out_list = [mdl.forward(x) for x in batchdata]
     out = torch.cat(out_list, dim=1)
@@ -180,7 +184,7 @@ def accuracy_fun_torch(data_file, mdl=None, batch_size_=128):
 
     return acc_str(class_hat, true_class)
 
-def acc_str(class_hat,class_true):
+def acc_str(class_hat, class_true):
     istrue = class_hat == int(class_true)
     return "{}/{}".format(str(istrue.sum().cpu().numpy()), str(istrue.shape[0]))
 
@@ -191,11 +195,12 @@ def test_acc_str():
     assert( acc_str(class_hat, 1) == "1/2" )
     assert( acc_str(class_hat, 2) == "1/2" )
 
-    class_hat = torch.FloatTensor([1,2,3,4,5,6,7,8,9,10])
+    class_hat = torch.FloatTensor([1, 2, 3, 4, 5, 6, 7, 8, 9, 10])
     assert( acc_str(class_hat, 1) == "1/10")
 
     class_hat = torch.FloatTensor([1, 2, 3, 4, 5, 6, 7, 8, 1, 10])
     assert (acc_str(class_hat, 1) == "2/10")
+
 
 def append_class(data_file, iclass):
     return data_file.replace(".pkl", "_" + str(iclass)+".pkl")
@@ -203,7 +208,6 @@ def append_class(data_file, iclass):
 
 def divide(res_int):
     return res_int[0] / res_int[1]
-
 
 
 def parse_(res_str):

### 2019-08-28 src/utils.py
Do not crash when class datafiles are missing. Skip the training instead.

@@ -142,7 +142,10 @@ def write_classmap(class2phn, folder):
 
 
 def accuracy_fun(data_file, mdl=None):
-    X = pkl.load(open(data_file, "rb"))
+    try:
+        X = pkl.load(open(data_file, "rb"))
+    except:
+        return "0/1"
     # Get the length of all the sequences
     l = [xx.shape[0] for xx in X]
     # zero pad data for batch training

### 2019-08-30 bin/compute_accuracy_class.py
minor

@@ -11,7 +11,8 @@ import time
 import numpy as np
 
 if __name__ == "__main__":
-    usage = "bin/compute_accuracy_class.py exp/gaus/13feats/models/epoch2_class1.mdlc exp/gaus/13feats/data/train.13.pkl exp/gaus/13feats/data/test.13.pkl"
+    usage = "Usage:\n" \
+            "bin/compute_accuracy_class.py exp/gaus/13feats/models/epoch2_class1.mdlc exp/gaus/13feats/data/train.13.pkl exp/gaus/13feats/data/test.13.pkl"
     if len(sys.argv) != 4 or sys.argv[1] == "-h" or sys.argv[1] == "--help":
         print(usage, file=sys.stderr)
         sys.exit(1)
@@ -33,6 +34,7 @@ if __name__ == "__main__":
     # Builds an array of string containing the train and test data sets for each class
     # size: nclass x 2 (train, test)
     data_files = [append_class(training_data_file, iclass), append_class(testing_data_file, iclass)]
+
     # Load Model
     device = 'cpu'
     if model_type == 'gaus':
@@ -40,6 +42,7 @@ if __name__ == "__main__":
             mdl = pkl.load(handle)
         mdl.device = 'cpu'
         f = lambda x: accuracy_fun(x, mdl=mdl)
+
     elif model_type == 'gen':
         mdl = load_model(mdl_file)
         if torch.cuda.is_available():
@@ -68,9 +71,7 @@ if __name__ == "__main__":
 
 
     # print("[Acc:] epoch:{}\tclass:{}\tPush model to {}. Done.".format(epoch,iclass, mdl.device), file=sys.stdout)
-    
-   
-    #f = lambda x: divide(parse_(accuracy_fun(x, mdl=mdl)))
+    # f = lambda x: divide(parse_(accuracy_fun(x, mdl=mdl)))
     
     results = list(map(f, data_files))
 

### 2019-08-30 bin/split_dataset.py
Fix 39labels issue, first attemp to be tested properly

@@ -1,8 +1,10 @@
 from parse import parse
+import sys
 import argparse
 import os
 import pickle as pkl
-from gm_hmm.src.utils import read_classmap,to_phoneme_level,flip, phn61_to_phn39,remove_label, getsubset, normalize
+from gm_hmm.src.utils import read_classmap, to_phoneme_level,\
+    flip, phn61_to_phn39, remove_label, getsubset, normalize
 from functools import partial
 
 if __name__ == "__main__":
@@ -17,8 +19,8 @@ if __name__ == "__main__":
     fname_dtrain = os.path.join(os.path.dirname(fname_dtest), "train.{}.pkl".format(nfeats))
 
     cmap = read_classmap(os.path.dirname(args.classmap))
-    te_DATA, te_keys, te_lengths, phn2int_61, te_PHN = pkl.load(open(fname_dtest, "rb"))
 
+    te_DATA, te_keys, te_lengths, phn2int_61, te_PHN = pkl.load(open(fname_dtest, "rb"))
     tr_DATA, tr_keys, tr_lengths, tr_PHN = pkl.load(open(fname_dtrain, "rb"))
 
     data_te, label_te = to_phoneme_level(te_DATA)
@@ -26,10 +28,12 @@ if __name__ == "__main__":
     phn2int = phn2int_61
     if args.totclass == 39:
         f = partial(phn61_to_phn39, int2phn_61=flip(phn2int_61), data_folder=os.path.dirname(fname_dtest))
-        label_tr, phn2int_39 = f(label_te)
+        label_tr, phn2int_39 = f(label_tr)
         label_te, _ = f(label_te, phn2int_39=phn2int_39)
 
         data_te, label_te = remove_label(data_te, label_te, phn2int_39)
+        data_tr, label_tr = remove_label(data_tr, label_tr, phn2int_39)
+
         phn2int_39.pop('-', None)
         phn2int = phn2int_39
 
@@ -50,4 +54,4 @@ if __name__ == "__main__":
         xtest_c = xtest[ytest == ic]
         pkl.dump(xtest_c, open(test_outfiles[i], "wb"))
 
-    print("here")
+    sys.exit(0)

### 2019-08-30 bin/split_dataset.py
fix 39class split

@@ -1,4 +1,5 @@
 from parse import parse
+import sys
 import argparse
 import os
 import pickle as pkl
@@ -26,10 +27,12 @@ if __name__ == "__main__":
     phn2int = phn2int_61
     if args.totclass == 39:
         f = partial(phn61_to_phn39, int2phn_61=flip(phn2int_61), data_folder=os.path.dirname(fname_dtest))
-        label_tr, phn2int_39 = f(label_te)
+        label_tr, phn2int_39 = f(label_tr)
         label_te, _ = f(label_te, phn2int_39=phn2int_39)
 
         data_te, label_te = remove_label(data_te, label_te, phn2int_39)
+        data_te, label_te = remove_label(data_te, label_te, phn2int_39)
+
         phn2int_39.pop('-', None)
         phn2int = phn2int_39
 
@@ -50,4 +53,5 @@ if __name__ == "__main__":
         xtest_c = xtest[ytest == ic]
         pkl.dump(xtest_c, open(test_outfiles[i], "wb"))
 
-    print("here")
+    
+    sys.exit(0)

### 2019-08-30 manuscript/main.tex
formate adjust

@@
   \caption{HMM model illustration.}\label{fig:hmm}
-  \vspace{0.1cm}
+  \vspace{-0.1cm}
 \end{figure}
@@
   \label{fig:gen-mix}
-  \vspace{0.1cm}
+  \vspace{-0.2cm}
 \end{figure}
@@
 For a latent variable $\bm{z}$ with density function $p_{s,\kappa}(\bm{z})$, the generated signal $\bm{x}$ follows an induced distribution with density function \eqref{eq:changel-variable}.
-We illustrate the signal flow between latent variable $\bm{z}$ and observable variable $\bm{x}$ as follows
+We illustrate the signal flow between latent variable $\bm{z}$ and observable variable $\bm{x}$ as
 \begin{equation*}
@@
 The efficient invertible mapping of flow model comes from following forward and inverse relations between $(l-1)$'th and $l$'th layers
-\begin{equation}\label{eq-gl}
-  \begin{array}{l}
-    \bm{h}_{l} =
+\begin{align}\label{eq-gl}
+    \bm{h}_{l} &=
     \begin{bmatrix}
@@
       \left(  \bm{h}_{l-1,b} - \bm{m}_b(\bm{h}_{l-1,a}) \right)\oslash \bm{m}_a(\bm{h}_{l-1,a}) 
-    \end{bmatrix}, \vspace{10pt} \\
-        \bm{h}_{l-1} =
+    \end{bmatrix}, \nonumber\\
+        \bm{h}_{l-1} \hspace{-0.1cm}&=\hspace{-0.1cm}
     \begin{bmatrix}
@@
     \end{bmatrix}
-    =
+    \hspace{-0.1cm}=\hspace{-0.1cm}
     \begin{bmatrix}
@@
       \bm{m}_a(\bm{h}_{l,a})\odot \bm{h}_{l,b} + \bm{m}_b(\bm{h}_{l,a})
-    \end{bmatrix},\\
-  \end{array}  
-\end{equation}
+    \end{bmatrix},
+\end{align}
 where $\odot$ denotes element-wise product, $\oslash$ denotes

### 2019-09-01 manuscript/main.tex
add noise testing

@@
     &= \EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p({s}_{1};\bm{H})  \right], \label{eq:init-distribution-update}\\
-  \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \sum_{t=1}^{T}\log\,p({s}_{t+1}|{s}_{t}; \bm{H}) \right], \label{eq:transition-update}\\
+  \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})}\hspace{-0.1cm}\left[ \sum_{t=1}^{T}\log\,p({s}_{t+1}|{s}_{t}; \bm{H}) \right], \label{eq:transition-update}\\
   \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) &= \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p(\ubm{x}, \ubm{\kappa}| \ubm{s}; \bm{H}) \right].
@@
 complex non-linear mappings (implemented by neural networks).
-For the flow-based neural network, the determinant of Jacobian matrix is
+For the flow model, the determinant of Jacobian matrix is
 \begin{equation}\label{eq:cat-jacobian}
@@
 \end{equation}
-where $\nabla{f_{s,\kappa}^{[l]}}$ is the Jacobian of the transformation from the $l$-th layer to the $(l-1)$-th layer, i.e., the inverse transformation.
-We compute the determinant of the Jacobian matrix as
+where $\nabla{f_{s,\kappa}^{[l]}}$ is the Jacobian of the mapping from the $l$-th layer to the $(l-1)$-th layer, i.e., the inverse transformation. We compute the determinant of the Jacobian matrix as
 \begin{align}\label{eq-hl-determinate}
@@
 \begin{align}
-  \bm{q} &= \uargmax{\bm{q}} \Qq(\bm{q}; \bm{H}^{\mathrm{old}}) \nonumber \\
-         &\mathrm{s.t.} \sum_{i=1}^{ |\Ss| }q_i = 1, q_i \geq 0, \forall i.
+  \bm{q} \hspace{-0.1cm} = \hspace{-0.1cm} \uargmax{\bm{q}} \Qq(\bm{q}; \bm{H}^{\mathrm{old}}),\; \mathrm{s.t.} \sum_{i=1}^{ |\Ss| }q_i = 1, q_i \geq 0, \forall i.
 \end{align}
@@
   \bm{A} = &\uargmax{\bm{A}} \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) \nonumber \\
-  \mathrm{s.t.} &\hspace{0.2cm} \bm{A} \cdot \bm{1} = \bm{1} \nonumber \\
-           & \bm{A}_{i,j} \geq 0 \,\, \forall i,j,
+  \mathrm{s.t.} &\hspace{0.2cm} \bm{A} \cdot \bm{1} = \bm{1}, \bm{A}_{i,j} \geq 0 \,\, \forall i,j,
 \end{align}
@@
 \subsection{Algorithm of GenHMM}
-
-To summarize the learning solution in subsection~\ref{subsec:optmGenHMM}, we wrap our algorithm into pseudocode as shown in Algorithm~\ref{algo:genhmm}.
-We use Adam \cite{DBLP:journals/corr/KingmaB14} optimizer for optimization w.r.t.
-the parameters of generators in GenHMM.
-As shown from line $6$ to $12$ in Algorithm~\ref{algo:genhmm}, the batch-size stochastic gradient decent can be naturally embedded into the learning algorithm of GenHMM.
-
-As described by the pseudocode in Algorithm~\ref{algo:genhmm}, the learning of GenHMM is divided into optimizations w.r.t.
-to generators' parameters $\bm{\Theta}$, initial probability $\bm{q}$ of hidden state, transition matrix $\bm{A}$, and generator mixture weights $\bm{\Pi}$.
-Different from the optimization w.r.t.
-to $\bm{q}$, $\bm{A}$ and $\bm{\Pi}$, which have optimal solutions, generator learning usually cannot give optimal solution to problem $\max_{\bm{\Theta}} \Qq(\bm{\Theta}; \bm{H}^{\mathrm{old}})$.
-In fact, given that no optimal $\bm{\Theta}$ is obtained, learning of GenHMM can still converge as long as quantity $\Qq(\bm{\Theta}; \bm{H})$ are improving in iterations in Algorithm~\ref{algo:genhmm}, where the inequalities in Proposition~\ref{proposition1} still hold.
-Therefore optimal $\bm{\Theta}$ in each iteration is not required for convergence of GenHMM as long as the loss in $\eqref{eq:obj-q-gen-mix}$ is getting improved.
-
 \begin{algorithm}[H]
@@
 
+To summarize the learning solution in subsection~\ref{subsec:optmGenHMM}, we wrap our algorithm into pseudocode as shown in Algorithm~\ref{algo:genhmm}.
+We use Adam \cite{DBLP:journals/corr/KingmaB14} optimizer for optimization w.r.t.
+the parameters of generators in GenHMM.
+As shown from line $6$ to $12$ in Algorithm~\ref{algo:genhmm}, the batch-size stochastic gradient decent can be naturally embedded into the learning algorithm of GenHMM.
+
+As described by the pseudocode in Algorithm~\ref{algo:genhmm}, the learning of GenHMM is divided into optimizations w.r.t.
+to generators' parameters $\bm{\Theta}$, initial probability $\bm{q}$ of hidden state, transition matrix $\bm{A}$, and generator mixture weights $\bm{\Pi}$.
+Different from the optimization w.r.t.
+to $\bm{q}$, $\bm{A}$ and $\bm{\Pi}$, which have optimal solutions, generator learning usually cannot give optimal solution to problem $\max_{\bm{\Theta}} \Qq(\bm{\Theta}; \bm{H}^{\mathrm{old}})$.
+In fact, given that no optimal $\bm{\Theta}$ is obtained, learning of GenHMM can still converge as long as quantity $\Qq(\bm{\Theta}; \bm{H})$ are improving in iterations in Algorithm~\ref{algo:genhmm}, where the inequalities in Proposition~\ref{proposition1} still hold.
+Therefore optimal $\bm{\Theta}$ in each iteration is not required for convergence of GenHMM as long as the loss in $\eqref{eq:obj-q-gen-mix}$ is getting improved.
+
 \section{Experiments}
@@
 
-The TIMIT dataset consists of $5300$ phoneme-labeled speech utterances which are partitioned into two sets: \textcolor{blue}{a train set consists of $4620$ utterance, and a test set consists of $1680$ utterances.} There are totally $61$ different types of phones in TIMIT labels.
+The TIMIT dataset consists of $5300$ phoneme-labeled speech utterances which are partitioned into two sets: {a train set consists of $4620$ utterance, and a test set consists of $1680$ utterances.} There are totally $61$ different types of phones in TIMIT labels.
 Our experiments in this section are carried out in two different cases: i) full $61$-phoneme classification case, where both training and testing use the $61$-phoneme data; ii) $39$-phonme classification case, where $61$ phonemes are folded into $39$ phonemes as described in \cite{Perdigao11}.
-\textcolor{blue}{The mapping table for the folding process is attached in appendix.}
 
@@
   \subcaption{Folded $39$ Phonemes}\label{tab:acc-classification39f_a}
+  \vspace{-0.2cm}
   \centering  
@@
   \end{tabular}
-  \bigskip
   \subcaption{$61$ Phonemes}\label{tab:acc-classification39f_b}
+  \vspace{-0.2cm}
   % \centering  
@@
   \end{tabular}
-  \vspace{0.3cm}
+  \vspace{0.2cm}
 \end{table}
@@
 Dataset usage for GenMM and GMM-HMM is the same, and number of states for GMM-HMM is the same as that for GenHMM in modeling each phoneme.
-Apart from setting the reference model, we also run the experiment comparisons with different total number of mixture components ($K$).
+Apart from setting the reference model, we also run the experiment comparisons with different total number of mixture components.
 
@@
   \subcaption{Folded $39$ Phonemes}
+  \vspace{-0.2cm}
   \centering  
@@
   \end{tabular}
-  \bigskip
   \subcaption{$61$ Phonemes}
+  \vspace{-0.2cm}
   \begin{tabular}{l|c|c|c}
@@
 The gain by using mixture generators is also presented in this set of experiments while the difference between $61$-phoneme and $39$-phoneme cases is also similar to the set of experiments in table~\ref{tab:acc-classification39f}.
+
+\begin{table}
+  \caption{Test accuracy table of perturbation with white noise ($K=3$, folded $39$ phonomes).}
+  \label{tab:acc-classification39f_noise_snr}
+  \vspace{-0.3cm}
+  \centering
+  \begin{tabular}{l|c|c|c|c}
+    \toprule
+    {SNR} &  15dB &  20dB &  25dB & 30 dB  \\  \midrule
+    GMM-HMM & $36.6\%$ &  $44.2\%$ &  $50.8\%$ & $57.1\%$  \\ \midrule
+    GenHMM & $52.4\%$ & $62.0\%$ &  $69.7\%$ & $74.3\%$ \\ 
+    \bottomrule                                                                  
+  \end{tabular}
+  \vspace{0.2cm}
+\end{table}
+\begin{table}
+  \caption{Test accuracy table of perturbation by different type of noise (SNR=$20$dB, $K=3$, folded $39$ phonomes).}
+  \label{tab:acc-classification39f_noise_type}
+  \vspace{-0.3cm}
+  \centering
+  \begin{tabular}{l|c|c|c|c}
+    \toprule
+    {Noise Type} &  White &  Pink &  Babble & Volvo  \\  \midrule
+    GMM-HMM & $44.2\%$ &  $48.8\%$ &  $57.7\%$ & $66.6\%$  \\ \midrule
+    GenHMM & $62.0\%$ & $65.1\%$ &  $70.0\%$ & $75.7\%$ \\ 
+    \bottomrule                                                                  
+  \end{tabular}
+  \vspace{0.2cm}
+\end{table}
+
+Apart from standard classification testing, we also test the robustness of our model to noise perturbations.
+We train GenHMM with $K=3$ by clean TIMIT training data in the case of folded $39$ phonemes with $39$ dimensional features.
+The testing dataset is perturbed by either the same type of noise with different signal-to-noise ratio (SNR) as shown in table~\ref{tab:acc-classification39f_noise_snr}, or different type of noises with the same SNR as shown in table~\ref{tab:acc-classification39f_noise_type}.
+The noise data is from NoiseX database.
+The baseline of these two sets of experiments is the accuracy testing of GenHMM and GMM-HMM on clean testing data in the same experimental condition, where GenHMM has $77.7\%$ and GMM-HMM gets $68.8\%$ as shown in table~\ref{tab:acc-classification39f}.
+It is shown in table~\ref{tab:acc-classification39f_noise_snr} that GMM-HMM's performance degenerates more than GenHMM's performance at them same level of noise perturbation, though the accuracy of both models increases along the increase of SNR.
+Especially, for SNR=$30$dB, the accuracy of GenHMM drops only about $3\%$ (from $77.7\%$ to $74.3\%$), while GMM-HMM encounters more than $10\%$ decrease (from $68.0\%$ to $57.1\%$) due to the noise perturbation.
+In table~\ref{tab:acc-classification39f_noise_type}, the SNR remains constant and GenHMM is tested with perturbation of different noise types.
+It is shown that GenHMM still has higher accuracy at different types of noise perturbations than GMM-HMM.
+Among these four types of noise, white noise shows most significant impact to GenHMM while the impact of volvo noise is negligible.
 

### 2019-09-01 manuscript/main.tex
dong proof reading

@@
 The typical way is to use a Gaussian mixture model (GMM) per state of HMM \cite{juang1986maximum}, where GMMs are used to represent the relationship between states of HMM and sequential data input.
-GMM based HMM (GMM-HMM) has become a standard model for sequential data modeling, and employed widely for practical applications, especially in speech recognition \cite{gales2008application,chatterjee2011auditory}.
+GMM based HMM (GMM-HMM) has become a standard model for sequential data modeling, and been employed widely for practical applications, especially in speech recognition \cite{gales2008application,chatterjee2011auditory}.
 
@@
   \item has high modeling capacity of sequential data, due to the neural network based generators;
-\item has advantage of easy to do training.
+\item has advantage of easy training.
 Training of GenHMM employs expectation maximization (EM) framework.
@@
 
-\begin{figure}[!h]
+\begin{figure}[!t]
   \centering
@@
   \caption{HMM model illustration.}\label{fig:hmm}
-  \vspace{-0.1cm}
+  \vspace{0.3cm}
 \end{figure}
@@
 A HMM $\bm{H}$ defined in a hypothesis space $\Hh$, i.e.
-$\bm{H} \in \Hh$, is capable to model time-span signal $\ubar{\bm{x}} = \left[ \bm{x}_1, \cdots, \bm{x}_T\right]^{\intercal}$, where $\bm{x}_t\in \RR^{N}$ is the signal at time $t$, $[\cdot]^{\intercal}$ denotes transpose, and $T$ denotes the time length\footnote{The length for difference sequential data can be different.}.
+$\bm{H} \in \Hh$, is capable to model time-span signal $\ubar{\bm{x}} = \left[ \bm{x}_1, \cdots, \bm{x}_T\right]^{\intercal}$, where $\bm{x}_t\in \RR^{N}$ is the signal at time $t$, $[\cdot]^{\intercal}$ denotes transpose, and $T$ denotes the time length\footnote{The length for  sequential data varies.}.
 Using HMM for signal representation is illustrated in Figure~\ref{fig:hmm}.
@@
 
-\begin{figure}[!th]
+\begin{figure}[!t]
   \centering
@@
   \label{fig:gen-mix}
-  \vspace{-0.2cm}
+  \vspace{0.3cm}
 \end{figure}
 
-In this section, we introduce a neural network based generator-mixed HMM model.
-We term it as GenHMM.
+In this section, we introduce a neural network based generator-mixed HMM model, i.e.
+GenHMM.
 GenHMM has a hidden markov process as a typical HMM does, while the probabilistic model of GenHMM for each hidden state is induced by mixture of neural network based generators.
@@
 
-\section{Practical Solution for GenHMM}
+\section{Solution for GenHMM}
 
@@
     \STATE $\bm{H}^{\mathrm{old}} \gets \bm{H}$
-    \STATE Set learning rate $\eta$, neural network optimization epochs $N$ per EM step
+    \STATE Set learning rate $\eta$, neural network optimization batches $N$ per EM step
     \FOR { $\bm{H}$ not converge}
@@
   \subcaption{Folded $39$ Phonemes}\label{tab:acc-classification39f_a}
-  \vspace{-0.2cm}
+  \vspace{-0.1cm}
   \centering  
@@
   \end{tabular}
+  \vspace{0.1cm}
   \subcaption{$61$ Phonemes}\label{tab:acc-classification39f_b}
-  \vspace{-0.2cm}
+  \vspace{-0.1cm}
   % \centering  
@@
   \end{tabular}
-  \vspace{0.2cm}
+  \vspace{0.3cm}
 \end{table}
@@
 The total training dataset size remains the same as $61$ phonemes are folded into $39$ phonemes.
-There are less training data available and more classes to be recognized in the $61$-phoneme case, which makes the task more challenging.
+There are less training data available per phonemes and more classes to be recognized in the $61$-phoneme case, which makes the task more challenging.
 
@@
   \subcaption{Folded $39$ Phonemes}
-  \vspace{-0.2cm}
+  \vspace{-0.1cm}
   \centering  
@@
   \end{tabular}
+  \vspace{0.1cm}
   \subcaption{$61$ Phonemes}
-  \vspace{-0.2cm}
+  \vspace{-0.1cm}
   \begin{tabular}{l|c|c|c}
@@
 The superior performance of GenHMM remains compared with reference model GMM-HMM.
-The gain by using mixture generators is also presented in this set of experiments while the difference between $61$-phoneme and $39$-phoneme cases is also similar to the set of experiments in table~\ref{tab:acc-classification39f}.
+The gain by using mixture generators is also presented in this set of experiments while the difference between $61$-phoneme and $39$-phoneme cases is similar to the set of experiments in table~\ref{tab:acc-classification39f}.
 
@@
   \label{tab:acc-classification39f_noise_snr}
-  \vspace{-0.3cm}
+  \vspace{-0.1cm}
   \centering
@@
   \label{tab:acc-classification39f_noise_type}
-  \vspace{-0.3cm}
+  \vspace{-0.1cm}
   \centering

### 2019-09-02 manuscript/main.tex
anonymous authors

@@
 
-\graphicspath{{./images/}}
-
 % \hypersetup{  
@@
 \title{Powering Hidden Markov Model by Neural Network based Generative Models}
+% \author{
+%   Dong Liu$^{\dagger}$, Antoine Honoré$^{\ast \dagger}$, Saikat Chatterjee$^{\dagger}$, and Lars~K. Rasmussen$^{\dagger}$\\
+%   $^{\dagger}$KTH Royal Institute of Technology, Stockholm, Sweden \\
+%   $^{\ast}$Karolinska Institute, Stockholm, Sweden \\
+%   E-mail: \{doli, honore, sach, lkra\}@kth.se}
 \author{
-  Dong Liu$^{\dagger}$, Antoine Honoré$^{\ast \dagger}$, Saikat Chatterjee$^{\dagger}$, and Lars~K. Rasmussen$^{\dagger}$\\
-  $^{\dagger}$KTH Royal Institute of Technology, Stockholm, Sweden \\
-  $^{\ast}$Karolinska Institute, Stockholm, Sweden \\
-  E-mail: \{doli, honore, sach, lkra\}@kth.se}
+  Author1, Author2, Author3, and Author4\\
+  Institute 1\\
+  Institute 2 \\
+  E-mail: emails}
 
@@
 The testing dataset is perturbed by either the same type of noise with different signal-to-noise ratio (SNR) as shown in table~\ref{tab:acc-classification39f_noise_snr}, or different type of noises with the same SNR as shown in table~\ref{tab:acc-classification39f_noise_type}.
-The noise data is from NoiseX database.
+The noise data is from NOISEX-92 database.
 The baseline of these two sets of experiments is the accuracy testing of GenHMM and GMM-HMM on clean testing data in the same experimental condition, where GenHMM has $77.7\%$ and GMM-HMM gets $68.8\%$ as shown in table~\ref{tab:acc-classification39f}.

### 2019-09-03 manuscript/main.tex
Revise abstract

@@
 \begin{abstract}
-  Hidden Markov model (HMM) has been successfully used for sequential data modeling problems.
-In this work, we propose to power the modeling capacity of HMM by bringing in neural network based generative models.
+Hidden Markov model (HMM) has been successfully used for sequential
+data modeling problems.
+In this work, we propose to power the modeling capacity of HMM by
+bringing in neural network based generative models.
 The proposed model is termed as GenHMM.
-In GenHMM, the hidden states of HMM and input data are related by a neural network based generative model that has tractability of exact likelihood and provides efficient likelihood computation.
-This is realized by a flow model.
-A learning algorithm for GenHMM is proposed in expectation-maximization framework.
+In the proposed GenHMM, each HMM hidden state is associated with a
+neural network based generative model that has
+tractability of exact likelihood and provides efficient likelihood
+computation.
+A generative model in GenHMM consists of mixture of generators
+that are realized by flow models.
+A learning algorithm for GenHMM is proposed in
+expectation-maximization framework.
 The convergence of the learning GenHMM is analyzed.

### 2019-09-03 manuscript/main.tex
revise 3.3 convergence analysis

@@
     \Qq(\bm{\Pi}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) &\geq \Qq(\bm{\Pi}^{\mathrm{old}}; \bm{H}^{\mathrm{old}}), \nonumber \\
-    \Qq(\bm{H}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) &\geq \Qq(\bm{H}^{\mathrm{old}}; \bm{H}^{\mathrm{old}})
+    \Qq(\bm{\Theta}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) &\geq \Qq(\bm{\Theta}^{\mathrm{old}}; \bm{H}^{\mathrm{old}}).
   \end{align*}
@@
   it gives
-  \begin{equation}
+  \begin{equation*}
     \Qq(\bm{H}^{\mathrm{new}}; \bm{H}^{\mathrm{old}}) \geq \Qq(\bm{H}^{\mathrm{old}}; \bm{H}^{\mathrm{old}}).
-  \end{equation}
-  Therefore we have the inequality 
+  \end{equation*}
+  With the above inequality, and the fact that $\EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})\right]$ is independent of $\bm{H}^{\mathrm{new}}$, we have the inequality
   \begin{align*}
@@
   \end{align*}
-  Due to the fact that $KL(p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})\|p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}}))=0$, we have
+  Due to the fact that $KL(p(\ubm{s}, \ubm{\kappa}|\ubm{x};
+  \bm{H}^{\mathrm{old}})\|p(\ubm{s}, \ubm{\kappa}|\ubm{x};
+  \bm{H}^{\mathrm{old}}))=0$, we have
   \begin{align*}

### 2019-09-03 manuscript/main.tex
revise introduction

@@
 Sequential data modeling is a challenging topic in pattern recognition and machine learning.
-For many applications, the assumption of independent and identically distributed data points is too strong to model data properly.
-Hidden Markov model (HMM) is a classic way to model sequential data.
-HMM has been widely used in different practical problems, and especially is known for applications in reinforcement learning \cite{ding2018reinforcementhmm,levine2018reinforcementReview}, natural language modeling \cite{khan2016survey,Hariyanti_2019}, biological sequence analysis such as proteins \cite{ASHWIN20172} and DNA \cite{ren2015dna}, etc.
+For many applications, the assumption of independent and identically distributed (i.i.d.) data points is too strong to model data properly.
+Hidden Markov model (HMM) is a classic way to model sequential data without the i.i.d.
+assumption.
+HMM has been widely used in different practical problems, including applications in reinforcement learning \cite{ding2018reinforcementhmm,levine2018reinforcementReview}, natural language modeling \cite{khan2016survey,Hariyanti_2019}, biological sequence analysis such as proteins \cite{ASHWIN20172} and DNA \cite{ren2015dna}, etc.
 
-HMM provides a framework for sequential data modeling.
-To connect states of HMM with input data, a probabilistic model is used to describe how each state of HMM represents a pattern of sequential data.
-The typical way is to use a Gaussian mixture model (GMM) per state of HMM \cite{juang1986maximum}, where GMMs are used to represent the relationship between states of HMM and sequential data input.
+A HMM is a statistical representation of sequential data generating process.
+Each state of a HMM is associated with a probabilistic model.
+The probabilistic model is used to represent the relationship between a state of HMM and sequential data input.
+The typical way is to use a Gaussian mixture model (GMM) per state of HMM \cite{juang1986maximum}, where GMMs are used to connect states of HMM to sequential data input.
 GMM based HMM (GMM-HMM) has become a standard model for sequential data modeling, and been employed widely for practical applications, especially in speech recognition \cite{gales2008application,chatterjee2011auditory}.
@@
 Given the success of GMM-HMM, it is not efficient for modeling data in nonlinear manifold.
-Research attempts of training HMM with neural networks have been made to boost the modeling capacity of HMM.
-A successful work is to bring restrictive Boltzmann machine (RBM) \cite{Hinton2012} into HMM based models \cite{hinton2012deepSpeech,li2013hybrid,Miao2013ImprovingLC}.
-In RBM based HMM, a hierarchical learning scheme is used: i) a stack of RBMs are trained one after another in unsupervised fashion, ii) then a final softmax layer is added to the stack of RBMs to represent the posterior of state of HMM given data input, iii) a discriminative training is followed to do the final tuning of the model.
+Research attempts at training HMM with neural networks have been made to boost the modeling capacity of HMM.
+A successful work has brought restrictive Boltzmann machine (RBM) \cite{Hinton2012} into HMM based models \cite{hinton2012deepSpeech,li2013hybrid,Miao2013ImprovingLC}.
+In RBM based HMM, a hierarchical learning scheme is used: i) RBMs are trained one after the other in unsupervised fashion, and are stacked together as one model, ii) then a final softmax layer is added to the stack of RBMs to represent the probability of a HMM state given a data input, iii) a discriminative training is performed for the final tuning of the model.
 
@@
 These hierarchical schemes are built based on intuition of domain knowledge.
-Training of these hierarchical models usually need specific expertise in relevant areas to be able to proceed with the hierarchical procedure of training and application usage.
+Training of these hierarchical models usually requires expertise in specific areas to be able to proceed with the hierarchical procedure of training and application usage.
 
-In this work, we propose a generative model based HMM.
-Specifically, we propose a generator-mixed HMM, termed as GenHMM, where generators are realized by neural networks to help the model gain high modeling capacity.
+In this work, we propose a generative model based HMM, termed as GenHMM.
+Specifically, a generative model in our GenHMM is generator-mixed, where a generator is realized by a neural network to help the model gain high modeling capacity.
 Our proposed model, GenHMM:
@@
   \item has high modeling capacity of sequential data, due to the neural network based generators;
-\item has advantage of easy training.
+\item is easy to train.
 Training of GenHMM employs expectation maximization (EM) framework.
-Training a GenHMM is as easy as training a GMM-HMM model, while configuration of GenHMM is flexible;
+Therefore, training a GenHMM is as easy as training a GMM-HMM model, while configuration of GenHMM is flexible;
 % \item has high modeling capability of sequence data, though its training and using are handy
@@
 Instead of using softmax for probability representation, our GenHMM has tractability of exact loglikelihood of given sequential data, which is based on change of variable formula.
-To make the loglikelihood computation efficient, neural network based generators of GenHMM are implemented as flow models.
+To make the loglikelihood computation efficient, neural network based generators of GenHMM are realized as flow models.
 

### 2019-09-03 manuscript/main.tex
revise section 2

@@
 $\bm{H} \in \Hh$, is capable to model time-span signal $\ubar{\bm{x}} = \left[ \bm{x}_1, \cdots, \bm{x}_T\right]^{\intercal}$, where $\bm{x}_t\in \RR^{N}$ is the signal at time $t$, $[\cdot]^{\intercal}$ denotes transpose, and $T$ denotes the time length\footnote{The length for  sequential data varies.}.
-Using HMM for signal representation is illustrated in Figure~\ref{fig:hmm}.
-The model assumption is that different instant signal of $\ubar{\bm{x}}$ is generated by a different signal source associated with a hidden state of HMM, while the source selection is done by a discrete hidden markov process.
-We define the hypothesis set of HMM as $\Hh := \{\bm{H} | \{\Ss, \bm{q}, \bm{A}, p(\bm{x}|{s}; \bm{\Phi}_{s})\}$, where
+We define the hypothesis set of HMM as $\Hh := \{\bm{H} | \bm{H}=\{\Ss, \bm{q}, \bm{A}, p(\bm{x}|{s}; \bm{\Phi}_{s})\}\}$, where
 \begin{itemize}
-\item $\Ss$ is the set of hidden states of HMM $\bm{H}$.
-\item $\bm{q} = \left[ q_1, q_2, \cdots, q_{|\Ss|}\right]^\intercal$ is the initial distribution of HMM $\bm{H}$ with $|\Ss|$ as cardinality of $\Ss$.
+\item $\Ss$ is the set of hidden states of $\bm{H}$.
+\item $\bm{q} = \left[ q_1, q_2, \cdots, q_{|\Ss|}\right]^\intercal$ is the initial distribution of $\bm{H}$ with $|\Ss|$ as cardinality of $\Ss$.
 For a state $s$, $q_s = p(s;\bm{H})$.
 In the following, we use $s_t$ to denote the state $s$ at time $t$.
-\item $\bm{A}$ is the transition matrix for the HMM $\bm{H}$ of size $|\Ss| \times |\Ss|$ and $\bm{A}_{i,j} = p(s_{t+1}=j|s_{t}=i; \bm{H})$.
+\item $\bm{A}$ matrix of size $|\Ss| \times |\Ss|$ is the transition matrix of $\bm{H}$. That is, $\forall i, j \in \Ss$,  $\bm{A}_{i,j} = p(s_{t+1}=j|s_{t}=i; \bm{H})$.
 \item For a given hidden state $s$, the density function of the observable signal is $p({\bm{x}}|{s};\bm{\Phi}_{s})$, where $\bm{\Phi}_{s}$ is the parameter set that defines this probabilistic model.
@@
 
-In the framework of HMM, at each time instance, signal $\bm{x}_t$ is assumed to be generated by a distribution with density function $p(\bm{x}_t| s_t; \bm{\Phi}_{s_t})$, and $s_t$ is decided by the hidden markov process.
+Using HMM for signal representation is illustrated in Figure~\ref{fig:hmm}.
+The model assumption is that different instant signal of $\ubar{\bm{x}}$ is generated by a different signal source associated with a hidden state of HMM, while the source selection is done by a discrete hidden markov process.
+In the framework of HMM, at each time instance $t$, signal $\bm{x}_t$ is assumed to be generated by a distribution with density function $p(\bm{x}_t| s_t; \bm{\Phi}_{s_t})$, and $s_t$ is decided by the hidden markov process.
 Putting these together gives us the probabilistic model $p(\ubm{x};\bm{H})$.

### 2019-09-03 manuscript/main.tex
revise section 2.1

@@
 
-In this section, we introduce a neural network based generator-mixed HMM model, i.e.
-GenHMM.
-GenHMM has a hidden markov process as a typical HMM does, while the probabilistic model of GenHMM for each hidden state is induced by mixture of neural network based generators.
-For a given state $s\in \Ss$, we define its probabilistic model in GenHMM as
-\begin{equation}
+In this section, we introduce the neural network based state probabilistic model of our GenHMM.
+Recall that $\bm{x}\in\mathbb{R}^N$.
+The probabilistic model of GenHMM for each hidden state is a mixture of $K$ neural network based generators, where $K$ standards for positive integer.
+The probabilistic model of a state $s\in\Ss$ is then given by
+\begin{equation}\label{eq:state-prob-model}
   p(\bm{x}| s; \bm{\Phi}_{s}) = \sum_{\kappa=1}^{K}\pi_{s, \kappa} p(\bm{x}| s, \kappa; \bm{\theta}_{s, \kappa}),
@@
 where $\kappa$ is a random variable following a categorical distribution, with probability $\pi_{s, \kappa} = p(\kappa | s; \bm{H})$.
-Naturally $\sum_{\kappa = 1}^{K} \pi_{s, \kappa}= 1$.
-Denote $\bm{\pi}_{s} = [\pi_{s,1}, \pi_{s,2}, \cdots, \pi_{s,K}]^{\intercal}$,
-$\bm{\theta}_s = \left\{ \bm{\theta}_{s, \kappa}| \kappa = 1, 2, \cdots, K \right\}$. 
-We define $p(\bm{x}| s, \kappa; \bm{\theta}_{s, \kappa})$ as induced distribution by a generator $\bm{g}_{s,\kappa}: \RR^{N}\rightarrow\RR^{N}$, such that $\bm{x}=\bm{g}_{s, \kappa}(\bm{z})$, where $\bm{z}$ is a latent variable following a distribution with density function $p_{s,\kappa}(\bm{z})$.
+Naturally $\sum_{\kappa = 1}^{K} \pi_{s, \kappa}= 1$. Denote $\bm{\pi}_{s} = [\pi_{s,1}, \pi_{s,2}, \cdots, \pi_{s,K}]^{\intercal}$. 
+In \eqref{eq:state-prob-model}, $p(\bm{x}| s, \kappa; \bm{\theta}_{s, \kappa})$ is defined as induced distribution by a generator $\bm{g}_{s,\kappa}: \RR^{N}\rightarrow\RR^{N}$, such that $\bm{x}=\bm{g}_{s, \kappa}(\bm{z})$, where $\bm{z}$ is a latent variable following a distribution with density function $p_{s,\kappa}(\bm{z})$.
+Generator $\bm{g}_{s,\kappa}$ is parameterized by $\bm{\theta}_{s, \kappa}$.
+Let us denote the collection of the parameter sets of generators for state $s$ as $\bm{\theta}_s = \left\{ \bm{\theta}_{s, \kappa}| \kappa = 1, 2, \cdots, K \right\}$.
 Assuming $\bm{g}_{s, \kappa}$ is invertible, by change of variable, we have
@@
 
-For a state $s$ of GenHMM, the induced probability distribution has the signal flow illustration as shown in Figure~\ref{fig:gen-mix}, in which the generator identity is up to the random variable $\kappa$.
+The signal flow of the probability distribution for a state $s$ of GenHMM is shown in Figure~\ref{fig:gen-mix}, in which the generator identity is up to the random variable $\kappa$.
 

### 2019-09-03 manuscript/main.tex
revise section 2.2

@@
 For the given dataset, we denote its empirical distribution by $\hat{p}(\ubm{x}) = \frac{1}{R}\sum_{r=1}^{R} \delta_{\ubmr{x}{r}}(\ubm{x})$, where superscipt $(\cdot)^{r}$ denotes the index of sequential signal.
-The KL divergence problem can be boiled down to a maximizing loglikelihood problem
+The KL divergence minimization problem can be reduced to a likelihood maximization problem
 \begin{equation}\label{eq:ml-of-hmm}
@@
 
-The first problem that we need to address is to deal with the hidden sequential variable $\ubm{s}$ and $\ubm{\kappa}$  in model $\bm{H}$, where $\ubm{s}=[ \bm{s}_1, \bm{s}_2, \cdots, \bm{s}_T ]^{\intercal}$ and $\ubm{\kappa} = [\bm{\kappa}_1, \bm{\kappa}_2, \cdots, \bm{\kappa}_T]^{\intercal}$.
+For the likelihood maximization, the first problem that we need to address is to deal with the hidden sequential variables of model $\bm{H}$, namely $\ubm{s}=[ \bm{s}_1, \bm{s}_2, \cdots, \bm{s}_T ]^{\intercal}$ and $\ubm{\kappa} = [\bm{\kappa}_1, \bm{\kappa}_2, \cdots, \bm{\kappa}_T]^{\intercal}$.
 For a sequential observable variable $\ubm{x}$, $\ubm{s}$ is the hidden state sequence corresponding to $\ubm{x}$, and $\ubm{\kappa}$ is the hidden variable sequence representing the generator identity sequence that actually generates $\ubm{x}$.
@@
 Since directly maximizing likelihood is not an option for our problem in \ref{eq:ml-of-hmm}, we address this problem in expectation maximization (EM) framework.
-This divides our problem into two steps: i) using the joint posterior of hidden variable sequences $\ubm{s}$ and $\ubm{\kappa}$ to obtain an ``expected likelihood'' of the observable variable sequence $\ubm{x}$, i.e.
+This divides our problem into  two iterative steps: i) using the joint posterior of hidden variable sequences $\ubm{s}$ and $\ubm{\kappa}$ to obtain an ``expected likelihood'' of the observable variable sequence $\ubm{x}$, i.e.
 the E-step; ii) maximizing the expected likelihood with regard to (w.r.t.) the model $\bm{H}$, i.e.
 the M-step.
-We formulate these two steps as follows:
+Assume model $\bm{H}$ is at a configuration of $\bm{H}^{\mathrm{old}}$, we formulate these two steps as follows:
 \begin{itemize}
@@
 \item how to realize GenHMM by neural network based generators such that likelihoond of their induced distributions can be computed explicitly and exactly?
-\item how to train GenHMM to sovle problem in \eqref{eq:ml-of-hmm} using practical algorithm?
+\item how to train GenHMM to solve problem in \eqref{eq:ml-of-hmm} using practical algorithm?
 \item would the training of GenHMM converge?
 \end{itemize}
-We would attempt these problems in the following section.
+We tackle these problems in the following section.
 

### 2019-09-03 manuscript/main.tex
revise section 3

@@
 \subsection{Realizing $\bm{g}_{s,\kappa}$ by a Flow Model}
-We use a feed-forward neural network to implement each generator $\bm{g}_{s,\kappa}$.
-Define  $\bm{g}_{s,\kappa}$ as a $L$-layer neural network and we formulate its mapping by layer-wise concatenation:
+Each generator $\bm{g}_{s,\kappa}$ is realized as a feed-forward neural netowrk.
+We define $\bm{g}_{s,\kappa}$ as a $L$-layer neural network and formulate its mapping by layer-wise concatenation:
 $\bm{g}_{s,\kappa}=\bm{g}_{s,\kappa}^{[L]}\circ \bm{g}_{s,\kappa}^{[L-1]}\circ \cdots
@@
 \end{equation}
-where $\nabla{f_{s,\kappa}^{[l]}}$ is the Jacobian of the mapping from the $l$-th layer to the $(l-1)$-th layer, i.e., the inverse transformation. We compute the determinant of the Jacobian matrix as
+where $\nabla{\bm{f}_{s,\kappa}^{[l]}}$ is the Jacobian of the mapping from the $l$-th layer to the $(l-1)$-th layer, i.e., the inverse transformation.
+We compute the determinant of the Jacobian matrix as
 \begin{align}\label{eq-hl-determinate}
@@
 \end{align}
-where $\bm{I}_a$ is identity matrix and $\mathrm{diag}(\cdot)$ returns a square matrix with the elements of $(\cdot)$ on the main diagnal.
+where $\bm{I}_a$ is identity matrix and $\mathrm{diag}(\cdot)$ returns a square matrix with the elements of $\cdot$ on the main diagnal.
 % \textcolor{red}{do I still need this?}
@@
 \subsubsection{Generative Model Learning}
-The generative model learning is actually to solve the problem in \eqref{eq:generative-model-update}, which can be further divided into two subproblems: i) generator learning; ii) weights of generators learning.
+The generative model learning is actually to solve the problem in \eqref{eq:generative-model-update}, which can be further divided into two subproblems: i) generator learning; ii) mixture weights of generators learning.
 Let us define notations:
@@
 \end{align}
-The generators of GenHMM can simply use standard Gaussian distribution for latent variables $\bm{z} \sim p_{s,\kappa}(\bm{z})$.
-It is straightforward to employ batch-size stochastic gradient decent to maximize $\Qq(\bm{\Theta; \bm{H}^{\mathrm{old}}})$ w.r.t.
-parameters of generators, i.e.
-$\bm{\Phi}$, since the dataset represented by $\hat{p}(\ubm{x})$ can be too large to do whole-dataset iterations.
+The generators of GenHMM simply use standard Gaussian distribution for latent variables $\bm{z} \sim p_{s,\kappa}(\bm{z})$.
+Since the dataset represented by $\hat{p}(\ubm{x})$ can be too large to do whole-dataset iterations, batch-size stochastic gradient decent can be used to maximize $\Qq(\bm{\Theta; \bm{H}^{\mathrm{old}}})$ w.r.t.
+parameters of generators.
 
@@
 In pursuit of representing a dataset by GenHMM,  we are interested if the learning solution discussed in subsection~\ref{subsec:optmGenHMM} would convergence.
-The properties on GenHMM's convergence is analyzed as follows.
+The properties on GenHMM's convergence are analyzed as follows.
 
@@
     \STATE {\bfseries Input:}{
+      Empirical distribution $\hat{p}(\bm{x})$ of dataset\\
       Initializing $\bm{H}^{\mathrm{old}}, \bm{H} \in \Hh$ gives: \\
-      $\bm{H}^{\mathrm{old}} = \{\Ss, \bm{q}^{\mathrm{old}}, A^{\mathrm{old}}, p(\bm{x}|s; \bm{\Phi}^{\mathrm{old}})\}$, $\bm{H} = \{\Ss, \bm{q}, A, p(\bm{x}|s; \bm{\Phi})\}$,\\
-      empirical distribution $\hat{p}(\bm{x})$ of dataset;} \\
-    \STATE Initialize $\bm{H}$
+      $\bm{H}^{\mathrm{old}} = \{\Ss, \bm{q}^{\mathrm{old}}, A^{\mathrm{old}}, p(\bm{x}|s; \bm{\Phi}_{s}^{\mathrm{old}})\}$ \\
+      $\bm{H} = \{\Ss, \bm{q}, A, p(\bm{x}|s; \bm{\Phi}_{s})\}$} \\
     \STATE $\bm{H}^{\mathrm{old}} \gets \bm{H}$

### 2019-09-03 manuscript/main.tex
revise section 4

@@
 To show the validity of our model, we implement our model in PyTorch and test it with sequential data.
-The code is available at github repository\footnote{Due to double-blind review, the author's repository is not present here temporarily.}.
-In the next subsection, we first discuss the experimental setups and then show the experimental results.
+The code is available at github repository\footnote{Due to double-blind review, the code is submitted in supplementary file.}.
+We first discuss the experimental setups and then show the experimental results.
 
 \subsection{Experimental Setups}
-The dataset used for sequential data modeling and classification is TIMIT dataset where the speech signal is sampled at $16$kHz.
+The dataset used for sequential data modeling and classification is TIMIT where the speech signal is sampled at $16$kHz.
+The TIMIT dataset consists of $5300$ phoneme-labeled speech utterances which are partitioned into two sets: {a train set consists of $4620$ utterance, and a test set consists of $1680$ utterances.} There are totally $61$ different types of phones in TIMIT.
+We performed experiments in two cases: i) full $61$-phoneme classification case; ii) $39$-phonme classification case, where $61$ phonemes are folded onto $39$ phonemes as described in \cite{Perdigao11}.
+
 For extraction of feature vectors, we use $25$ms frame length and $10$ms frame shift to convert sound track into standard Mel-frequency cepstral coefficients (MFCCs) features.
-The deltas and delta-deltas features are also concatenated and used in our experiments.
+Experiments using the deltas and delta-deltas of the features are also carried out.
 
-The TIMIT dataset consists of $5300$ phoneme-labeled speech utterances which are partitioned into two sets: {a train set consists of $4620$ utterance, and a test set consists of $1680$ utterances.} There are totally $61$ different types of phones in TIMIT labels.
-Our experiments in this section are carried out in two different cases: i) full $61$-phoneme classification case, where both training and testing use the $61$-phoneme data; ii) $39$-phonme classification case, where $61$ phonemes are folded into $39$ phonemes as described in \cite{Perdigao11}.
+Our experiments are performed for: i) standard classification tasks (table~\ref{tab:acc-classification39f},\ref{tab:acc-classification13f}), ii) classification under noise perturbation (table~\ref{tab:acc-classification39f_noise_snr}, \ref{tab:acc-classification39f_noise_type}).
+In all experiments, generators $\left\{\bm{g}_{s,\kappa}|s\in \Ss, \kappa=1, 2, \cdots, K \right\}$ of GenHMM are implemented as flow models.
+Specifically, our generator structure follows that of a RealNVP described in \cite{2016arXiv160508803D}.
+As discussed, the coupling layer shown in \eqref{eq-gl} maps a part of its input signal identically.
+The implementation is such that layer $l+1$ would alternate the input signal order of layer $l$ such that no signal remains the same after two consecutive coupling layers.
+We term such a pair of consecutive coupling layers as a \textit{flow block}.
+In our experiments, each generator $\bm{g}_{s,\kappa}$ consists of four \textit{flow blocks}.
+The density of samples in the latent space is defined as Normal, i.e. $p_{s,\kappa}(\bm{z})$ is the density function of standard Gaussian.
 
-In all the following experiments, all generators $\left\{\bm{g}_{s,\kappa}|s\in \Ss, \kappa=1, 2, \cdots, K \right\}$ of GenHMM are implemented as flow models.
-To be more specific, our generator structure follows the structure of flow model in RealNVP \cite{2016arXiv160508803D}.
-The coupling layer shown as \eqref{eq-gl} maps a part of its input signal identically.
-The following coupling layer would alternate the input signal order such that after two consequential coupling layers, there is no signal remains the same any more.
-We term such two consequential coupling layers as a \textit{flow block}.
-In our experiments, a generator $\bm{g}_{s,\kappa}$ consists of four flow blocks.
-The latent source signal is assumed as standard Gaussian, i.e.
-$p_{s,\kappa}(\bm{z})$ is density function of standard Gaussian.
-
-In learning of GenHMM, the number of states is set as $|\Ss| \in \left\{ 3,4,5 \right\}$.
-The exact number of states for a phoneme is decided by computing the average length of MFCC frames per phone in training dataset, and clipping the average length into $\left\{ 3,4,5 \right\}$.
+For each GenHMM, the number of states is adapted to the training dataset.
+The exact number of states is decided by computing the average length of MFCC frames per phone in training dataset, and clipping the average length into $\left\{ 3,4,5 \right\}$.
 Transition matrix $\bm{A}$ is initialized as upper triangular matrix for GenHMM.
@@
   \centering
-  \subcaption{Folded $39$ Phonemes}
+  \subcaption{Folded $39$ Phonemes}\label{tab:acc-classification13f_a}
   \vspace{-0.1cm}
@@
   \vspace{0.1cm}
-  \subcaption{$61$ Phonemes}
+  \subcaption{$61$ Phonemes}\label{tab:acc-classification13f_b}
   \vspace{-0.1cm}

### 2019-09-03 manuscript/main.tex
format spacing and some typo fix

@@
 \author{
-  Author1, Author2, Author3, and Author4\\
-  Institute 1\\
-  Institute 2 \\
-  E-mail: emails}
+  Author1, Author2, Author3, and Author4}
 
@@
 \item $\bm{q} = \left[ q_1, q_2, \cdots, q_{|\Ss|}\right]^\intercal$ is the initial distribution of $\bm{H}$ with $|\Ss|$ as cardinality of $\Ss$.
-For a state $s$, $q_s = p(s;\bm{H})$.
-In the following, we use $s_t$ to denote the state $s$ at time $t$.
+For $i \in \Ss$, $q_i = p(s_{1}=i;\bm{H})$.
+We use $s_t$ to denote the state $s$ at time $t$.
 \item $\bm{A}$ matrix of size $|\Ss| \times |\Ss|$ is the transition matrix of $\bm{H}$. That is, $\forall i, j \in \Ss$,  $\bm{A}_{i,j} = p(s_{t+1}=j|s_{t}=i; \bm{H})$.
@@
 Using HMM for signal representation is illustrated in Figure~\ref{fig:hmm}.
-The model assumption is that different instant signal of $\ubar{\bm{x}}$ is generated by a different signal source associated with a hidden state of HMM, while the source selection is done by a discrete hidden markov process.
+The model assumption is that different instant signal of $\ubar{\bm{x}}$ is generated by a different signal source associated with a hidden state of HMM.
 In the framework of HMM, at each time instance $t$, signal $\bm{x}_t$ is assumed to be generated by a distribution with density function $p(\bm{x}_t| s_t; \bm{\Phi}_{s_t})$, and $s_t$ is decided by the hidden markov process.
@@
 Recall that $\bm{x}\in\mathbb{R}^N$.
-The probabilistic model of GenHMM for each hidden state is a mixture of $K$ neural network based generators, where $K$ standards for positive integer.
+The probabilistic model of GenHMM for each hidden state is a mixture of $K$ neural network based generators, where $K$ is a positive integer.
 The probabilistic model of a state $s\in\Ss$ is then given by
@@
 
-Since directly maximizing likelihood is not an option for our problem in \ref{eq:ml-of-hmm}, we address this problem in expectation maximization (EM) framework.
+Since directly maximizing likelihood is not an option for our problem in \eqref{eq:ml-of-hmm}, we address this problem in expectation maximization (EM) framework.
 This divides our problem into  two iterative steps: i) using the joint posterior of hidden variable sequences $\ubm{s}$ and $\ubm{\kappa}$ to obtain an ``expected likelihood'' of the observable variable sequence $\ubm{x}$, i.e.
@@
 The generators of GenHMM simply use standard Gaussian distribution for latent variables $\bm{z} \sim p_{s,\kappa}(\bm{z})$.
-Since the dataset represented by $\hat{p}(\ubm{x})$ can be too large to do whole-dataset iterations, batch-size stochastic gradient decent can be used to maximize $\Qq(\bm{\Theta; \bm{H}^{\mathrm{old}}})$ w.r.t.
+Since training dataset can be too large to do whole-dataset iterations, batch-size stochastic gradient decent can be used to maximize $\Qq(\bm{\Theta; \bm{H}^{\mathrm{old}}})$ w.r.t.
 parameters of generators.
@@
 \begin{equation}
-  \pi_{s,\kappa} = \frac{1}{\lambda_s}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}} p(\smtr{s}{t}{r}=s, \smtr{\kappa}{t}{r} =\kappa| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})
+  \pi_{s,\kappa} = \frac{1}{\lambda_s}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}} p(\smtr{s}{t}{r}=s, \smtr{\kappa}{t}{r} =\kappa| \ubmr{x}{r}; \bm{H}^{\mathrm{old}}).
 \end{equation}
@@
 where $p(s, \kappa | \ubm{x}; \bm{H}^{\mathrm{old}}) = p(s| \ubm{x}; \bm{H}^{\mathrm{old}}) p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$.
-$p(s| \ubm{x}; \bm{H}^{\mathrm{old}})$ that can be computed by results of forward-backward algorithm, while $p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$ is given by \eqref{eq:kappa-posterior}.
+$p(s| \ubm{x}; \bm{H}^{\mathrm{old}})$ can be computed by forward-backward algorithm, while $p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$ is given by \eqref{eq:kappa-posterior}.
 
@@
   &= \sum_{r=1}^{R} \sum_{\ubmr{s}{r}}{p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \sum_{t=1}^{T^{r}-1}\log\,p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; \bm{H}) \nonumber \\
-  &= \sum_{r=1}^{R} \hspace{-0.1cm}\sum_{t=1}^{T^{r}-1}\hspace{-0.1cm} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}\hspace{-0.05cm}\sum_{\smtr{s}{t+1}{r}=1}^{|\Ss|}\hspace{-0.2cm}{p(\smtr{s}{t}{r}, \smtr{s}{t+1}{r}| \ubmr{x}{r};\hspace{-0.05cm} \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; \bm{H})
+  &= \sum_{r=1}^{R} \hspace{-0.1cm}\sum_{t=1}^{T^{r}-1}\hspace{-0.1cm} \sum_{\smtr{s}{t}{r}=1}^{|\Ss|}\hspace{-0.05cm}\sum_{\smtr{s}{t+1}{r}=1}^{|\Ss|}\hspace{-0.2cm}{p(\smtr{s}{t}{r}, \smtr{s}{t+1}{r}| \ubmr{x}{r};\hspace{-0.05cm} \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{t+1}{r}|\smtr{s}{t}{r}; \bm{H}).
 \end{align}
@@
   \end{align*}
-  Due to the fact that $KL(p(\ubm{s}, \ubm{\kappa}|\ubm{x};
+  Due to $KL(p(\ubm{s}, \ubm{\kappa}|\ubm{x};
   \bm{H}^{\mathrm{old}})\|p(\ubm{s}, \ubm{\kappa}|\ubm{x};
@@
   \end{tabular}
-  \vspace{0.3cm}
+  \vspace{0.2cm}
 \end{table}
-
 We firstly show the phoneme classification using 39 dimensional MFCC features (MFCC coefficients, deltas, and delta-deltas), to validate one possible usage of our proposed model.
@@
   \end{tabular}
-  \vspace{0.3cm}
+  \vspace{0.1cm}
 \end{table}
@@
 The baseline of these two sets of experiments is the accuracy testing of GenHMM and GMM-HMM on clean testing data in the same experimental condition, where GenHMM has $77.7\%$ and GMM-HMM gets $68.8\%$ as shown in table~\ref{tab:acc-classification39f}.
-It is shown in table~\ref{tab:acc-classification39f_noise_snr} that GMM-HMM's performance degenerates more than GenHMM's performance at them same level of noise perturbation, though the accuracy of both models increases along the increase of SNR.
+It is shown in table~\ref{tab:acc-classification39f_noise_snr} that GMM-HMM's performance degenerates more than GenHMM's performance at the same level of noise perturbation, though the accuracy of both models increases along the increase of SNR.
 Especially, for SNR=$30$dB, the accuracy of GenHMM drops only about $3\%$ (from $77.7\%$ to $74.3\%$), while GMM-HMM encounters more than $10\%$ decrease (from $68.0\%$ to $57.1\%$) due to the noise perturbation.

### 2019-09-04 src/genHMM.py
rm comments with my name

@@ -340,7 +340,7 @@ class GenHMM(torch.nn.Module):
 
  #       print(loglh_sk.shape, self.n_states, self.n_prob_components)
         # max_loglh = torch.max(torch.max(loglh_sk, dim=1)[0],dim=1)[0]
-        ### dong: original update here was not correct
+        
 
         local_loglh_sk = loglh_sk
         max_loglh = torch.max(local_loglh_sk, dim=3, keepdim=True)[0]

### 2019-10-13 manuscript/main.tex
capitalize introduction
for arxiv

@@
 \title{Powering Hidden Markov Model by Neural Network based Generative Models}
+\author{
+  Dong Liu$^{\dagger}$, Antoine Honor\'e$^{\ast \dagger}$, Saikat Chatterjee$^{\dagger}$, and Lars~K. Rasmussen$^{\dagger}$\\
+  $^{\dagger}$KTH Royal Institute of Technology, Stockholm, Sweden \\
+  $^{\ast}$Karolinska Institute, Stockholm, Sweden \\
+  E-mail: \{doli, honore, sach, lkra\}@kth.se}
 % \author{
-%   Dong Liu$^{\dagger}$, Antoine Honoré$^{\ast \dagger}$, Saikat Chatterjee$^{\dagger}$, and Lars~K. Rasmussen$^{\dagger}$\\
-%   $^{\dagger}$KTH Royal Institute of Technology, Stockholm, Sweden \\
-%   $^{\ast}$Karolinska Institute, Stockholm, Sweden \\
-%   E-mail: \{doli, honore, sach, lkra\}@kth.se}
-\author{
-  Author1, Author2, Author3, and Author4}
-
-% \name{
-% Dong Liu,
-% Minh Thành Vu,
-% Saikat Chatterjee,
-% and Lars~K. Rasmussen
-% }
-
-%   \address{
-%   KTH Royal Institute of Technology, Stockholm, Sweden \\
-%   E-mail: \{doli, mtvu, sach, lkra\}@kth.se}
+%   Author1, Author2, Author3, and Author4}
 
@@
 
-\section{introduction}
+\section{Introduction}
 Sequential data modeling is a challenging topic in pattern recognition and machine learning.

### 2019-11-13 manuscript/main.tex
disable link to code
fix processing problem by arxiv
ecai formating and revising

@@
-% \def\year{2020}\relax
-% File: formatting-instruction.tex
-\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
-\usepackage{aaai20}  % DO NOT CHANGE THIS
-\usepackage{times}  % DO NOT CHANGE THIS
-\usepackage{helvet} % DO NOT CHANGE THIS
-\usepackage{courier}  % DO NOT CHANGE THIS
-\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
-\usepackage{graphicx} % DO NOT CHANGE THIS
-\urlstyle{rm} % DO NOT CHANGE THIS
-\def\UrlFont{\rm}  % DO NOT CHANGE THIS
-\usepackage{graphicx}  % DO NOT CHANGE THIS
-\frenchspacing  % DO NOT CHANGE THIS
-\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
-\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
-% \nocopyright
-% PDF Info Is REQUIRED.
-% For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
-% For /Title, add Title in Mixed Case. No accents or commands. Retain the parentheses.
-\pdfinfo{/Title gm_hmm
-} %Leave this	
-\newcommand{\citet}[1]{\citeauthor{#1} \shortcite{#1}}
-\newcommand{\citep}{\cite}
-\newcommand{\citealp}[1]{\citeauthor{#1} \citeyear{#1}}
+\documentclass{ecai}
+\usepackage{times}
+\usepackage{graphicx}
+\usepackage{latexsym}
+
+%\ecaisubmission   % inserts page numbers. Use only for submission of paper.
+                  % Do NOT use for camera-ready version of paper.
+
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 \usepackage{amsmath,graphicx}
-% format A4
-% \usepackage{vmargin}
-% \setpapersize{A4}
-
 \RequirePackage{algorithm}
 \RequirePackage{algorithmic}
+%%%%%%%%%%%%%%%%%%%%%%%%
 
-% Attempt to make hyperref and algorithmic work together better:
-% \newcommand{\theHalgorithm}{\arabic{algorithm}}
-
-% \hypersetup{  
-% bookmarks=true,
-% backref=true,
-% pagebackref=false,
-% colorlinks=true,
-% linkcolor=blue,
-% citecolor=red,
-% urlcolor=blue,
-% pdftitle={Generative Model HMM},
-% pdfauthor={Dong Liu},
-% pdfsubject={}
-% }
-
-\setcounter{secnumdepth}{2} %May be changed to 1 or 2 if section numbers are desired.
-
-% The file aaai20.sty is the style file for AAAI Press 
-% proceedings, working notes, and technical reports.
-% 
-\setlength\titlebox{2.5in} % If your paper contains an overfull \vbox too high warning at the beginning of the document, use this
-% command to correct it. You may not alter the value below 2.5 in
+\begin{document}
 
@@
 \author{
-  Dong Liu$^{\dagger}$, Antoine Honor\'e$^{\ast \dagger}$, Saikat Chatterjee$^{\dagger}$, and Lars~K. Rasmussen$^{\dagger}$\\
-  $^{\dagger}$KTH Royal Institute of Technology, Stockholm, Sweden \\
-  $^{\ast}$Karolinska Institute, Stockholm, Sweden \\
-  E-mail: \{doli, honore, sach, lkra\}@kth.se}
-% \author{
-%   Author1, Author2, Author3, and Author4}
+  Dong Liu$^{\dagger}$ \and
+  Antoine Honor\'e$^{\ast \dagger}$ \and
+  Saikat Chatterjee$^{\dagger}$ \and
+  Lars~K. Rasmussen$^{\dagger}$
+  \institute{
+    $^{\dagger}$KTH Royal Institute of Technology, Stockholm, Sweden \newline
+    $^{\ast}$Karolinska Institute, Stockholm, Sweden \newline
+  E-mail: \{doli, honore, sach, lkra\}@kth.se}}
 
-\begin{document}
+% \title{Guidelines for Preparing a Paper for the \\
+% European Conference on Artificial Intelligence}
+
+% \author{Name1 Surname1 \and Name2 Surname2 \and Name3 Surname3\institute{University,
+% Country, email: somename@university.edu} }
 
 \maketitle
+\bibliographystyle{ecai}
+
+% \name{
+% Dong Liu,
+% Minh Thành Vu,
+% Saikat Chatterjee,
+% and Lars~K. Rasmussen
+% }
+
+%   \address{
+%   KTH Royal Institute of Technology, Stockholm, Sweden \\
+%   E-mail: \{doli, mtvu, sach, lkra\}@kth.se}
+
 \begin{abstract}
@@
 computation.
-A generative model in GenHMM consists of mixture of generators
+A generative model in GenHMM consists of a mixture of generators
 that are realized by flow models.
@@
 In RBM based HMM, a hierarchical learning scheme is used: i) RBMs are trained one after the other in unsupervised fashion, and are stacked together as one model, ii) then a final softmax layer is added to the stack of RBMs to represent the probability of a HMM state given a data input, iii) a discriminative training is performed for the final tuning of the model.
+\textcolor{red}{explain more here..}
 
@@
 This hierarchical training procedure is motivated by the intuition of using LSTM or RNN to fill in the gap where HMM can not learn.
+\textcolor{red}{explain more here...}
 
@@
 \end{itemize}
-Instead of using softmax for probability representation, our GenHMM has tractability of exact loglikelihood of given sequential data, which is based on change of variable formula.
+Instead of using softmax for probability representation, our GenHMM has tractability of exact loglikelihood of given sequential data, which is based on the change of variable formula.
 To make the loglikelihood computation efficient, neural network based generators of GenHMM are realized as flow models.
@@
 Therefore, we employ forward-backward algorithm \cite{Bishop:2006:PRM:1162264} to do the posterior computation efficiently.
-As we would detail in next section, what are needed to formulate the problem, are actually the $p(s| \ubm{x}; \bm{H})$ and $p(s, \kappa| \ubm{x}; \bm{H})$.
+As we would detail in the next section, what are needed to formulate the problem, are actually the $p(s| \ubm{x}; \bm{H})$ and $p(s, \kappa| \ubm{x}; \bm{H})$.
 For the joint posterior $p(s, \kappa| \ubm{x}; \bm{H})$, it can be computed by the Bayesian rule when posterior of hidden state is available.
@@
 \end{equation*}
-where $\bm{f}_{s,\kappa}^{[l]}$ is the $l$-th layer of $\bm{f}_{s,\kappa}$. We have $\bm{z}=\bm{f}_{s,\kappa}(\bm{x})$. If every layer of $\bm{g}_{s,\kappa}$ is invertible,
-the full feed-forward neural network is invertible. Flow model, proposed in \cite{DBLP:journals/corr/DinhKB14} as an image generating model, is such an invertible
-feed-forward neural network.
+where $\bm{f}_{s,\kappa}^{[l]}$ is the $l$-th layer of $\bm{f}_{s,\kappa}$.
+We have $\bm{z}=\bm{f}_{s,\kappa}(\bm{x})$.
+If every layer of $\bm{g}_{s,\kappa}$ is invertible, the full feed-forward neural network is invertible.
+Flow model, proposed in \cite{DBLP:journals/corr/DinhKB14} as an image generating model, is such an invertible feed-forward neural network.
 It is further improved in subsquential works \cite{2016arXiv160508803D,kingma2018glow} for high-fidelity and high-resolution image generating and representation.
-Flow model also has advantages as efficient Jacobian computation and low computational complexity.
-%Computation of inverse mapping for flow-based network and determinant of Jacobin were addressed in the literature.
+As shown in \eqref{eq:changel-variable}, the challenge lies at the computation of Jacobin determinant.
+Flow model has advantages as efficient Jacobian computation and low computational complexity.
+\textcolor{blue}{Another track of flow model uses a continuous -depth model and use an ordinary differential equation (ODE) specified by a neural network to define the change of variable \cite{NIPS2018_7892}, where the key is to solve the ODE problem.}
 
@@
 
-Our experiments are performed for: i) standard classification tasks (table~\ref{tab:acc-classification39f},\ref{tab:acc-classification13f}), ii) classification under noise perturbation (table~\ref{tab:acc-classification39f_noise_snr}, \ref{tab:acc-classification39f_noise_type}).
+Our experiments are performed for: i) standard classification tasks (Table~\ref{tab:acc-classification39f_a}, \ref{tab:acc-classification39f_b}, \ref{tab:acc-classification13f_a}, \ref{tab:acc-classification13f_b}), ii) classification under noise perturbation (table~\ref{tab:acc-classification39f_noise_snr}, \ref{tab:acc-classification39f_noise_type}).
 In all experiments, generators $\left\{\bm{g}_{s,\kappa}|s\in \Ss, \kappa=1, 2, \cdots, K \right\}$ of GenHMM are implemented as flow models.
@@
 \begin{table}
-  \caption{Test accuracy table for $39$ dimensional features }\label{tab:acc-classification39f}
-  \vspace{-0.3cm}
-  \centering
-  \subcaption{Folded $39$ Phonemes}\label{tab:acc-classification39f_a}
-  \vspace{-0.1cm}
+  \caption{Test accuracy table for $39$ dimensional features and folded $39$ Phonemes }\label{tab:acc-classification39f_a}
   \centering  
@@
   \end{tabular}
-  \vspace{0.1cm}
-  \subcaption{$61$ Phonemes}\label{tab:acc-classification39f_b}
-  \vspace{-0.1cm}
-  % \centering  
+\end{table}
+
+\begin{table}
+  \caption{Test accuracy table for $39$ dimensional features and $61$ Phonemes}\label{tab:acc-classification39f_b}
+  \centering  
   \begin{tabular}{l|c|c|c} \toprule
@@
 
-Table \ref{tab:acc-classification39f} shows the results for this experiments, in which we test both the folded $39$-phoneme classification case (the conventional way) in table \ref{tab:acc-classification39f_a} and the $61$-phoneme classification case in table \ref{tab:acc-classification39f_b}.
+Table \ref{tab:acc-classification39f_a} and \ref{tab:acc-classification39f_b} shows the results for this experiments, in which we test both the folded $39$-phoneme classification case (the conventional way) in table \ref{tab:acc-classification39f_a} and the $61$-phoneme classification case in table \ref{tab:acc-classification39f_b}.
 As shown in both $61$-phoneme and $39$-phoneme cases, GenHMM gets significant higher accuracy than GMM-HMM for the same number of mixture components.
@@
 \begin{table}
-  \caption{Test accuracy table for $13$ dimensional features }
-  \label{tab:acc-classification13f}
-  \vspace{-0.3cm}
-  \centering
-  \subcaption{Folded $39$ Phonemes}\label{tab:acc-classification13f_a}
-  \vspace{-0.1cm}
+  \caption{Test accuracy table for $13$ dimensional features and Folded $39$ Phonemes}\label{tab:acc-classification13f_a}
   \centering  
@@
   \end{tabular}
-  \vspace{0.1cm}
-  \subcaption{$61$ Phonemes}\label{tab:acc-classification13f_b}
-  \vspace{-0.1cm}
+\end{table}
+\begin{table}
+  \caption{Test accuracy table for $13$ dimensional features and $61$
+    Phonemes}\label{tab:acc-classification13f_b}
+  \centering
   \begin{tabular}{l|c|c|c}
@@
   \end{tabular}
-  \vspace{0.2cm}
 \end{table}
 
-Similar experiments in table~\ref{tab:acc-classification13f} are carried out by using only the MFCC coefficients as feature input (excluding deltas and delta-deltas).
-The results are shown in table~\ref{tab:acc-classification13f}.
+Similar experiments are carried out by using only the MFCC coefficients as feature input (excluding deltas and delta-deltas).
+The results are shown in table in Table~\ref{tab:acc-classification13f_a} and \ref{tab:acc-classification13f_b}.
 The superior performance of GenHMM remains compared with reference model GMM-HMM.
-The gain by using mixture generators is also presented in this set of experiments while the difference between $61$-phoneme and $39$-phoneme cases is similar to the set of experiments in table~\ref{tab:acc-classification39f}.
+The gain by using mixture generators is also presented in this set of experiments while the difference between $61$-phoneme and $39$-phoneme cases is similar to the set of experiments in Table~\ref{tab:acc-classification39f_a} and \ref{tab:acc-classification39f_b}.
 
@@
 The noise data is from NOISEX-92 database.
-The baseline of these two sets of experiments is the accuracy testing of GenHMM and GMM-HMM on clean testing data in the same experimental condition, where GenHMM has $77.7\%$ and GMM-HMM gets $68.8\%$ as shown in table~\ref{tab:acc-classification39f}.
+The baseline of these two sets of experiments is the accuracy testing of GenHMM and GMM-HMM on clean testing data in the same experimental condition, where GenHMM has $77.7\%$ and GMM-HMM gets $68.0\%$ as shown in table~\ref{tab:acc-classification39f_a}.
 It is shown in table~\ref{tab:acc-classification39f_noise_snr} that GMM-HMM's performance degenerates more than GenHMM's performance at the same level of noise perturbation, though the accuracy of both models increases along the increase of SNR.
@@
 \bibliography{myref}
-\bibliographystyle{aaai}
+
 \end{document}

### 2019-11-14 manuscript/main.tex
revising consider comments

@@
 Research attempts at training HMM with neural networks have been made to boost the modeling capacity of HMM.
-A successful work has brought restrictive Boltzmann machine (RBM) \cite{Hinton2012} into HMM based models \cite{hinton2012deepSpeech,li2013hybrid,Miao2013ImprovingLC}.
-In RBM based HMM, a hierarchical learning scheme is used: i) RBMs are trained one after the other in unsupervised fashion, and are stacked together as one model, ii) then a final softmax layer is added to the stack of RBMs to represent the probability of a HMM state given a data input, iii) a discriminative training is performed for the final tuning of the model.
-\textcolor{red}{explain more here..}
+A successful work of this track has brought deep neural network (DNN) that is defined by restrictive Boltzmann machines (RBMs) \cite{Hinton2012} into HMM based models \cite{hinton2012deepSpeech,li2013hybrid,Miao2013ImprovingLC}.
+RBM based HMM is trained with a hierarchical scheme consisting of multiple-step unsupervised learning, formatting of a classification network and then supervised learning.
+The hierarchical procedure comes from the empirical expertise in this domain.
+To be more specific, the hierarchical learning scheme of RBM/DNN based HMM consists of: i) RBMs are trained one after the other in unsupervised fashion, and are stacked together as one deep neural network model, ii) then a final softmax layer is added to the stack of RBMs to represent the probability of a HMM state given a data input, iii) a discriminative training is performed for the final tuning of the model at the final stage.
 
@@
 A HMM $\bm{H}$ defined in a hypothesis space $\Hh$, i.e.
-$\bm{H} \in \Hh$, is capable to model time-span signal $\ubar{\bm{x}} = \left[ \bm{x}_1, \cdots, \bm{x}_T\right]^{\intercal}$, where $\bm{x}_t\in \RR^{N}$ is the signal at time $t$, $[\cdot]^{\intercal}$ denotes transpose, and $T$ denotes the time length\footnote{The length for  sequential data varies.}.
+$\bm{H} \in \Hh$, is capable to model time-span signal $\ubar{\bm{x}} = \left[ \bm{x}_1, \cdots, \bm{x}_T\right]^{\intercal}$, where $\bm{x}_t\in \RR^{N}$ is the $N$-dimensional signal at time $t$, $[\cdot]^{\intercal}$ denotes transpose, and $T$ denotes the time length\footnote{The length for  sequential data varies.}.
 We define the hypothesis set of HMM as $\Hh := \{\bm{H} | \bm{H}=\{\Ss, \bm{q}, \bm{A}, p(\bm{x}|{s}; \bm{\Phi}_{s})\}\}$, where
@@
 the dataset of this distribution.
-For the given dataset, we denote its empirical distribution by $\hat{p}(\ubm{x}) = \frac{1}{R}\sum_{r=1}^{R} \delta_{\ubmr{x}{r}}(\ubm{x})$, where superscipt $(\cdot)^{r}$ denotes the index of sequential signal.
+For the given dataset, we denote its empirical distribution by $\hat{p}(\ubm{x}) = \frac{1}{R}\sum_{r=1}^{R} \delta_{\ubmr{x}{r}}(\ubm{x})$, where $R$ denotes the total number of sequential samples and superscipt $(\cdot)^{r}$ denotes the index of $r$-th sequential signal.
 The KL divergence minimization problem can be reduced to a likelihood maximization problem
@@
 If every layer of $\bm{g}_{s,\kappa}$ is invertible, the full feed-forward neural network is invertible.
-Flow model, proposed in \cite{DBLP:journals/corr/DinhKB14} as an image generating model, is such an invertible feed-forward neural network.
+Flow model, proposed in \cite{DBLP:journals/corr/DinhKB14} as an image generating model, is such an invertible feed-forward layer-wise neural network.
 It is further improved in subsquential works \cite{2016arXiv160508803D,kingma2018glow} for high-fidelity and high-resolution image generating and representation.
 As shown in \eqref{eq:changel-variable}, the challenge lies at the computation of Jacobin determinant.
-Flow model has advantages as efficient Jacobian computation and low computational complexity.
-\textcolor{blue}{Another track of flow model uses a continuous -depth model and use an ordinary differential equation (ODE) specified by a neural network to define the change of variable \cite{NIPS2018_7892}, where the key is to solve the ODE problem.}
+{Another track of flow models uses a continuous-depth models instead.
+The variable change is defined by an ordinary differential equation implemented by a neural network \cite{NIPS2018_7892, DBLP:grathwohl2018FFJORD}, where the key becomes to solve the ODE problem.
+We use the layer-wise flow model to model the variable change in \eqref{eq:changel-variable} in which the efficient Jacobian computation is available.}
 

### 2019-11-14 manuscript/main.tex
modify table
minor change

@@
 \begin{table}
-  \caption{Test accuracy table for $39$ dimensional features and folded $39$ Phonemes }\label{tab:acc-classification39f_a}
+  \caption{Test accuracy table for $39$ dimensional features and folded $39$ phonemes.}\label{tab:acc-classification39f_a}
   \centering  
@@
 \begin{table}
-  \caption{Test accuracy table for $39$ dimensional features and $61$ Phonemes}\label{tab:acc-classification39f_b}
+  \caption{Test accuracy table for $39$ dimensional features and $61$ phonemes.}\label{tab:acc-classification39f_b}
   \centering  
@@
 \begin{table}
-  \caption{Test accuracy table for $13$ dimensional features and Folded $39$ Phonemes}\label{tab:acc-classification13f_a}
+  \caption{Test accuracy table for $13$ dimensional features and folded $39$ phonemes.}\label{tab:acc-classification13f_a}
   \centering  
@@
   \caption{Test accuracy table for $13$ dimensional features and $61$
-    Phonemes}\label{tab:acc-classification13f_b}
+    phonemes.}\label{tab:acc-classification13f_b}
   \centering

### 2019-11-16 bin/aggregate_accuracy.py
include detail metrics in accuracy aggregating

@@ -7,7 +7,7 @@ from gm_hmm.src.ref_hmm import GaussianHMMclassifier
 import numpy as np
 from parse import parse
 from gm_hmm.src.utils import divide, acc_str, append_class, parse_, accuracy_fun
-
+from gm_hmm.src.eval_utils import write_eval_line, to_onehot
 import torch
 from torch.utils.data import DataLoader
 
@@ -53,17 +53,37 @@ if __name__ == "__main__":
 
     # rebuild information
     mdl_file = os.path.join(models_dir,"epoch{}.mdl".format(epoch))
+    out_results = os.path.join(models_dir,"epoch{}.report".format(epoch))
     data_files = np.array([[append_class(training_data_file, iclass+1), append_class(testing_data_file, iclass+1)]
                    for iclass in range(nclasses)])
 
     # Read results
+    def parse_out_list(out_list_str):
+        return [list(map(np.float64, x.split(","))) for x in out_list_str.split(";")]
+    
     def file2str(fname):
-        with open(fname,"r") as f:
-            line = f.read().strip().split("\n")[0]
-        _, train_res, test_res = parse("{}accc train: {} test: {}", line)
-        return train_res, test_res
+        with open(fname, "r") as f:
+            lines = f.read().strip().split("\n")
+        _, train_res, test_res = parse("{}accc train: {} test: {}", lines[0])
+        _, raw_train, raw_test = parse("{}llh train: {} test:{}", lines[1])
+        return train_res, test_res, parse_out_list(raw_train), parse_out_list(raw_test)
 
-    results = np.array([list(file2str(x)) for x in accc_files])
+    out = [list(file2str(x)) for x in accc_files]
+    userdata = [o[2:] for o in out]
+
+    mdl_userdata = {"xtrain_hat": np.concatenate([np.array(userdata[i][0]) for i in range(nclasses)]),
+             "ttrain": np.concatenate([i*np.ones(len(userdata[i][0])) for i in range(nclasses)]),
+             "xtest_hat": np.concatenate([np.array(userdata[i][1]) for i in range(nclasses)]),
+             "ttest": np.concatenate([i * np.ones(len(userdata[i][1])) for i in range(nclasses)]),
+             }
+    mdl_userdata["ttest"] = to_onehot(mdl_userdata["ttest"])
+    mdl_userdata["ttrain"] = to_onehot(mdl_userdata["ttrain"])
+
[…]

### 2019-11-16 bin/compute_accuracy_class.py

@@ -72,7 +72,10 @@ if __name__ == "__main__":
    
     #f = lambda x: divide(parse_(accuracy_fun(x, mdl=mdl)))
     
-    results = list(map(f, data_files))
+    results_ = list(map(f, data_files))
+    results = [r[0] for r in results_]
+    userdata = [r[1] for r in results_]
 
     print("epoch: {} class: {} accc train: {} test: {}".format(epoch, iclass, results[0], results[1]), file=sys.stdout)
+    print("epoch: {} class: {} llh train: {} test: {}".format(epoch, iclass, userdata[0], userdata[1]), file=sys.stdout)
     sys.exit(0)

### 2019-11-16 src/utils.py

@@ -158,7 +158,7 @@ def accuracy_fun(data_file, mdl=None):
     class_hat = np.argmax(out, axis=0) + 1
     istrue = class_hat == int(true_class)
     print(data_file, "Done ...", "{}/{}".format(str(istrue.sum()), str(istrue.shape[0])), file=sys.stderr)
-    return "{}/{}".format(str(istrue.sum()), str(istrue.shape[0]))
+    return "{}/{}".format(str(istrue.sum()), str(istrue.shape[0])), format_out_list(out_list)
 
 
 def accuracy_fun_torch(data_file, mdl=None, batch_size_=128):
@@ -185,12 +185,14 @@ def accuracy_fun_torch(data_file, mdl=None, batch_size_=128):
     class_hat = torch.argmax(out, dim=0) + 1
     print(data_file, "Done ...", "{}".format(acc_str(class_hat, true_class)), file=sys.stderr)
 
-    return acc_str(class_hat, true_class)
+    return acc_str(class_hat, true_class), format_out_list(out.cpu().numpy().T.tolist())
 
 def acc_str(class_hat, class_true):
     istrue = class_hat == int(class_true)
     return "{}/{}".format(str(istrue.sum().cpu().numpy()), str(istrue.shape[0]))
 
+def format_out_list(out_list):
+    return ";".join([",".join(list(map(str, o))) for o in out_list])
 
 def test_acc_str():
     class_hat = torch.FloatTensor([1, 2])

### 2019-11-16 src/eval_utils.py
utils functions for extra evaluation

@@ -1,0 +1,91 @@
+from sklearn.metrics import mean_squared_error, accuracy_score,\
+                            roc_auc_score, precision_score,\
+                            recall_score, average_precision_score
+from parse import parse
+from sklearn.preprocessing import OneHotEncoder
+import numpy as np
+from functools import partial
+import subprocess
+
+
+eval_metrics = {"MSE": lambda y_true, y_hat: mean_squared_error(y_true, y_hat),
+                "Acc": lambda y_true, y_hat: accuracy_score(y_true, cont_to_binary(y_hat,n=y_true.shape[1])),
+                "AUROC": lambda y_true, y_hat: roc_auc_score(y_true, y_hat),
+                "p": lambda y_true, y_hat: partial(precision_score, average="macro")(y_true, cont_to_binary(y_hat,n=y_true.shape[1])),
+                "r": lambda y_true, y_hat: partial(recall_score, average="macro")(y_true, cont_to_binary(y_hat,n=y_true.shape[1])),
+                "avgPr": lambda y_true, y_hat: partial(average_precision_score, average="macro")(y_true, y_hat)
+                }
+
+
+
+
+def githash():
+    """Get the current git commit short hash."""
+    return subprocess.check_output(["git", "rev-parse", "--short", "HEAD"]).strip().decode("utf-8")
+
+def to_onehot(Y,n=None):
+    enc = OneHotEncoder(handle_unknown='ignore')
+    if n is None:
+        enc.fit(np.unique(Y).reshape(-1, 1))
+    else:
+        enc.fit(np.arange(n).reshape(-1,1))
+
+    Y_ = enc.transform(Y[:, np.newaxis]).toarray()
+    return Y_
+
+def cont_to_binary(y,n=None):
+    return to_onehot(np.argmax(y, axis=1),n=n)
+
+
+def test_cont_to_binary():
+    y = np.array([[-1, -2], [-1000, -999], [0.3, 0.7]])
+    assert ((cont_to_binary(y) == np.array([[1, 0], [0, 1], [0, 1]])).all())
+
+
+def test_format_score():
+    assert(format_score("abc", 1.23456789) == "abc:1.23457")
+
+
+def format_score(k, s):
+    return "{}:{}".format(k, round(s, 5))
+
+
+def test_parse_score():
+    assert(parse_score("abc:1.23457") == ["abc", 1.23457])
+
+
+def parse_score(s_str):
[…]

### 2019-11-17 manuscript/main.tex
minor change
include precision and f1 in experiment section

@@
 
-%\ecaisubmission   % inserts page numbers. Use only for submission of paper.
+% \ecaisubmission   % inserts page numbers. Use only for submission of paper.
 % Do NOT use for camera-ready version of paper.
@@
 This hierarchical training procedure is motivated by the intuition of using LSTM or RNN to fill in the gap where HMM can not learn.
-\textcolor{red}{explain more here...}
 
@@
 \item Giving convergence analysis for GenHMM under the proposed learning algorithm.
-\item Verifying the efficiency of the proposed model on practical sequential data.
+\item Verifying the proposed model on practical sequential data.
 \end{itemize}
@@
 \section{Experiments}
-To show the validity of our model, we implement our model in PyTorch and test it with sequential data.
-The code is available at github repository\footnote{Due to double-blind review, the code is submitted in supplementary file.}.
-We first discuss the experimental setups and then show the experimental results.
+To show the validity of our model, we implement our model in PyTorch and test it with sequential data. We first discuss the experimental setups and then show the experimental results.
 
@@
 Our experiments are performed for: i) standard classification tasks (Table~\ref{tab:acc-classification39f_a}, \ref{tab:acc-classification39f_b}, \ref{tab:acc-classification13f_a}, \ref{tab:acc-classification13f_b}), ii) classification under noise perturbation (table~\ref{tab:acc-classification39f_noise_snr}, \ref{tab:acc-classification39f_noise_type}).
+The criterion used to report the results includes accuracy, precision and F1 score.
 In all experiments, generators $\left\{\bm{g}_{s,\kappa}|s\in \Ss, \kappa=1, 2, \cdots, K \right\}$ of GenHMM are implemented as flow models.
@@
   \centering  
-  \begin{tabular}{l|c|c|c}
+  \begin{tabular}{l|l|c|c|c}
     \toprule
-    {Model} &  K=1 &  K=3 &  K=5  \\  \midrule
-    GMM-HMM & $62.3\%$ &  $68.0\%$ &  $68.7\%$  \\ \midrule
-    GenHMM & $76.7\%$ & $77.7\%$ &  $77.7\%$ \\ 
+    {Model} & Criterion &  K=1 &  K=3 &  K=5  \\  \midrule
+    \multirow{3}{*}{GMM-HMM}
+            & Accuracy    & $62.3\%$ &  $68.0\%$ &  $68.7\%$  \\
+            & {Precision} & $67.9\%$ &  $72.6\%$ &  $73.0\%$  \\
+            & {F1}        & $63.7\%$ &  $69.1\%$ &  $69.7\%$ \\
+    \midrule
+    \multirow{3}{*}{GenHMM}
+            & Accuracy    & $76.7\%$   & $77.7\%$ &  $77.7\%$ \\ 
+            & {Precision} & $76.9\%$   & $78.1\%$ &  $78.0\%$ \\
+            & {F1}        & $76.1\%$   & $77.1\%$ &  $77.0\%$\\
     \bottomrule                                                                  
@@
   \centering  
-  \begin{tabular}{l|c|c|c} \toprule
-    {Model} &  K=1 &  K=3 &  K=5
+  \begin{tabular}{l|l|c|c|c} \toprule
+    {Model} & Criterion & K=1 &  K=3 &  K=5
     \\ \midrule
-    GMM-HMM & $53.6\%$ &  $59.6\%$ & $61.9\%$  \\ \midrule
-    GenHMM & $69.5\%$ & $70.6\%$ & $70.7\%$   \\ 
+    \multirow{4}{*}{GMM-HMM}
+            & Accuracy    & $53.6\%$ &  $59.6\%$ & $61.9\%$  \\
+            & {Precision} & $59.1\%$ &  $63.9\%$ & $65.7\%$ \\
+            & {F1}        & $54.7\%$ &  $60.5\%$ & $62.7\%$\\
+    \midrule
+    \multirow{3}{*}{GenHMM}
+            & Accuracy    & $69.5\%$ & $70.6\%$ & $70.7\%$   \\
+            & {Precision} & $69.2\%$ & $70.5\%$ & $71.0\%$ \\
+            & {F1}        & $68.6\%$ & $69.6\%$ & $69.6\%$\\
     \bottomrule
@@
 As shown in both $61$-phoneme and $39$-phoneme cases, GenHMM gets significant higher accuracy than GMM-HMM for the same number of mixture components.
+The comparisons with regarding to precision and F1 scores show similar trends and also demonstrate significant improvement of GenHMM's performance.
 As our expectation, GenHMM should have better modeling capacity of sequential data since we bring in the neural network based generators into GenHMM, which should be able to represent complex relationship between states of HMM and sequential data.
@@
   \centering  
-  \begin{tabular}{l|c|c|c}
+  \begin{tabular}{l|l|c|c|c}
     \toprule
-    {Model} &  K=1 &  K=3 &  K=5  \\  \midrule
-    GMM-HMM & $48.5\%$ &  $51.2\%$ &  $52.4\%$  \\ \midrule
-    GenHMM & $61.1\%$ & $62.1\%$ & $62.1\%$   \\ 
+    {Model} & Criterion & K=1 &  K=3 &  K=5  \\  \midrule
+    \multirow{3}{*}{GMM-HMM}
+            & Accuracy & $48.5\%$ &  $51.2\%$ &  $52.4\%$  \\
+            & Precision& $56.2\%$ &  $58.3\%$ &  $59.5\%$  \\
+            & F1       & $50.3\%$ &  $53.0\%$ &  $54.2\%$  \\
+    \midrule
+    \multirow{3}{*}{GenHMM}
+            & Accuracy & $61.1\%$ &  $62.1\%$ & $62.1\%$   \\ 
+            & Precision& $61.1\%$ &  $61.9\%$ &  $62.1\%$  \\
+            & F1       & $59.7\%$ &  $60.7\%$ &  $60.2\%$  \\
+
     \bottomrule                                                                  
@@
 \end{table}
+
 \begin{table}
@@
   \centering
-  \begin{tabular}{l|c|c|c}
+  \begin{tabular}{l|l|c|c|c}
     \toprule
-    {Model} &  K=1 &  K=3 &  K=5
-    \\ \midrule
-    GMM-HMM & $37.1\%$ &  $40.6\%$ & $42.2\%$  \\ \midrule
-    GenHMM & $50.3\%$ & $51.0\%$ & $52.3\%$   \\ 
+    {Model} & Criterion &  K=1 &  K=3 &  K=5 \\ \midrule
+    \multirow{3}{*}{GMM-HMM}
+            & Accuracy & $37.1\%$ &  $40.6\%$ & $42.2\%$  \\
+            &Precision & $44.6\%$ &  $47.4\%$ & $48.8\%$  \\
+            & F1       & $38.8\%$ &  $42.1\%$ & $43.7\%$ \\
+    \midrule
+    \multirow{3}{*}{GenHMM}
+            & Accuracy & $50.3\%$ & $50.8\%$ & $52.3\%$   \\
+            & Precision& $49.3\%$ & $50.9\%$ & $52.1\%$ \\
+            &F1        & $47.8\%$ & $48.3\%$ & $49.3\%$ \\
     \bottomrule
@@
   \centering
-  \begin{tabular}{l|c|c|c|c}
+  \begin{tabular}{l|l|c|c|c|c}
     \toprule
-    {SNR} &  15dB &  20dB &  25dB & 30 dB  \\  \midrule
-    GMM-HMM & $36.6\%$ &  $44.2\%$ &  $50.8\%$ & $57.1\%$  \\ \midrule
-    GenHMM & $52.4\%$ & $62.0\%$ &  $69.7\%$ & $74.3\%$ \\ 
+    \multirow{2}{*}{Model} & \multirow{2}{*}{Criterion} &
+                              \multicolumn{4}{c}{White Noise SNR} \\
+    \cline{3-6}
+                          && 15dB &  20dB &  25dB & 30 dB  \\
+    \midrule
+    \multirow{3}{*}{GMM-HMM}
+            & Accuracy & $36.6\%$ &  $44.2\%$ &  $50.8\%$ & $57.1\%$
+ \\
+            &Precision & $59.2\%$ &  $64.2\%$ &  $68.4\%$ & $70.6\%$  \\
+            & F1       & $39.9\%$ &  $47.7\%$ &  $53.9\%$ & $59.9\%$ \\
+    \midrule
+    \multirow{3}{*}{GenHMM}
+            & Accuracy & $52.4\%$ & $62.0\%$ &  $69.7\%$ & $74.3\%$ \\
+            &Precision & $60.0\%$ &  $65.9\%$ &  $71.7\%$ & $74.8\%$  \\
+            & F1       & $52.5\%$ &  $62.0\%$ &  $69.3\%$ & $73.5\%$ \\
     \bottomrule                                                                  
@@
   \centering
-  \begin{tabular}{l|c|c|c|c}
+  \begin{tabular}{l|l|c|c|c|c}
     \toprule
-    {Noise Type} &  White &  Pink &  Babble & Volvo  \\  \midrule
-    GMM-HMM & $44.2\%$ &  $48.8\%$ &  $57.7\%$ & $66.6\%$  \\ \midrule
-    GenHMM & $62.0\%$ & $65.1\%$ &  $70.0\%$ & $75.7\%$ \\ 
+    \multirow{2}{*}{Model} & \multirow{2}{*}{Criterion} &
+                              \multicolumn{4}{c}{Noise Type} \\
+    \cline{3-6}
+
+                           &  &  White &  Pink &  Babble & Volvo  \\
+    \midrule
+    \multirow{3}{*}{GMM-HMM}
+            & Accuracy & $44.2\%$ &  $48.8\%$ &  $57.7\%$ & $66.6\%$
+    \\
+            &Precision & $64.2\%$ &  $66.1\%$ &  $67.0\%$ & $71.9\%$  \\
+            & F1       & $47.7\%$ &  $52.3\%$ &  $59.7\%$ & $67.8\%$ \\
+    \midrule
+    \multirow{3}{*}{GenHMM}
+            & Accuracy & $62.0\%$ &  $65.1\%$ &  $70.0\%$ & $75.7\%$ \\
+            &Precision & $65.9\%$ &  $67.8\%$ &  $70.4\%$ & $75.9\%$  \\
+            & F1       & $62.0\%$ &  $64.6\%$ &  $69.0\%$ & $75.3\%$ \\
     \bottomrule                                                                  
@@
 The baseline of these two sets of experiments is the accuracy testing of GenHMM and GMM-HMM on clean testing data in the same experimental condition, where GenHMM has $77.7\%$ and GMM-HMM gets $68.0\%$ as shown in table~\ref{tab:acc-classification39f_a}.
+Similar superior performance of GenHMM with regarding to precision and F1 scores is also shown.
 It is shown in table~\ref{tab:acc-classification39f_noise_snr} that GMM-HMM's performance degenerates more than GenHMM's performance at the same level of noise perturbation, though the accuracy of both models increases along the increase of SNR.

### 2019-11-18 manuscript/main.tex
update content

@@
 A successful work of this track has brought deep neural network (DNN) that is defined by restrictive Boltzmann machines (RBMs) \cite{Hinton2012} into HMM based models \cite{hinton2012deepSpeech,li2013hybrid,Miao2013ImprovingLC}.
-RBM based HMM is trained with a hierarchical scheme consisting of multiple-step unsupervised learning, formatting of a classification network and then supervised learning.
+RBM based HMM is trained with a hierarchical scheme consisting of multiple steps of unsupervised learning, formatting of a classification network and then supervised learning.
 The hierarchical procedure comes from the empirical expertise in this domain.
@@
 Specifically, a generative model in our GenHMM is generator-mixed, where a generator is realized by a neural network to help the model gain high modeling capacity.
-Our proposed model, GenHMM:
+Our proposed model, GenHMM,
 \begin{itemize}
@@
 
-Our contributions in the paper are as follows:
+Our contributions in the paper are as follows.
 \begin{itemize}
@@
 \item $\Ss$ is the set of hidden states of $\bm{H}$.
-\item $\bm{q} = \left[ q_1, q_2, \cdots, q_{|\Ss|}\right]^\intercal$ is the initial distribution of $\bm{H}$ with $|\Ss|$ as cardinality of $\Ss$.
+\item $\bm{q} = \left[ q_1, q_2, \cdots, q_{|\Ss|}\right]^\intercal$ is the initial state distribution of $\bm{H}$ with $|\Ss|$ as cardinality of $\Ss$.
 For $i \in \Ss$, $q_i = p(s_{1}=i;\bm{H})$.
 We use $s_t$ to denote the state $s$ at time $t$.
-\item $\bm{A}$ matrix of size $|\Ss| \times |\Ss|$ is the transition matrix of $\bm{H}$. That is, $\forall i, j \in \Ss$,  $\bm{A}_{i,j} = p(s_{t+1}=j|s_{t}=i; \bm{H})$.
+\item $\bm{A}$ matrix of size $|\Ss| \times |\Ss|$ is the transition matrix of states in $\bm{H}$. That is, $\forall i, j \in \Ss$,  $\bm{A}_{i,j} = p(s_{t+1}=j|s_{t}=i; \bm{H})$.
 \item For a given hidden state $s$, the density function of the observable signal is $p({\bm{x}}|{s};\bm{\Phi}_{s})$, where $\bm{\Phi}_{s}$ is the parameter set that defines this probabilistic model.
@@
 In this section, we introduce the neural network based state probabilistic model of our GenHMM.
-Recall that $\bm{x}\in\mathbb{R}^N$.
+Recall that $\bm{x}\in\mathbb{R}^N$. Subscript is omitted when it does not cause ambiguity.
 The probabilistic model of GenHMM for each hidden state is a mixture of $K$ neural network based generators, where $K$ is a positive integer.
@@
 the M-step.
-Assume model $\bm{H}$ is at a configuration of $\bm{H}^{\mathrm{old}}$, we formulate these two steps as follows:
+Assume model $\bm{H}$ is at a configuration of $\bm{H}^{\mathrm{old}}$, we formulate these two steps as follows.
 \begin{itemize}
@@
 \end{align}
-where
+where the decomposed optimization problems are
 \begin{align}
@@
     &= \EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p({s}_{1};\bm{H})  \right], \label{eq:init-distribution-update}\\
-  \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})}\hspace{-0.1cm}\left[ \sum_{t=1}^{T}\log\,p({s}_{t+1}|{s}_{t}; \bm{H}) \right], \label{eq:transition-update}\\
+  \Qq(\bm{A}; \bm{H}^{\mathrm{old}}) &=\EE_{\hat{p}(\ubm{x}),p(\ubm{s}| \ubm{x}; \bm{H}^{\mathrm{old}})}\hspace{-0.1cm}\left[ \sum_{t=1}^{T-1}\log\,p({s}_{t+1}|{s}_{t}; \bm{H}) \right], \label{eq:transition-update}\\
   \Qq(\bm{\Phi}; \bm{H}^{\mathrm{old}}) &= \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})} \left[ \log\,p(\ubm{x}, \ubm{\kappa}| \ubm{s}; \bm{H}) \right].
@@
 \begin{equation}
-  \lambda_s = \sum_{\kappa=1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{(r)}} p(\smtr{s}{t}{r}=s, \smtr{\kappa}{t}{r} =\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}).
+  \lambda_s = \sum_{\kappa=1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}} p(\smtr{s}{t}{r}=s, \smtr{\kappa}{t}{r} =\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}).
 \end{equation}
@@
 \begin{equation}\label{eq:mix-latent-parameter-solution}
-  \pi_{s, \kappa} = \frac{\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }{\sum_{k =1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}-1} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=k | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) },
+  \pi_{s, \kappa} = \frac{\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=\kappa | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) }{\sum_{k =1}^{K}\sum_{r=1}^{R} \sum_{t=1}^{{T}^{r}} p(\smtr{s}{t}{r} =s, \smtr{\kappa}{t}{r}=k | \ubmr{x}{r}; \bm{H}^{\mathrm{old}}) },
 \end{equation}
-where $p(s, \kappa | \ubm{x}; \bm{H}^{\mathrm{old}}) = p(s| \ubm{x}; \bm{H}^{\mathrm{old}}) p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$.
-$p(s| \ubm{x}; \bm{H}^{\mathrm{old}})$ can be computed by forward-backward algorithm, while $p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$ is given by \eqref{eq:kappa-posterior}.
+where
+\begin{equation}
+  p(s, \kappa | \ubm{x}; \bm{H}^{\mathrm{old}}) = p(s| \ubm{x}; \bm{H}^{\mathrm{old}}) p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}}).
+\end{equation}
+Here $p(s| \ubm{x}; \bm{H}^{\mathrm{old}})$ can be computed by forward-backward algorithm, while $p(\kappa | s, \ubm{x}; \bm{H}^{\mathrm{old}})$ is given by \eqref{eq:kappa-posterior}.
 
@@
 \begin{align}
-  \Qq(\bm{q}; \bm{H}^{\mathrm{old}})
+  &\Qq(\bm{q}; \bm{H}^{\mathrm{old}}) \nonumber \\
   &=\frac{1}{R} \sum_{r=1}^{R}\sum_{\ubmr{s}{r}} {p(\ubmr{s}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r};\bm{H}) \nonumber \\
-  % & = \frac{1}{R}\sum_{r=1}^{R}\sum_{\smtr{s}{1}{r}=1}^{|\Ss|}\sum_{\smtr{s}{2}{r}=1}^{|\Ss|}\cdots \sum_{\smtr{s}{T^{r}}{r}}^{{|\Ss|}} {p(\smtr{s}{1}{r}, \smtr{s}{2}{r}, \cdots, \smtr{s}{T^{r}}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r}) \\
+  & = \frac{1}{R}\sum_{r=1}^{R}\sum_{\smtr{s}{1}{r}=1}^{|\Ss|}\sum_{\smtr{s}{2}{r}=1}^{|\Ss|}\cdots \sum_{\smtr{s}{T^{r}}{r}}^{{|\Ss|}} {p(\smtr{s}{1}{r}, \smtr{s}{2}{r}, \cdots, \smtr{s}{T^{r}}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r}) \nonumber\\
   & = \frac{1}{R}\sum_{r=1}^{R}\sum_{\smtr{s}{1}{r}=1}^{|\Ss|}{p(\smtr{s}{1}{r}| \ubmr{x}{r}; \bm{H}^{\mathrm{old}})} \log\,p(\smtr{s}{1}{r};\bm{H}).
@@
     &\EE_{\hat{p}(\ubm{x})}\left[ \log\,p(\ubm{x};\bm{H}^{\mathrm{new}}) \right] \nonumber \\
-    \geq & \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x}; \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x}, \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{old}})}{p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right] \nonumber \\
+    \geq & \EE_{\hat{p}(\ubm{x}),p(\ubm{s},\ubm{\kappa}| \ubm{x};
+           \bm{H}^{\mathrm{old}})}\left[ \log\,\frac{p(\ubm{x},
+           \ubm{s}, \ubm{\kappa}; \bm{H}^{\mathrm{old}})}{p(\ubm{s},
+           \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})}\right]
+           \nonumber \\
+    &+ \EE_{\hat{p}(\ubm{x})}\left[ KL(p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})\|p(\ubm{s}, \ubm{\kappa}|\ubm{x}; \bm{H}^{\mathrm{old}})) \right]
+      \nonumber \\
     = & \EE_{\hat{p}(\ubm{x})}\left[ \log\,p(\ubm{x};\bm{H}^{\mathrm{old}}) \right].
@@
     \STATE {\bfseries Input:}{
-      Empirical distribution $\hat{p}(\bm{x})$ of dataset\\
-      Initializing $\bm{H}^{\mathrm{old}}, \bm{H} \in \Hh$ gives: \\
-      $\bm{H}^{\mathrm{old}} = \{\Ss, \bm{q}^{\mathrm{old}}, A^{\mathrm{old}}, p(\bm{x}|s; \bm{\Phi}_{s}^{\mathrm{old}})\}$ \\
-      $\bm{H} = \{\Ss, \bm{q}, A, p(\bm{x}|s; \bm{\Phi}_{s})\}$} \\
+      Empirical distribution $\hat{p}(\bm{x})$ of dataset}\\
+    \STATE Initializing $\bm{H}^{\mathrm{old}}, \bm{H} \in \Hh$ gives: \\
+    $\bm{H}^{\mathrm{old}} = \{\Ss, \bm{q}^{\mathrm{old}}, A^{\mathrm{old}}, p(\bm{x}|s; \bm{\Phi}_{s}^{\mathrm{old}})\}$, \\
+    $\bm{H} = \{\Ss, \bm{q}, A, p(\bm{x}|s; \bm{\Phi}_{s})\}$, \\
+    in which generators $\left\{\bm{g}_{s,\kappa}|s\in \Ss, \kappa=1,
+      2, \cdots, K \right\}$ are all initialized randomly.
     \STATE $\bm{H}^{\mathrm{old}} \gets \bm{H}$
@@
 the parameters of generators in GenHMM.
-As shown from line $6$ to $12$ in Algorithm~\ref{algo:genhmm}, the batch-size stochastic gradient decent can be naturally embedded into the learning algorithm of GenHMM.
+As shown from line $6$ to $10$ in Algorithm~\ref{algo:genhmm}, the batch-size stochastic gradient decent can be naturally embedded into the learning algorithm of GenHMM.
 
@@
 Our experiments are performed for: i) standard classification tasks (Table~\ref{tab:acc-classification39f_a}, \ref{tab:acc-classification39f_b}, \ref{tab:acc-classification13f_a}, \ref{tab:acc-classification13f_b}), ii) classification under noise perturbation (table~\ref{tab:acc-classification39f_noise_snr}, \ref{tab:acc-classification39f_noise_type}).
-The criterion used to report the results includes accuracy, precision and F1 score.
+The criterion used to report the results includes accuracy, precision and F1 scores.
 In all experiments, generators $\left\{\bm{g}_{s,\kappa}|s\in \Ss, \kappa=1, 2, \cdots, K \right\}$ of GenHMM are implemented as flow models.
@@
 In our experiments, each generator $\bm{g}_{s,\kappa}$ consists of four \textit{flow blocks}.
-The density of samples in the latent space is defined as Normal, i.e. $p_{s,\kappa}(\bm{z})$ is the density function of standard Gaussian.
+The density of samples in the latent space is defined as Normal,
+i.e. $p_{s,\kappa}(\bm{z})$ is the density function of standard
+Gaussian. The configuration for each generator is shown as Table~\ref{table:generator-setting}.
+
+\begin{table}
+  \caption{Configuration of generators in Experiments}\label{table:generator-setting}
+  \centering
+  \begin{tabular}{l|c}
+    \toprule
+    \makecell{Latent distribution $p_{s,\kappa}(\bm{z})$
+    \\$s\in \Ss, \kappa=1, 2, \cdots, K $} & Standard Gaussian \\
+    \hline
+    Number of flow blocks & $4$ \\
+    \hline
+    Non-linear mapping $\bm{m}_a$, $\bm{m}_b$ & \makecell{Multiple
+                                                layer perception, \\
+    $3$ layers and with hidden dimension $24$}\\
+    \bottomrule
+  \end{tabular}
+\end{table}
 
@@
   \end{tabular}
-  \vspace{0.2cm}
+  \vspace{0.5cm}
 \end{table}
@@
 
-Table \ref{tab:acc-classification39f_a} and \ref{tab:acc-classification39f_b} shows the results for this experiments, in which we test both the folded $39$-phoneme classification case (the conventional way) in table \ref{tab:acc-classification39f_a} and the $61$-phoneme classification case in table \ref{tab:acc-classification39f_b}.
+Table \ref{tab:acc-classification39f_a} and \ref{tab:acc-classification39f_b} shows the results for this experiments, in which we test both the folded $39$-phoneme classification case (the conventional way) in Table~\ref{tab:acc-classification39f_a} and the $61$-phoneme classification case in Table~\ref{tab:acc-classification39f_b}.
 As shown in both $61$-phoneme and $39$-phoneme cases, GenHMM gets significant higher accuracy than GMM-HMM for the same number of mixture components.
@@
 As our expectation, GenHMM should have better modeling capacity of sequential data since we bring in the neural network based generators into GenHMM, which should be able to represent complex relationship between states of HMM and sequential data.
-This is some increase of accuracy as total number of mixture components in GenHMM is increased from $K=1$ to $K=5$.
+There are increases of accuracy, precision and F1 scores as  number of mixture components in GenHMM is increased from $K=1$ to $K=5$.
 The sequential dependency of data is modeled by HMM itself, while each state of HMM can have better representation using a mixture probabilistic model if data represented by the state is multi-mode.
@@
   \end{tabular}
+  \vspace{0.5cm}
 \end{table}
@@
 Similar experiments are carried out by using only the MFCC coefficients as feature input (excluding deltas and delta-deltas).
-The results are shown in table in Table~\ref{tab:acc-classification13f_a} and \ref{tab:acc-classification13f_b}.
-The superior performance of GenHMM remains compared with reference model GMM-HMM.
+The results are shown in Table~\ref{tab:acc-classification13f_a} and \ref{tab:acc-classification13f_b}.
+The superior performance of GenHMM remains compared with reference model GMM-HMM, with regarding to accuracy, precision and F1 scores.
 The gain by using mixture generators is also presented in this set of experiments while the difference between $61$-phoneme and $39$-phoneme cases is similar to the set of experiments in Table~\ref{tab:acc-classification39f_a} and \ref{tab:acc-classification39f_b}.
@@
     \cline{3-6}
-                          && 15dB &  20dB &  25dB & 30 dB  \\
+
+                           && \makecell{15dB} &  \makecell{20dB} &  \makecell{25dB} & \makecell{30dB}  \\
     \midrule
@@
 We train GenHMM with $K=3$ by clean TIMIT training data in the case of folded $39$ phonemes with $39$ dimensional features.
-The testing dataset is perturbed by either the same type of noise with different signal-to-noise ratio (SNR) as shown in table~\ref{tab:acc-classification39f_noise_snr}, or different type of noises with the same SNR as shown in table~\ref{tab:acc-classification39f_noise_type}.
+The testing dataset is perturbed by either the same type of noise with different signal-to-noise ratio (SNR) as shown in Table~\ref{tab:acc-classification39f_noise_snr}, or different type of noises with the same SNR as shown in Table~\ref{tab:acc-classification39f_noise_type}.
 The noise data is from NOISEX-92 database.
-The baseline of these two sets of experiments is the accuracy testing of GenHMM and GMM-HMM on clean testing data in the same experimental condition, where GenHMM has $77.7\%$ and GMM-HMM gets $68.0\%$ as shown in table~\ref{tab:acc-classification39f_a}.
+The baseline of these two sets of experiments is the accuracy testing of GenHMM and GMM-HMM on clean testing data in the same experimental condition, where GenHMM has $77.7\%$ and GMM-HMM gets $68.0\%$ as shown in Table~\ref{tab:acc-classification39f_a}.
 Similar superior performance of GenHMM with regarding to precision and F1 scores is also shown.
-It is shown in table~\ref{tab:acc-classification39f_noise_snr} that GMM-HMM's performance degenerates more than GenHMM's performance at the same level of noise perturbation, though the accuracy of both models increases along the increase of SNR.
+It is shown in Table~\ref{tab:acc-classification39f_noise_snr} that GMM-HMM's performance degenerates more than GenHMM's performance at the same level of noise perturbation, though the accuracy of both models increases along the increase of SNR.
 Especially, for SNR=$30$dB, the accuracy of GenHMM drops only about $3\%$ (from $77.7\%$ to $74.3\%$), while GMM-HMM encounters more than $10\%$ decrease (from $68.0\%$ to $57.1\%$) due to the noise perturbation.
-In table~\ref{tab:acc-classification39f_noise_type}, the SNR remains constant and GenHMM is tested with perturbation of different noise types.
+In Table~\ref{tab:acc-classification39f_noise_type}, the SNR remains constant and GenHMM is tested with perturbation of different noise types.
 It is shown that GenHMM still has higher accuracy at different types of noise perturbations than GMM-HMM.

### 2019-11-18 src/eval_utils.py
expriment for score report

@@ -1,6 +1,6 @@
 from sklearn.metrics import mean_squared_error, accuracy_score,\
                             roc_auc_score, precision_score,\
-                            recall_score, average_precision_score
+                            recall_score, average_precision_score, f1_score
 from parse import parse
 from sklearn.preprocessing import OneHotEncoder
 import numpy as np
@@ -10,10 +10,11 @@ import subprocess
 
 eval_metrics = {"MSE": lambda y_true, y_hat: mean_squared_error(y_true, y_hat),
                 "Acc": lambda y_true, y_hat: accuracy_score(y_true, cont_to_binary(y_hat,n=y_true.shape[1])),
-                "AUROC": lambda y_true, y_hat: roc_auc_score(y_true, y_hat),
-                "p": lambda y_true, y_hat: partial(precision_score, average="macro")(y_true, cont_to_binary(y_hat,n=y_true.shape[1])),
-                "r": lambda y_true, y_hat: partial(recall_score, average="macro")(y_true, cont_to_binary(y_hat,n=y_true.shape[1])),
-                "avgPr": lambda y_true, y_hat: partial(average_precision_score, average="macro")(y_true, y_hat)
+                "AUROC": lambda y_true, y_hat: roc_auc_score(y_true, y_hat, "weighted"),
+                "p": lambda y_true, y_hat: partial(precision_score, average="weighted")(y_true, cont_to_binary(y_hat,n=y_true.shape[1])),
+                "r": lambda y_true, y_hat: partial(recall_score, average="weighted")(y_true, cont_to_binary(y_hat,n=y_true.shape[1])),
+                "avgPr": lambda y_true, y_hat: partial(average_precision_score, average="weighted")(y_true, y_hat),
+                "f1": lambda y_true, y_hat: partial(f1_score, average="weighted")(y_true, cont_to_binary(y_hat,n=y_true.shape[1]))
                 }
 
 
